---
title: "0_abares_species_states"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture
#Most figures done here are part of Marina Christofidis' PhD Thesis and helped to give shape to the Manuscript named Mapping Australia’s aquaculture footprint. 
date: "2025-10-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#install.packages("docxtractr")
```

#upload libraries
```{r}
library(tmaptools)
library(gridExtra)
library(cowplot)
library(tidyverse)
library(here)
library(tmap)
library(sf)
library(ggplot2)
library(dplyr)
library(docxtractr)
library(tidyr)
```
#upload the australian map, reproject and make valid and with states as well data from Australia and states was downloaded from the Australian Bureau of Statistics  https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs-edition-3/jul2021-jun2026/access-and-downloads/digital-boundary-files
```{r}
#upload, make valid and reproject data
Australia<-st_read(here("raw_data/Australia_country_SHP/AUS_2021_AUST_GDA2020.shp"))%>%
  st_transform(, 3577)%>%
  st_make_valid()

AUS_states<-st_read(here("raw_data/AustralianStates_SHP/STE_2021_AUST_GDA2020.shp"))%>%
  st_transform(, 3577)%>%
  st_make_valid()

```
#upload abares table that we did using the species data we colledcted and from the ABARES table

```{r}

# Read your file
Abares_doc <- read_docx(here("raw_data/Aquaculture_Production_ABARES_DAFF/table_states_species_aquaculture_Australia.docx"))

# Extract the first table
Abares_data <- docx_extract_tbl(Abares_doc, tbl_number = 1)

# Check it out
head(Abares_data)

```


#ensure that if a species is present at any given state it will be in AUS figured as also Y present
```{r}
Abares_data_clean <- Abares_data %>%
  # Ensure AUS = "Y" if any state has "Y"
  rowwise() %>%
  mutate(AUS = ifelse(any(c_across(c(NT, QLD, NSW, VIC, TAS, SA, WA)) == "Y"), "Y", AUS)) %>%
  ungroup()
```

# pivot to long format for mapping

```{r}
Abares_data_long <- Abares_data_clean %>%
  pivot_longer(
    cols = c(NT, QLD, NSW, VIC, TAS, SA, WA),
    names_to = "State",
    values_to = "Presence"
  ) %>%
  filter(Presence == "Y")

```

#make map

```{r}

# Add matching abbreviations
AUS_states <- AUS_states %>%
  mutate(State = case_when(
    STE_NAME21 == "Queensland" ~ "QLD",
    STE_NAME21 == "New South Wales" ~ "NSW",
    STE_NAME21 == "Victoria" ~ "VIC",
    STE_NAME21 == "Tasmania" ~ "TAS",
    STE_NAME21 == "South Australia" ~ "SA",
    STE_NAME21 == "Western Australia" ~ "WA",
    STE_NAME21 == "Northern Territory" ~ "NT",
    TRUE ~ NA_character_
  ))


```
#merge datasets
```{r}
Abares_map <- AUS_states %>%
  left_join(Abares_data_long, by = "State")

```

#map
```{r}

group_aquaculture<- ggplot(Abares_map) +
  geom_sf(aes(fill = group)) +
  facet_wrap(~ group) +
  theme_minimal() +
  labs(title = "Aquaculture Species Distribution by State in Australia",
       fill = "Species group")
plot(group_aquaculture)

ggsave(
  filename = here("Figures", "Aquaculture_groups_map.png"),
  plot = group_aquaculture,
  width = 10,       # adjust size to your liking
  height = 6,
  dpi = 300         # good for publication quality
)

```

#Presence of species in australia anywhere
```{r}
AUS_presence <- Abares_data_clean %>%
  filter(AUS == "Y") %>%
  dplyr:::select(Species, group, aquaculturerealm)

```
#calculate the number of species per state
```{r}
# Count how many species occur in each state
state_species_counts <- Abares_data_long %>%
  group_by(State) %>%
  summarise(species_count = n_distinct(Species)) %>%
  arrange(desc(species_count))

```

#join spatial data for counts
```{r}
AUS_states_richness <- AUS_states %>%
  left_join(state_species_counts, by = "State")

```

#make richness map
```{r}
library(ggplot2)

richness_map<- ggplot(AUS_states_richness) +
  geom_sf(aes(fill = species_count)) +
  scale_fill_viridis_c(option = "C", na.value = "grey90") +
  theme_minimal() +
  labs(
    title = "Aquaculture Species Richness by State",
    fill = "No. of Species"
  )
plot(richness_map)
```

#save richness map
```{r}
ggsave(
  filename = here("figures", "Richness_map.png"),
  plot = richness_map,
  width = 10,       # adjust size to your liking
  height = 6,
  dpi = 300         # good for publication quality
)

```

#read production
#The ABARES production table was downloaded from : https://view.officeapps.live.com/op/view.aspx?src=https%3A%2F%2Fdaff.ent.sirsidynix.net.au%2Fclient%2Fen_AU%2Fsearch%2Fasset%2F1037341%2F2&wdOrigin=BROWSELINK
```{r}
# Read the production data
library(readxl)

production_data_AU <- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 4"  # select the sheets in the excell file
)

production_data_NSW <- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 5"  # select the sheets in the excell file
)

production_data_VIC<- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 6"  # select the sheets in the excell file
)

production_data_QLD <- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 7"  # select the sheets in the excell file
)
production_data_SA <- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 8"  # select the sheets in the excell file
)
production_data_WA <- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 9"  # select the sheets in the excell file
)
production_data_TAS <- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 10"  # select the sheets in the excell file
)
production_data_NT <- read_excel(
  here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx"),
  sheet = "Table 11"  # select the sheets in the excell file
)
```

#filter only aquaculture (there is fishery data as well in each layer)
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)

# Improved function that finds "Aquaculture" regardless of the letter
extract_aquaculture_data <- function(file_path, sheet_name, state_abbr) {
  
  # Read raw data without headers
  raw_data <- read_excel(file_path, sheet = sheet_name, col_names = FALSE)
  
  # Find ALL rows that start with "Aquaculture" (followed by any letter or space)
  # This regex matches "Aquaculture", "Aquaculture a", "Aquaculture h", etc.
  aqua_rows <- which(str_detect(raw_data[[1]], "^Aquaculture\\s*[a-z]?\\s*$"))
  
  if(length(aqua_rows) < 2) {
    warning(paste("Expected 2 Aquaculture sections in", sheet_name, 
                  "but found", length(aqua_rows)))
    # Print what was found for debugging
    print(paste("Found in", sheet_name, ":"))
    print(raw_data[[1]][str_detect(raw_data[[1]], "Aquaculture")])
    return(NULL)
  }
  
  # First aquaculture section = VALUE data ($'000)
  value_start <- aqua_rows[1] + 1
  value_end <- aqua_rows[2] - 1
  
  # Second aquaculture section = QUANTITY (tonnes) data
  tonnes_start <- aqua_rows[2] + 1
  
  # Find where tonnes section ends (look for "Total" or end of data)
  total_rows_after_tonnes <- which(str_detect(
    raw_data[[1]][(tonnes_start):nrow(raw_data)], 
    "^Total"
  ))
  
  if(length(total_rows_after_tonnes) > 0) {
    tonnes_end <- tonnes_start + total_rows_after_tonnes[1] - 2
  } else {
    tonnes_end <- nrow(raw_data)
  }
  
  # Find the header row with years (usually row 2 or 3)
  # Look for cells containing year patterns like "1998-99", "2023-24", etc.
  year_pattern <- "[0-9]{4}[-–][0-9]{2}"
  header_candidates <- which(apply(raw_data, 1, function(row) {
    any(str_detect(as.character(row), year_pattern))
  }))
  
  if(length(header_candidates) > 0) {
    header_row <- header_candidates[1]
  } else {
    # Fallback to row 2
    header_row <- 2
  }
  
  # Extract year headers
  year_headers <- as.character(raw_data[header_row, ])
  year_headers[1] <- "Species"  # First column is commodity/species names
  year_headers[2] <- "Unit"      # Second column is unit (often present)
  
  # Clean up year headers (remove NA, clean whitespace)
  year_headers <- ifelse(is.na(year_headers) | year_headers == "NA", 
                         paste0("Col_", seq_along(year_headers)), 
                         year_headers)
  
  # --- Extract VALUE section ---
  value_section <- raw_data[value_start:value_end, ]
  colnames(value_section) <- year_headers
  
  value_clean <- value_section %>%
    filter(!is.na(Species), 
           Species != "", 
           !str_detect(Species, "^Total"),
           !str_detect(Species, "^Aquaculture")) %>%
    select(-any_of("Unit")) %>%  # Remove unit column if present
    mutate(Species = str_trim(Species)) %>%
    pivot_longer(
      cols = -Species,
      names_to = "Financial_Year",
      values_to = "Value_AUD_thousands"
    ) %>%
    filter(!str_detect(Financial_Year, "^Col_")) %>%  # Remove placeholder columns
    mutate(
      State = state_abbr,
      Value_AUD_thousands = as.numeric(Value_AUD_thousands)
    ) %>%
    filter(!is.na(Value_AUD_thousands))
  
  # --- Extract QUANTITY (tonnes) section ---
  tonnes_section <- raw_data[tonnes_start:tonnes_end, ]
  colnames(tonnes_section) <- year_headers
  
  tonnes_clean <- tonnes_section %>%
    filter(!is.na(Species), 
           Species != "", 
           !str_detect(Species, "^Total"),
           !str_detect(Species, "^Aquaculture")) %>%
    select(-any_of("Unit")) %>%
    mutate(Species = str_trim(Species)) %>%
    pivot_longer(
      cols = -Species,
      names_to = "Financial_Year",
      values_to = "Tonnes"
    ) %>%
    filter(!str_detect(Financial_Year, "^Col_")) %>%
    mutate(
      State = state_abbr,
      Tonnes = as.numeric(Tonnes)
    ) %>%
    filter(!is.na(Tonnes))
  
  # --- Combine value and tonnes data ---
  combined_data <- value_clean %>%
    full_join(tonnes_clean, by = c("State", "Species", "Financial_Year"))
  
  return(combined_data)
}

# Test on a couple of states first to verify it works
file_path <- here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx")

# Test NSW (has "Aquaculture h")
test_nsw <- extract_aquaculture_data(file_path, "Table 5", "NSW")
View(test_nsw)

# Test QLD (has "Aquaculture c")
test_qld <- extract_aquaculture_data(file_path, "Table 7", "QLD")
View(test_qld)

# If those look good, process all states
production_data_AUS <- extract_aquaculture_data(file_path, "Table 4", "AUS")
production_data_NSW <- extract_aquaculture_data(file_path, "Table 5", "NSW")
production_data_VIC <- extract_aquaculture_data(file_path, "Table 6", "VIC")
production_data_QLD <- extract_aquaculture_data(file_path, "Table 7", "QLD")
production_data_SA <- extract_aquaculture_data(file_path, "Table 8", "SA")
production_data_WA <- extract_aquaculture_data(file_path, "Table 9", "WA")
production_data_TAS <- extract_aquaculture_data(file_path, "Table 10", "TAS")
production_data_NT <- extract_aquaculture_data(file_path, "Table 11", "NT")

# Combine all states
production_data_all <- bind_rows(
  production_data_AUS,
  production_data_NSW,
  production_data_VIC,
  production_data_QLD,
  production_data_SA,
  production_data_WA,
  production_data_TAS,
  production_data_NT
)

View(production_data_all)

# Quick summary
summary_stats <- production_data_all %>%
  group_by(State) %>%
  summarise(
    n_species = n_distinct(Species),
    n_years = n_distinct(Financial_Year),
    total_records = n()
  )

print(summary_stats)

# Check unique species names
unique_species <- production_data_all %>%
  distinct(Species) %>%
  arrange(Species)

View(unique_species)
```
#function 3
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)

# Corrected function - extract BETWEEN "Total wild-caught" and "Total" rows
extract_aquaculture_data_v3 <- function(file_path, sheet_name, state_abbr) {
  
  # Read raw data without headers
  raw_data <- read_excel(file_path, sheet = sheet_name, col_names = FALSE)
  
  # Find "Total wild-caught" rows (should be 2: one in Value, one in Quantity)
  wild_caught_rows <- which(str_detect(raw_data[[1]], "^Total wild-caught"))
  
  # Find "Total production" or just "Total" rows that come AFTER aquaculture data
  # We need to find rows that say exactly "Total" and are after "Total wild-caught"
  total_rows <- which(str_detect(raw_data[[1]], "^Total$"))
  
  if(length(wild_caught_rows) < 2) {
    warning(paste("Expected 2 'Total wild-caught' markers in", sheet_name, 
                  "but found", length(wild_caught_rows)))
    return(NULL)
  }
  
  # VALUE section: rows between first "Total wild-caught" and first "Total" after it
  value_start <- wild_caught_rows[1] + 1
  
  # Find the first "Total" row that comes after the aquaculture section starts
  # It should be labeled as "Aquaculture" followed by species, then "Total"
  value_end_candidates <- total_rows[total_rows > value_start & total_rows < wild_caught_rows[2]]
  if(length(value_end_candidates) > 0) {
    value_end <- value_end_candidates[1] - 1
  } else {
    value_end <- wild_caught_rows[2] - 1
  }
  
  # QUANTITY section: rows between second "Total wild-caught" and next "Total" after it
  tonnes_start <- wild_caught_rows[2] + 1
  
  tonnes_end_candidates <- total_rows[total_rows > tonnes_start]
  if(length(tonnes_end_candidates) > 0) {
    tonnes_end <- tonnes_end_candidates[1] - 1
  } else {
    tonnes_end <- nrow(raw_data)
  }
  
  # Find the header row with years
  year_pattern <- "[0-9]{4}[-–][0-9]{2}"
  header_candidates <- which(apply(raw_data, 1, function(row) {
    any(str_detect(as.character(row), year_pattern))
  }))
  
  if(length(header_candidates) > 0) {
    header_row <- header_candidates[1]
  } else {
    header_row <- 2
  }
  
  # Extract year headers
  year_headers <- as.character(raw_data[header_row, ])
  year_headers[1] <- "Species"
  year_headers[2] <- "Unit"
  year_headers <- ifelse(is.na(year_headers) | year_headers == "NA", 
                         paste0("Col_", seq_along(year_headers)), 
                         year_headers)
  
  # --- Extract VALUE section (aquaculture only) ---
  value_section <- raw_data[value_start:value_end, ]
  colnames(value_section) <- year_headers
  
  value_clean <- value_section %>%
    filter(!is.na(Species), 
           Species != "", 
           !str_detect(Species, "^Total"),
           !str_detect(Species, "^Aquaculture")) %>%
    select(-any_of("Unit")) %>%
    mutate(Species = str_trim(Species)) %>%
    pivot_longer(
      cols = -Species,
      names_to = "Financial_Year",
      values_to = "Value_AUD_thousands"
    ) %>%
    filter(!str_detect(Financial_Year, "^Col_")) %>%
    mutate(
      State = state_abbr,
      Value_AUD_thousands = as.numeric(Value_AUD_thousands)
    ) %>%
    filter(!is.na(Value_AUD_thousands))
  
  # --- Extract QUANTITY section (aquaculture only) ---
  tonnes_section <- raw_data[tonnes_start:tonnes_end, ]
  colnames(tonnes_section) <- year_headers
  
  tonnes_clean <- tonnes_section %>%
    filter(!is.na(Species), 
           Species != "", 
           !str_detect(Species, "^Total"),
           !str_detect(Species, "^Aquaculture")) %>%
    select(-any_of("Unit")) %>%
    mutate(Species = str_trim(Species)) %>%
    pivot_longer(
      cols = -Species,
      names_to = "Financial_Year",
      values_to = "Tonnes"
    ) %>%
    filter(!str_detect(Financial_Year, "^Col_")) %>%
    mutate(
      State = state_abbr,
      Tonnes = as.numeric(Tonnes)
    ) %>%
    filter(!is.na(Tonnes))
  
  # --- Combine value and tonnes ---
  combined_data <- value_clean %>%
    full_join(tonnes_clean, by = c("State", "Species", "Financial_Year"))
  
  return(combined_data)
}

# Test on Victoria (VIC) 
file_path <- here("raw_data/Aquaculture_Production_ABARES_DAFF/02_DataTables_AustFishAquacStats_2024_v1.0.0.xlsx")

test_vic <- extract_aquaculture_data_v3(file_path, "Table 6", "VIC")

cat("\n=== VIC Test Results ===\n")
cat("Total records:", nrow(test_vic), "\n")
cat("Unique species:", n_distinct(test_vic$Species), "\n")
cat("\nSpecies found:\n")
print(sort(unique(test_vic$Species)))

# Check for wild-caught species (should be NONE)
wild_caught_check <- c("Southern Rock Lobster", "Prawns", "Crabs", "Abalone", 
                       "Scallops", "Squids", "Octopus", "Australian Sardine",
                       "Black Bream", "Southern Garfish", "Sharks & Rays",
                       "Snapper", "Freshwater Eels", "Australian Salmons",
                       "King George Whiting", "Other Finfish")

found_wild <- test_vic %>%
  filter(Species %in% wild_caught_check)

if(nrow(found_wild) > 0) {
  cat("\n⚠️ WARNING: Found wild-caught species:\n")
  print(unique(found_wild$Species))
} else {
  cat("\n✓ ✓ ✓ PERFECT! No wild-caught species found!\n")
}

View(test_vic)
# Process all states
production_data_AUS <- extract_aquaculture_data_v3(file_path, "Table 4", "AUS")
production_data_NSW <- extract_aquaculture_data_v3(file_path, "Table 5", "NSW")
production_data_VIC <- extract_aquaculture_data_v3(file_path, "Table 6", "VIC")
production_data_QLD <- extract_aquaculture_data_v3(file_path, "Table 7", "QLD")
production_data_SA <- extract_aquaculture_data_v3(file_path, "Table 8", "SA")
production_data_WA <- extract_aquaculture_data_v3(file_path, "Table 9", "WA")
production_data_TAS <- extract_aquaculture_data_v3(file_path, "Table 10", "TAS")
production_data_NT <- extract_aquaculture_data_v3(file_path, "Table 11", "NT")

# Combine all
production_data_all_clean <- bind_rows(
  production_data_AUS,
  production_data_NSW,
  production_data_VIC,
  production_data_QLD,
  production_data_SA,
  production_data_WA,
  production_data_TAS,
  production_data_NT
)


cat("\n=== All species (should be AQUACULTURE ONLY) ===\n")
all_species_clean <- production_data_all_clean %>%
  distinct(Species) %>%
  arrange(Species)
print(all_species_clean$Species)

View(production_data_all_clean)
```

```{r}
# Create simplified species lookup
species_lookup_final <- tribble(
  ~Production_Name,           ~Species,                    ~popular.name,
  
  # === BIVALVES ===
  "Oysters",                  "Oyster spp",               "Oysters (all species)",
  "Oysters e",                "Oyster spp",               "Oysters (all species)",
  "Oysters i",                "Oyster spp",               "Oysters (all species)",
  "Oystersi",                 "Oyster spp",               "Oysters (all species)",
  
  "Mussels",                  "Mytilus galloprovincialis", "blue mussel",
  "Blue Mussel",              "Mytilus galloprovincialis", "blue mussel",
  
  "Abalone",                  "Haliotis spp",             "Abalone (all species)",
  "Abalone e",                "Haliotis spp",             "Abalone (all species)",
  "Blacklip Abalones",        "Haliotis spp",             "Abalone (all species)",
  
  # === FISH ===
  "Barramundi",               "Lates calcifer",           "Barramundi",
  "Golden Perch",             "Macquarie ambigua",        "Golden Perch",
  "Silver Perch",             "Bidyanis bidyanus",        "Silver Perch",
  "Barcoo Grunter",           "Scortum barcoo",           "Jade perch",
  "Murray Cod",               "Maccullochella mariensis", "Mary River cod",
  
  # === SALMONIDS - ===
  "Trouts",                   "Oncorhynchus mykiss",      "Rainbow Trout",
  "Trout",                    "Oncorhynchus mykiss",      "Rainbow Trout",
  
   "Salmonids d",              "Salmo salar",              "Salmon",
  "Salmon",                   "Salmo salar",              "Salmon",
  "Salmonids c",              "Salmo salar",              "Salmon",
  
  # Other fish
  "Yellowtail Kingfish",      "Seriola lalandi",          "Yellowtail Kingfish",
  "Snapper",                  "Pagrus auratus",           "Snapper",
  "Mulloway",                 "Argyrosomus japonicus",    "Mulloway",
  "Coral Trouts",             "Epinephelinae",            "Grouper",
  "Southern Bluefin Tuna d",  "Thunnus maccoyii",         "Southern Bluefin Tuna",
  "Freshwater Eels",          "Anguilla spp",             "Freshwater Eels",
  
  # === CRUSTACEANS ===
  "Prawns",                   "Penaeus spp",              "Prawns (all species)",
  "Tiger Prawns",             "Penaeus spp",              "Prawns (all species)",
  "King Prawns",              "Penaeus spp",              "Prawns (all species)",
  "Banana Prawns",            "Penaeus spp",              "Prawns (all species)",
  "School Prawns",            "Penaeus spp",              "Prawns (all species)",
  "Endeavour Prawns",         "Penaeus spp",              "Prawns (all species)",
  
  "Yabbies",                  "cherax destructor",        "Yabbies",
  "Marron",                   "cherax tenuimanus",        "Maroon",
  "Marron c",                 "cherax tenuimanus",        "Maroon",
  "Redclaw",                  "Cherax quadricarinatus",   "Red Claw crayfish",
  
  # === OTHER ===
  "Pearls",                   "pearl",                    "pearl",
  "Pearl Oyster",             "pearl",                    "pearl",
  "Ornamental species",       "Ornamental species",       "Ornamental species",
  "Aquarium species d",       "Ornamental species",       "Ornamental species",
  "Warmwater Finfish d",      "Warmwater finfish",        "Warmwater finfish"
)

# Update your presence data to match these categories
Abares_data_long_updated <- Abares_data_long %>%
  mutate(
    Species = case_when(
      # Combine oyster species
      Species %in% c("Crassostrea gigas", "Saccostrea glomerata", 
                     "Saccostrea Lineage G/Saccostrea solaris") ~ "Oyster spp",
      
      # Combine abalone species
      Species %in% c("Haliotis laevigata", "Haliotis rubra") ~ "Haliotis spp",
      
      # Combine eel species
      Species %in% c("Anguilla australis", "Anguilla dieffenbachii") ~ "Anguilla spp",
      
      # Combine prawn species
      Species %in% c("Penaeus monodon", "Marsupenaus japonicus") ~ "Penaeus spp",
      
      # Keep others as is
      TRUE ~ Species
    ),
    # Update popular names
    popular.name = case_when(
      Species == "Oyster spp" ~ "Oysters (all species)",
      Species == "Haliotis spp" ~ "Abalone (all species)",
      Species == "Anguilla spp" ~ "Freshwater Eels",
      Species == "Penaeus spp" ~ "Prawns (all species)",
      TRUE ~ popular.name
    )
  ) %>%
  # Remove duplicates created by combining - fix the column name here
  distinct(group, popular.name, Species, aquaculturerealm, AUS, State, Presence)

View(Abares_data_long_updated)

# Check what changed
cat("\n=== Species combining summary ===\n")
cat("Original species:", n_distinct(Abares_data_long$Species), "\n")
cat("Updated species:", n_distinct(Abares_data_long_updated$Species), "\n")

## CORRECTED aggregation
production_aggregated <- production_data_all_clean %>%
  left_join(species_lookup_final, 
            by = c("Species" = "Production_Name"),
            relationship = "many-to-many") %>%
  rename(Production_Species = Species,
         Species = Species.y) %>%
  select(Species, popular.name, Production_Species, State, Financial_Year, 
         Value_AUD_thousands, Tonnes) %>%
  filter(!is.na(Species)) %>%
  group_by(Species, popular.name, State, Financial_Year) %>%
  summarise(
    Value_AUD_thousands = sum(Value_AUD_thousands, na.rm = TRUE),
    Tonnes = sum(Tonnes, na.rm = TRUE),
    Production_Species_list = paste(unique(Production_Species), collapse = " + "),
    n_source_records = n(),
    .groups = "drop"
  ) %>%
  mutate(
    Value_AUD_thousands = ifelse(Value_AUD_thousands == 0, NA_real_, Value_AUD_thousands),
    Tonnes = ifelse(Tonnes == 0, NA_real_, Tonnes)
  )

# Check Tasmania salmon specifically
tas_salmon_check <- production_aggregated %>%
  filter(State == "TAS", Species == "Salmo salar") %>%
  arrange(desc(Financial_Year))

cat("\n=== Tasmania Salmon (should be HUGE!) ===\n")
print(head(tas_salmon_check, 10))

# Check aggregation results
agg_summary <- production_aggregated %>%
  group_by(Species, popular.name) %>%
  summarise(
    n_states = n_distinct(State),
    n_years = n_distinct(Financial_Year),
    has_value = sum(!is.na(Value_AUD_thousands)) > 0,
    has_tonnes = sum(!is.na(Tonnes)) > 0,
    .groups = "drop"
  )

cat("\n=== Production Data Summary ===\n")
print(agg_summary, n = 50)

# Final join
Abares_data_enriched <- Abares_data_long_updated %>%
  left_join(
    production_aggregated,
    by = c("State", "Species", "popular.name"),
    relationship = "many-to-many"
  )

View(Abares_data_enriched)

# Summary statistics
enrichment_summary <- Abares_data_enriched %>%
  group_by(Species, popular.name, State) %>%
  summarise(
    has_data = any(!is.na(Financial_Year)),
    n_years = sum(!is.na(Financial_Year)),
    .groups = "drop"
  ) %>%
  left_join(Abares_data_long_updated %>% distinct(Species, group), by = "Species")

cat("\n=== Final Integration Summary ===\n")
cat("Total species-state combinations:", nrow(enrichment_summary), "\n")
cat("With production data:", sum(enrichment_summary$has_data), "\n")
cat("Without production data:", sum(!enrichment_summary$has_data), "\n")
cat("Match rate:", round(sum(enrichment_summary$has_data) / nrow(enrichment_summary) * 100, 1), "%\n")

# Species without data
no_data <- enrichment_summary %>%
  filter(!has_data) %>%
  arrange(group, popular.name, State)

cat("\n=== Species without production data ===\n")
print(no_data, n = 100)

# Get most recent year for each species-state
production_latest <- Abares_data_enriched %>%
  filter(!is.na(Financial_Year)) %>%
  group_by(Species, popular.name, State, group, aquaculturerealm) %>%
  filter(Financial_Year == max(Financial_Year)) %>%
  slice(1) %>%
  ungroup()

View(production_latest)

# State-level totals (latest year)
state_totals <- production_latest %>%
  group_by(State) %>%
  summarise(
    n_species = n_distinct(Species),
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    total_tonnes = sum(Tonnes, na.rm = TRUE),
    latest_year = max(Financial_Year),
    .groups = "drop"
  ) %>%
  arrange(desc(total_value_millions))

cat("\n=== ✓✓✓ CORRECTED STATE TOTALS (TASMANIA SHOULD BE #1!) ✓✓✓ ===\n")
print(state_totals)

# Top species by value
top_species <- production_latest %>%
  group_by(Species, popular.name, group) %>%
  summarise(
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    total_tonnes = sum(Tonnes, na.rm = TRUE),
    n_states = n_distinct(State),
    .groups = "drop"
  ) %>%
  arrange(desc(total_value_millions)) %>%
  head(15)

cat("\n=== Top 15 Species (SALMON SHOULD BE #1!) ===\n")
print(top_species)

# Create the processed_data folder if it doesn't exist
if (!dir.exists(here("processed_data"))) {
  dir.create(here("processed_data"), recursive = TRUE)
  cat("Created processed_data folder\n")
}

# Top species across multiple years for comparison
# USE Abares_data_enriched (which has all years), not production_latest
top_species_multi_year <- Abares_data_enriched %>%
  filter(Financial_Year %in% c("2020–21", "2021–22", "2022–23")) %>%
  filter(!is.na(Value_AUD_thousands)) %>%  # Only rows with production data
  group_by(Species, popular.name, group, Financial_Year) %>%
  summarise(
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    total_tonnes = sum(Tonnes, na.rm = TRUE),
    n_states = n_distinct(State),
    .groups = "drop"
  ) %>%
  arrange(Species, Financial_Year)

cat("\n=== Checking filtered data ===\n")
cat("Rows after filtering:", nrow(top_species_multi_year), "\n")

# Get top 15 species by average value across all three years
top_species_list <- top_species_multi_year %>%
  group_by(Species, popular.name, group) %>%
  summarise(avg_value = mean(total_value_millions, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(avg_value)) %>%
  head(15) %>%
  pull(Species)

# Filter to just those top 15 species
top_species_comparison <- top_species_multi_year %>%
  filter(Species %in% top_species_list) %>%
  arrange(desc(total_value_millions))

cat("\n=== Top 15 Species Comparison: 2020–21, 2021–22, 2022–23 ===\n")
print(top_species_comparison, n = 50)

# Create a wide format for easier comparison
top_species_wide <- top_species_comparison %>%
  select(Species, popular.name, group, Financial_Year, total_value_millions, total_tonnes) %>%
  pivot_wider(
    names_from = Financial_Year,
    values_from = c(total_value_millions, total_tonnes),
    names_sep = "_"
  )

cat("\n=== Wide Format (easier to read!) ===\n")
print(top_species_wide, n = 20)

# Save everything - ALL WITH row.names NOT row_names
write.csv(Abares_data_enriched, 
          here("processed_data", "aquaculture_complete_dataset.csv"), 
          row.names = FALSE)

write.csv(production_latest, 
          here("processed_data", "aquaculture_latest_year.csv"), 
          row.names = FALSE)

write.csv(state_totals, 
          here("processed_data", "state_totals_latest_CORRECTED.csv"), 
          row.names = FALSE)

write.csv(top_species,
          here("processed_data", "top_species_CORRECTED.csv"),
          row.names = FALSE)

write.csv(top_species_comparison,
          here("processed_data", "top_species_multi_year_long.csv"),
          row.names = FALSE)

write.csv(top_species_wide,
          here("processed_data", "top_species_multi_year_wide.csv"),
          row.names = FALSE)

cat("\n✓ All files saved successfully!\n")
cat("Files location:", here("processed_data"), "\n")
```


```{r}
# Create the processed_data folder if it doesn't exist
if (!dir.exists(here("processed_data"))) {
  dir.create(here("processed_data"), recursive = TRUE)
  cat("Created processed_data folder\n")
}

# Top species across multiple years for comparison
top_species_multi_year <- Abares_data_enriched %>%
  filter(Financial_Year %in% c("2020–21", "2021–22", "2022–23")) %>%  # USE EN-DASHES!
  filter(!is.na(Value_AUD_thousands)) %>%  # Only rows with production data
  group_by(Species, popular.name, group, Financial_Year) %>%
  summarise(
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    total_tonnes = sum(Tonnes, na.rm = TRUE),
    n_states = n_distinct(State),
    .groups = "drop"
  ) %>%
  arrange(Species, Financial_Year)

# Get top 15 species by average value across all three years
top_species_list <- top_species_multi_year %>%
  group_by(Species, popular.name, group) %>%
  summarise(avg_value = mean(total_value_millions, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(avg_value)) %>%
  head(15) %>%
  pull(Species)

# Filter to just those top 15 species
top_species_comparison <- top_species_multi_year %>%
  filter(Species %in% top_species_list) %>%
  arrange(desc(total_value_millions))

cat("\n=== Top 15 Species Comparison: 2020–21, 2021–22, 2022–23 ===\n")
print(top_species_comparison, n = 50)

# Create a wide format for easier comparison
top_species_wide <- top_species_comparison %>%
  select(Species, popular.name, group, Financial_Year, total_value_millions, total_tonnes) %>%
  pivot_wider(
    names_from = Financial_Year,
    values_from = c(total_value_millions, total_tonnes),
    names_sep = "_"
  )

cat("\n=== Wide Format (easier to read!) ===\n")
print(top_species_wide, n = 20)

# Save the new multi-year files
write.csv(top_species_comparison,
          here("processed_data", "top_species_multi_year_long.csv"),
          row.names = FALSE)

write.csv(top_species_wide,
          here("processed_data", "top_species_multi_year_wide.csv"),
          row.names = FALSE)

cat("\n✓ Multi-year comparison files saved!\n")
```

#MAP
# scale_fill_viridis_c() = for continuous variables (like your first map with numerical values)
#scale_fill_viridis_d() = for discrete categories (like groups: Bivalves, Fish, Crustacean)
#viridis options:
#option = "A" (magma)
#option = "B" (inferno)
#option = "C" (plasma) ← default
#option = "D" (viridis)
#option = "E" (cividis)

```{r}
library(ggplot2)

# Map: State production value
state_map <- AUS_states %>%
  left_join(state_totals, by = "State") %>%
  ggplot() +
  geom_sf(aes(fill = total_value_millions), color = "white", size = 0.5) +
  scale_fill_viridis_c(
    option = "C",
    na.value = "grey90",
    labels = scales::comma,
    name = "Value\n(AUD$ millions)"
  ) +
  theme_minimal() +
  labs(
    title = "Australian Aquaculture Production Value by State",
    subtitle = paste("Financial year:", unique(state_totals$latest_year)),
    caption = "Source: ABARES 2024"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10)
  )

print(state_map)

ggsave(
  filename = here("Figures", "State_production_value_map.png"),
  plot = state_map,
  width = 10,
  height = 6,
  dpi = 300
)

# Bar chart: Top 10 species
top_species_plot <- production_latest %>%
  group_by(Species, popular.name, group) %>%
  summarise(
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    .groups = "drop"
  ) %>%
  arrange(desc(total_value_millions)) %>%
  slice(1:10) %>%  # Top 10 species
  ggplot(aes(x = reorder(popular.name, total_value_millions), 
             y = total_value_millions,
             fill = group)) +
  geom_col() +
  coord_flip() +
   scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Top 10 Aquaculture Species by Production Value",
    subtitle = paste("Australia, Financial year:", unique(production_latest$Financial_Year)),
    x = NULL,
    y = "Production Value ($ millions)",
    fill = "Species Group",
    caption = "Source: ABARES 2024"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )

print(top_species_plot)

ggsave(
  filename = here("Figures", "Top10_species_by_value.png"),
  plot = top_species_plot,
  width = 10,
  height = 8,
  dpi = 300
)

# Stacked bar: Production by group and state
group_state_plot <- production_latest %>%
  group_by(State, group) %>%
  summarise(total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
            .groups = "drop") %>%
  ggplot(aes(x = reorder(State, total_value_millions), 
             y = total_value_millions, 
             fill = group)) +
  geom_col() +
  coord_flip() +
  scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Aquaculture Production Value by State and Species Group",
    subtitle = paste("Financial year:", unique(production_latest$Financial_Year)),
    x = "State",
    y = "Production Value (AUD$ millions)",
    fill = "Species Group",
    caption = "Source: ABARES 2024"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )

print(group_state_plot)

ggsave(
  filename = here("Figures", "Production_by_state_and_group.png"),
  plot = group_state_plot,
  width = 10,
  height = 6,
  dpi = 300
)

cat("\n✓ All visualizations saved!\n")
```
#CHECK production of other years
```{r}
# Define the years you want to analyze
selected_years <- c("2020–21", "2021–22", "2022–23", "2023–24p")

# Filter for specific years instead of just latest
production_selected_years <- Abares_data_enriched %>%
  filter(Financial_Year %in% selected_years) %>%
  filter(!is.na(Financial_Year))

# State-level totals BY YEAR
state_totals_by_year <- production_selected_years %>%
  group_by(State, Financial_Year) %>%
  summarise(
    n_species = n_distinct(Species),
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    total_tonnes = sum(Tonnes, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(Financial_Year, desc(total_value_millions))

cat("\n=== State Production Totals BY YEAR ===\n")
print(state_totals_by_year, n = 100)

# Top species by value BY YEAR
top_species_by_year <- production_selected_years %>%
  group_by(Financial_Year, Species, popular.name, group) %>%
  summarise(
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    total_tonnes = sum(Tonnes, na.rm = TRUE),
    n_states = n_distinct(State),
    .groups = "drop"
  ) %>%
  group_by(Financial_Year) %>%
  arrange(Financial_Year, desc(total_value_millions)) %>%
  slice(1:10) %>%  # Top 10 per year
  ungroup()

cat("\n=== Top 10 Species by Value BY YEAR ===\n")
print(top_species_by_year, n = 100)

# Group-level summary BY YEAR
group_summary_by_year <- production_selected_years %>%
  group_by(Financial_Year, group) %>%
  summarise(
    n_species = n_distinct(Species),
    total_value_millions = sum(Value_AUD_thousands, na.rm = TRUE) / 1000,
    total_tonnes = sum(Tonnes, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(Financial_Year, desc(total_value_millions))

cat("\n=== Production by Group BY YEAR ===\n")
print(group_summary_by_year, n = 100)
```
#trends over time
```{r}
# Line plot: State production over time
state_trend_plot <- state_totals_by_year %>%
  ggplot(aes(x = Financial_Year, y = total_value_millions, 
             color = State, group = State)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_viridis_d(option = "D") +
  theme_minimal() +
  labs(
    title = "Aquaculture Production Value Trends by State",
    x = "Financial Year",
    y = "Production Value ($ millions)",
    color = "State"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(state_trend_plot)
```


