---
title: "Australian Land-Based Aquaculture Sites & Pond Detection"
subtitle: "Spatial Analysis of NDWI-Derived Ponds within Official Aquaculture Leases"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: today
format: 
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    theme: cosmo
    embed-resources: true
    smooth-scroll: true
    fig-width: 10
    fig-height: 8
    df-print: paged
execute:
  warning: false
  message: false
  cache: false
---

## Executive Summary

This analysis identifies and maps aquaculture ponds across Australia using Google Earth Engine-derived NDWI (Normalized Difference Water Index) data from 2021-2022. The study focuses exclusively on **land-based aquaculture sites** across seven Australian states.

:::{.callout-note}
## Key Findings
- Analysis covers: QLD, NSW, SA, WA, TAS, VIC, and NT
- Data source: NDWI satellite imagery (2021-2022)
- Focus: Land-based aquaculture sites only
:::

---

## 1. Setup & Data Preparation

### Load Required Libraries

```{r setup}
#| code-fold: false

library(here)
library(sf)
library(tmap)
library(dplyr)
library(viridis)
library(knitr)
library(DT)

# Set random seed for reproducibility
set.seed(123)
```

### Define Project Parameters

```{r parameters}
#| code-fold: false

# Australian states to process
STATES <- c("QLD", "WA", "SA", "TAS", "VIC", "NT", "NSW")

# Coordinate Reference System - Australian Albers
AUS_ALBERS <- "EPSG:3577"

# Output directory
OUTPUT_DIR <- here("output_data", "ponds_in_sites")

# Create output directory if needed
if (!dir.exists(OUTPUT_DIR)) {
  dir.create(OUTPUT_DIR, recursive = TRUE)
}
```

---

## 2. Data Processing Pipeline

### 2.1 Processing Function

This function processes each state individually to:
1. Load NDWI-derived pond polygons from Google Earth Engine
2. Load land-based aquaculture sites
3. Transform to Australian Albers projection
4. Dissolve pond geometries
5. Intersect ponds with land sites
6. Export results as shapefiles

```{r process-function}
process_ponds_by_state <- function() {
  
  # Create output directory if it doesn't exist
  output_dir <- here("output_data", "ponds_in_sites")
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }
  
  # Initialize results tracking
  results <- data.frame(
    State = character(),
    Ponds_Found = integer(),
    Land_Sites = integer(),
    Ponds_In_Sites = integer(),
    Status = character(),
    stringsAsFactors = FALSE
  )
  
  # Process each state
  for (state in STATES) {
    
    cat("\n", rep("=", 60), "\n")
    cat("Processing state:", state, "\n")
    cat(rep("=", 60), "\n")
    
    # Define file paths
    pond_file <- here("output_data", "Australian_NDWI_aqua_ponds", "2021_2022", 
                      paste0("ndwi_sum_thresh_vector_", state, "_aqua_21_22.shp"))
    
    land_sites_file <- here("output_data",  
                            "Australian_Aqua_sites",
                            paste0(state, "_land_sites.shp"))
    
    # Check if pond file exists
    if (!file.exists(pond_file)) {
      cat("  ⚠️  Warning: Pond file not found - skipping\n")
      results <- rbind(results, data.frame(
        State = state, Ponds_Found = 0, Land_Sites = 0, 
        Ponds_In_Sites = 0, Status = "No pond file"
      ))
      next
    }
    
    # Check if land sites file exists
    if (!file.exists(land_sites_file)) {
      cat("  ⚠️  Warning: No land sites file found - skipping\n")
      results <- rbind(results, data.frame(
        State = state, Ponds_Found = NA, Land_Sites = 0, 
        Ponds_In_Sites = 0, Status = "No land sites file"
      ))
      next
    }
    
    tryCatch({
      # Load ponds
      cat("  📥 Loading pond shapefile...\n")
      ponds <- st_read(pond_file, quiet = TRUE)
      
      # Transform ponds to Australian Albers
      cat("  🔄 Transforming ponds to Australian Albers...\n")
      ponds <- st_transform(ponds, crs = AUS_ALBERS)
      
      # Check if there are any pond features
      if (nrow(ponds) == 0) {
        cat("  ⚠️  Warning: No pond features found - skipping\n")
        results <- rbind(results, data.frame(
          State = state, Ponds_Found = 0, Land_Sites = NA, 
          Ponds_In_Sites = 0, Status = "Empty pond file"
        ))
        next
      }
      
      # Load ONLY land-based sites
      cat("  📥 Loading land-based sites...\n")
      land_sites <- st_read(land_sites_file, quiet = TRUE) %>%
        st_transform(crs = AUS_ALBERS) %>%
        st_make_valid() %>%
        mutate(lease_type = "land_based")
      
      cat("    ✓ Found", nrow(land_sites), "land-based sites\n")
      
      # Check if there are any site features
      if (nrow(land_sites) == 0) {
        cat("  ⚠️  Warning: No land site features found - skipping\n")
        results <- rbind(results, data.frame(
          State = state, Ponds_Found = nrow(ponds), Land_Sites = 0, 
          Ponds_In_Sites = 0, Status = "Empty land sites file"
        ))
        next
      }
      
      # Dissolve ponds (union all pond geometries)
      cat("  🔗 Dissolving ponds...\n")
      ponds_dissolved <- ponds %>%
        summarise(geometry = st_union(geometry)) %>%
        st_cast("POLYGON")
      
      # Intersect ponds with land-based sites
      cat("  ✂️  Intersecting ponds with land-based sites...\n")
      ponds_in_sites <- st_intersection(ponds_dissolved, land_sites)
      
      # Check if intersection produced any results
      if (nrow(ponds_in_sites) == 0) {
        cat("  ⚠️  Warning: No ponds found within land sites\n")
        results <- rbind(results, data.frame(
          State = state, Ponds_Found = nrow(ponds), 
          Land_Sites = nrow(land_sites), 
          Ponds_In_Sites = 0, Status = "No intersection"
        ))
      } else {
        cat("  ✅ Found", nrow(ponds_in_sites), "pond features within land sites\n")
        results <- rbind(results, data.frame(
          State = state, Ponds_Found = nrow(ponds), 
          Land_Sites = nrow(land_sites), 
          Ponds_In_Sites = nrow(ponds_in_sites), 
          Status = "Success"
        ))
      }
      
      # Define output file path
      output_file <- here(output_dir, paste0("ponds_in_sites_in_", state, ".shp"))
      
      # Save as shapefile
      cat("  💾 Saving results...\n")
      st_write(ponds_in_sites, output_file, delete_layer = TRUE, quiet = TRUE)
      
      cat("  ✅ Successfully processed", state, "\n")
      
      # Calculate and display total area
      if (nrow(ponds_in_sites) > 0) {
        total_area <- sum(st_area(ponds_in_sites))
        cat("  📊 Total pond area:", format(total_area, scientific = FALSE), "sq meters\n")
      }
      
    }, error = function(e) {
      cat("  ❌ Error processing", state, ":", e$message, "\n")
      results <- rbind(results, data.frame(
        State = state, Ponds_Found = NA, Land_Sites = NA, 
        Ponds_In_Sites = 0, Status = paste("Error:", e$message)
      ))
    })
  }
  
  cat("\n", rep("=", 60), "\n")
  cat("Processing complete!\n")
  cat(rep("=", 60), "\n\n")
  
  return(results)
}
```

### 2.2 Execute Processing

```{r run-processing}
#| output: true

# Run the processing function
processing_results <- process_ponds_by_state()

# Display results table
kable(processing_results, 
      caption = "Processing Summary by State",
      col.names = c("State", "Ponds Found", "Land Sites", 
                    "Ponds in Sites", "Status"))
```

---

## 3. Interactive Mapping

### 3.1 Load Spatial Data for Mapping

```{r load-spatial-data}
#| output: false

cat("Loading Australia base layers...\n")

# Load Australian state boundaries
AUS_states <- st_read(here("raw_data/AustralianStates_SHP/STE_2021_AUST_GDA2020.shp")) %>%
  st_make_valid() %>%
  st_transform(3577) %>%
  filter(!st_is_empty(geometry))

cat("✓ State boundaries loaded\n\n")

# Function to load land sites
load_land_sites <- function(state) {
  land_sites_file <- here("raw_data/copy_aquaculture_leases/Australian_Aqua_leases",
                          paste0(state, "_land_sites.shp"))
  
  if (!file.exists(land_sites_file)) {
    return(NULL)
  }
  
  tryCatch({
    land_sites <- st_read(land_sites_file, quiet = TRUE) %>%
      st_make_valid() %>%
      st_transform(3577) %>%
      filter(!st_is_empty(geometry)) %>%
      mutate(lease_type = "land_based", State = state) %>%
      select(lease_type, State, geometry)
    
    cat("  ✓", state, ":", nrow(land_sites), "land sites\n")
    return(land_sites)
    
  }, error = function(e) {
    cat("  ✗", state, ": Error loading\n")
    return(NULL)
  })
}

# Function to load ponds
load_ponds <- function(state) {
  ponds_file <- here("output_data/ponds_in_sites", 
                     paste0("ponds_in_sites_in_", state, ".shp"))
  
  if (!file.exists(ponds_file)) {
    return(NULL)
  }
  
  tryCatch({
    ponds <- st_read(ponds_file, quiet = TRUE) %>%
      st_make_valid() %>%
      st_transform(3577) %>%
      filter(!st_is_empty(geometry)) %>%
      mutate(State = state) %>%
      select(geometry, any_of(c("lease_type", "State")))
    
    cat("  ✓", state, ":", nrow(ponds), "ponds\n")
    return(ponds)
    
  }, error = function(e) {
    return(NULL)
  })
}

# Load all data
cat("\nLoading land-based sites:\n")
land_sites_list <- setNames(lapply(STATES, load_land_sites), STATES)
land_sites_list <- land_sites_list[!sapply(land_sites_list, is.null)]

cat("\nLoading ponds:\n")
ponds_list <- setNames(lapply(STATES, load_ponds), STATES)
ponds_list <- ponds_list[!sapply(ponds_list, is.null)]

cat("\n✓ Spatial data loading complete\n")
```

### 3.2 Create Interactive Map

```{r create-map}
#| fig-cap: "Interactive map of Australian land-based aquaculture sites and detected ponds"
#| column: page

# Set tmap to interactive mode
tmap_mode("view")

# Define color palette
state_colors <- viridis(7)
names(state_colors) <- STATES

# Start with basemap
comprehensive_map <- tm_basemap("Esri.WorldImagery") +
  tm_shape(AUS_states) +
  tm_borders(col = 'white', alpha = 0.3, lwd = 0.5)

# Add land-based sites for each state
for (state in STATES) {
  if (state %in% names(land_sites_list) && 
      !is.null(land_sites_list[[state]]) && 
      nrow(land_sites_list[[state]]) > 0) {
    
    comprehensive_map <- comprehensive_map +
      tm_shape(land_sites_list[[state]]) +
      tm_borders(col = state_colors[state], lwd = 2, alpha = 0.8) +
      tm_fill(col = state_colors[state], alpha = 0.2, 
              id = paste(state, "Land Sites"),
              popup.vars = c("State", "lease_type"))
  }
}

# Add ponds with high visibility
pond_color <- "#FFFF00"  # Bright yellow

for (state in STATES) {
  if (state %in% names(ponds_list) && 
      !is.null(ponds_list[[state]]) && 
      nrow(ponds_list[[state]]) > 0) {
    
    comprehensive_map <- comprehensive_map +
      tm_shape(ponds_list[[state]]) +
      tm_fill(col = pond_color, alpha = 0.8, 
              id = paste(state, "Ponds"),
              popup.vars = "State") +
      tm_borders(col = "red", lwd = 1)
  }
}

# Add layout
comprehensive_map <- comprehensive_map +
  tm_layout(
    title = "Australia's Land-Based Aquaculture Sites & Detected Ponds",
    legend.outside = TRUE,
    legend.outside.position = "right"
  ) +
  tm_view(set.view = c(lon = 133.7751, lat = -25.2744, zoom = 5))

# Display map
comprehensive_map
```

:::{.callout-tip}
## How to Use the Map
- **Zoom**: Scroll or use +/- buttons
- **Pan**: Click and drag
- **Layers**: Toggle basemap using layer control (top right)
- **Details**: Click on features to see popup information
:::

---

## 4. Results & Statistics

### 4.1 Summary by State

```{r summary-stats}
#| column: page

# Calculate statistics
summary_stats <- data.frame(
  State = STATES,
  Land_Sites = sapply(STATES, function(state) {
    if (state %in% names(land_sites_list) && !is.null(land_sites_list[[state]])) {
      nrow(land_sites_list[[state]])
    } else {
      0
    }
  }),
  Ponds = sapply(STATES, function(state) {
    if (state %in% names(ponds_list) && !is.null(ponds_list[[state]])) {
      nrow(ponds_list[[state]])
    } else {
      0
    }
  })
)

# Add totals
summary_stats <- rbind(
  summary_stats,
  data.frame(
    State = "TOTAL",
    Land_Sites = sum(summary_stats$Land_Sites),
    Ponds = sum(summary_stats$Ponds)
  )
)

# Display interactive table
datatable(summary_stats, 
          caption = "Summary Statistics: Land-Based Sites and Ponds by State",
          options = list(
            pageLength = 10,
            dom = 't',
            columnDefs = list(list(className = 'dt-center', targets = 1:2))
          ),
          rownames = FALSE) %>%
  formatStyle('State',
              target = 'row',
              fontWeight = styleEqual('TOTAL', 'bold')) %>%
  formatStyle(c('Land_Sites', 'Ponds'),
              background = styleColorBar(summary_stats$Ponds, 'lightblue'),
              backgroundSize = '100% 90%',
              backgroundRepeat = 'no-repeat',
              backgroundPosition = 'center')
```

### 4.2 Visual Summary

```{r plot-summary}
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Comparison of land sites and ponds across Australian states"

library(ggplot2)
library(tidyr)

# Prepare data for plotting (exclude TOTAL row)
plot_data <- summary_stats %>%
  filter(State != "TOTAL") %>%
  pivot_longer(cols = c(Land_Sites, Ponds), 
               names_to = "Type", 
               values_to = "Count")

# Create bar plot
ggplot(plot_data, aes(x = State, y = Count, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  scale_fill_viridis_d(labels = c("Land Sites", "Ponds")) +
  labs(
    title = "Land-Based Aquaculture Sites and Ponds by State",
    subtitle = "Comparison across Australian jurisdictions",
    x = "State",
    y = "Count",
    fill = "Feature Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(size = 11),
    legend.position = "bottom"
  )
```

### 4.3 Coverage Analysis

```{r coverage-analysis}
# Calculate coverage percentages
coverage_data <- summary_stats %>%
  filter(State != "TOTAL") %>%
  mutate(
    Has_Data = Land_Sites > 0 | Ponds > 0,
    Has_Both = Land_Sites > 0 & Ponds > 0
  )

total_states <- nrow(coverage_data)
states_with_data <- sum(coverage_data$Has_Data)
states_with_both <- sum(coverage_data$Has_Both)
```

:::{.callout-important}
## Coverage Statistics
- **States analyzed**: `r total_states`
- **States with data**: `r states_with_data` (`r round(100*states_with_data/total_states, 1)`%)
- **States with both sites and ponds**: `r states_with_both` (`r round(100*states_with_both/total_states, 1)`%)
:::

---

## 5. Export & Save Results

### 5.1 Save Interactive Map

```{r save-map}
#| output: true

# Create output directory
output_dir <- here("output_data", "comprehensive_maps")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

# Save map
tryCatch({
  tmap_save(comprehensive_map, 
            filename = here(output_dir, "australia_land_sites_and_ponds.html"),
            selfcontained = TRUE)
  cat("✅ Interactive map saved successfully to:\n")
  cat("   ", here(output_dir, "australia_land_sites_and_ponds.html"), "\n")
}, error = function(e) {
  cat("❌ Error saving HTML map:", conditionMessage(e), "\n")
  cat("Trying alternative save method...\n")
  
  # Alternative: save using htmlwidgets
  tryCatch({
    htmlwidgets::saveWidget(
      tmap_leaflet(comprehensive_map),
      file = here(output_dir, "australia_land_sites_and_ponds.html"),
      selfcontained = TRUE
    )
    cat("✅ Map saved using alternative method!\n")
  }, error = function(e2) {
    cat("❌ Alternative save also failed:", conditionMessage(e2), "\n")
  })
})
```

### 5.2 Save Summary Statistics

```{r save-stats}
#| output: true

# Save summary statistics as CSV
stats_file <- here(output_dir, "summary_statistics.csv")
write.csv(summary_stats, stats_file, row.names = FALSE)

cat("✅ Summary statistics saved to:\n")
cat("   ", stats_file, "\n")
```

---

## 6. Methods & Data Sources

### Data Sources

1. **Pond Detection**
   - Source: Google Earth Engine NDWI analysis
   - Time period: 2021-2022
   - Method: NDWI sum threshold vectorization
   - Location: `output_data/Australian_NDWI_aqua_ponds/2021_2022/`

2. **Aquaculture Sites**
   - Source: Australian state government aquaculture lease databases
   - Type: Land-based aquaculture sites only
   - Location: `raw_data/copy_aquaculture_leases/Australian_Aqua_leases/`

3. **State Boundaries**
   - Source: Australian Bureau of Statistics
   - Year: 2021
   - CRS: GDA2020

### Methodology

1. **Projection**: All spatial data transformed to Australian Albers (EPSG:3577)
2. **Pond Processing**: Dissolved pond geometries to create unified features
3. **Intersection**: Spatial intersection of ponds with land-based sites
4. **Validation**: Geometries validated using `st_make_valid()`

### Software

- **R version**: `r R.version.string`
- **Key packages**:
  - sf: `r packageVersion("sf")`
  - tmap: `r packageVersion("tmap")`
  - dplyr: `r packageVersion("dplyr")`

---

## 7. Conclusions

This analysis successfully identified and mapped aquaculture ponds within official land-based aquaculture sites across Australia. The results provide a comprehensive view of the spatial distribution of aquaculture operations and can inform:

- Resource management decisions
- Environmental monitoring
- Industry development planning
- Research on aquaculture extent and growth

:::{.callout-note}
## Next Steps
- Temporal analysis comparing multiple years
- Pond size distribution analysis
- Species-specific site analysis
- Integration with production data
:::

---

## Appendix: Session Information

```{r session-info}
sessionInfo()
```

---

**Report generated**: `r Sys.time()`

**Contact**: [marina.christofidis@griffithuni.edu.au/ marina.christofidis@gmail.com]

**License**: [to be filled when we know type]
