---
title: "04-QLD_marine_aquaculture_dataset"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: "2025-11-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ==============================================================================
# Create Polygon Buffers for Queensland Marine Aquaculture Farms
# Using Australian Albers Equal Area Projection (EPSG:3577)
# ==============================================================================
# Author: Marina
# Date: 2025-11-10
# Description: Convert point locations to polygon areas using farm size data sourced by DAF/2023
# ==============================================================================
```

# ==============================================================================
# 0. upload packages or install if needed 
# ==============================================================================
```{r}
#Install or upload packages

library(readxl)
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(units)
library(here)
```

# ==============================================================================
# 1. READ AND PARSE DATA
# 2. CREATE POLYGONS USING AUSTRALIAN ALBERS (EPSG:3577)
# 3. SAVE OUTPUTS IN MULTIPLE FORMATS
# 4. SUMMARY STATISTICS
# ==============================================================================
```{r}
# ==============================================================================
# 1. READ AND PARSE DATA
# ==============================================================================
# Set your file path here
file_path <- here("raw_data/QLD/Qld_Marine_Aqua_DAF.xlsx")

cat("\n=== Reading Queensland Marine Aquaculture Data ===\n")

# Read the Excel file
raw_data <- read_excel(file_path, sheet = 1)

# Parse the comma-separated data
qld_aqua <- raw_data %>%
  separate(col = 1, 
           into = c("Authority_RAA", "Current_Status", "Species", "Oyster_area", 
                    "Size_ha", "Latitude", "Longitude", "Depth", "System"),
           sep = ",", extra = "merge") %>%
  mutate(
    # Convert numeric columns
    Size_ha = as.numeric(Size_ha),
    Latitude = as.numeric(Latitude),
    Longitude = as.numeric(Longitude),
    Depth = as.numeric(Depth),
    # Clean text columns
    Authority_RAA = trimws(Authority_RAA),
    Current_Status = trimws(Current_Status),
    Species = trimws(Species),
    Oyster_area = trimws(Oyster_area),
    System = trimws(System),
    # Standardize species names
    Species = case_when(
      Species == "" ~ NA_character_,
      Species == "Oyster" ~ "Oysters",
      TRUE ~ Species
    )
  ) %>%
  # Remove rows with missing coordinates
  filter(!is.na(Latitude) & !is.na(Longitude))

cat("Records loaded:", nrow(qld_aqua), "\n")
cat("Records with size data:", sum(!is.na(qld_aqua$Size_ha)), "\n")
cat("Records missing size data:", sum(is.na(qld_aqua$Size_ha)), "\n")

# Create point spatial object in WGS84
qld_aqua_points_wgs84 <- st_as_sf(qld_aqua, 
                                  coords = c("Longitude", "Latitude"),
                                  crs = 4326,
                                  remove = FALSE)

# ==============================================================================
# 2. CREATE POLYGONS USING AUSTRALIAN ALBERS (EPSG:3577)
# ==============================================================================

cat("\n=== Creating polygon buffers using Australian Albers (EPSG:3577) ===\n")

# Transform to Australian Albers Equal Area projection
qld_aqua_albers <- st_transform(qld_aqua_points_wgs84, crs = 3577)

# Calculate buffer radius from area
# For a circle: Area = π × r²
# Therefore: r = sqrt(Area / π)
# Convert hectares to square meters (1 ha = 10,000 m²)

qld_aqua_albers <- qld_aqua_albers %>%
  mutate(
    area_m2 = Size_ha * 10000,
    buffer_radius_m = sqrt(area_m2 / pi)
  )

# Create circular polygon buffers for farms WITH size data
qld_aqua_polygons_with_size <- qld_aqua_albers %>%
  filter(!is.na(Size_ha)) %>%
  mutate(
    geometry = st_buffer(geometry, dist = buffer_radius_m),
    size_source = "actual"
  )

# Create default buffers for farms WITHOUT size data (use 1 hectare default)
default_area_ha <- 1.0
default_radius_m <- sqrt(default_area_ha * 10000 / pi)

qld_aqua_polygons_no_size <- qld_aqua_albers %>%
  filter(is.na(Size_ha)) %>%
  mutate(
    Size_ha = default_area_ha,
    area_m2 = default_area_ha * 10000,
    buffer_radius_m = default_radius_m,
    geometry = st_buffer(geometry, dist = default_radius_m),
    size_source = "default"
  )

# Combine all polygons
qld_aqua_polygons_albers <- bind_rows(
  qld_aqua_polygons_with_size,
  qld_aqua_polygons_no_size
)

# Calculate actual polygon areas for validation
qld_aqua_polygons_albers <- qld_aqua_polygons_albers %>%
  mutate(
    polygon_area_m2 = as.numeric(st_area(geometry)),
    polygon_area_ha = polygon_area_m2 / 10000,
    area_difference_pct = if_else(
      size_source == "actual",
      ((polygon_area_ha - Size_ha) / Size_ha) * 100,
      NA_real_
    )
  )

cat("\nPolygon creation complete!\n")
cat("Total farms with polygons:", nrow(qld_aqua_polygons_albers), "\n")
cat("  - With actual size data:", sum(qld_aqua_polygons_albers$size_source == "actual"), "\n")
cat("  - With default size:", sum(qld_aqua_polygons_albers$size_source == "default"), "\n")

# Validation check
mean_diff <- mean(abs(qld_aqua_polygons_albers$area_difference_pct), na.rm = TRUE)
cat("\nArea validation (circular buffers):\n")
cat("Mean area difference from input:", round(mean_diff, 4), "%\n")

# ==============================================================================
# 3. SAVE OUTPUTS IN MULTIPLE FORMATS
# ==============================================================================

cat("\n=== Saving outputs ===\n")

# Create output directory
output_dir <- here("output_data", "QLD_marine_sites")
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

# Save in Australian Albers (EPSG:3577) - for analysis
st_write(qld_aqua_polygons_albers, 
         file.path(output_dir, "qld_aquaculture_polygons_albers.shp"),
         delete_dsn = TRUE, quiet = FALSE)

st_write(qld_aqua_polygons_albers, 
         file.path(output_dir, "qld_aquaculture_polygons_albers.geojson"),
         delete_dsn = TRUE, quiet = FALSE)

saveRDS(qld_aqua_polygons_albers, 
        file.path(output_dir, "qld_aquaculture_polygons_albers.rds"))

# Also save in WGS84 (EPSG:4326) - for web mapping and GEE
qld_aqua_polygons_wgs84 <- st_transform(qld_aqua_polygons_albers, crs = 4326)

st_write(qld_aqua_polygons_wgs84, 
         file.path(output_dir, "qld_aquaculture_polygons_wgs84.shp"),
         delete_dsn = TRUE, quiet = FALSE)

st_write(qld_aqua_polygons_wgs84, 
         file.path(output_dir, "qld_aquaculture_polygons_wgs84.geojson"),
         delete_dsn = TRUE, quiet = FALSE)

st_write(qld_aqua_polygons_wgs84, 
         file.path(output_dir, "qld_aquaculture_polygons_wgs84.kml"),
         delete_dsn = TRUE, quiet = FALSE)

saveRDS(qld_aqua_polygons_wgs84, 
        file.path(output_dir, "qld_aquaculture_polygons_wgs84.rds"))

# Also save the original points for comparison
st_write(st_transform(qld_aqua_points_wgs84, crs = 3577), 
         file.path(output_dir, "qld_aquaculture_points_albers.shp"),
         delete_dsn = TRUE, quiet = FALSE)

st_write(qld_aqua_points_wgs84, 
         file.path(output_dir, "qld_aquaculture_points_wgs84.geojson"),
         delete_dsn = TRUE, quiet = FALSE)

cat("\nFiles saved:\n")
cat("Australian Albers (EPSG:3577) - recommended for analysis:\n")
cat("  - qld_aquaculture_polygons_albers.shp\n")
cat("  - qld_aquaculture_polygons_albers.geojson\n")
cat("  - qld_aquaculture_polygons_albers.rds\n")
cat("  - qld_aquaculture_points_albers.shp\n")
cat("\nWGS84 (EPSG:4326) - for web/GEE:\n")
cat("  - qld_aquaculture_polygons_wgs84.shp\n")
cat("  - qld_aquaculture_polygons_wgs84.geojson\n")
cat("  - qld_aquaculture_polygons_wgs84.kml\n")
cat("  - qld_aquaculture_polygons_wgs84.rds\n")
cat("  - qld_aquaculture_points_wgs84.geojson\n")

# ==============================================================================
# 4. SUMMARY STATISTICS
# ==============================================================================

cat("\n=== Summary Statistics ===\n\n")

# Species distribution
cat("Species distribution:\n")
print(table(qld_aqua_polygons_albers$Species, useNA = "ifany"))

# Total area by species
area_by_species <- qld_aqua_polygons_albers %>%
  st_drop_geometry() %>%
  group_by(Species) %>%
  summarise(
    n_farms = n(),
    total_area_ha = sum(polygon_area_ha, na.rm = TRUE),
    mean_area_ha = mean(polygon_area_ha, na.rm = TRUE),
    median_area_ha = median(polygon_area_ha, na.rm = TRUE)
  ) %>%
  arrange(desc(total_area_ha))

cat("\n\nTotal area by species:\n")
print(area_by_species, n = 20)

# Overall statistics
total_area_ha <- sum(qld_aqua_polygons_albers$polygon_area_ha, na.rm = TRUE)
total_area_km2 <- total_area_ha / 100

cat("\n\nTotal Queensland marine aquaculture coverage:\n")
cat("  Total area:", round(total_area_ha, 2), "hectares\n")
cat("  Total area:", round(total_area_km2, 2), "km²\n")

# Size distribution
cat("\n\nFarm size distribution (hectares):\n")
print(summary(qld_aqua_polygons_albers$polygon_area_ha))
```

# ==============================================================================
# 5. VISUALIZATIONS
# 6. SPATIAL ANALYSIS EXAMPLES
# ==============================================================================
```{r}

# 5. VISUALIZATIONS
# ==============================================================================

cat("\n=== Creating visualizations ===\n")

# Map 1: Points vs Polygons comparison
p1 <- ggplot() +
  geom_sf(data = qld_aqua_polygons_albers, 
          aes(fill = Species), 
          alpha = 0.4, color = "gray30", size = 0.2) +
  geom_sf(data = st_transform(qld_aqua_points_wgs84, crs = 3577), 
          aes(color = Species), size = 0.8) +
  scale_fill_brewer(palette = "Set2", na.value = "gray80", 
                    name = "Species (polygons)") +
  scale_color_brewer(palette = "Set2", na.value = "gray50", 
                     name = "Species (points)") +
  theme_minimal() +
  labs(
    title = "Queensland Marine Aquaculture: Points vs Polygons",
    subtitle = paste0("N = ", nrow(qld_aqua_polygons_albers), " farms | Australian Albers projection"),
    x = "Easting (m)",
    y = "Northing (m)"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 8)
  )

ggsave(file.path(output_dir, "map_1_points_vs_polygons.png"),
       p1, width = 14, height = 10, dpi = 300)

# Map 2: Polygons colored by size (log scale)
p2 <- ggplot() +
  geom_sf(data = qld_aqua_polygons_albers %>% filter(!is.na(polygon_area_ha)), 
          aes(fill = polygon_area_ha), 
          color = "gray30", size = 0.1) +
  scale_fill_viridis_c(trans = "log10", 
                       name = "Size (ha)", 
                       option = "plasma",
                       labels = scales::comma) +
  theme_minimal() +
  labs(
    title = "Queensland Marine Aquaculture Farms by Size",
    subtitle = paste0("N = ", nrow(qld_aqua_polygons_albers), " farms | Australian Albers projection"),
    x = "Easting (m)",
    y = "Northing (m)"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 8)
  )

ggsave(file.path(output_dir, "map_2_polygons_by_size.png"),
       p2, width = 14, height = 10, dpi = 300)

# Map 3: Polygons by species
p3 <- ggplot() +
  geom_sf(data = qld_aqua_polygons_albers, 
          aes(fill = Species), 
          color = "gray20", size = 0.2, alpha = 0.7) +
  scale_fill_brewer(palette = "Set2", na.value = "gray80") +
  theme_minimal() +
  labs(
    title = "Queensland Marine Aquaculture Farms by Species",
    subtitle = paste0("N = ", nrow(qld_aqua_polygons_albers), " farms | Australian Albers projection"),
    x = "Easting (m)",
    y = "Northing (m)"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 8)
  )

ggsave(file.path(output_dir, "map_3_polygons_by_species.png"),
       p3, width = 14, height = 10, dpi = 300)

# Map 4: Density by region (simple lat bins)
qld_aqua_summary <- qld_aqua_polygons_albers %>%
  st_drop_geometry() %>%
  mutate(lat_bin = cut(Latitude, breaks = 10)) %>%
  group_by(lat_bin) %>%
  summarise(
    n_farms = n(),
    total_area_ha = sum(polygon_area_ha, na.rm = TRUE)
  )

p4 <- ggplot(qld_aqua_summary, aes(x = lat_bin, y = n_farms)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Farm Distribution by Latitude",
    x = "Latitude bin",
    y = "Number of farms"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 14)
  )

ggsave(file.path(output_dir, "map_4_farm_distribution_latitude.png"),
       p4, width = 10, height = 6, dpi = 300)

cat("\nMaps saved:\n")
cat("  - map_1_points_vs_polygons.png\n")
cat("  - map_2_polygons_by_size.png\n")
cat("  - map_3_polygons_by_species.png\n")
cat("  - map_4_farm_distribution_latitude.png\n")

# ==============================================================================
# 6. SPATIAL ANALYSIS EXAMPLES
# ==============================================================================

cat("\n=== Spatial Analysis Examples ===\n\n")

# Check for overlapping farms
cat("Checking for overlapping farms...\n")
intersections <- st_intersects(qld_aqua_polygons_albers)
overlaps <- sapply(intersections, function(x) length(x) > 1)
n_overlaps <- sum(overlaps)

cat("Farms with potential overlaps:", n_overlaps, "\n")

if (n_overlaps > 0) {
  cat("\nNote: Some overlaps are expected due to circular buffer approximation\n")
  cat("For precise overlap analysis, obtain actual lease boundaries\n")
}

# Regional summary (simple latitude-based regions)
regional_summary <- qld_aqua_polygons_albers %>%
  st_drop_geometry() %>%
  mutate(
    region = case_when(
      Latitude >= -16 ~ "Far North Queensland",
      Latitude >= -20 ~ "North Queensland",
      Latitude >= -24 ~ "Central Queensland",
      Latitude >= -26 ~ "South Queensland",
      TRUE ~ "Southeast Queensland"
    )
  ) %>%
  group_by(region) %>%
  summarise(
    n_farms = n(),
    total_area_ha = sum(polygon_area_ha, na.rm = TRUE),
    dominant_species = names(which.max(table(Species)))
  ) %>%
  arrange(desc(total_area_ha))

cat("\n\nRegional summary:\n")
print(regional_summary, n = 20)

# ==============================================================================
# 7. QUICK RELOAD INSTRUCTIONS
# ==============================================================================

cat("\n=== Processing Complete! ===\n\n")
cat("To quickly reload data in future R sessions:\n\n")
cat('# Load polygons in Australian Albers (for analysis):\n')
cat('qld_polygons <- readRDS("output/qld_aquaculture_polygons_albers.rds")\n\n')
cat('# Or load in WGS84 (for web mapping):\n')
cat('qld_polygons_wgs84 <- readRDS("output/qld_aquaculture_polygons_wgs84.rds")\n\n')
cat('# Load points:\n')
cat('qld_points <- st_read("output/qld_aquaculture_points_wgs84.geojson")\n\n')

cat("\n=== IMPORTANT NOTES ===\n\n")
cat("1. Polygons are CIRCULAR BUFFERS based on farm area\n")
cat("   - This is a reasonable approximation for spatial analysis\n")
cat("   - Actual farm shapes may vary\n\n")
cat("2. Coordinate systems:\n")
cat("   - Australian Albers (EPSG:3577): Use for area calculations and analysis\n")
cat("   - WGS84 (EPSG:4326): Use for web mapping and Google Earth Engine\n\n")
cat("3. Default size:\n")
cat("   - Farms without size data: assigned 1 ha (default)\n")
cat("   - These are flagged in 'size_source' column\n\n")
cat("4. For Google Earth Engine:\n")
cat("   - Upload: qld_aquaculture_polygons_wgs84.geojson\n")
cat("   - Or: qld_aquaculture_points_wgs84.geojson for point-based analysis\n\n")
```
#if I want the food aquaculture apart from the other aquaculture such as pearls that are for jewelery etc...
```{r}
# ==============================================================================
# Filter Queensland Aquaculture Data - Food Species Only
# ==============================================================================
# Create a food category column (recommended for analysis)
# ==============================================================================
qld_all <- readRDS(here('output_data/QLD_marine_sites/qld_aquaculture_polygons_albers.rds'))

qld_categorized <- qld_all %>%
  mutate(
    aquaculture_type = case_when(
      Species %in% c("Oysters", "Scallops", "Sea cucumbers") ~ "Food",
      Species %in% c("Pearls", "Coral", "Sponges") ~ "Non-food",
       is.na(Species) ~ "Unknown",
      TRUE ~ "Other"
    )
  )

# Summary by category
cat("\n=== Aquaculture by Category ===\n")
print(table(qld_categorized$aquaculture_type))

#  easilier to filter what is food and what is not:
qld_food <- qld_categorized %>%
  filter(aquaculture_type == "Food")

# ==============================================================================
#Save filtered datasets
# ==============================================================================

 Shapefile (Albers)
st_write(qld_food, 
         here("output_data", "QLD_marine_sites", "qld_aquaculture_FOOD_ONLY_albers.shp"),
         delete_dsn = TRUE, quiet = FALSE)

# GeoJSON (Albers)
st_write(qld_food, 
         here("output_data", "QLD_marine_sites", "qld_aquaculture_FOOD_ONLY_albers.geojson"),
         delete_dsn = TRUE, quiet = FALSE)

# RDS (Albers)
saveRDS(qld_food, 
        here("output_data", "QLD_marine_sites","qld_aquaculture_FOOD_ONLY_albers.rds"))

# Also save in WGS84 for GEE
qld_food_wgs84 <- st_transform(qld_food, crs = 4326)

st_write(qld_food_wgs84, 
         here("output_data", "QLD_marine_sites", "qld_aquaculture_FOOD_ONLY_wgs84.geojson"),
         delete_dsn = TRUE, quiet = FALSE)

st_write(qld_food_wgs84, 
         here("output_data", "QLD_marine_sites", "qld_aquaculture_FOOD_ONLY_wgs84.kml"),
         delete_dsn = TRUE, quiet = FALSE)

saveRDS(qld_food_wgs84,
        here("output_data","QLD_marine_sites","qld_aquaculture_FOOD_ONLY_wgs84.rds"))


# ==============================================================================
# STATISTICS: Food aquaculture summary
# ==============================================================================

cat("\n=== Food Aquaculture Summary ===\n\n")

# Total area
total_food_area <- sum(qld_food$polygon_area_ha, na.rm = TRUE)
cat("Total food aquaculture area:", round(total_food_area, 2), "hectares\n")
cat("Total food aquaculture area:", round(total_food_area / 100, 2), "km²\n\n")

# By species
food_summary <- qld_food %>%
  st_drop_geometry() %>%
  group_by(Species) %>%
  summarise(
    n_farms = n(),
    total_area_ha = sum(polygon_area_ha, na.rm = TRUE),
    mean_area_ha = mean(polygon_area_ha, na.rm = TRUE),
    median_area_ha = median(polygon_area_ha, na.rm = TRUE),
    min_area_ha = min(polygon_area_ha, na.rm = TRUE),
    max_area_ha = max(polygon_area_ha, na.rm = TRUE)
  ) %>%
  arrange(desc(total_area_ha))

cat("Summary by food species:\n")
print(food_summary)

# Regional distribution of food aquaculture
food_regional <- qld_food %>%
  st_drop_geometry() %>%
  mutate(
    region = case_when(
      Latitude >= -16 ~ "Far North Queensland",
      Latitude >= -20 ~ "North Queensland",
      Latitude >= -24 ~ "Central Queensland",
      Latitude >= -26 ~ "South Queensland",
      TRUE ~ "Southeast Queensland"
    )
  ) %>%
  group_by(region, Species) %>%
  summarise(
    n_farms = n(),
    total_area_ha = sum(polygon_area_ha, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(region, desc(total_area_ha))

cat("\n\nFood aquaculture by region:\n")
print(food_regional, n = 30)
```


```{r}
# ==============================================================================
# VISUALIZATION: Food aquaculture map
# ==============================================================================
# Load Australian states shapefile
AUS_states <- st_read(here("raw_data/AustralianStates_SHP/STE_2021_AUST_GDA2020.shp")) %>%
  st_transform(3577) %>%  # Transform to Australian Albers
  st_make_valid()

# Check the column name - it might be STE_NAME21 not STE_NAME1
cat("Available columns:", paste(names(AUS_states), collapse = ", "), "\n")

# Filter for Queensland
# Note: The column is likely STE_NAME21, adjust if different
QLD_boundary <- AUS_states %>%
  filter(STE_NAME21 == "Queensland")  # Use STE_NAME21 or check your column name

cat("Queensland boundary loaded\n")

# Create figures folder if it doesn't exist
if (!dir.exists(here("figures"))) {
  dir.create(here("figures"), recursive = TRUE)
}

# MAP 1: Food aquaculture with Queensland boundary
p1 <- ggplot() +
  # Queensland boundary (base layer)
  geom_sf(data = QLD_boundary, 
          fill = "gray95", 
          color = "gray30", 
          size = 0.8) +
  # Food aquaculture polygons
  geom_sf(data = qld_food, 
          aes(fill = Species, size = polygon_area_ha),
          alpha = 0.6, 
          color = "gray20",
          stroke = 0.2) +
  scale_fill_manual(values = c("Oysters" = "#2E86AB", "Scallops" = "#A23B72"),
                    name = "Food Species") +
  scale_size_continuous(name = "Area (ha)", 
                       range = c(0.5, 5),
                       trans = "log10") +
  theme_minimal() +
  labs(
    title = "Queensland Food Aquaculture Farms",
    subtitle = paste0("Oysters and Scallops | N = ", nrow(qld_food), " farms"),
    x = "Easting (m)",
    y = "Northing (m)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12),
    legend.position = "right",
    panel.grid.major = element_line(color = "gray90", size = 0.3),
    axis.text = element_text(size = 9)
  )

ggsave(here("figures", "qld_food_aquaculture_with_boundary.png"),
       plot = p1, 
       width = 14, 
       height = 12, 
       dpi = 300)

cat("Map 1 saved: qld_food_aquaculture_with_boundary.png\n")

# MAP 2: Just polygons, colored by species (cleaner version)
p2 <- ggplot() +
  geom_sf(data = QLD_boundary, 
          fill = "aliceblue", 
          color = "gray40", 
          size = 0.6) +
  geom_sf(data = qld_food, 
          aes(fill = Species),
          alpha = 0.7, 
          color = "gray30",
          size = 0.3) +
  scale_fill_manual(values = c("Oysters" = "#2E86AB", "Scallops" = "#A23B72"),
                    name = "Species") +
  theme_minimal() +
  labs(
    title = "Queensland Food Aquaculture Distribution",
    subtitle = paste0("N = ", nrow(qld_food), " farms"),
    x = "Easting (m)",
    y = "Northing (m)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    legend.position = "right",
    panel.grid.major = element_line(color = "gray90", size = 0.3)
  )

ggsave(here("figures", "qld_food_aquaculture_distribution.png"),
       plot = p2, 
       width = 14, 
       height = 12, 
       dpi = 300)
```

#other filtering tips so I can use to write the manuscript
```{r}
# ==============================================================================
# QUICK REFERENCE GUIDE
# ==============================================================================

cat("\n=== QUICK REFERENCE: Filtering Commands ===\n\n")

cat("# Load all data:\n")
cat("qld_all <- readRDS('output/qld_aquaculture_polygons_albers.rds')\n\n")

cat("# Filter for oysters only:\n")
cat("oysters <- qld_all %>% filter(Species == 'Oysters')\n\n")

cat("# Filter for all food species:\n")
cat("food <- qld_all %>% filter(Species %in% c('Oysters', 'Scallops'))\n\n")

cat("# Filter and save:\n")
cat("st_write(food, 'output/food_only.geojson', delete_dsn = TRUE)\n\n")

cat("# Filter by size (e.g., farms > 10 ha):\n")
cat("large_oysters <- qld_all %>%\n")
cat("  filter(Species == 'Oysters' & polygon_area_ha > 10)\n\n")

cat("# Filter by region AND species:\n")
cat("se_qld_oysters <- qld_all %>%\n")
cat("  filter(Species == 'Oysters' & Latitude < -26)\n\n")

```

# Improved Map - Making Aquaculture Farms Visible
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(here)
library(cowplot)  # For combining plots

# Load data
QLD_boundary <- readRDS(here("output_data","QLD_marine_sites", "qld_boundary_albers.rds"))
qld_food <- readRDS(here("output_data","QLD_marine_sites", "qld_aquaculture_FOOD_ONLY_albers.rds"))
# Or if you haven't saved it separately:
qld_all <- readRDS(here("output_data","QLD_marine_sites", "qld_aquaculture_polygons_albers.rds"))
qld_food <- qld_all %>% filter(Species %in% c("Oysters", "Scallops"))

# Create figures folder
if (!dir.exists(here("figures"))) {
  dir.create(here("figures"), recursive = TRUE)
}

# ==============================================================================
# SOLUTION 1: INCREASE POINT/POLYGON SIZE
# ==============================================================================

# Convert polygons to points for better visibility at state scale
qld_food_points <- st_centroid(qld_food)

p1 <- ggplot() +
  geom_sf(data = QLD_boundary, 
          fill = "gray98", 
          color = "gray30", 
          size = 0.6) +
  # Use points instead of polygons for overview map
  geom_sf(data = qld_food_points, 
          aes(color = Species),
          size = 3,  # Larger points
          alpha = 0.8) +
  scale_color_manual(values = c("Oysters" = "#2E86AB", "Scallops" = "#A23B72"),
                     name = "Species") +
  theme_minimal() +
  labs(
    title = "Queensland Food Aquaculture Distribution",
    subtitle = paste0("N = ", nrow(qld_food), " farms (shown as points)"),
    x = "Easting (m)",
    y = "Northing (m)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    legend.position = "right",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    panel.grid.major = element_line(color = "gray90", size = 0.3)
  )

ggsave(here("figures", "qld_food_aquaculture_POINTS.png"),
       plot = p1, 
       width = 14, 
       height = 12, 
       dpi = 300)

cat("Map 1 saved: Points version for visibility\n")

# ==============================================================================
# SOLUTION 2: REGIONAL ZOOM MAPS
# ==============================================================================

# Identify main aquaculture regions
regions <- qld_food %>%
  st_drop_geometry() %>%
  mutate(
    region = case_when(
      Latitude >= -12 & Latitude <= -10 ~ "Far North (Torres Strait)",
      Latitude >= -17 & Latitude <= -15 ~ "North (Cairns area)",
      Latitude >= -20 & Latitude <= -18 ~ "Central North",
      Latitude >= -25 & Latitude <= -23 ~ "Central (Gladstone area)",
      Latitude >= -28 & Latitude <= -26 ~ "South (Moreton Bay)",
      TRUE ~ "Other"
    )
  ) %>%
  group_by(region) %>%
  summarise(n = n()) %>%
  filter(n > 5)  # Only regions with >5 farms

print("Main aquaculture regions:")
print(regions)

# Create a faceted map by latitude bands
qld_food_regions <- qld_food %>%
  mutate(
    lat_region = case_when(
      Latitude >= -12 ~ "Far North (10-12°S)",
      Latitude >= -20 ~ "North (12-20°S)",
      Latitude >= -27 ~ "South (20-27°S)",
      TRUE ~ "Southeast (27-28°S)"
    ),
    lat_region = factor(lat_region, 
                       levels = c("Far North (10-12°S)", 
                                 "North (12-20°S)", 
                                 "South (20-27°S)", 
                                 "Southeast (27-28°S)"))
  )

p2 <- ggplot() +
  geom_sf(data = QLD_boundary, 
          fill = "gray98", 
          color = "gray30", 
          size = 0.3) +
  geom_sf(data = qld_food_regions, 
          aes(fill = Species),
          alpha = 0.7, 
          color = "black",
          size = 0.3) +
  scale_fill_manual(values = c("Oysters" = "#2E86AB", "Scallops" = "#A23B72"),
                    name = "Species") +
  facet_wrap(~lat_region, ncol = 2) +
  theme_minimal() +
  labs(
    title = "Queensland Food Aquaculture by Region",
    subtitle = paste0("N = ", nrow(qld_food), " farms"),
    x = "Easting (m)",
    y = "Northing (m)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom"
  )

ggsave(here("figures", "qld_food_aquaculture_BY_REGION.png"),
       plot = p2, 
       width = 16, 
       height = 14, 
       dpi = 300)

cat("Map 2 saved: Regional faceted maps\n")

# ==============================================================================
# SOLUTION 3: DETAILED ZOOM ON MAIN AREAS
# ==============================================================================

# Find the 3 main concentration areas
# Group farms by proximity
farm_clusters <- qld_food %>%
  mutate(
    cluster = case_when(
      Latitude >= -11 & Latitude <= -10 ~ "Torres Strait",
      Latitude >= -17 & Latitude <= -15.5 ~ "Cairns Region",
      Latitude >= -27.5 & Latitude <= -26.5 ~ "Moreton Bay",
      TRUE ~ "Other"
    )
  ) %>%
  filter(cluster != "Other")

# Get bounding boxes for each cluster with buffer
create_cluster_map <- function(data, cluster_name, buffer_km = 50) {
  
  cluster_data <- data %>% filter(cluster == cluster_name)
  
  if(nrow(cluster_data) == 0) return(NULL)
  
  bbox <- st_bbox(cluster_data)
  buffer_m <- buffer_km * 1000
  
  # Clip QLD boundary to area of interest
  bbox_poly <- st_as_sfc(bbox) %>%
    st_buffer(buffer_m)
  qld_clip <- st_intersection(QLD_boundary, bbox_poly)
  
  ggplot() +
    geom_sf(data = qld_clip, 
            fill = "aliceblue", 
            color = "gray40", 
            size = 0.5) +
    geom_sf(data = cluster_data, 
            aes(fill = Species, size = polygon_area_ha),
            alpha = 0.7, 
            color = "black",
            stroke = 0.3) +
    scale_fill_manual(values = c("Oysters" = "#2E86AB", "Scallops" = "#A23B72"),
                      name = "Species") +
    scale_size_continuous(name = "Area (ha)", 
                         range = c(1, 8),
                         trans = "log10") +
    coord_sf(xlim = c(bbox["xmin"] - buffer_m, bbox["xmax"] + buffer_m),
             ylim = c(bbox["ymin"] - buffer_m, bbox["ymax"] + buffer_m)) +
    theme_minimal() +
    labs(
      title = paste("Food Aquaculture:", cluster_name),
      subtitle = paste0("N = ", nrow(cluster_data), " farms"),
      x = "Easting (m)",
      y = "Northing (m)"
    ) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      legend.position = "right"
    )
}

# Create individual zoom maps
p3a <- create_cluster_map(farm_clusters, "Torres Strait", buffer_km = 30)
p3b <- create_cluster_map(farm_clusters, "Cairns Region", buffer_km = 40)
p3c <- create_cluster_map(farm_clusters, "Moreton Bay", buffer_km = 40)

if(!is.null(p3a)) {
  ggsave(here("figures", "zoom_torres_strait.png"),
         plot = p3a, width = 12, height = 10, dpi = 300)
  cat("Map 3a saved: Torres Strait zoom\n")
}

if(!is.null(p3b)) {
  ggsave(here("figures", "zoom_cairns_region.png"),
         plot = p3b, width = 12, height = 10, dpi = 300)
  cat("Map 3b saved: Cairns region zoom\n")
}

if(!is.null(p3c)) {
  ggsave(here("figures", "zoom_moreton_bay.png"),
         plot = p3c, width = 12, height = 10, dpi = 300)
  cat("Map 3c saved: Moreton Bay zoom\n")
}

# ==============================================================================
# SOLUTION 4: OVERVIEW + INSET MAPS COMBINED
# ==============================================================================

# Main overview map with boxes showing zoom areas
bbox_torres <- if(nrow(filter(farm_clusters, cluster == "Torres Strait")) > 0) {
  st_bbox(filter(farm_clusters, cluster == "Torres Strait")) %>%
    st_as_sfc() %>% st_buffer(30000)
} else NULL

bbox_moreton <- if(nrow(filter(farm_clusters, cluster == "Moreton Bay")) > 0) {
  st_bbox(filter(farm_clusters, cluster == "Moreton Bay")) %>%
    st_as_sfc() %>% st_buffer(40000)
} else NULL

p4_main <- ggplot() +
  geom_sf(data = QLD_boundary, 
          fill = "gray98", 
          color = "gray30", 
          size = 0.6) +
  geom_sf(data = qld_food_points, 
          aes(color = Species),
          size = 2.5,
          alpha = 0.8) +
  # Add boxes for inset areas
  {if(!is.null(bbox_torres)) geom_sf(data = bbox_torres, fill = NA, color = "red", size = 1, linetype = "dashed")} +
  {if(!is.null(bbox_moreton)) geom_sf(data = bbox_moreton, fill = NA, color = "blue", size = 1, linetype = "dashed")} +
  scale_color_manual(values = c("Oysters" = "#2E86AB", "Scallops" = "#A23B72"),
                     name = "Species") +
  theme_minimal() +
  labs(
    title = "Queensland Food Aquaculture",
    subtitle = "Overview with detailed zoom areas highlighted",
    x = NULL, y = NULL
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right",
    axis.text = element_text(size = 8)
  )

ggsave(here("figures", "qld_overview_with_zoom_boxes.png"),
       plot = p4_main, 
       width = 14, 
       height = 12, 
       dpi = 300)

cat("Map 4 saved: Overview with zoom area boxes\n")


```

#Use the output of this script to find aquaculture ifrastructure in Google Earth Engine in the Marine Farms of QLD (most other States have already datasets made that can be downloaded from the GOv web page)- THis file we used here can be ordered from DAF DEpartment of Aquaculture and Fisheries Queensland.


