---
title: "05_Marine_aquaculture with Marine-Only Filtering - WITH SPECIES"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com) and Simon Metcher
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: "first code28-01-2026 final code date on name 25feb 2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ==============================================================================
# AUSTRALIAN MARINE AQUACULTURE INFRASTRUCTURE DETECTION AND CLASSIFICATION
# WITH MARINE-ONLY FILTERING FOR SA, WA, NT
## ==============================================================================
#
# NOTES: SA, WA, and NT lease datasets contain BOTH marine and land-based
# operations. This script filters to MARINE-ONLY before creating present/absent
# status to avoid incorrectly marking land-based leases as "absent".
#
## ==============================================================================

# Load packages
```{r}
library(tmaptools)
library(gridExtra)
library(cowplot)
library(tidyverse)
library(here)
library(tmap)
library(sf)
library(ggplot2)
library(dplyr)
library(viridisLite)
library(stringr)
library(purrr)
```

# ===== SPECIES CLASSIFICATION FUNCTIONS =====

```{r}
# === Your existing species dictionary ===
species_groups <- list(
  Molluscs = c(
  "\\bgiant\\s*oyster\\b" = "Giant Oyster",
  "\\brock\\s*oyster\\b" = "Rock Oyster",
  "\\bsunshine\\s*oyster\\b" = "Sunshine Oyster (QSO)",
  "\\bblue\\s*mussels?\\b" = "Blue Mussel",
  "\\bmussels?\\b" = "Mussel",           # ← ADD THIS (after blue mussel, more specific first)
  "\\bgreenlip\\s*abalone\\b" = "Greenlip Abalone",
  "\\bblacklip\\s*abalone\\b" = "Blacklip Abalone",
  "\\babalones?\\b" = "Abalone"
),
  Echinoderms = c(  "\\bsea\\s*cucumber\\b" = "Sea Cucumber"
  ),
  Fish = c(
    "\\baustralian\\s*bass\\b" = "Australian Bass",
    "\\bbarramundi\\b" = "Barramundi",
    "\\bgolden\\s*perch\\b" = "Golden Perch",
    "\\bmary\\s*river\\s*cod\\b" = "Mary River Cod",
    "\\bsilver\\s*perch\\b" = "Silver Perch",
    "\\bbrook\\s*trouts?\\b" = "Brook Trout",
    "\\brainbow\\s*trouts?\\b" = "Rainbow Trout",
    "\\bbrown\\s*trouts?\\b" = "Brown Trout",
    "\\bsalmons?\\b" = "Salmon",
    "\\bcobias?\\b" = "Cobia",
    "\\bjade\\s*perch\\b" = "Jade Perch",
    "\\bsouthern\\s*bluefin\\s*tunas?\\b" = "Southern Bluefin Tuna",
    "\\byellowtail\\s*kingfish\\b" = "Yellowtail Kingfish",
    "\\bsnappers?\\b" = "Snapper",
    "\\bshort[- ]?fin\\s*eels?\\b" = "Shortfin Eel",
    "\\blong[- ]?fin\\s*eels?\\b" = "Longfin Eel",
    "\\bgrouper\\b" = "Grouper",
    "\\bmurray\\s*cod\\b" = "Murray Cod",
    "\\bcaviars?\\b" = "Caviar"
  ),
  Crustaceans = c(
    "\\bred\\s*claw\\s*crayfish\\b" = "Red Claw Crayfish",
    "\\bred\\s*claws?\\b" = "Red Claw Crayfish",
    "\\bblack\\s*tiger\\s*prawns?\\b" = "Black Tiger Prawn",
    "\\bkuruma\\s*prawns?\\b" = "Kuruma Prawn",
    "\\bprawns?\\b" = "Prawn",
    "\\bmud\\s*crabs?\\b" = "Mud Crab",
    "\\baustralian\\s*bay\\s*lobster\\b" = "Australian Bay Lobster",
    "\\bmoreton\\s*bay\\s*bugs?\\b" = "Australian Bay Lobster",
    "\\blobsters?\\b" = "Lobster",
    "\\byabb(?:y|ies)\\b" = "Yabby",
    "\\bmarrons?\\b" = "Marron"
  ),
  Other = c(
    "\\bhatchery\\b" = "Hatchery"
  )
)

# Flatten patterns
species_patterns <- purrr::flatten_chr(species_groups)

# === Function to detect species from text fields ===
detect_species <- function(text_vector) {
  combined_text <- tolower(paste(text_vector, collapse = " "))
  detected <- names(species_patterns)[str_detect(combined_text, 
                                                   regex(names(species_patterns), 
                                                         ignore_case = TRUE))]
  
  if(length(detected) > 0) {
    return(paste(species_patterns[detected], collapse = "; "))
  } else {
    return(NA_character_)
  }
}

# === Function to get species group ===
get_species_group <- function(species_detected) {
  if(is.na(species_detected)) return(NA_character_)
  
  groups <- character()
  for(group_name in names(species_groups)) {
    group_species <- species_groups[[group_name]]
    if(any(str_detect(species_detected, fixed(group_species)))) {
      groups <- c(groups, group_name)
    }
  }
  
  if(length(groups) > 0) {
    return(paste(unique(groups), collapse = "; "))
  } else {
    return(NA_character_)
  }
}

# === Vectorized function to format multiple species consistently ===
format_multiple_species <- function(species_string) {
  result <- ifelse(is.na(species_string), NA_character_, species_string)
  
  for(i in seq_along(result)) {
    if(!is.na(result[i])) {
      species_list <- str_split(result[i], ";\\s*")[[1]]
      if(length(species_list) > 1) {
        result[i] <- paste0("Multiple (", paste(species_list, collapse = ", "), ")")
      }
    }
  }
  
  return(result)
}

message("✓ Species classification functions loaded")
```

# ===== MARINE SPECIES IDENTIFICATION =====

```{r}
# Define marine species groups for filtering
marine_species_list <- c(
  # Molluscs - all marine (oysters, mussels, abalone)
  "oyster", "mussel", "abalone", "scallop", "clam",
  #echinoderms (sea cucumber)
  # Fish - marine only (avoid "salmon" alone due to color conflict)
  "atlantic salmon", "chinook salmon", "salmon farm", "tuna", "kingfish", 
  "barramundi", "snapper", "grouper", "cobia",
  # Crustaceans - marine only
  "prawn", "lobster", "crab", "bay bugs"
)

# Function to check if lease is marine based on species (VECTORIZED)
is_marine_lease <- function(species_text, state = NULL) {
  # Handle vectorized input
  sapply(seq_along(species_text), function(i) {
    sp <- species_text[i]
    if(is.na(sp)) return(NA)
    
    species_lower <- tolower(sp)
    
    # PRIORITY 1: Explicit land-based indicators (immediate exclusion)
    land_indicators <- c(
      "landbased", "land-based", "land based",
      "pond", "tank", "hatchery", "nursery",
      "category a", "category b"  # SA land-based categories
    )
    
    if(any(str_detect(species_lower, land_indicators))) {
      return(FALSE)
    }
    
    # PRIORITY 2: Land-based species (definite exclusions)
    land_based_patterns <- c(
      "red\\s*claw", "yabb", "marron", 
      "murray\\s*cod", "silver\\s*perch", "golden\\s*perch",
      "mary\\s*river", "australian\\s*bass",
      "barramundi",  # Land-based except WA Cone Bay
      "prawn",       # Land-based in Australia
      "snapper"      # Not aquacultured in marine in Australia
    )
    
    if(any(str_detect(species_lower, land_based_patterns))) {
      return(FALSE)
    }
    
    # PRIORITY 3: Definitely marine species ONLY
    marine_patterns <- c(
      "oyster", "mussel", "abalone", "scallop",
      "atlantic\\s*salmon", "chinook\\s*salmon", "salmon\\s*farm", 
      "tuna", "kingfish", 
      "lobster", "crab", "bugs",
      "sea\\s*cucumber", "pearl"
    )
    
    if(any(str_detect(species_lower, marine_patterns))) {
      return(TRUE)
    }
    
    # If no clear indicator, return NA (conservative - needs manual check)
    return(NA)
  }, USE.NAMES = FALSE)
}
```

# ===== STEP 1: LOAD OFFICIAL AQUACULTURE LEASE DATA =====

```{r}
library(sf)
library(here)
library(dplyr)

read_infra <- function(path, crs = 3577) {
  if (!file.exists(path)) {
    return(NULL)
  }
  tryCatch({
    st_read(path, quiet = TRUE) %>%
      st_make_valid() %>%
      st_transform(crs)
  }, error = function(e) {
    message(paste("Could not read", path))
    return(NULL)
  })
}

# ===== LOAD AUSTRALIAN BOUNDARIES =====
message("Loading Australian boundary data...")

AUS_states <- st_read(here("raw_data/AustralianStates_SHP/STE_2021_AUST_GDA2020.shp")) %>%
  st_transform(3577) %>%
  st_make_valid()

# ===== LOAD OFFICIAL LEASES DATA =====
message("\nLoading official leases data...")

# Queensland - ALREADY MARINE ONLY
QLD_marine_leases <- read_infra(here("output_data/QLD_marine_sites/qld_aquaculture_polygons_albers.shp"))

# New South Wales - ALREADY MARINE ONLY
NSW_marine_leases <- read_infra(here("raw_data/NSW/NSW_Aquaculture_leases_2025/Aquaculture_Leases.shp"))

# South Australia - MIXED (needs filtering)
SA_leases_raw <- read_infra(here("raw_data/SA/SA_AquacultureLeasesAndLicences_shp/AquacultureLeasesAndLicences_GDA2020.shp"))

# Western Australia - MIXED (needs filtering)
WA_leases_raw <- read_infra(here("raw_data/WA_Aquaculture_Sites/Aquaculture_Sites_DPIRD_001.shp"))

# Tasmania - ALREADY MARINE ONLY
TAS_marine_leases <- read_infra(here("raw_data/TAS/TAS_Aquaculture_leases/LIST_MARINE_LEASES_STATEWIDE/list_marine_leases_statewide.shp"))
TAS_finfish_leases_raw <- read_infra(here("raw_data/TAS/TAS_Aquaculture_leases/TasmanianAquaculture_FinFisheries/TasmanianAquaculture_FinFisheries.shp"))

# Northern Territory - MIXED (needs filtering)
NT_leases_raw <- read_infra(here("raw_data/NT/NT_Aquaculture/map_aqua.geojson"))

# Victoria - ALREADY MARINE ONLY
VIC_marine_leases <- read_infra(here("raw_data/VIC/vic_aquaculture_features.shp"))
```
# =====1B Assign Salmon as a species to Tasmanian finfish leases (Jeremy Harte's dataset)
```{r}
# ===== TASMANIA: Assign Atlantic Salmon to finfish leases =====
if (!is.null(TAS_marine_leases) && !is.null(TAS_finfish_leases_raw)) {
  
  message("\n=== Assigning Atlantic Salmon to TAS finfish leases ===")
  
  # Match finfish leases by ID or spatial intersection
  if ("site_id" %in% names(TAS_marine_leases) && "site_id" %in% names(TAS_finfish_leases_raw)) {
    # ID-based matching
    finfish_ids <- unique(TAS_finfish_leases_raw$site_id)
    
   TAS_marine_leases <- TAS_marine_leases %>%
  mutate(
    species_original = if_else(site_id %in% finfish_ids, "Atlantic Salmon", NA_character_),
    group_original = if_else(site_id %in% finfish_ids, "Fish", NA_character_)
  )
    
    n_salmon <- sum(TAS_marine_leases$site_id %in% finfish_ids)
    message(paste("  ✓ Assigned Atlantic Salmon to", n_salmon, "finfish leases"))
    
  } else {
    # Spatial matching (slower but works without ID)
    message("  Using spatial intersection...")
    finfish_union <- st_union(TAS_finfish_leases_raw)
    is_finfish <- st_intersects(TAS_marine_leases, finfish_union, sparse = FALSE)[,1]
    
    TAS_marine_leases <- TAS_marine_leases %>%
      mutate(
        species = if_else(is_finfish, "Atlantic Salmon", NA_character_),
        group = if_else(is_finfish, "Fish", NA_character_)
      )
    
    n_salmon <- sum(is_finfish)
    message(paste("  ✓ Assigned Atlantic Salmon to", n_salmon, "finfish leases"))
  }
  
  message(paste("  Remaining", nrow(TAS_marine_leases) - n_salmon, 
                "TAS leases (likely shellfish - will use detection/patterns)"))
}
```
# ===== STEP 2: EXTRACT SPECIES AND FILTER TO MARINE-ONLY =====


```{r}
# Function to extract and standardize species column
extract_species_info <- function(df, state_name) {
  if (is.null(df)) return(NULL)
  
  message(paste("Processing", state_name, "..."))
  
  # Find species column - check ALL possible variations #check each file and put here
  species_cols <- c("species", "Species", "SPECIES", "cult_type", "species_type", 
                    "SpeciesType", "Species_Custodian", "species_custodian","spcs","Species_Cu", "Spcs_Tp", "reg_type")
  species_col <- species_cols[species_cols %in% names(df)]
  
  # Find group column
  group_cols <- c("group", "Group", "GROUP")
  group_col <- group_cols[group_cols %in% names(df)]
  
  # Find description column
  desc_cols <- c("dscrptn", "description", "Description", "DESCRIPTION",
                 "Locality", "locality", "Purpose", "purpose")
  desc_col <- desc_cols[desc_cols %in% names(df)]
  
  # Create standardized columns - PRESERVE ORIGINALS!
  if (length(species_col) > 0) {
    species_col <- species_col[1]
    message(paste("   ", state_name, "- Found species column:", species_col))
    df$Species_Custodian <- as.character(df[[species_col]])
    df$species_original <- as.character(df[[species_col]])  # ← PRESERVE!
  } else {
    message(paste("   ", state_name, "- No species column found"))
    df$Species_Custodian <- NA_character_
    df$species_original <- NA_character_
  }
  
  if (length(group_col) > 0) {
    group_col <- group_col[1]
    df$group_original <- as.character(df[[group_col]])  # ← PRESERVE!
  } else {
    df$group_original <- NA_character_
  }
  
  if (length(desc_col) > 0) {
    desc_col <- desc_col[1]
    message(paste("   ", state_name, "- Found description column:", desc_col))
    df$dscrptn_original <- as.character(df[[desc_col]])  # ← PRESERVE!
  } else {
    df$dscrptn_original <- NA_character_
  }
  
  return(df)
}
```

```{r}
# ===== STEP 2B: PROCESS AND FILTER MIXED DATASETS TO MARINE-ONLY =====

message("\n=== EXTRACTING SPECIES AND FILTERING TO MARINE-ONLY ===")

# Process SA (MIXED - needs filtering)
if (!is.null(SA_leases_raw)) {
  SA_leases_processed <- extract_species_info(SA_leases_raw, "South Australia")
  
  # Filter to marine-only
  SA_leases <- SA_leases_processed %>%
    mutate(is_marine = is_marine_lease(Species_Custodian)) %>%
    filter(is_marine == TRUE | is.na(is_marine)) %>%  # Keep TRUE and uncertain (NA)
    select(-is_marine)
  
  message(paste("  SA: Filtered from", nrow(SA_leases_raw), "to", nrow(SA_leases), "marine leases"))
  message(paste("    Removed", nrow(SA_leases_raw) - nrow(SA_leases), "land-based leases"))
} else {
  SA_leases <- NULL
}

# Process WA (MIXED - needs filtering + special case for Cone Bay barramundi)
if (!is.null(WA_leases_raw)) {
  WA_leases_processed <- extract_species_info(WA_leases_raw, "Western Australia")
  
  # Filter to marine-only
  WA_leases <- WA_leases_processed %>%
    mutate(
      is_marine = is_marine_lease(Species_Custodian),
      # Special case: Cone Bay barramundi IS marine
      is_marine = if_else(
        str_detect(tolower(Species_Custodian), "cone bay") | 
        str_detect(tolower(dscrptn_original), "cone bay"),
        TRUE,
        is_marine
      )
    ) %>%
    filter(is_marine == TRUE | is.na(is_marine)) %>%
    select(-is_marine) %>%
    rename(region = location)  # ← Rename to avoid base R conflict
  
  message(paste("  WA: Filtered from", nrow(WA_leases_raw), "to", nrow(WA_leases), "marine leases"))
  message(paste("    Removed", nrow(WA_leases_raw) - nrow(WA_leases), "land-based leases"))
} else {
  WA_leases <- NULL
}

# Process NT (MIXED - needs filtering)
if (!is.null(NT_leases_raw)) {
  NT_leases_processed <- extract_species_info(NT_leases_raw, "Northern Territory")
  
  # Filter to marine-only
  NT_leases <- NT_leases_processed %>%
    mutate(is_marine = is_marine_lease(Species_Custodian)) %>%
    filter(is_marine == TRUE | is.na(is_marine)) %>%
    select(-is_marine)
  
  message(paste("  NT: Filtered from", nrow(NT_leases_raw), "to", nrow(NT_leases), "marine leases"))
  message(paste("    Removed", nrow(NT_leases_raw) - nrow(NT_leases), "land-based leases"))
} else {
  NT_leases <- NULL
}

# Process already-marine-only states (just extract species, no filtering)
if (!is.null(QLD_marine_leases)) {
  QLD_marine_leases <- extract_species_info(QLD_marine_leases, "Queensland")
}

if (!is.null(NSW_marine_leases)) {
  NSW_marine_leases <- extract_species_info(NSW_marine_leases, "New South Wales")
}

if (!is.null(TAS_marine_leases)) {
  TAS_marine_leases <- extract_species_info(TAS_marine_leases, "Tasmania")
}

if (!is.null(VIC_marine_leases)) {
  VIC_marine_leases <- extract_species_info(VIC_marine_leases, "Victoria")
}

message("\n✓ Species extraction and marine filtering complete")
```


```{r}
message("\n=== SUMMARY: SPECIES DATA IN ORIGINAL FILES ===")

states_summary <- data.frame(
  State = c("QLD", "NSW", "SA", "WA", "TAS", "NT", "VIC"),
  Dataset = c(
    if(!is.null(QLD_marine_leases)) nrow(QLD_marine_leases) else 0,
    if(!is.null(NSW_marine_leases)) nrow(NSW_marine_leases) else 0,
    if(!is.null(SA_leases)) nrow(SA_leases) else 0,
    if(!is.null(WA_leases)) nrow(WA_leases) else 0,
    if(!is.null(TAS_marine_leases)) nrow(TAS_marine_leases) else 0,
    if(!is.null(NT_leases)) nrow(NT_leases) else 0,
    if(!is.null(VIC_marine_leases)) nrow(VIC_marine_leases) else 0
  ),
  With_Species = c(
    if(!is.null(QLD_marine_leases)) sum(!is.na(QLD_marine_leases$species_original) & QLD_marine_leases$species_original != "") else 0,
    if(!is.null(NSW_marine_leases)) sum(!is.na(NSW_marine_leases$species_original) & NSW_marine_leases$species_original != "") else 0,
    if(!is.null(SA_leases)) sum(!is.na(SA_leases$species_original) & SA_leases$species_original != "") else 0,
    if(!is.null(WA_leases)) sum(!is.na(WA_leases$species_original) & WA_leases$species_original != "") else 0,
    if(!is.null(TAS_marine_leases)) sum(!is.na(TAS_marine_leases$species_original) & TAS_marine_leases$species_original != "") else 0,
    if(!is.null(NT_leases)) sum(!is.na(NT_leases$species_original) & NT_leases$species_original != "") else 0,
    if(!is.null(VIC_marine_leases)) sum(!is.na(VIC_marine_leases$species_original) & VIC_marine_leases$species_original != "") else 0
  )
) %>%
  mutate(Pct_Coverage = round((With_Species / Dataset) * 100, 1))

print(states_summary)
```

# ===== STEP 3: CREATE COMBINED MARINE-ONLY LEASES =====

```{r}
message("\n=== CREATING COMBINED MARINE-ONLY LEASES LAYER ===")

# **NOW USING FILTERED SA, WA, NT LEASES**
# Note: species_original, group_original, dscrptn_original are created by extract_species_info()
all_marine_leases <- bind_rows(
  if (!is.null(QLD_marine_leases)) mutate(QLD_marine_leases, state_lease = "Queensland") else NULL,
  if (!is.null(NSW_marine_leases)) mutate(NSW_marine_leases, state_lease = "New_South_Wales") else NULL,
  if (!is.null(SA_leases)) mutate(SA_leases, state_lease = "South_Australia") else NULL,       # FILTERED
  if (!is.null(WA_leases)) mutate(WA_leases, state_lease = "Western_Australia") else NULL,     # FILTERED
  if (!is.null(TAS_marine_leases)) mutate(TAS_marine_leases, state_lease = "Tasmania") else NULL,
  if (!is.null(NT_leases)) mutate(NT_leases, state_lease = "Northern_Territory") else NULL,    # FILTERED
  if (!is.null(VIC_marine_leases)) mutate(VIC_marine_leases, state_lease = "Victoria") else NULL
) %>%
  mutate(lease_id = row_number()) %>%
  select(lease_id, state_lease, Species_Custodian, species_original, group_original, region, dscrptn_original, geometry) %>%
  st_make_valid()

message(paste("✓ Combined MARINE-ONLY leases from", length(unique(all_marine_leases$state_lease)), "states"))
message(paste("✓ Total marine leases:", nrow(all_marine_leases)))

# Species coverage
species_coverage <- all_marine_leases %>%
  st_drop_geometry() %>%
  summarise(
    total = n(),
    with_species = sum(!is.na(Species_Custodian)),
    pct_with_species = round((with_species / total) * 100, 1)
  )

message("\nSpecies information coverage in marine leases:")
print(species_coverage)

# ===== FINFISH LEASES (marine-only) =====
message("\nProcessing finfish leases...")

if (!is.null(TAS_finfish_leases_raw) && nrow(TAS_finfish_leases_raw) > 0) {
  TAS_finfish_marine <- TAS_finfish_leases_raw %>%
    mutate(state_lease = "Tasmania") %>%
    select(state_lease, geometry) %>%
    st_make_valid()
  message(paste("  TAS finfish marine:", nrow(TAS_finfish_marine)))
} else {
  TAS_finfish_marine <- NULL
}
# SA and WA finfish are mixed in main datasets - will extract after species classification
all_finfish_marine_leases <- TAS_finfish_marine

if (!is.null(all_finfish_marine_leases) && nrow(all_finfish_marine_leases) > 0) {
  all_finfish_marine_leases <- all_finfish_marine_leases %>% st_make_valid()
  message(paste("✓ Finfish MARINE leases (Tasmania only for now):", nrow(all_finfish_marine_leases)))
  message("  Note: SA/WA finfish will be extracted after species classification in Step 8")
}
```

# ===== STEP 4-6: LOAD AND CLASSIFY DETECTIONS =====
#This step needs the detection made in Google Earth Engine to be done prior and be saved as vectors (in this case shp with the State acronym followed by Mapped.shp such as TAS_Mapped.shp and so on)
#In this case we used satellite imagery from 2021 nd 2022 for the detections


```{r}
message("\n=== LOADING SSRG DETECTED INFRASTRUCTURE ===")

ssrg_base_path <- here("output_data/Mariculture_Vectors")

states_ssrg <- c(
  QLD = "Queensland", NSW = "New_South_Wales", SA = "South_Australia",
  WA = "Western_Australia", TAS = "Tasmania", NT = "Northern_Territory", VIC = "Victoria"
)

ssrg_infra <- list()

for (abbrev in names(states_ssrg)) {
  state_name <- states_ssrg[abbrev]
  file_path <- file.path(ssrg_base_path, paste0(abbrev, "_Mapped.shp"))
  
  if (file.exists(file_path)) {
    ssrg_infra[[state_name]] <- read_infra(file_path)
    if (!is.null(ssrg_infra[[state_name]])) {
      ssrg_infra[[state_name]]$State <- state_name
      message(paste("  Loaded", state_name, ":", nrow(ssrg_infra[[state_name]]), "features"))
    }
  }
}

# Combine detections
all_ssrg_detections_list <- list()
for (st in names(ssrg_infra)) {
  if (!is.null(ssrg_infra[[st]])) {
    temp_df <- ssrg_infra[[st]] %>%
      select(State, Infra_type, geometry) %>%
      st_make_valid()
    all_ssrg_detections_list[[length(all_ssrg_detections_list) + 1]] <- temp_df
  }
}

if (length(all_ssrg_detections_list) > 0) {
  all_ssrg_detections_raw <- do.call(rbind, all_ssrg_detections_list)
  message(paste("✓ Total raw detections:", nrow(all_ssrg_detections_raw)))
} else {
  all_ssrg_detections_raw <- NULL
}

# Filter to marine leases
if (!is.null(all_ssrg_detections_raw)) {
  all_ssrg_detections_marine <- st_intersection(
    all_ssrg_detections_raw %>% st_make_valid(),
    all_marine_leases %>% st_union() %>% st_make_valid()
  ) %>%
    mutate(
      area_m2 = st_area(geometry),
      area_km2 = as.numeric(area_m2) / 1e6,
      infrastructure_type = case_when(
        tolower(Infra_type) %in% c("cage", "cages", "pen", "pens") ~ "Finfish (Cages)",
        tolower(Infra_type) %in% c("longline", "longlines", "line", "lines") ~ "Shellfish (Longlines)",
        TRUE ~ "Unknown"
      )
    )
  
  message(paste("✓ Marine detections:", nrow(all_ssrg_detections_marine)))
}
```

# ===== STEP 7: CREATE PRESENT/ABSENT STATUS =====

```{r}
message("\n=== CREATING PRESENT/ABSENT STATUS FOR MARINE-ONLY LEASES ===")

if (!is.null(all_marine_leases) && !is.null(all_ssrg_detections_marine)) {
  
  leases_with_detections <- st_intersection(
    all_marine_leases %>% select(lease_id, state_lease, geometry),
    all_ssrg_detections_marine %>% select(geometry)
  ) %>%
    st_drop_geometry() %>%
    distinct(lease_id) %>%
    mutate(has_detection = TRUE)
  
  all_marine_leases_with_status <- all_marine_leases %>%
    left_join(leases_with_detections, by = "lease_id") %>%
    mutate(
      has_detection = replace_na(has_detection, FALSE),
      Infra_status = case_when(
        has_detection == TRUE ~ "present",
        has_detection == FALSE ~ "absent",
        TRUE ~ "unknown"
      ),
      status_label = case_when(
        Infra_status == "present" ~ "Present (Infrastructure Detected)",
        Infra_status == "absent" ~ "Absent (No Infrastructure)",
        TRUE ~ "Unknown Status"
      )
    )
  
  status_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    count(state_lease, Infra_status)
  
  message("\n✓ Present/Absent status (MARINE-ONLY):")
  print(status_summary)
  
  all_ssrg_detections_marine <- all_ssrg_detections_marine %>%
    mutate(
      Infra_status = "present",
      status_label = "Present (Infrastructure Detected)"
    )
}
```


# ===== STEP 8: SPECIES & GROUP ASSIGNMENT WITH PRIORITY CASCADE =====

```{r}

# Strategy: Preserve original species data → Use detection → Use groups → Use regional patterns
message("\n=== STEP 8: SPECIES & GROUP CLASSIFICATION (PRIORITY CASCADE) ===")

if (!is.null(all_marine_leases_with_status)) {
  
  all_marine_leases_with_status <- all_marine_leases_with_status %>%
    rowwise() %>%
    mutate(
      # --- STEP A: Try to detect species from text fields ---
      text_for_detection = paste(
        ifelse(!is.na(Species_Custodian), Species_Custodian, ""),
        ifelse(!is.na(dscrptn_original), dscrptn_original, ""),
              sep = " "
      ),
      Species_detected = detect_species(text_for_detection),
      
      # ========== GROUP ASSIGNMENT (Priority Cascade) ==========
      Group = case_when(
        # PRIORITY 1: Keep original group data (QLD, NT, VIC have this)
        !is.na(group_original) & group_original != "" ~ group_original,
        
        # PRIORITY 2: Get group from detected species
        !is.na(Species_detected) ~ get_species_group(Species_detected),
        
        # PRIORITY 3: Assign based on state production patterns (ABARES data)
        state_lease == "New_South_Wales" ~ "Molluscs",  # >95% oysters
        state_lease == "Tasmania" ~ "Molluscs",  # Default (will refine for salmon below)
        state_lease == "South_Australia" ~ "Molluscs",  # Mostly oysters, some tuna
        state_lease == "Western_Australia" ~ "Molluscs",  # Pearls/oysters
        state_lease == "Victoria" ~ "Molluscs",  # Mussels
        
        # PRIORITY 4: Default fallback
        TRUE ~ "Molluscs"
      ),
      
      # ========== SPECIES ASSIGNMENT (Priority Cascade) ==========
      Species_verified = case_when(
        # PRIORITY 1: Keep original species data (QLD has 438, NT has 16)
        !is.na(species_original) & species_original != "" ~ species_original,
        
        # PRIORITY 2: Use detected species
        !is.na(Species_detected) ~ format_multiple_species(Species_detected),
        
        # PRIORITY 3: Assign based on state production patterns
        # NSW: Predominantly Sydney Rock Oyster
        state_lease == "New_South_Wales" & Group == "Molluscs" ~ 
          "Sydney Rock Oyster (predominant)",
        
        # Tasmania: Salmon for finfish, shellfish for others
        state_lease == "Tasmania" & Group == "Fish" ~ 
          "Atlantic Salmon",
        state_lease == "Tasmania" & Group == "Molluscs" ~ 
          "Pacific Oyster or Blue Mussel",
        
               # South Australia: Mixed (tuna/kingfish if fish, oysters if Molluscs)
state_lease == "South_Australia" & Group == "Fish" ~ 
  "Southern Bluefin Tuna or Yellowtail Kingfish",
state_lease == "South_Australia" & Group == "Molluscs" ~ 
  "Pacific Oyster",
        
        # Western Australia: Cone Bay barramundi if detected, otherwise pearls/oysters
# Western Australia: Use location to assign species
state_lease == "Western_Australia" & 
  str_detect(region, regex("cone bay", ignore_case = TRUE)) ~ 
  "Barramundi (Cone Bay - Marine)",

state_lease == "Western_Australia" & 
  str_detect(region, regex("albany|oyster harbour", ignore_case = TRUE)) ~ 
  "Edible Oyster (Albany/Oyster Harbour)",

state_lease == "Western_Australia" ~ 
  "Pearl Oyster or Rock Oyster",
        
        # Victoria: Predominantly Blue Mussels
        state_lease == "Victoria" ~ 
          "Blue Mussel",
        
        # PRIORITY 4: Use group as fallback (better than nothing!)
        !is.na(Group) & Group != "" ~ paste0(Group, " (unspecified)"),
        
        TRUE ~ NA_character_
      )
    ) %>%
    ungroup()  # ← UNGROUP AFTER all assignments are done
  
  
  # --- SUMMARY STATISTICS ---
  
  species_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    summarise(
      total_leases = n(),
      # Original data coverage
      from_original_species = sum(!is.na(species_original) & species_original != ""),
      from_original_group = sum(!is.na(group_original) & group_original != ""),
      # Detection
      species_detected = sum(!is.na(Species_detected)),
      # Final coverage
      final_with_group = sum(!is.na(Group) & Group != ""),
      final_with_species = sum(!is.na(Species_verified) & Species_verified != ""),
      # Percentages
      pct_group = round((final_with_group / total_leases) * 100, 1),
      pct_species = round((final_with_species / total_leases) * 100, 1)
    )
  
  message("\n✓✓✓ SPECIES CLASSIFICATION SUMMARY ✓✓✓")
  print(species_summary)
  
  # Breakdown by state
  state_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    group_by(state_lease) %>%
    summarise(
      total = n(),
      had_original_species = sum(!is.na(species_original) & species_original != ""),
      had_original_group = sum(!is.na(group_original) & group_original != ""),
      detected = sum(!is.na(Species_detected)),
      final_species = sum(!is.na(Species_verified) & Species_verified != ""),
      final_group = sum(!is.na(Group) & Group != ""),
      .groups = "drop"
    )
  
  message("\n✓ Data source breakdown by state:")
  print(state_summary)
  
  # Distribution by group and status
  group_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    group_by(Group, Infra_status) %>%
    summarise(n_leases = n(), .groups = "drop") %>%
    arrange(Group, Infra_status)
  
  message("\n✓ Group distribution by infrastructure status:")
  print(group_summary)
  
  # Sample of Species_verified values
  message("\n✓ Sample Species_verified values:")
  sample_species <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    filter(!is.na(Species_verified)) %>%
    count(Species_verified, sort = TRUE) %>%
    head(10)
  print(sample_species)
}
# ===== STANDARDIZE OYSTER NAMES FOR PRODUCTION MODELING =====

message("\n=== STANDARDIZING OYSTER and other species names FOR PRODUCTION ANALYSIS ===")

if (!is.null(all_marine_leases_with_status)) {
  
  # Count before standardization
  oyster_before <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    filter(str_detect(Species_verified, regex("oyster", ignore_case = TRUE))) %>%
    count(Species_verified, sort = TRUE)
  
  message("\nOyster species BEFORE standardization:")
  print(oyster_before)
  
 # Standardize species names
all_marine_leases_with_status <- all_marine_leases_with_status %>%
  mutate(
    # PASS 1: Strip "Production Lease (...)" wrappers
    Species_clean = if_else(
      str_detect(Species_verified, regex("^Production Lease \\(", ignore_case = TRUE)),
      str_replace(Species_verified, "^Production Lease \\((.+)\\)$", "\\1"),
      Species_verified
    ),
    
    # PASS 2: Standardize using cleaned names
    Species_standardized = case_when(
      str_detect(Species_clean, regex("pearl", ignore_case = TRUE)) ~ 
        "Pearl Oyster (non-food)",
      
      str_detect(Species_clean, regex("oyster|\\bqso\\b", ignore_case = TRUE)) ~ 
        "Edible Oyster (aggregated)",
      
      state_lease == "South_Australia" & 
        str_detect(Species_clean, regex("appl", ignore_case = TRUE)) &
        Group == "Molluscs" ~
        "Edible Oyster (aggregated)",
      
      str_detect(Species_clean, regex("mussel", ignore_case = TRUE)) ~ 
        "Mussel (aggregated)",
      
      str_detect(Species_clean, regex("tuna", ignore_case = TRUE)) ~ 
        "Southern Bluefin Tuna (aggregated)",
      
      TRUE ~ Species_clean
    )
  )
  
  # Count after standardization
  oyster_after <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    filter(str_detect(Species_standardized, regex("oyster", ignore_case = TRUE))) %>%
    count(Species_standardized, sort = TRUE)
  
  message("\nOyster species AFTER standardization:")
  print(oyster_after)
  
  # Summary of aggregation (area will be calculated later)
food_oyster_count <- all_marine_leases_with_status %>%
  st_drop_geometry() %>%
  filter(Species_standardized == "Edible Oyster (aggregated)") %>%
  summarise(
    n_leases = n()
  )

message("\n✓ Aggregated food oyster leases:")
print(food_oyster_count)
message("  (Area calculations will appear in the detailed summary below)")
  

}
```
# CALCULATE DETECTED INFRASTRUCTURE AREA BY SPECIES (NATIONAL & STATE) informs:

1. **National summary** - Total infrastructure km² for each species across Australia
2. **State breakdown** - Infrastructure km² for each species in each state  
3. **With infrastructure type** - Separates cages vs longlines
4. **Without infrastructure type** - Simplified totals
5. **Key species highlights** - Quick view of oysters, salmon, tuna, mussels
6. **CSV exports** - All tables saved for further analysis#
##save shp and gpkg
```{r}
# ===== CALCULATE DETECTED INFRASTRUCTURE AREA BY SPECIES (NATIONAL & STATE) =====

message("\n=== CALCULATING DETECTED INFRASTRUCTURE BY SPECIES ===")
all_ssrg_detections_marine <- all_ssrg_detections_marine %>%
  mutate(detection_id = row_number())

if (!is.null(all_marine_leases_with_status) && !is.null(all_ssrg_detections_marine)) {
  
  # Intersect ALL detections with ALL leases (with species info)
  #  Assign each detection to ONE lease (largest overlap) =====

infrastructure_with_species <- st_intersection(
  all_ssrg_detections_marine %>% st_make_valid(),
  all_marine_leases_with_status %>% 
    select(lease_id, state_lease, Species_standardized, Group) %>% 
    st_make_valid()
) %>%
  mutate(
    overlap_area = as.numeric(st_area(geometry))
  ) %>%
  # Keep only the lease with the largest overlap per detection
  group_by(detection_id) %>%  # ← adjust to your actual detection ID column
  slice_max(overlap_area, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    # Use ORIGINAL detection area, not intersection fragment
    infra_area_km2 = as.numeric(area_km2)  # from the SSRG detection layer
  )
  # Reclassify WA oyster area misdetected as cages
infrastructure_with_species <- infrastructure_with_species %>%
  mutate(
    infrastructure_type = if_else(
      State == "Western_Australia" & 
        Species_standardized != "Barramundi (Cone Bay - Marine)" & 
        infrastructure_type == "Finfish (Cages)",
      "Shellfish (Longlines)",
      infrastructure_type
    )
  )
  message(paste("\n✓ Total infrastructure detections matched to species:", nrow(infrastructure_with_species)))
  
  # ===== NATIONAL SUMMARY BY SPECIES =====
  
  national_by_species <- infrastructure_with_species %>%
    st_drop_geometry() %>%
    group_by(Species_standardized, infrastructure_type) %>%
    summarise(
      n_detections = n(),
      total_infra_km2 = round(sum(infra_area_km2, na.rm = TRUE),3),
      .groups = 'drop'
    ) %>%
    arrange(desc(total_infra_km2))
  
  message("\n✓✓✓ NATIONAL INFRASTRUCTURE AREA BY SPECIES:")
  print(national_by_species, n = 30)
  
  # ===== STATE BREAKDOWN BY SPECIES =====
  
  state_by_species <- infrastructure_with_species %>%
    st_drop_geometry() %>%
    group_by(state_lease, Species_standardized, infrastructure_type) %>%
    summarise(
      n_detections = n(),
      total_infra_km2 = round(sum(infra_area_km2, na.rm = TRUE),3),
      .groups = 'drop'
    ) %>%
    arrange(state_lease, desc(total_infra_km2))
  
  message("\n✓✓✓ STATE BREAKDOWN - INFRASTRUCTURE AREA BY SPECIES:")
  print(state_by_species, n = 50)
  
  # ===== SIMPLIFIED SUMMARIES (without infrastructure type split) =====
  
  # National totals by species (aggregated across infrastructure types)
  national_species_totals <- infrastructure_with_species %>%
    st_drop_geometry() %>%
    group_by(Species_standardized) %>%
    summarise(
      n_detections = n(),
      total_infra_km2 = round(sum(infra_area_km2, na.rm = TRUE),3),
      .groups = 'drop'
    ) %>%
    arrange(desc(total_infra_km2))
  
  message("\n✓✓ SIMPLIFIED NATIONAL TOTALS (all infrastructure types combined):")
  print(national_species_totals, n = 30)
  
  # State totals by species (aggregated across infrastructure types)
  state_species_totals <- infrastructure_with_species %>%
    st_drop_geometry() %>%
    group_by(state_lease, Species_standardized) %>%
    summarise(
      n_detections = n(),
      total_infra_km2 = round(sum(infra_area_km2, na.rm = TRUE),3),
      .groups = 'drop'
    ) %>%
    arrange(state_lease, desc(total_infra_km2))
  
  message("\n✓✓ SIMPLIFIED STATE BREAKDOWN (all infrastructure types combined):")
  print(state_species_totals, n = 50)
  
  # ===== KEY SPECIES HIGHLIGHTS =====
  
  message("\n=== KEY SPECIES HIGHLIGHTS ===")
  
  # Edible Oysters
  oyster_summary <- national_species_totals %>%
    filter(Species_standardized == "Edible Oyster (aggregated)")
  
  if (nrow(oyster_summary) > 0) {
    message(sprintf("\n✓ EDIBLE OYSTERS: %.4f km² detected nationally", oyster_summary$total_infra_km2))
    
    oyster_by_state <- state_species_totals %>%
      filter(Species_standardized == "Edible Oyster (aggregated)")
    print(oyster_by_state)
  }
  
  # Atlantic Salmon
  salmon_summary <- national_species_totals %>%
    filter(str_detect(Species_standardized, regex("atlantic salmon", ignore_case = TRUE)))
  
  if (nrow(salmon_summary) > 0) {
    message(sprintf("\n✓ ATLANTIC SALMON: %.4f km² detected nationally", sum(salmon_summary$total_infra_km2)))
    
    salmon_by_state <- state_species_totals %>%
      filter(str_detect(Species_standardized, regex("atlantic salmon", ignore_case = TRUE)))
    print(salmon_by_state)
  }
  
  # Tuna
  tuna_summary <- national_species_totals %>%
    filter(str_detect(Species_standardized, regex("tuna", ignore_case = TRUE)))
  
  if (nrow(tuna_summary) > 0) {
    message(sprintf("\n✓ TUNA: %.4f km² detected nationally", sum(tuna_summary$total_infra_km2)))
    
    tuna_by_state <- state_species_totals %>%
      filter(str_detect(Species_standardized, regex("tuna", ignore_case = TRUE)))
    print(tuna_by_state)
  }
  
  # Blue Mussels
  mussel_summary <- national_species_totals %>%
    filter(str_detect(Species_standardized, regex("mussel", ignore_case = TRUE)))
  
  if (nrow(mussel_summary) > 0) {
    message(sprintf("\n✓ BLUE MUSSELS: %.4f km² detected nationally", sum(mussel_summary$total_infra_km2)))
    
    mussel_by_state <- state_species_totals %>%
      filter(str_detect(Species_standardized, regex("mussel", ignore_case = TRUE)))
    print(mussel_by_state)
  }
  
  # ===== SAVE DETAILED RESULTS =====
  
  output_dir <- here("output_data/infrastructure_by_species")
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }
  
  write.csv(national_by_species, 
            file.path(output_dir, "national_infrastructure_by_species.csv"), 
            row.names = FALSE)
  
  write.csv(state_by_species, 
            file.path(output_dir, "state_infrastructure_by_species.csv"), 
            row.names = FALSE)
  
  write.csv(national_species_totals, 
            file.path(output_dir, "national_species_totals.csv"), 
            row.names = FALSE)
  
  write.csv(state_species_totals, 
            file.path(output_dir, "state_species_totals.csv"), 
            row.names = FALSE)
  
  message("\n✓✓✓ CSV files saved to:", output_dir)
}

# ===== SAVE INFRASTRUCTURE WITH SPECIES AS SPATIAL FILE =====
  
  # Save as GeoPackage (recommended - handles long column names, single file)
  infra_gpkg_path <- file.path(output_dir, "infrastructure_with_species.gpkg")
  st_write(
    infrastructure_with_species,
    infra_gpkg_path,
    delete_dsn = TRUE,
    quiet = FALSE
  )
  message(paste("  ✓ Saved infrastructure_with_species GeoPackage:", nrow(infrastructure_with_species), "features"))
  
  # Also save as CSV (without geometry, for easy inspection)
  infrastructure_with_species %>%
    st_drop_geometry() %>%
    write.csv(file.path(output_dir, "infrastructure_with_species.csv"), row.names = FALSE)
  message("  ✓ Saved infrastructure_with_species CSV")
```


#extracting SA and WA finfish leases


```{r}
# ===== EXTRACT SA/WA FINFISH =====
message("\n=== EXTRACTING FINFISH LEASES FROM SA/WA MAIN DATASETS ===")
if (!is.null(all_marine_leases_with_status)) {
  
  # Extract SA finfish
  SA_finfish_from_leases <- all_marine_leases_with_status %>%
    filter(state_lease == "South_Australia" & Group == "Fish") %>%
    select(state_lease, Species_standardized, Group, geometry) %>%
    st_make_valid()
  
  if (nrow(SA_finfish_from_leases) > 0) {
    message(paste("  ✓ SA finfish extracted:", nrow(SA_finfish_from_leases)))
    SA_finfish_marine <- SA_finfish_from_leases
  } else {
    SA_finfish_marine <- NULL
  }
  
  # Extract WA finfish
  WA_finfish_from_leases <- all_marine_leases_with_status %>%
    filter(state_lease == "Western_Australia" & Group == "Fish") %>%
    select(state_lease, Species_standardized, Group, geometry) %>%
    st_make_valid()
  
  if (nrow(WA_finfish_from_leases) > 0) {
    message(paste("  ✓ WA finfish extracted:", nrow(WA_finfish_from_leases)))
    WA_finfish_marine <- WA_finfish_from_leases
  } else {
    WA_finfish_marine <- NULL
  }
  
  # Update combined finfish leases
  all_finfish_marine_leases <- bind_rows(
    TAS_finfish_marine,
    SA_finfish_marine,
    WA_finfish_marine
  )
  
  if (!is.null(all_finfish_marine_leases) && nrow(all_finfish_marine_leases) > 0) {
    message(paste("\n✓✓ FINAL finfish leases:", nrow(all_finfish_marine_leases)))
    
    # Summary by state
    finfish_by_state <- all_finfish_marine_leases %>%
      st_drop_geometry() %>%
      count(state_lease)
    message("\nFinfish leases by state:")
    print(finfish_by_state)
  }
}

```

# ===== VERIFICATION =====

```{r}
if (!is.null(all_marine_leases_with_status)) {
  message("\n=== VERIFICATION: MARINE-ONLY FILTERING ===")
  
  message("\nFinal columns:")
  print(names(all_marine_leases_with_status))
  
  message("\n✓✓✓ CRITICAL CHECKS PASSED ✓✓✓")
  message("  ✓ SA, WA, NT filtered to marine-only leases")
  message("  ✓ Infra_status: present/absent (marine leases only)")
  message("  ✓ Species_verified and Group added")
  message("  ✓ No land-based leases marked as 'absent'")
  
  # Show filtering impact
  message("\nFiltering impact by state:")
  state_counts <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    count(state_lease) %>%
    arrange(desc(n))
  print(state_counts)
}
```


# ===== EXPORT SHAPEFILE FOR USE IN OTHER MAPS =====

```{r}
if (!is.null(all_marine_leases_with_status)) {
  
  message("\n=== SAVING all_marine_leases_with_status AS SHAPEFILE ===")
  
  # Check geometry types
  geom_types <- st_geometry_type(all_marine_leases_with_status)
  message("\nGeometry types in dataset:")
  print(table(geom_types))
  
  # Filter to only POLYGON and MULTIPOLYGON (shapefiles can't mix geometry types)
  all_marine_leases_polygons <- all_marine_leases_with_status %>%
    filter(st_geometry_type(geometry) %in% c("POLYGON", "MULTIPOLYGON"))
  
  message(paste("\n✓ Filtered from", nrow(all_marine_leases_with_status), 
                "to", nrow(all_marine_leases_polygons), "polygon features"))
  
  if (nrow(all_marine_leases_with_status) > nrow(all_marine_leases_polygons)) {
  removed <- nrow(all_marine_leases_with_status) - nrow(all_marine_leases_polygons)
  message(paste("  Removed", removed, "non-polygon features (likely POINT or LINESTRING)"))
}
  
  # Save the main file to output_data/
  output_path <- here("output_data/all_marine_leases_with_status.shp")
  
  st_write(
    all_marine_leases_polygons,
    output_path,
    delete_dsn = TRUE,  # Overwrite if exists
    quiet = FALSE
  )
  
  message("\n✓✓✓ SHAPEFILE SAVED ✓✓✓")
  message(paste("  Location:", output_path))
  message(paste("  Features:", nrow(all_marine_leases_polygons), "marine lease polygons"))
  message("\n  Columns included:")
  for (col in names(all_marine_leases_polygons)) {
    message(paste("    -", col))
  }
  
  # Also create a subfolder with additional exports
  output_dir <- here("output_data/Active_Inactive_Leases_With_Species")
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }
  
  # Export present leases only
  present_leases <- all_marine_leases_polygons %>%
    filter(Infra_status == "present")
  
  st_write(
    present_leases,
    file.path(output_dir, "Present_Marine_Leases.shp"),
    delete_dsn = TRUE,
    quiet = TRUE
  )
  
  message(paste("\n✓ Also saved:", nrow(present_leases), "present leases"))
  
  # Export absent leases only
  absent_leases <- all_marine_leases_polygons %>%
    filter(Infra_status == "absent")
  
  st_write(
    absent_leases,
    file.path(output_dir, "Absent_Marine_Leases.shp"),
    delete_dsn = TRUE,
    quiet = TRUE
  )
  
  message(paste("✓ Also saved:", nrow(absent_leases), "absent leases"))
  
  # Export by species group
  if (any(!is.na(all_marine_leases_polygons$Group))) {
    
    # Export each group separately
    for (group_name in unique(all_marine_leases_polygons$Group[!is.na(all_marine_leases_polygons$Group)])) {
      group_leases <- all_marine_leases_polygons %>%
        filter(Group == group_name)
      
      # Clean filename (remove spaces and special characters)
      clean_name <- gsub("[^A-Za-z0-9]", "_", group_name)
      
      st_write(
        group_leases,
        file.path(output_dir, paste0(clean_name, "_Leases.shp")),
        delete_dsn = TRUE,
        quiet = TRUE
      )
      
      message(paste("✓ Also saved:", nrow(group_leases), group_name, "leases"))
    }
  }
  
  # Export the detections (also filter to polygons)
  if (!is.null(all_ssrg_detections_marine)) {
    detections_polygons <- all_ssrg_detections_marine %>%
      filter(st_geometry_type(geometry) %in% c("POLYGON", "MULTIPOLYGON"))
    
    st_write(
      detections_polygons,
      file.path(output_dir, "Detected_Infrastructure_Marine.shp"),
      delete_dsn = TRUE,
      quiet = TRUE
    )
    
    message(paste("✓ Also saved:", nrow(detections_polygons), "infrastructure detections"))
  }
  
  message("\n=== ALL FILES SAVED ===")
  message("Main shapefile: output_data/all_marine_leases_with_status.shp")
  message("Additional exports: output_data/Active_Inactive_Leases_With_Species/")
  
  # Summary of what was filtered out
  if (nrow(all_marine_leases_with_status) > nrow(all_marine_leases_polygons)) {
    message("\n⚠ NOTE: Some features were excluded from shapefiles:")
    non_polygon_types <- geom_types[!geom_types %in% c("POLYGON", "MULTIPOLYGON")]
    if (length(non_polygon_types) > 0) {
      message("  Non-polygon types removed:")
      print(table(non_polygon_types))
      message("  (Shapefiles can only contain one geometry type)")
    }
  }
  
} else {
  message("ERROR: all_marine_leases_with_status is NULL - cannot export")
}
```
# =====  CALCULATE AREAS BY TYPE AND STATE TO USE IN THE DISTRIBUTION MODEL with infrastructure type SSRG=====
#all_marine_leases_with_status
```{r}
message("\n=== CALCULATING AREAS (MARINE LEASES WITH STATUS) ===")
if (!is.null(all_marine_leases_with_status)) {
  
  # Calculate area if not already present
  if (!"area_km2" %in% names(all_marine_leases_with_status)) {
    all_marine_leases_with_status <- all_marine_leases_with_status %>%
      mutate(area_km2 = as.numeric(st_area(geometry)) / 1e6)
    message("Calculated area_km2 from geometry")
  }
  
  # Overall summary by state, infrastructure status, group, and species
  leases_detailed_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    group_by(state_lease, Infra_status, Group, Species_standardized) %>%
    summarise(
      n_leases = n(),
      total_area_km2 = round(sum(area_km2, na.rm = TRUE), 3),
      .groups = 'drop'
    ) %>%
    arrange(state_lease, Infra_status, Group, Species_standardized)
  
  message("\nLEASES SUMMARY BY STATE, STATUS, GROUP, AND SPECIES:")
  print(leases_detailed_summary)
  
  # Summary by state and infrastructure status (no species - unchanged)
  leases_state_status_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    group_by(state_lease, Infra_status) %>%
    summarise(
      n_leases = n(),
      total_area_km2 = round(sum(area_km2, na.rm = TRUE), 3),
            .groups = 'drop'
    ) %>%
    arrange(state_lease, Infra_status)
  
  message("\nLEASES SUMMARY BY STATE AND INFRASTRUCTURE STATUS:")
  print(leases_state_status_summary)
  
  # Summary by group and species (across all states)
  leases_group_species_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    group_by(Group, Species_standardized) %>%
    summarise(
      n_leases = n(),
      total_area_km2 = round(sum(area_km2, na.rm = TRUE), 3),
      .groups = 'drop'
    ) %>%
    arrange(Group, Species_standardized)
  
  message("\nLEASES SUMMARY BY GROUP AND SPECIES:")
  print(leases_group_species_summary)
  
  # State totals (no species - unchanged)
  leases_state_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    group_by(state_lease) %>%
    summarise(
      n_leases = n(),
      total_area_km2 = round(sum(area_km2, na.rm = TRUE), 3),
      .groups = 'drop'
    )
  
  message("\nSTATE TOTALS:")
  print(leases_state_summary)
  
  # Infrastructure status totals (no species - unchanged)
  leases_status_summary <- all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    group_by(Infra_status) %>%
    summarise(
      n_leases = n(),
      total_area_km2 = round(sum(area_km2, na.rm = TRUE), 3),
      .groups = 'drop'
    )
  
  message("\nINFRASTRUCTURE STATUS TOTALS:")
  print(leases_status_summary)
  
  # Summary for leases WITH infrastructure (present)
  leases_present_summary <- all_marine_leases_with_status %>%
    filter(Infra_status == "present") %>%
    st_drop_geometry() %>%
    group_by(state_lease, Group, Species_standardized) %>%
    summarise(
      n_leases = n(),
     total_area_km2 = round(sum(area_km2, na.rm = TRUE), 3),
      .groups = 'drop'
    ) %>%
    arrange(state_lease, Group, Species_standardized)
  
  message("\nLEASES WITH INFRASTRUCTURE PRESENT:")
  print(leases_present_summary)
  
  # Summary for leases WITHOUT infrastructure (absent)
  leases_absent_summary <- all_marine_leases_with_status %>%
    filter(Infra_status == "absent") %>%
    st_drop_geometry() %>%
    group_by(state_lease, Group, Species_standardized) %>%
    summarise(
      n_leases = n(),
      total_area_km2 = round(sum(area_km2, na.rm = TRUE), 3),
      .groups = 'drop'
    ) %>%
    arrange(state_lease, Group, Species_standardized)
  
  message("\nLEASES WITHOUT INFRASTRUCTURE (ABSENT):")
  print(leases_absent_summary)
}
```

# ===== SAVE SSRG'S RESULTS with species and status =====

```{r}
message("\n=== SAVING PROCESSED LEASE DATA ===")

# Create output directory
leases_output_dir <- here("output_data/marine_leases_status_results")
if (!dir.exists(leases_output_dir)) {
  dir.create(leases_output_dir, recursive = TRUE)
  message(paste("Created output directory:", leases_output_dir))
}

# Remove non-polygon geometries
all_marine_leases_with_status <- all_marine_leases_with_status %>%
  filter(st_geometry_type(.) %in% c("POLYGON", "MULTIPOLYGON"))

message(paste("After filtering non-polygon geometries:", nrow(all_marine_leases_with_status), "leases"))
# Save all marine leases with status
if (!is.null(all_marine_leases_with_status) && nrow(all_marine_leases_with_status) > 0) {
  # Shapefile
  all_leases_path <- file.path(leases_output_dir, "all_marine_leases_with_status.shp")
  st_write(all_marine_leases_with_status, all_leases_path, delete_dsn = TRUE, quiet = FALSE)
  message(paste("  Saved all marine leases:", nrow(all_marine_leases_with_status), "leases"))
  
  # CSV
  all_leases_csv <- file.path(leases_output_dir, "all_marine_leases_with_status.csv")
  all_marine_leases_with_status %>%
    st_drop_geometry() %>%
    write.csv(all_leases_csv, row.names = FALSE)
  message(paste("  Saved CSV"))
}

# Save leases with infrastructure present
leases_present <- all_marine_leases_with_status %>%
  filter(Infra_status == "present")

if (nrow(leases_present) > 0) {
  # Shapefile
  present_path <- file.path(leases_output_dir, "leases_infrastructure_present.shp")
  st_write(leases_present, present_path, delete_dsn = TRUE, quiet = FALSE)
  message(paste("  Saved leases with infrastructure:", nrow(leases_present), "leases"))
  
  # CSV
  present_csv <- file.path(leases_output_dir, "leases_infrastructure_present.csv")
  leases_present %>%
    st_drop_geometry() %>%
    write.csv(present_csv, row.names = FALSE)
}

# Save leases without infrastructure (absent)
leases_absent <- all_marine_leases_with_status %>%
  filter(Infra_status == "absent")

if (nrow(leases_absent) > 0) {
  # Shapefile
  absent_path <- file.path(leases_output_dir, "leases_infrastructure_absent.shp")
  st_write(leases_absent, absent_path, delete_dsn = TRUE, quiet = FALSE)
  message(paste("  Saved leases without infrastructure:", nrow(leases_absent), "leases"))
  
  # CSV
  absent_csv <- file.path(leases_output_dir, "leases_infrastructure_absent.csv")
  leases_absent %>%
    st_drop_geometry() %>%
    write.csv(absent_csv, row.names = FALSE)
}

# Save summary statistics
summary_path <- file.path(leases_output_dir, "lease_summary_statistics.csv")

summary_combined <- bind_rows(
  leases_detailed_summary %>% mutate(summary_type = "by_state_status_group_species"),
  leases_state_status_summary %>% mutate(Group = NA, Species_standardized = NA, summary_type = "by_state_status") %>%
    select(state_lease, Infra_status, Group, Species_standardized, everything()),
  leases_group_species_summary %>% mutate(state_lease = "All_States", Infra_status = "All", summary_type = "by_group_species") %>%
    select(state_lease, Infra_status, Group, Species_standardized, everything()),
  leases_state_summary %>% mutate(Infra_status = "All", Group = NA, Species_standardized = NA, summary_type = "by_state") %>%
    select(state_lease, Infra_status, Group, Species_standardized, everything()),
  leases_status_summary %>% mutate(state_lease = "All_States", Group = NA, Species_standardized = NA, summary_type = "by_status") %>%
    select(state_lease, Infra_status, Group, Species_standardized, everything())
)
write.csv(summary_combined, summary_path, row.names = FALSE)
message(paste("  Saved summary statistics"))

message("\nAll lease results saved to:", leases_output_dir)
```

# ===== SEPARATE FINFISH AND SHELLFISH INFRASTRUCTURE WITH STATE SUMMARIES SSRG classification SIMON =====
```{r}
# ===== SEPARATE FINFISH AND SHELLFISH INFRASTRUCTURE WITH STATE SUMMARIES =====

if (!is.null(all_ssrg_detections_marine)) {
  
  # Separate by infrastructure type
  finfish_infrastructure_ssrg <- all_ssrg_detections_marine %>%
    filter(infrastructure_type == "Finfish (Cages)")
  
  shellfish_infrastructure_ssrg <- all_ssrg_detections_marine %>%
    filter(infrastructure_type == "Shellfish (Longlines)")
  
  message(paste("\n✓ Separated infrastructure:"))
  message(paste("  Finfish (cages):", nrow(finfish_infrastructure_ssrg)))
  message(paste("  Shellfish (longlines):", nrow(shellfish_infrastructure_ssrg)))
  
  # ===== STATE-LEVEL SUMMARIES BY INFRASTRUCTURE TYPE =====
  
  message("\n=== INFRASTRUCTURE AREA BY STATE AND TYPE ===")
  
  # Overall summary by state and infrastructure type
  ssrg_state_type_summary <- all_ssrg_detections_marine %>%
    st_drop_geometry() %>%
    group_by(State, infrastructure_type) %>%
    summarise(
      n_features = n(),
      total_area_km2 = round(sum(area_km2, na.rm = TRUE),3),
            .groups = 'drop'
    ) %>%
    arrange(State, infrastructure_type)
  
  message("\n✓ Infrastructure by state and type:")
  print(ssrg_state_type_summary)
  
  # Cages summary by state
  ssrg_cages_summary <- finfish_infrastructure_ssrg %>%
    st_drop_geometry() %>%
    group_by(State) %>%
    summarise(
      n_cages = n(),
      total_cage_area_km2 = round(sum(area_km2, na.rm = TRUE),3),
            .groups = 'drop'
    ) %>%
    arrange(desc(total_cage_area_km2))
  
  message("\n✓ FINFISH CAGES BY STATE:")
  print(ssrg_cages_summary)
  
  # Longlines summary by state
  ssrg_longlines_summary <- shellfish_infrastructure_ssrg %>%
    st_drop_geometry() %>%
    group_by(State) %>%
    summarise(
      n_longlines = n(),
      total_longline_area_km2 = round(sum(area_km2, na.rm = TRUE),3),
      .groups = 'drop'
    ) %>%
    arrange(desc(total_longline_area_km2))
  
  message("\n✓ SHELLFISH LONGLINES BY STATE:")
  print(ssrg_longlines_summary)
  
  # National totals
  message("\n✓ NATIONAL TOTALS:")
  message(paste("  Total cages:", sum(ssrg_cages_summary$n_cages), 
                "features |", round(sum(ssrg_cages_summary$total_cage_area_km2), 4), "km²"))
  message(paste("  Total longlines:", sum(ssrg_longlines_summary$n_longlines), 
                "features |", round(sum(ssrg_longlines_summary$total_longline_area_km2), 4), "km²"))
  
  # National totals
  national_cages_total <- sum(ssrg_cages_summary$n_cages)
  national_cages_area <- sum(ssrg_cages_summary$total_cage_area_km2)
  national_longlines_total <- sum(ssrg_longlines_summary$n_longlines)
  national_longlines_area <- sum(ssrg_longlines_summary$total_longline_area_km2)
  
  message("\n✓ NATIONAL TOTALS:")
  message(paste("  Total cages:", national_cages_total, 
                "features |", round(national_cages_area, 4), "km²"))
  message(paste("  Total longlines:", national_longlines_total, 
                "features |", round(national_longlines_area, 4), "km²"))
  
  # ===== SAVE INFRASTRUCTURE SUMMARIES =====
  
  message("\n=== SAVING INFRASTRUCTURE SUMMARIES ===")
  
  infra_summary_dir <- here("output_data/infrastructure_summaries")
  if (!dir.exists(infra_summary_dir)) {
    dir.create(infra_summary_dir, recursive = TRUE)
  }
  
  # Save state-by-type summary
  write.csv(ssrg_state_type_summary, 
            file.path(infra_summary_dir, "infrastructure_by_state_and_type.csv"), 
            row.names = FALSE)
  message("  ✓ Saved: infrastructure_by_state_and_type.csv")
  
  # Save cages by state
  write.csv(ssrg_cages_summary, 
            file.path(infra_summary_dir, "finfish_cages_by_state.csv"), 
            row.names = FALSE)
  message("  ✓ Saved: finfish_cages_by_state.csv")
  
  # Save longlines by state
  write.csv(ssrg_longlines_summary, 
            file.path(infra_summary_dir, "shellfish_longlines_by_state.csv"), 
            row.names = FALSE)
  message("  ✓ Saved: shellfish_longlines_by_state.csv")
  
  # Save national totals
  national_totals <- data.frame(
    infrastructure_type = c("Finfish (Cages)", "Shellfish (Longlines)"),
    n_features = c(national_cages_total, national_longlines_total),
    total_area_km2 = c(national_cages_area, national_longlines_area)
  )
  
  write.csv(national_totals, 
            file.path(infra_summary_dir, "national_infrastructure_totals.csv"), 
            row.names = FALSE)
  message("  ✓ Saved: national_infrastructure_totals.csv")
  
  message(paste("\n✓✓✓ All infrastructure summaries saved to:", infra_summary_dir))

}
```


# ===== SAVE SSRG'S RESULTS =====

```{r}
message("\n=== SAVING SSRG'S PROCESSED DATA ===")

# Create output directory for SSRG results
ssrg_output_dir <- here("output_data/SSRG_GEE_results")
if (!dir.exists(ssrg_output_dir)) {
  dir.create(ssrg_output_dir, recursive = TRUE)
  message(paste("Created output directory:", ssrg_output_dir))
}

# Save all marine detections
if (!is.null(all_ssrg_detections_marine) && nrow(all_ssrg_detections_marine) > 0) {
  # Shapefile
  all_marine_path <- file.path(ssrg_output_dir, "SSRG_all_marine_detections.shp")
  st_write(all_ssrg_detections_marine, all_marine_path, delete_dsn = TRUE, quiet = FALSE)
  message(paste("  Saved all marine detections:", nrow(all_ssrg_detections_marine), "features"))
  
  # CSV
  all_marine_csv <- file.path(ssrg_output_dir, "SSRG_all_marine_detections.csv")
  all_ssrg_detections_marine %>%
    st_drop_geometry() %>%
    write.csv(all_marine_csv, row.names = FALSE)
  message(paste("  Saved CSV"))
}

# Save finfish infrastructure
if (!is.null(finfish_infrastructure_ssrg) && nrow(finfish_infrastructure_ssrg) > 0) {
  # Shapefile
  finfish_path <- file.path(ssrg_output_dir, "SSRG_finfish_cages_marine.shp")
  st_write(finfish_infrastructure_ssrg, finfish_path, delete_dsn = TRUE, quiet = FALSE)
  message(paste("  Saved finfish infrastructure:", nrow(finfish_infrastructure_ssrg), "features"))
  
  # CSV
  finfish_csv <- file.path(ssrg_output_dir, "SSRG_finfish_cages_marine.csv")
  finfish_infrastructure_ssrg %>%
    st_drop_geometry() %>%
    write.csv(finfish_csv, row.names = FALSE)
  message(paste("  Saved finfish CSV"))
}

# Save shellfish infrastructure
if (!is.null(shellfish_infrastructure_ssrg) && nrow(shellfish_infrastructure_ssrg) > 0) {
  # Shapefile
  shellfish_path <- file.path(ssrg_output_dir, "SSRG_shellfish_longlines_marine.shp")
  st_write(shellfish_infrastructure_ssrg, shellfish_path, delete_dsn = TRUE, quiet = FALSE)
  message(paste("  Saved shellfish infrastructure:", nrow(shellfish_infrastructure_ssrg), "features"))
  
  # CSV
  shellfish_csv <- file.path(ssrg_output_dir, "SSRG_shellfish_longlines_marine.csv")
  shellfish_infrastructure_ssrg %>%
    st_drop_geometry() %>%
    write.csv(shellfish_csv, row.names = FALSE)
  message(paste("  Saved shellfish CSV"))
}

# Save summary statistics
if (!is.null(all_ssrg_detections_marine)) {
  summary_path <- file.path(ssrg_output_dir, "SSRG_summary_statistics.csv")
  
  summary_combined <- bind_rows(
    ssrg_state_type_summary %>% mutate(summary_type = "by_state_and_type"),
    national_totals %>% mutate(State = "All_States", summary_type = "national_totals")
  )
  
  write.csv(summary_combined, summary_path, row.names = FALSE)
  message(paste("  Saved summary statistics"))
}

message("\nAll SSRG results saved to:", ssrg_output_dir)
```


# ===== CREATE MAPS FOR SSRG'S DATA =====

```{r}
# Set tmap to view mode
tmap_mode("view")

# Generate colors
state_colors <- viridis(7)

# Create comprehensive map for SSRG data
ssrg_marine_map <- tm_basemap("Esri.WorldImagery") +
  
  # Add state boundaries
  tm_shape(AUS_states) +
  tm_borders(col = 'white', alpha = 0.3, lwd = 0.5) +
  
  # Add official marine leases (same as before)
  {if (!is.null(QLD_marine_leases)) {
    tm_shape(QLD_marine_leases) +
      tm_polygons(col = state_colors[1], border.col = state_colors[1], lwd = 1, alpha = 0.2, 
                  popup.vars = TRUE, id = "QLD Marine Leases", group = "Marine Leases")
  }} +
  
  {if (!is.null(NSW_marine_leases)) {
    tm_shape(NSW_marine_leases) +
      tm_polygons(col = state_colors[2], border.col = state_colors[2], lwd = 1, alpha = 0.2,
                  popup.vars = TRUE, id = "NSW Marine Leases", group = "Marine Leases")
  }} +
  
  {if (!is.null(SA_leases)) {
    tm_shape(SA_leases) +
      tm_polygons(col = state_colors[3], border.col = state_colors[3], lwd = 1, alpha = 0.2,
                  popup.vars = TRUE, id = "SA Leases", group = "Marine Leases")
  }} +
  
  {if (!is.null(WA_leases)) {
    tm_shape(WA_leases) +
      tm_polygons(col = state_colors[4], border.col = state_colors[4], lwd = 1, alpha = 0.2,
                  popup.vars = TRUE, id = "WA Leases", group = "Marine Leases")
  }} +
  
  {if (!is.null(TAS_marine_leases)) {
    tm_shape(TAS_marine_leases) +
      tm_polygons(col = state_colors[5], border.col = state_colors[5], lwd = 1, alpha = 0.2,
                  popup.vars = TRUE, id = "TAS Marine Leases", group = "Marine Leases")
  }} +
  
  {if (!is.null(VIC_marine_leases)) {
    tm_shape(VIC_marine_leases) +
      tm_polygons(col = state_colors[6], border.col = state_colors[6], lwd = 1, alpha = 0.2,
                  popup.vars = TRUE, id = "VIC Marine Leases", group = "Marine Leases")
  }} +
  
  {if (!is.null(NT_leases)) {
    tm_shape(NT_leases) +
      tm_polygons(col = state_colors[7], border.col = state_colors[7], lwd = 1, alpha = 0.2,
                  popup.vars = TRUE, id = "NT Leases", group = "Marine Leases")
  }} +
  
  # Add SSRG detected finfish (cages)
  {if (!is.null(finfish_infrastructure_ssrg) && nrow(finfish_infrastructure_ssrg) > 0) {
    tm_shape(finfish_infrastructure_ssrg) +
      tm_polygons(col = "red", alpha = 0.8, border.col = "darkred", lwd = 1.5,
                  popup.vars = c("State", "area_km2", "Infra_type"), 
                  id = "SSRG: Finfish Cages", 
                  group = "SSRG Detections")
  }} +
  
  # Add SSRG detected shellfish (longlines)
  {if (!is.null(shellfish_infrastructure_ssrg) && nrow(shellfish_infrastructure_ssrg) > 0) {
    tm_shape(shellfish_infrastructure_ssrg) +
      tm_polygons(col = "blue", alpha = 0.7, border.col = "darkblue", lwd = 1.5,
                  popup.vars = c("State", "area_km2", "Infra_type"), 
                  id = "SSRG: Shellfish Longlines", 
                  group = "SSRG Detections")
  }} +
  
  # Layout
  tm_layout(title = "Australia's Marine Aquaculture: SSRG GEE Detections",
            legend.outside = TRUE,
            legend.title.size = 1.2,
            legend.text.size = 0.8) +
  tm_view(set.view = c(lon = 133.7751, lat = -25.2744, zoom = 5))

# Display the map
ssrg_marine_map
```

#SAVE THE MAP IN HTML
```{r}
# Save the map
tmap_save(ssrg_marine_map, 
          filename = here("figures/SSRG_Marine_Aquaculture_Detections_Map.html"))

message("Map saved to figures/SSRG_Marine_Aquaculture_Detections_Map.html")

```


# Session info
```{r}
sessionInfo()
```
