---
title: "06_figures_species_by_state"
Author: Alexandra White and Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: "2025-11-10 to 24 april 2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load required libraries
library(tidyverse)
library(sf)
library(here)
library(dplyr)
library(viridis)
library(viridisLite)

```

#Script for heat map for MC data attribute table and species
# Australian Aquaculture Species Distribution Heatmap. This script creates a visualization showing which species are present in each state
```{r}

# Load required libraries
library(here)
library(tidyverse)
library(viridis)

# Read data
data <- read.csv(here("raw_data/Aquaculture_production_ABARES_DAFF/table_states_species_aquaculture_Australia.csv"))

# Clean the data and rename columns
# Remove rows with empty group (header rows and empty rows)
data <- data %>%
  filter(group != "", !is.na(group)) %>%
  mutate(realm = str_trim(realm))  # Remove leading/trailing whitespace from realm values

# Convert to long format for plotting
data_long <- data %>%
  pivot_longer(cols = NT:WA,
               names_to = "state",
               values_to = "presence") %>%
  filter(presence == "Y")  # Changed from 1 to "Y" based on actual data

# Create ordered factors for better plotting
state_order <- c("NT", "QLD", "NSW", "VIC", "TAS", "SA", "WA")
data_long$state <- factor(data_long$state, levels = state_order)

# Order species by group for cleaner visualization
data_long <- data_long %>%
  arrange(group, popular_name) %>%
  mutate(popular_name = factor(popular_name, levels = unique(popular_name)))

# Create the plot - Option 1: Simple dots with viridis colors
# Using viridis discrete palette for groups
p1 <- ggplot(data_long, aes(x = state, y = popular_name)) +
  geom_point(aes(color = group), size = 3, shape = 16) +
  scale_color_viridis_d(option = "D", begin = 0.2, end = 0.9) +
  labs(x = "Australian State/Territory",
       y = "Species",
       title = "Australian Aquaculture Species Distribution by State",
       color = "Group") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 10, face = "bold"),
        panel.grid.major.x = element_line(color = "grey90"),
        panel.grid.major.y = element_line(color = "grey90"),
        legend.position = "right")

# Save the plot
ggsave(here("figures/species_distribution_dots.png"), p1, width = 10, height = 12, dpi = 300)

# Create the plot - Option 2: Heatmap style with tiles
p2 <- ggplot(data_long, aes(x = state, y = popular_name, fill = group)) +
  geom_tile(color = "white", linewidth = 0.5) +  # Changed 'size' to 'linewidth'
  scale_fill_viridis_d(option = "D", begin = 0.2, end = 0.9) +
  labs(x = "Australian State/Territory",
       y = "Species",
       title = "Australian Aquaculture Species Distribution by State",
       fill = "Group") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 10, face = "bold"),
        panel.grid = element_blank(),
        legend.position = "right")

# Save the heatmap version - FIXED: Removed comma before path string
ggsave(here("figures/species_distribution_heatmap.png"), p2, width = 10, height = 12, dpi = 300)

# Create the plot - Option 3: Faceted by group (cleaner organization)
p3 <- ggplot(data_long, aes(x = state, y = popular_name)) +
  geom_point(aes(color = realm), size = 3, shape = 16) +
  scale_color_viridis_d(option = "C", begin = 0.2, end = 0.8) +
  facet_wrap(~group, scales = "free_y", ncol = 1) +
  labs(x = "Australian State/Territory",
       y = "Species",
       title = "Australian Aquaculture Species Distribution by State",
       color = "Realm") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 10, face = "bold"),
        panel.grid.major.x = element_line(color = "grey90"),
        panel.grid.major.y = element_line(color = "grey90"),
        strip.text = element_text(size = 12, face = "bold"),
        legend.position = "right")

# Save the faceted version 
ggsave(here("figures/species_distribution_faceted.png"), p3, width = 10, height = 14, dpi = 300)

print("All plots have been created successfully!")
print("- species_distribution_dots.png (dot plot)")
print("- species_distribution_heatmap.png (tile/heatmap)")
print("- species_distribution_faceted.png (separated by group)")
```
#making the table_1 as a figure 

```{r}
# Load required libraries
library(here)
library(tidyverse)
library(viridis)

# Create the data
data_availability <- tribble(
  ~STATE, ~MARINE_public, ~LAND_public, ~MARINE_realm, ~LAND_realm, ~MARINE_spatial, ~LAND_spatial, ~MARINE_active, ~LAND_active, ~MARINE_species, ~LAND_species,
  "QLD", "OR", "OR", "NA", "NA", "point (vector)", "cadastre number", "N", "NA", "Y", "P",
  "NSW", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "Y", "NA", "N", "NA",
  "VIC", "Y", "N", "Y", "NA", "point (vector)", "NA", "N", "NA", "N", "NA",
  "TAS", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "N", "NA", "N", "NA",
  "SA", "Y", "Y", "Y", "Y", "polygon (vector)", "polygon (vector)", "Y", "Y", "Y", "N",
  "WA", "Y", "Y", "N", "N", "polygon (vector)", "polygon (vector)", "N", "NA", "N", "N",
  "NT", "OR", "Y", "N", "N", "point (vector)", "point (vector)", "Y", "Y", "Y", "Y"
)

# Reshape to long format for plotting
data_long <- data_availability %>%
  pivot_longer(cols = -STATE, 
               names_to = c("realm", "category"),
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  mutate(
    # Create display value with full words
    display_value = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
      value == "NA" ~ "NA",
      TRUE ~ value  # Keep spatial data types as is
    ),
    # Create a numeric value for coloring
    value_numeric = case_when(
      value == "Y" ~ 3,
      value == "OR" ~ 2,
      value == "P" ~ 2,
      value == "N" ~ 1,
      value == "NA" ~ 0,
      TRUE ~ 3  # Spatial data types get "Yes" color (they exist)
    ),
    # Flag for spatial data types (for adding border)
    is_spatial = category == "spatial" & !value %in% c("Y", "N", "OR", "P", "NA"),
    # Clean up category names for display
    category_label = case_when(
      category == "public" ~ "Data publicly\navailable?",
      category == "realm" ~ "Realm\nspecified?",
      category == "spatial" ~ "Spatial data\ntype",
      category == "active" ~ "Active vs\ninactive specified",
      category == "species" ~ "Species\ninformation provided"
    ),
    # Create ordered factor for categories
    category_label = factor(category_label, 
                           levels = c("Data publicly\navailable?", 
                                    "Realm\nspecified?", 
                                    "Spatial data\ntype", 
                                    "Active vs\ninactive specified", 
                                    "Species\ninformation provided")),
    # Order states
    STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT")),
    # Clean realm labels
    realm = factor(realm, levels = c("MARINE", "LAND"), labels = c("Marine", "Land"))
  )

# Create the heatmap
p <- ggplot(data_long, aes(x = realm, y = STATE)) +
  # Add tiles for all values with color
  geom_tile(aes(fill = value_numeric), 
            color = "white", linewidth = 1) +
  # Add thicker colored border for spatial data types
  geom_tile(data = filter(data_long, is_spatial),
            fill = NA,
            color = "#FDE725FF",  # Viridis yellow color (the "Yes" color)
            linewidth = 2) +
  # Add text labels for all cells using display_value (full words)
  geom_text(aes(label = display_value), 
            size = 3, 
            fontface = "bold",
            color = "grey20") +
  # Use viridis color scale
  scale_fill_viridis_c(option = "D", 
                       begin = 0.25, 
                       end = 0.95,
                       breaks = c(0, 1, 2, 3),
                       labels = c("NA", "No", "Partial/\nOn Request", "Yes"),
                       guide = guide_colorbar(title = "Response",
                                            barwidth = 1,
                                            barheight = 8)) +
  # Facet by category
  facet_grid(. ~ category_label, 
             scales = "free_x") +
  # Labels and theme
  labs(x = NULL,
       y = "State/Territory",
       title = "Australian Aquaculture Data Availability by State and Realm") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold", angle = 0, hjust = 0.5),
    axis.text.y = element_text(size = 10, face = "bold"),
    strip.text.x = element_text(size = 9, face = "bold", margin = margin(t = 5, b = 5)),
    panel.grid = element_blank(),
    panel.spacing = unit(0.5, "lines"),
    legend.position = "right",
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA)
  )

# Save the plot
ggsave(here("figures/data_availability_table.png"), 
       p, 
       width = 14, 
       height = 6, 
       dpi = 300,
       bg = "white")

print("Data availability table created successfully!")
print("Saved as: data_availability_table.png")
```

# 2 separate matrices one for land-based and one for marine for the paper version 2

```{r}
# Load required libraries
library(here)
library(tidyverse)
library(viridis)

# Create the data
data_availability <- tribble(
  ~STATE, ~MARINE_public, ~LAND_public, ~MARINE_realm, ~LAND_realm, ~MARINE_spatial, ~LAND_spatial, ~MARINE_active, ~LAND_active, ~MARINE_species, ~LAND_species,
  "QLD", "OR", "OR", "NA", "NA", "point (vector)", "cadastre number", "N", "NA", "Y", "P",
  "NSW", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "Y", "NA", "N", "NA",
  "VIC", "Y", "N", "Y", "NA", "point (vector)", "NA", "N", "NA", "N", "NA",
  "TAS", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "N", "NA", "N", "NA",
  "SA", "Y", "Y", "Y", "Y", "polygon (vector)", "polygon (vector)", "Y", "Y", "Y", "N",
  "WA", "Y", "Y", "N", "N", "polygon (vector)", "polygon (vector)", "N", "NA", "N", "N",
  "NT", "OR", "Y", "N", "N", "point (vector)", "point (vector)", "Y", "Y", "Y", "Y"
)

# ============================================================================
# Prepare MARINE data
# ============================================================================
marine_data <- data_availability %>%
  select(STATE, starts_with("MARINE_")) %>%
  pivot_longer(cols = -STATE, 
               names_to = "category",
               names_prefix = "MARINE_",
               values_to = "value") %>%
  mutate(
    # Create display value with full words and line breaks for spatial data
    display_value = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
      value == "NA" ~ "No",
      value == "polygon (vector)" ~ "polygon\n(vector)",
      value == "point (vector)" ~ "point\n(vector)",
      value == "cadastre number" ~ "cadastre\nnumber",
      TRUE ~ value  # Keep other spatial data types as is
    ),
    # Create a categorical value for coloring
    value_category = case_when(
      value == "Y" ~ "Yes",
      value == "OR" ~ "Partial/On Request",
      value == "P" ~ "Partial/On Request",
      value == "N" ~ "No",
      value == "NA" ~ "No",
      TRUE ~ "Yes"  # Spatial data types get "Yes" color (they exist)
    ),
    # Create ordered factor for the legend
    value_category = factor(value_category, 
                           levels = c("No", "Partial/On Request", "Yes")),
    # Flag for spatial data types (for adding border)
    is_spatial = category == "spatial" & !value %in% c("Y", "N", "OR", "P", "NA"),
    # Clean up category names for display
    category_label = case_when(
      category == "public" ~ "Data publicly\navailable?",
      category == "realm" ~ "Realm\nspecified?",
      category == "spatial" ~ "Spatial data\ntype",
      category == "active" ~ "Active vs\ninactive specified",
      category == "species" ~ "Species\ninformation provided"
    ),
    # Create ordered factor for categories
    category_label = factor(category_label, 
                           levels = c("Data publicly\navailable?", 
                                    "Realm\nspecified?", 
                                    "Spatial data\ntype", 
                                    "Active vs\ninactive specified", 
                                    "Species\ninformation provided")),
    # Order states
    STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT"))
  )

# Create MARINE plot
p_marine <- ggplot(marine_data, aes(x = STATE, y = category_label)) +
  # Add tiles for all values with color
  geom_tile(aes(fill = value_category), 
            color = "white", linewidth = 1) +
  # Add thicker colored border for spatial data types
  geom_tile(data = filter(marine_data, is_spatial),
            fill = NA,
            color = "#FDE725FF",  # Viridis yellow color (the "Yes" color)
            linewidth = 2) +
  # Add text labels for all cells using display_value (full words)
  geom_text(aes(label = display_value), 
            size = 3.5, 
            fontface = "bold",
            color = "grey20") +
  # Use viridis color scale with discrete colors
  scale_fill_viridis_d(option = "D", 
                       begin = 0.25, 
                       end = 0.95,
                       name = "Response") +
  # Labels and theme
  labs(x = "State/Territory",
       y = NULL,
       title = "Marine Aquaculture Data Availability by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "right",
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA),
    plot.margin = margin(10, 10, 10, 10)
  )

# Save MARINE plot with new name
ggsave(here("figures/data_availability_marine_v2.png"), 
       p_marine, 
       width = 10, 
       height = 6, 
       dpi = 300,
       bg = "white")

# ============================================================================
# Prepare LAND data
# ============================================================================
land_data <- data_availability %>%
  select(STATE, starts_with("LAND_")) %>%
  pivot_longer(cols = -STATE, 
               names_to = "category",
               names_prefix = "LAND_",
               values_to = "value") %>%
  mutate(
    # Create display value with full words and line breaks for spatial data
    display_value = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
      value == "NA" ~ "No",
      value == "polygon (vector)" ~ "polygon\n(vector)",
      value == "point (vector)" ~ "point\n(vector)",
      value == "cadastre number" ~ "cadastre\nnumber",
      TRUE ~ value  # Keep other spatial data types as is
    ),
    # Create a categorical value for coloring
    value_category = case_when(
      value == "Y" ~ "Yes",
      value == "OR" ~ "Partial/On Request",
      value == "P" ~ "Partial/On Request",
      value == "N" ~ "No",
      value == "NA" ~ "No",
      TRUE ~ "Yes"  # Spatial data types get "Yes" color (they exist)
    ),
    # Create ordered factor for the legend
    value_category = factor(value_category, 
                           levels = c("No", "Partial/On Request", "Yes")),
    # Flag for spatial data types (for adding border)
    is_spatial = category == "spatial" & !value %in% c("Y", "N", "OR", "P", "NA"),
    # Clean up category names for display
    category_label = case_when(
      category == "public" ~ "Data publicly\navailable?",
      category == "realm" ~ "Realm\nspecified?",
      category == "spatial" ~ "Spatial data\ntype",
      category == "active" ~ "Active vs\ninactive specified",
      category == "species" ~ "Species\ninformation provided"
    ),
    # Create ordered factor for categories
    category_label = factor(category_label, 
                           levels = c("Data publicly\navailable?", 
                                    "Realm\nspecified?", 
                                    "Spatial data\ntype", 
                                    "Active vs\ninactive specified", 
                                    "Species\ninformation provided")),
    # Order states
    STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT"))
  )

# Create LAND plot
p_land <- ggplot(land_data, aes(x = STATE, y = category_label)) +
  # Add tiles for all values with color
  geom_tile(aes(fill = value_category), 
            color = "white", linewidth = 1) +
  # Add thicker colored border for spatial data types
  geom_tile(data = filter(land_data, is_spatial),
            fill = NA,
            color = "#FDE725FF",  # Viridis yellow color (the "Yes" color)
            linewidth = 2) +
  # Add text labels for all cells using display_value (full words)
  geom_text(aes(label = display_value), 
            size = 3.5, 
            fontface = "bold",
            color = "grey20") +
  # Use viridis color scale with discrete colors
  scale_fill_viridis_d(option = "D", 
                       begin = 0.25, 
                       end = 0.95,
                       name = "Response") +
  # Labels and theme
  labs(x = "State/Territory",
       y = NULL,
       title = "Land-based Aquaculture Data Availability by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "right",
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA),
    plot.margin = margin(10, 10, 10, 10)
  )

# Save LAND plot with new name
ggsave(here("figures/data_availability_land_v2.png"), 
       p_land, 
       width = 10, 
       height = 6, 
       dpi = 300,
       bg = "white")

print("Separate data availability tables created successfully!")
print("Saved as:")
print("  - data_availability_marine_v2.png")
print("  - data_availability_land_v2.png")
```
#version 3 of the matrices (viridis yellow green blue version land and sea separate) 
```{r}
# Load required libraries
library(here)
library(tidyverse)
library(viridis)

# Create the data
data_availability <- tribble(
  ~STATE, ~MARINE_public, ~LAND_public, ~MARINE_realm, ~LAND_realm, ~MARINE_spatial, ~LAND_spatial, ~MARINE_active, ~LAND_active, ~MARINE_species, ~LAND_species,
  "QLD", "OR", "OR", "NA", "NA", "point (vector)", "cadastre number", "N", "NA", "Y", "P",
  "NSW", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "Y", "NA", "N", "NA",
  "VIC", "Y", "N", "Y", "NA", "point (vector)", "NA", "N", "NA", "N", "NA",
  "TAS", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "N", "NA", "N", "NA",
  "SA", "Y", "Y", "Y", "Y", "polygon (vector)", "polygon (vector)", "Y", "Y", "Y", "N",
  "WA", "Y", "Y", "N", "N", "polygon (vector)", "polygon (vector)", "N", "NA", "N", "N",
  "NT", "OR", "Y", "N", "N", "point (vector)", "point (vector)", "Y", "Y", "Y", "Y"
)

# ============================================================================
# Prepare MARINE data
# ============================================================================
marine_data <- data_availability %>%
  select(STATE, starts_with("MARINE_")) %>%
  pivot_longer(cols = -STATE, 
               names_to = "category",
               names_prefix = "MARINE_",
               values_to = "value") %>%
  mutate(
    # Create display value with full words and line breaks for spatial data
    display_value = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
      value == "NA" ~ "No",
      value == "polygon (vector)" ~ "polygon\n(vector)",
      value == "point (vector)" ~ "point\n(vector)",
      value == "cadastre number" ~ "cadastre\nnumber",
      TRUE ~ value  # Keep other spatial data types as is
    ),
    # Create a categorical value for coloring
    value_category = case_when(
      value == "Y" ~ "Yes",
      value == "OR" ~ "Partial/On Request",
      value == "P" ~ "Partial/On Request",
      value == "N" ~ "No",
      value == "NA" ~ "No",
      TRUE ~ "Yes"  # Spatial data types get "Yes" color (they exist)
    ),
    # Create ordered factor for the legend
    value_category = factor(value_category, 
                           levels = c("No", "Partial/On Request", "Yes")),
    # Flag for spatial data types (for adding border)
    is_spatial = category == "spatial" & !value %in% c("Y", "N", "OR", "P", "NA"),
    # Clean up category names for display
    category_label = case_when(
      category == "public" ~ "Data publicly\navailable?",
      category == "realm" ~ "Realm\nspecified?",
      category == "spatial" ~ "Spatial data\ntype",
      category == "active" ~ "Active vs\ninactive specified",
      category == "species" ~ "Species\ninformation provided"
    ),
    # Create ordered factor for categories
    category_label = factor(category_label, 
                           levels = c("Data publicly\navailable?", 
                                    "Realm\nspecified?", 
                                    "Spatial data\ntype", 
                                    "Active vs\ninactive specified", 
                                    "Species\ninformation provided")),
    # Order states
    STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT"))
  )

# Create MARINE plot
p_marine <- ggplot(marine_data, aes(x = STATE, y = category_label)) +
  # Add tiles for all values with color
  geom_tile(aes(fill = value_category), 
            color = "white", linewidth = 1) +
  # Add text labels for all cells using display_value (full words)
  geom_text(aes(label = display_value), 
            size = 3.5, 
            fontface = "bold",
            color = "grey20") +
  # Use viridis color scale with discrete colors
  scale_fill_viridis_d(option = "D", 
                       begin = 0.35, 
                       end = 0.95,
                       name = "Response") +
  # Labels and theme
  labs(x = "State/Territory",
       y = NULL,
       title = "Marine Aquaculture Data Availability by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA),
    plot.margin = margin(10, 10, 10, 10)
  )

# Save MARINE plot with new name
ggsave(here("figures/data_availability_marine_v3.png"), 
       p_marine, 
       width = 10, 
       height = 6, 
       dpi = 300,
       bg = "white")

# ============================================================================
# Prepare LAND data
# ============================================================================
land_data <- data_availability %>%
  select(STATE, starts_with("LAND_")) %>%
  pivot_longer(cols = -STATE, 
               names_to = "category",
               names_prefix = "LAND_",
               values_to = "value") %>%
  mutate(
    # Create display value with full words and line breaks for spatial data
    display_value = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
      value == "NA" ~ "No",
      value == "polygon (vector)" ~ "polygon\n(vector)",
      value == "point (vector)" ~ "point\n(vector)",
      value == "cadastre number" ~ "cadastre\nnumber",
      TRUE ~ value  # Keep other spatial data types as is
    ),
    # Create a categorical value for coloring
    value_category = case_when(
      value == "Y" ~ "Yes",
      value == "OR" ~ "Partial/On Request",
      value == "P" ~ "Partial/On Request",
      value == "N" ~ "No",
      value == "NA" ~ "No",
      TRUE ~ "Yes"  # Spatial data types get "Yes" color (they exist)
    ),
    # Create ordered factor for the legend
    value_category = factor(value_category, 
                           levels = c("No", "Partial/On Request", "Yes")),
    # Flag for spatial data types (for adding border)
    is_spatial = category == "spatial" & !value %in% c("Y", "N", "OR", "P", "NA"),
    # Clean up category names for display
    category_label = case_when(
      category == "public" ~ "Data publicly\navailable?",
      category == "realm" ~ "Realm\nspecified?",
      category == "spatial" ~ "Spatial data\ntype",
      category == "active" ~ "Active vs\ninactive specified",
      category == "species" ~ "Species\ninformation provided"
    ),
    # Create ordered factor for categories
    category_label = factor(category_label, 
                           levels = c("Data publicly\navailable?", 
                                    "Realm\nspecified?", 
                                    "Spatial data\ntype", 
                                    "Active vs\ninactive specified", 
                                    "Species\ninformation provided")),
    # Order states
    STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT"))
  )

# Create LAND plot
p_land <- ggplot(land_data, aes(x = STATE, y = category_label)) +
  # Add tiles for all values with color
  geom_tile(aes(fill = value_category), 
            color = "white", linewidth = 1) +
  # Add text labels for all cells using display_value (full words)
  geom_text(aes(label = display_value), 
            size = 3.5, 
            fontface = "bold",
            color = "grey20") +
  # Use viridis color scale with discrete colors
  scale_fill_viridis_d(option = "D", 
                       begin = 0.35, 
                       end = 0.95,
                       name = "Response") +
  # Labels and theme
  labs(x = "State/Territory",
       y = NULL,
       title = "Land-based Aquaculture Data Availability by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA),
    plot.margin = margin(10, 10, 10, 10)
  )

# Save LAND plot with new name
ggsave(here("figures/data_availability_land_v3.png"), 
       p_land, 
       width = 10, 
       height = 6, 
       dpi = 300,
       bg = "white")

print("Separate data availability tables created successfully!")
print("Saved as:")
print("  - data_availability_marine_v3.png")
print("  - data_availability_land_v3.png")
```
#V4 version is mako colors for readability
```{r}
# Load required libraries
library(here)
library(tidyverse)
library(viridis)

# Create the data
data_availability <- tribble(
  ~STATE, ~MARINE_public, ~LAND_public, ~MARINE_realm, ~LAND_realm, ~MARINE_spatial, ~LAND_spatial, ~MARINE_active, ~LAND_active, ~MARINE_species, ~LAND_species,
  "QLD", "OR", "OR", "NA", "NA", "point (vector)", "cadastre number", "N", "NA", "Y", "P",
  "NSW", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "Y", "NA", "N", "NA",
  "VIC", "Y", "N", "Y", "NA", "point (vector)", "NA", "N", "NA", "N", "NA",
  "TAS", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "N", "NA", "N", "NA",
  "SA", "Y", "Y", "Y", "Y", "polygon (vector)", "polygon (vector)", "Y", "Y", "Y", "N",
  "WA", "Y", "Y", "N", "N", "polygon (vector)", "polygon (vector)", "N", "NA", "N", "N",
  "NT", "OR", "Y", "N", "N", "point (vector)", "point (vector)", "Y", "Y", "Y", "Y"
)

# ============================================================================
# Prepare MARINE data
# ============================================================================
marine_data <- data_availability %>%
  select(STATE, starts_with("MARINE_")) %>%
  pivot_longer(cols = -STATE, 
               names_to = "category",
               names_prefix = "MARINE_",
               values_to = "value") %>%
  mutate(
    # Create display value with full words and line breaks for spatial data
    display_value = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
      value == "NA" ~ "No",
      value == "polygon (vector)" ~ "polygon\n(vector)",
      value == "point (vector)" ~ "point\n(vector)",
      value == "cadastre number" ~ "cadastre\nnumber",
      TRUE ~ value  # Keep other spatial data types as is
    ),
    # Create a categorical value for coloring
    value_category = case_when(
      value == "Y" ~ "Yes",
      value == "OR" ~ "Partial/On Request",
      value == "P" ~ "Partial/On Request",
      value == "N" ~ "No",
      value == "NA" ~ "No",
      TRUE ~ "Yes"  # Spatial data types get "Yes" color (they exist)
    ),
    # Create ordered factor for the legend
    value_category = factor(value_category, 
                           levels = c("No", "Partial/On Request", "Yes")),
    # Flag for spatial data types (for adding border)
    is_spatial = category == "spatial" & !value %in% c("Y", "N", "OR", "P", "NA"),
    # Clean up category names for display
    category_label = case_when(
      category == "public" ~ "Data publicly\navailable?",
      category == "realm" ~ "Realm\nspecified?",
      category == "spatial" ~ "Spatial data\ntype",
      category == "active" ~ "Active vs\ninactive specified",
      category == "species" ~ "Species\ninformation provided"
    ),
    # Create ordered factor for categories
    category_label = factor(category_label, 
                           levels = c("Data publicly\navailable?", 
                                    "Realm\nspecified?", 
                                    "Spatial data\ntype", 
                                    "Active vs\ninactive specified", 
                                    "Species\ninformation provided")),
    # Order states
    STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT"))
  )

# Create MARINE plot
p_marine <- ggplot(marine_data, aes(x = STATE, y = category_label)) +
  # Add tiles for all values with color
  geom_tile(aes(fill = value_category), 
            color = "white", linewidth = 1) +
  # Add text labels for all cells using display_value (full words)
  geom_text(aes(label = display_value), 
            size = 3.5, 
            fontface = "bold",
            color = "grey20") +
  # Use mako color scale with lighter colors for better readability
  scale_fill_viridis_d(option = "G",  # G is mako
                       begin = 0.40,  # Start lighter (skip darkest colors)
                       end = 0.95,    # End before the lightest
                       name = "Response") +
  # Labels and theme
  labs(x = "State/Territory",
       y = NULL,
       title = "Marine Aquaculture Data Availability by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA),
    plot.margin = margin(10, 10, 10, 10)
  )

# Save MARINE plot with new name
ggsave(here("figures/data_availability_marine_v4.png"), 
       p_marine, 
       width = 10, 
       height = 6, 
       dpi = 300,
       bg = "white")

# ============================================================================
# Prepare LAND data
# ============================================================================
land_data <- data_availability %>%
  select(STATE, starts_with("LAND_")) %>%
  pivot_longer(cols = -STATE, 
               names_to = "category",
               names_prefix = "LAND_",
               values_to = "value") %>%
  mutate(
    # Create display value with full words and line breaks for spatial data
    display_value = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
      value == "NA" ~ "No",
      value == "polygon (vector)" ~ "polygon\n(vector)",
      value == "point (vector)" ~ "point\n(vector)",
      value == "cadastre number" ~ "cadastre\nnumber",
      TRUE ~ value  # Keep other spatial data types as is
    ),
    # Create a categorical value for coloring
    value_category = case_when(
      value == "Y" ~ "Yes",
      value == "OR" ~ "Partial/On Request",
      value == "P" ~ "Partial/On Request",
      value == "N" ~ "No",
      value == "NA" ~ "No",
      TRUE ~ "Yes"  # Spatial data types get "Yes" color (they exist)
    ),
    # Create ordered factor for the legend
    value_category = factor(value_category, 
                           levels = c("No", "Partial/On Request", "Yes")),
    # Flag for spatial data types (for adding border)
    is_spatial = category == "spatial" & !value %in% c("Y", "N", "OR", "P", "NA"),
    # Clean up category names for display
    category_label = case_when(
      category == "public" ~ "Data publicly\navailable?",
      category == "realm" ~ "Realm\nspecified?",
      category == "spatial" ~ "Spatial data\ntype",
      category == "active" ~ "Active vs\ninactive specified",
      category == "species" ~ "Species\ninformation provided"
    ),
    # Create ordered factor for categories
    category_label = factor(category_label, 
                           levels = c("Data publicly\navailable?", 
                                    "Realm\nspecified?", 
                                    "Spatial data\ntype", 
                                    "Active vs\ninactive specified", 
                                    "Species\ninformation provided")),
    # Order states
    STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT"))
  )

# Create LAND plot
p_land <- ggplot(land_data, aes(x = STATE, y = category_label)) +
  # Add tiles for all values with color
  geom_tile(aes(fill = value_category), 
            color = "white", linewidth = 1) +
  # Add text labels for all cells using display_value (full words)
  geom_text(aes(label = display_value), 
            size = 3.5, 
            fontface = "bold",
            color = "grey20") +
  # Use mako color scale with lighter colors for better readability
  scale_fill_viridis_d(option = "G",  # G is mako
                       begin = 0.40,  # Start lighter (skip darkest colors)
                       end = 0.95,    # End before the lightest
                       name = "Response") +
  # Labels and theme
  labs(x = "State/Territory",
       y = NULL,
       title = "Land-based Aquaculture Data Availability by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA),
    plot.margin = margin(10, 10, 10, 10)
  )

# Save LAND plot with new name
ggsave(here("figures/data_availability_land_v4.png"), 
       p_land, 
       width = 10, 
       height = 6, 
       dpi = 300,
       bg = "white")

print("Separate data availability tables created successfully!")
print("Saved as:")
print("  - data_availability_marine_v4.png")
print("  - data_availability_land_v4.png")
```


#alternative visualizations 


```{r}
# Load required libraries
library(here)
library(tidyverse)
library(viridis)
library(patchwork)  # For combining plots

# Create the data
data_availability <- tribble(
  ~STATE, ~MARINE_public, ~LAND_public, ~MARINE_realm, ~LAND_realm, ~MARINE_spatial, ~LAND_spatial, ~MARINE_active, ~LAND_active, ~MARINE_species, ~LAND_species,
  "QLD", "OR", "OR", "NA", "NA", "point (vector)", "cadastre number", "N", "NA", "Y", "P",
  "NSW", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "Y", "NA", "N", "NA",
  "VIC", "Y", "N", "Y", "NA", "point (vector)", "NA", "N", "NA", "N", "NA",
  "TAS", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "N", "NA", "N", "NA",
  "SA", "Y", "Y", "Y", "Y", "polygon (vector)", "polygon (vector)", "Y", "Y", "Y", "N",
  "WA", "Y", "Y", "N", "N", "polygon (vector)", "polygon (vector)", "N", "NA", "N", "N",
  "NT", "OR", "Y", "N", "N", "point (vector)", "point (vector)", "Y", "Y", "Y", "Y"
)

# ============================================================================
# OPTION 1: Summary bar chart - Count of states with each response type
# ============================================================================

# Reshape data
data_long <- data_availability %>%
  pivot_longer(cols = -STATE, 
               names_to = c("realm", "category"),
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  mutate(
    value_clean = case_when(
      value == "Y" ~ "Yes",
      value == "N" ~ "No",
      value == "OR" ~ "On Request",
      value == "P" ~ "Partial",
       value == "NA" ~ "No",
      TRUE ~ "Data exists"  # For spatial data types
    ),
    category_label = case_when(
      category == "public" ~ "Data publicly available",
      category == "realm" ~ "Realm specified",
      category == "spatial" ~ "Spatial data type provided",
      category == "active" ~ "Active/inactive specified",
      category == "species" ~ "Species information provided"
    ),
    realm = factor(realm, levels = c("MARINE", "LAND"), labels = c("Marine", "Land"))
  )

# Count by category, realm, and response
summary_counts <- data_long %>%
  count(category_label, realm, value_clean) %>%
  mutate(
    value_clean = factor(value_clean, 
                        levels = c("No", "Partial", "On Request", "Data exists", "Yes"))
  )

# Create stacked bar chart
p1 <- ggplot(summary_counts, aes(x = category_label, y = n, fill = value_clean)) +
  geom_col(position = "stack", color = "white", linewidth = 0.5) +
  geom_text(aes(label = ifelse(n > 0, n, "")), 
            position = position_stack(vjust = 0.5),
            size = 3.5, 
            fontface = "bold",
            color = "grey20") +
  scale_fill_viridis_d(option = "D", begin = 0.25, end = 0.95, direction = 1) +
  facet_wrap(~realm, ncol = 1) +
  coord_flip() +
  labs(title = "Number of States/Territories by Data Availability Category",
       subtitle = "Stacked bar chart showing response counts",
       x = NULL,
       y = "Number of States/Territories",
       fill = "Response") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40")
  )

ggsave(here("figures/data_availability_summary_bars.png"), 
       p1, 
       width = 12, 
       height = 8, 
       dpi = 300,
       bg = "white")

# ============================================================================
# OPTION 2: Dot plot showing which states have each feature
# ============================================================================

# Filter for "Yes" responses and create labels
yes_data <- data_long %>%
  filter(value_clean %in% c("Yes", "Data exists", "Partial", "On Request")) %>%
  mutate(
    STATE = factor(STATE, levels = c("NT", "WA", "SA", "TAS", "VIC", "NSW", "QLD"))
  )

p2 <- ggplot(yes_data, aes(x = category_label, y = STATE)) +
  geom_point(aes(color = value_clean, shape = realm), size = 4, stroke = 1.5) +
  scale_color_viridis_d(option = "D", begin = 0.4, end = 0.95) +
  scale_shape_manual(values = c("Marine" = 16, "Land" = 17)) +
  labs(title = "Which States Provide Each Type of Data?",
       subtitle = "Showing only positive responses (Yes, Partial, On Request, or data exists)",
       x = NULL,
       y = "State/Territory",
       color = "Response Type",
       shape = "Realm") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y = element_text(size = 10, face = "bold"),
    panel.grid.major = element_line(color = "grey90"),
    legend.position = "right",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40")
  )

ggsave(here("figures/data_availability_dot_plot.png"), 
       p2, 
       width = 12, 
       height = 6, 
       dpi = 300,
       bg = "white")

# ============================================================================
# OPTION 3: Completeness score - Percentage of "Yes" responses by state
# ============================================================================

# Calculate completeness scores
completeness <- data_long %>%
  mutate(
    score = case_when(
      value_clean == "Yes" | value_clean == "Data exists" ~ 1,
      value_clean == "Partial" | value_clean == "On Request" ~ 0.5,
      value_clean == "No" ~ 0,
      value_clean == "NA" ~ NA_real_
    )
  ) %>%
  group_by(STATE, realm) %>%
  summarise(
    completeness_pct = mean(score, na.rm = TRUE) * 100,
    .groups = "drop"
  ) %>%
  mutate(
    STATE = factor(STATE, levels = c("NT", "WA", "SA", "TAS", "VIC", "NSW", "QLD"))
  )

p3 <- ggplot(completeness, aes(x = STATE, y = completeness_pct, fill = completeness_pct)) +
  geom_col(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.0f%%", completeness_pct)), 
            vjust = -0.5,
            size = 3.5, 
            fontface = "bold") +
  scale_fill_viridis_c(option = "D", begin = 0.25, end = 0.95) +
  facet_wrap(~realm, ncol = 1) +
  labs(title = "Data Completeness Score by State/Territory",
       subtitle = "Percentage based on: Yes = 100%, Partial/On Request = 50%, No = 0%",
       x = "State/Territory",
       y = "Completeness Score (%)",
       fill = "Score (%)") +
  ylim(0, 105) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 10),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "right",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40")
  )

ggsave(here("figures/data_availability_completeness.png"), 
       p3, 
       width = 12, 
       height = 8, 
       dpi = 300,
       bg = "white")

# ============================================================================
# OPTION 4: Grouped bar chart - Direct comparison of Marine vs Land
# ============================================================================

# Count Yes/Partial/On Request responses
positive_counts <- data_long %>%
  filter(value_clean %in% c("Yes", "Data exists", "Partial", "On Request")) %>%
  count(category_label, realm) %>%
  complete(category_label, realm, fill = list(n = 0))

p4 <- ggplot(positive_counts, aes(x = category_label, y = n, fill = realm)) +
  geom_col(position = "dodge", color = "white", linewidth = 0.5) +
  geom_text(aes(label = n), 
            position = position_dodge(width = 0.9),
            vjust = -0.5,
            size = 3.5, 
            fontface = "bold") +
  scale_fill_viridis_d(option = "D", begin = 0.4, end = 0.8) +
  labs(title = "Marine vs Land Data Availability Comparison",
       subtitle = "Count of states with positive responses (Yes, Partial, On Request, or data exists)",
       x = NULL,
       y = "Number of States/Territories",
       fill = "Realm") +
  ylim(0, 8) +
  coord_flip() +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40")
  )

ggsave(here("figures/data_availability_marine_vs_land.png"), 
       p4, 
       width = 12, 
       height = 6, 
       dpi = 300,
       bg = "white")

# ============================================================================
# OPTION 5: Radial/spider plot - Data quality profile by state
# ============================================================================

# Calculate category scores for each state
state_scores <- data_long %>%
  mutate(
    score = case_when(
      value_clean == "Yes" | value_clean == "Data exists" ~ 1,
      value_clean == "Partial" | value_clean == "On Request" ~ 0.5,
      value_clean == "No" ~ 0,
      value_clean == "NA" ~ 0
    )
  ) %>%
  group_by(STATE, realm, category_label) %>%
  summarise(score = mean(score, na.rm = TRUE), .groups = "drop")

p5 <- ggplot(state_scores, aes(x = category_label, y = score, group = STATE, color = STATE)) +
  geom_polygon(aes(fill = STATE), alpha = 0.1, linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  scale_color_viridis_d(option = "D", begin = 0.1, end = 0.9) +
  scale_fill_viridis_d(option = "D", begin = 0.1, end = 0.9) +
  facet_wrap(~realm, ncol = 2) +
  coord_polar() +
  labs(title = "Data Quality Profile by State/Territory",
       subtitle = "Radar chart showing completeness across all categories",
       x = NULL,
       y = "Completeness Score",
       color = "State",
       fill = "State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 8),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40")
  )

ggsave(here("figures/data_availability_radar.png"), 
       p5, 
       width = 12, 
       height = 8, 
       dpi = 300,
       bg = "white")

# ============================================================================
# Print summary statistics
# ============================================================================

print("Alternative visualizations created successfully!")
print("")
print("Files created:")
print("1. data_availability_summary_bars.png - Stacked bar chart showing counts by response type")
print("2. data_availability_dot_plot.png - Dot plot showing which states have each feature")
print("3. data_availability_completeness.png - Completeness score by state")
print("4. data_availability_marine_vs_land.png - Direct comparison of Marine vs Land")
print("5. data_availability_radar.png - Radar chart showing quality profiles")
print("")

# Print some useful summary stats
cat("\n=== SUMMARY STATISTICS ===\n\n")

cat("States with publicly available MARINE data:\n")
marine_public <- data_availability %>% filter(MARINE_public == "Y") %>% pull(STATE)
cat(paste(marine_public, collapse = ", "), "\n\n")

cat("States with publicly available LAND data:\n")
land_public <- data_availability %>% filter(LAND_public == "Y") %>% pull(STATE)
cat(paste(land_public, collapse = ", "), "\n\n")

cat("States that specify if sites are ACTIVE/INACTIVE for MARINE:\n")
marine_active <- data_availability %>% filter(MARINE_active == "Y") %>% pull(STATE)
cat(paste(marine_active, collapse = ", "), "\n\n")

cat("States that specify if sites are ACTIVE/INACTIVE for LAND:\n")
land_active <- data_availability %>% filter(LAND_active == "Y") %>% pull(STATE)
cat(paste(land_active, collapse = ", "), "\n\n")

cat("States with complete species information for MARINE:\n")
marine_species <- data_availability %>% filter(MARINE_species == "Y") %>% pull(STATE)
cat(paste(marine_species, collapse = ", "), "\n\n")

cat("Most complete state for MARINE aquaculture data: SA\n")
cat("Most complete state for LAND aquaculture data: SA and NT (tie)\n")
```

#extra plots 
#heatmap
```{r}
# ============================================================================
# OPTION 5A: Heatmap showing completeness scores
# ============================================================================
p5a <- ggplot(state_scores, aes(x = STATE, y = category_label, fill = score)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = scales::percent(score, accuracy = 1)), 
            color = "grey20", fontface = "bold", size = 3.5) +
  scale_fill_viridis_c(option = "D", 
                       begin = 0.1, 
                       end = 0.95,
                       labels = scales::percent,
                       limits = c(0, 1)) +
  facet_wrap(~realm, ncol = 2, scales = "free_y") +
  labs(title = "Data Quality Profile by State/Territory",
       subtitle = "Completeness score for each category",
       x = "State/Territory",
       y = NULL,
       fill = "Completeness\nScore") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "right",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40"),
    panel.grid = element_blank()
  )

ggsave(here("figures/data_availability_heatmap.png"), 
       p5a, 
       width = 12, 
       height = 7, 
       dpi = 300,
       bg = "white")
```

#grouped chart by category 
```{r}
# ============================================================================
# OPTION 5B: Grouped bar chart - easier to compare across states
# ============================================================================
p5b <- ggplot(state_scores, aes(x = STATE, y = score, fill = STATE)) +
  geom_col(color = "white", linewidth = 0.3) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25), 
                     labels = scales::percent) +
  scale_fill_viridis_d(option = "D", begin = 0.1, end = 0.9) +
  facet_grid(realm ~ category_label) +
  labs(title = "Data Quality Profile by State/Territory",
       subtitle = "Completeness score across data categories",
       x = "State/Territory",
       y = "Completeness Score",
       fill = "State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8, face = "bold"),
    axis.text.y = element_text(size = 8),
    strip.text = element_text(size = 9, face = "bold"),
    legend.position = "none",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40"),
    panel.grid.major.x = element_blank()
  )

ggsave(here("figures/data_availability_grouped_bars.png"), 
       p5b, 
       width = 14, 
       height = 8, 
       dpi = 300,
       bg = "white")
```

#faceted mini radars per state
```{r}
# ============================================================================
# OPTION 5C: Small multiples radar - one per state (cleaner than overlapping)
# ============================================================================
p5c <- ggplot(state_scores, aes(x = category_label, y = score, 
                                 group = realm, color = realm, fill = realm)) +
  geom_polygon(alpha = 0.2, linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.5)) +
  scale_color_viridis_d(option = "D", begin = 0.2, end = 0.8, 
                        name = "Realm") +
  scale_fill_viridis_d(option = "D", begin = 0.2, end = 0.8,
                       name = "Realm") +
  facet_wrap(~STATE, ncol = 4) +
  coord_polar() +
  labs(title = "Data Quality Profile by State/Territory",
       subtitle = "Radar charts comparing Marine vs Land data completeness",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 6),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40")
  )

ggsave(here("figures/data_availability_radar_faceted.png"), 
       p5c, 
       width = 14, 
       height = 10, 
       dpi = 300,
       bg = "white")
```

#lollipop chart 
```{r}
# ============================================================================
# OPTION 5D: Lollipop chart showing scores
# ============================================================================
p5d <- ggplot(state_scores, aes(x = score, y = STATE, color = STATE)) +
  geom_segment(aes(x = 0, xend = score, y = STATE, yend = STATE), 
               color = "grey70", linewidth = 0.5) +
  geom_point(size = 4) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25),
                     labels = scales::percent) +
  scale_color_viridis_d(option = "D", begin = 0.1, end = 0.9) +
  facet_grid(realm ~ category_label, scales = "free_y", space = "free_y") +
  labs(title = "Data Quality Profile by State/Territory",
       subtitle = "Completeness across categories",
       x = "Completeness Score",
       y = NULL) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9, face = "bold"),
    axis.text.x = element_text(size = 8),
    strip.text = element_text(size = 9, face = "bold"),
    legend.position = "none",
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(here("figures/data_availability_lollipop.png"), 
       p5d, 
       width = 14, 
       height = 8, 
       dpi = 300,
       bg = "white")
```

#version v5 combined matrices
```{r}
# v5 — combined marine + land matrix with shared middle labels and shared legend
library(here)
library(tidyverse)
library(viridis)
library(patchwork)

# Create the data (same as v4)
data_availability <- tribble(
  ~STATE, ~MARINE_public, ~LAND_public, ~MARINE_realm, ~LAND_realm, ~MARINE_spatial, ~LAND_spatial, ~MARINE_active, ~LAND_active, ~MARINE_species, ~LAND_species,
  "QLD", "OR", "OR", "NA", "NA", "point (vector)", "cadastre number", "N", "NA", "Y", "P",
  "NSW", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "Y", "NA", "N", "NA",
  "VIC", "Y", "N", "Y", "NA", "point (vector)", "NA", "N", "NA", "N", "NA",
  "TAS", "Y", "N", "Y", "NA", "polygon (vector)", "NA", "N", "NA", "N", "NA",
  "SA", "Y", "Y", "Y", "Y", "polygon (vector)", "polygon (vector)", "Y", "Y", "Y", "N",
  "WA", "Y", "Y", "N", "N", "polygon (vector)", "polygon (vector)", "N", "NA", "N", "N",
  "NT", "OR", "Y", "N", "N", "point (vector)", "point (vector)", "Y", "Y", "Y", "Y"
)

# Helper: same prep logic used for both realms — avoids duplicating your v4 block twice
prep_realm_data <- function(data, realm_prefix) {
  data %>%
    select(STATE, starts_with(paste0(realm_prefix, "_"))) %>%
    pivot_longer(cols = -STATE,
                 names_to = "category",
                 names_prefix = paste0(realm_prefix, "_"),
                 values_to = "value") %>%
    mutate(
      display_value = case_when(
  value == "Y"                ~ "Yes",
  value == "N"                ~ "No",
  value == "OR"               ~ "On\nRequest",    # <- two lines
  value == "P"                ~ "Partial",
  value == "NA"               ~ "No",
  value == "polygon (vector)" ~ "polygon\n(vector)",
  value == "point (vector)"   ~ "point\n(vector)",
  value == "cadastre number"  ~ "cadastre\nnumber",
  TRUE                        ~ value
),
value_category = case_when(
  value == "Y"  ~ "Yes",
  value == "OR" ~ "On Request",        # <- renamed
  value == "P"  ~ "On Request",        # <- folded in
  value == "N"  ~ "No",
  value == "NA" ~ "No",
  TRUE          ~ "Yes"
),
value_category = factor(value_category,
                        levels = c("No", "On Request", "Yes")), 
      category_label = case_when(
        category == "public"  ~ "Data publicly\navailable?",
        category == "realm"   ~ "Realm\nspecified?",
        category == "spatial" ~ "Spatial data\ntype",
        category == "active"  ~ "Active vs\ninactive specified",
        category == "species" ~ "Species\ninformation provided"
      ),
      # Ordered so "Data publicly available?" sits at the BOTTOM of the matrix,
      # matching your PPT mockup
      category_label = factor(category_label,
                              levels = rev(c("Data publicly\navailable?",
                                             "Realm\nspecified?",
                                             "Spatial data\ntype",
                                             "Active vs\ninactive specified",
                                             "Species\ninformation provided"))),
      STATE = factor(STATE, levels = c("QLD", "NSW", "VIC", "TAS", "SA", "WA", "NT"))
    )
}

marine_data <- prep_realm_data(data_availability, "MARINE")
land_data   <- prep_realm_data(data_availability, "LAND")

# Shared theme
matrix_theme <- theme_minimal() +
  theme(
    axis.text.x      = element_text(size = 10, face = "bold"),
    axis.text.y      = element_text(size = 9,  face = "bold"),
    panel.grid       = element_blank(),
    plot.title       = element_text(size = 13, face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "grey95", color = NA),
    plot.margin      = margin(10, 5, 10, 10)
  )

# 1. Drop y-axis labels from BOTH plots, bump cell text size
p_land <- ggplot(land_data, aes(x = STATE, y = category_label)) +
  geom_tile(aes(fill = value_category), color = "white", linewidth = 1) +
  geom_text(aes(label = display_value), size = 4.5,   # was 3.5
            fontface = "bold", color = "grey20", lineheight = 0.9) +
  scale_fill_viridis_d(option = "G", begin = 0.40, end = 0.95,
                       name = "Response",
                       limits = c("No", "On Request", "Yes")) +
  labs(x = "State/Territory", y = NULL, title = "Land-based Realm") +
  matrix_theme +
  theme(axis.text.y = element_blank(),
        legend.position = "bottom", legend.direction = "horizontal")

p_marine <- ggplot(marine_data, aes(x = STATE, y = category_label)) +
  geom_tile(aes(fill = value_category), color = "white", linewidth = 1) +
  geom_text(aes(label = display_value), size = 4.5,
            fontface = "bold", color = "grey20", lineheight = 0.9) +
  scale_fill_viridis_d(option = "G", begin = 0.40, end = 0.95,
                       name = "Response",
                       limits = c("No", "On Request", "Yes")) +
  labs(x = "State/Territory", y = NULL, title = "Marine Realm") +
  matrix_theme +
  theme(axis.text.y = element_blank(),
        legend.position = "bottom", legend.direction = "horizontal")

# 2. A dedicated middle plot that just holds the category labels.
#    Using the same factor + levels as the matrices, so patchwork aligns rows.
label_levels <- rev(c("Data publicly\navailable?",
                      "Realm\nspecified?",
                      "Spatial data\ntype",
                      "Active vs\ninactive specified",
                      "Species\ninformation provided"))

label_df <- data.frame(
  y = factor(label_levels, levels = label_levels),
  x = 1
)

p_labels <- ggplot(label_df, aes(x = x, y = y)) +
  geom_text(aes(label = y), size = 4.2, fontface = "bold",
            color = "grey20", lineheight = 0.95) +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  theme_void() +
  theme(plot.margin = margin(10, 0, 10, 0))

# 3. Combine: land | labels | marine, with shared legend at the bottom
p_combined <- p_land + p_labels + p_marine +
  plot_layout(ncol = 3, widths = c(7, 2, 7), guides = "collect") +
  plot_annotation(
    theme = theme(
      legend.position  = "bottom",
      legend.title     = element_text(size = 14, face = "bold"),
      legend.text      = element_text(size = 13),
      legend.key.size  = unit(0.9, "cm"),      # bigger colour swatches
      legend.spacing.x = unit(0.4, "cm")       # more breathing room between entries
    )
  )

ggsave(here("figures/data_availability_combined_v5.png"),
       p_combined,
       width = 17, height = 6, dpi = 300, bg = "white")
```

