---
title: "07_atribute_table_setting_land_sites"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: "2025-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#ponds shapefiles are at "output_data/ponds_in_sites/ponds_in_sites_in_
#land sites shapefiles are at: "output_data/Australian_Aqua_sites" and are "state"_land_sites.shp
```{r}
library(terra)
library(here)
library(dplyr)
```


```{r}
state_shapes <- list(
  "QLD" = here("output_data", "Australian_Aqua_sites", "QLD_land_sites.shp"),
  "NT"  = here("output_data", "Australian_Aqua_sites", "NT_land_sites.shp"),
  "NSW" = here("output_data", "Australian_Aqua_sites", "NSW_land_sites.shp"),
  "TAS" = here("output_data", "Australian_Aqua_sites", "TAS_land_sites.shp"),
  "SA"  = here("output_data", "Australian_Aqua_sites", "SA_land_sites.shp"),
  "WA"  = here("output_data", "Australian_Aqua_sites", "WA_land_sites.shp"),
  "VIC" = here("output_data", "Australian_Aqua_sites", "VIC_land_sites.shp")
)
 
for (state in names(state_shapes)) {
  path <- state_shapes[[state]]
  if (file.exists(path)) {
    v <- vect(path)
    print(paste("----------------", state, "----------------"))
    print("COLUMN NAMES:")
    print(names(v))
    print("FIRST ROW DATA:")
    print(head(v, 1))
    print(" ")
  } else {
    print(paste("Warning: File not found for", state))
  }
}
```

 
#loop to make the table
#the shp created has pond polygons for actove but lease polygons for inactive, should we maintain that or not? Australia_Land_Aquaculture_National 

```{r}

library(terra)
library(sf)
library(here)
library(dplyr)

# --- 1. CONFIGURATION ---

# Ponds shapefiles are in output_data/ponds_in_sites/
input_ponds_dir <- here("output_data", "ponds_in_sites")

# Output directory for final shapefiles
output_dir <- here("output_data", "Land_Aquaculture_Vectors")

if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
 
# Land-based aquaculture site shapefiles (ALL Australian states)
state_shapes <- list(
  "QLD" = here("output_data", "Australian_Aqua_sites", "QLD_land_sites.shp"),
  "NT"  = here("output_data", "Australian_Aqua_sites", "NT_land_sites.shp"),
  "NSW" = here("output_data", "Australian_Aqua_sites", "NSW_land_sites.shp"),
  "TAS" = here("output_data", "Australian_Aqua_sites", "TAS_land_sites.shp"),
  "SA"  = here("output_data", "Australian_Aqua_sites", "SA_land_sites.shp"),
  "WA"  = here("output_data", "Australian_Aqua_sites", "WA_land_sites.shp"),
  "VIC" = here("output_data", "Australian_Aqua_sites", "VIC_land_sites.shp")
)
 
# Maps State Shapefile ID -> Our 'SiteID_Custodian'
id_col_map <- list(
  "QLD" = "site_id",
  "NT"  = "site_id",
  "NSW" = "site_id",
  "TAS" = "site_id",
  "SA"  = "reg_no",
  "WA"  = "aqua_id",
  "VIC" = "site_id"
)

# Define target CRS - Australian Albers (EPSG:3577)
target_crs <- "EPSG:3577"

# Custodian organization mapping
custodian_map <- list(
  "QLD" = "Queensland Department of Agriculture and Fisheries",
  "NT"  = "Northern Territory Department of Industry, Tourism and Trade",
  "NSW" = "NSW Department of Primary Industries",
  "TAS" = "Tasmania Department of Natural Resources and Environment",
  "SA"  = "South Australia Primary Industries and Regions",
  "WA"  = "Western Australia Department of Primary Industries and Regional Development",
  "VIC" = "Victoria Department of Energy, Environment and Climate Action"
)
# --- 2. MAIN PROCESSING LOOP ---
 
for (state in names(state_shapes)) {

  # --- A. Setup & Checks ---
  if (!file.exists(state_shapes[[state]])) {
    message(paste("Skipping", state, "- Shapefile not found."))
    next
  }

  message(paste("--- Processing State:", state, "---"))

  # Load Sites using sf
  sites_sf <- st_read(state_shapes[[state]], quiet = TRUE)
  
  # Check and reproject sites to Australian Albers if needed
  message(paste("Sites CRS:", st_crs(sites_sf)$input))
  if (st_crs(sites_sf) != st_crs(target_crs)) {
    message(paste("Reprojecting sites to Australian Albers (EPSG:3577)"))
    sites_sf <- st_transform(sites_sf, target_crs)
  }

  # PRE-CALC PARENT AREA (Used for both Active and Inactive)
  # Converting square meters to square kilometers
  sites_sf$PARENT_AREA_KM <- as.numeric(st_area(sites_sf)) / 1e6

  # Identify the ID column for this state
  target_id_col <- id_col_map[[state]]
  
  # Create a unique numeric ID for joining
  sites_sf$TEMP_JOIN_ID <- 1:nrow(sites_sf)

  # --- B. Process Ponds Shapefiles (Find Active Sites) ---
  
  # Look for pond shapefile for this state
  pond_file <- here(input_ponds_dir, paste0("ponds_in_sites_in_", state, ".shp"))

  # Initialize empty objects
  active_sf <- NULL
  active_df <- NULL
  found_ids <- c() # Keep track of IDs we found ponds in

  if (file.exists(pond_file)) {
    message(paste("Found pond shapefile:", basename(pond_file)))
    
    # Read pond vector using sf
    ponds_sf <- st_read(pond_file, quiet = TRUE)
    
    # Check and reproject ponds to Australian Albers if needed
    message(paste("Ponds CRS:", st_crs(ponds_sf)$input))
    if (st_crs(ponds_sf) != st_crs(target_crs)) {
      message(paste("Reprojecting ponds to Australian Albers (EPSG:3577)"))
      ponds_sf <- st_transform(ponds_sf, target_crs)
    }

    message(paste("Intersecting", nrow(ponds_sf), "ponds with sites..."))
    
    # SPATIAL INTERSECTION - This will keep only ponds within sites
    # and attach site attributes to each pond
    ponds_in_sites <- st_intersection(ponds_sf, sites_sf)
    
    message(paste("Found", nrow(ponds_in_sites), "pond-site intersections"))
    
  if (nrow(ponds_in_sites) > 0) {

      # Calculate pond areas AFTER intersection
      ponds_in_sites$pond_area_m2 <- as.numeric(st_area(ponds_in_sites))

      # Aggregate pond areas by site
      site_summaries <- ponds_in_sites %>%
        st_drop_geometry() %>%
        group_by(TEMP_JOIN_ID) %>%
        summarise(
          total_pond_area_km = sum(pond_area_m2, na.rm = TRUE) / 1e6,
          n_ponds = n(),
          .groups = "drop"
        ) %>%
        as.data.frame()

      message(paste("Found", nrow(site_summaries), "sites with ponds"))

      # ---- Build active output with LEASE geometry (not pond union) ----
      active_sites <- sites_sf %>%
        inner_join(site_summaries, by = "TEMP_JOIN_ID")

      found_ids <- active_sites[[target_id_col]]

      active_df <- data.frame(
        SiteID            = paste0(state, "_A_", sprintf("%04d", seq_len(nrow(active_sites)))),
        SiteID_Custodian  = found_ids,
        Production_realm  = "Land-based",
        Taxon             = if("group"   %in% names(active_sites)) active_sites$group   else NA_character_,
        Species_custodian = if("species" %in% names(active_sites)) active_sites$species else NA_character_,
        Species_verified  = if("species" %in% names(active_sites)) active_sites$species else NA_character_,
        Status_custodian  = if("status"  %in% names(active_sites)) active_sites$status  else NA_character_,
        Production_t      = NA_real_,
        Infra_status      = "active",
        Infra_type        = "ponds",
        Lease_Area_KM     = round(active_sites$PARENT_AREA_KM, 6),
        Infra_Area_KM     = round(active_sites$total_pond_area_km, 6),
        LGA               = if("LGA" %in% names(active_sites)) active_sites$LGA else NA_character_,
        State             = state,
        Year              = "2021-2022",
        Custodian_ID      = custodian_map[[state]],
        Method            = "NDWI Sentinel2 Stack",
        stringsAsFactors  = FALSE
      )

     active_sf <- st_sf(active_df, geometry = st_geometry(active_sites))
     
     # ---- Save pond-level detections (separate from leases) ----
      # Link each pond to the SiteID of its parent lease
      siteid_lookup <- data.frame(
        TEMP_JOIN_ID = active_sites$TEMP_JOIN_ID,
        SiteID       = active_df$SiteID
      )
      ponds_linked <- ponds_in_sites %>%
        left_join(siteid_lookup, by = "TEMP_JOIN_ID")

      n_ponds <- nrow(ponds_linked)
      ponds_df <- data.frame(
        DetectID         = paste0(state, "_POND_", sprintf("%06d", seq_len(n_ponds))),
        SiteID           = ponds_linked$SiteID,
        SiteID_Custodian = as.character(ponds_linked[[target_id_col]]),
        Production_realm = "Land-based",
        State            = state,
        Infra_type       = "Pond",
        Infra_Area_KM    = round(ponds_linked$pond_area_m2 / 1e6, 6),
        Year             = "2021-2022",
        Method           = "NDWI Sentinel2 Stack",
        stringsAsFactors = FALSE
      )
      active_ponds_sf <- st_sf(ponds_df, geometry = st_geometry(ponds_linked))

      out_ponds_gpkg <- here(output_dir, paste0(state, "_Land_Ponds.gpkg"))
      st_write(active_ponds_sf, out_ponds_gpkg, delete_dsn = TRUE, quiet = TRUE)
    }   # closes: if (nrow(ponds_in_sites) > 0)

  } else {
    message(paste("No pond shapefile found for", state))
  }     # closes: if (file.exists(pond_file))

  # --- C. Process Empty Sites (Find Inactive Sites) ---
 
  # Identify which IDs were NOT found in the intersection
  all_ids <- sites_sf[[target_id_col]]
  missing_ids <- setdiff(all_ids, found_ids)

  inactive_sf <- NULL
  inactive_df <- NULL

  if (length(missing_ids) > 0) {
    
    message(paste("Found", length(missing_ids), "sites without ponds (inactive)"))
    
    # Filter the original sites to only the missing ones
    inactive_sf <- sites_sf[sites_sf[[target_id_col]] %in% missing_ids, ]

    # Get status if available
    status_val <- rep(NA, nrow(inactive_sf))
    if ("status" %in% tolower(names(inactive_sf))) {
      real_col <- names(inactive_sf)[grep("^status$", names(inactive_sf), ignore.case=TRUE)[1]]
      status_val <- inactive_sf[[real_col]]
    }

    # Build Inactive DataFrame
   inactive_df <- data.frame(
  SiteID            = paste0(state, "_I_", sprintf("%04d", 1:nrow(inactive_sf))),
  SiteID_Custodian  = inactive_sf[[target_id_col]],  # Original ID from state government
  Production_realm  = "Land-based",
  Taxon             = if("group" %in% names(inactive_sf)) inactive_sf$group else NA_character_,
  Species_custodian = if("species" %in% names(inactive_sf)) inactive_sf$species else NA_character_,
  Species_verified  = if("species" %in% names(inactive_sf)) inactive_sf$species else NA_character_,
  Status_custodian  = if("status" %in% names(inactive_sf)) inactive_sf$status else NA_character_,
  Production_t      = NA_real_,
  Infra_status      = "inactive",
  Infra_type        = NA_character_,
  Lease_Area_KM     = round(inactive_sf$PARENT_AREA_KM, 6),
  Infra_Area_KM     = 0,
  LGA               = if("LGA" %in% names(inactive_sf)) inactive_sf$LGA else NA_character_,
  State             = state,
  Year              = "2021-2022",
  Custodian_ID      = custodian_map[[state]],  # Organization name
  Method            = "No water detected with NDWI in Sentinel2 Stack",
  stringsAsFactors  = FALSE
)

    # Create new sf object with just the geometry and new attributes
    inactive_sf <- st_sf(inactive_df, geometry = st_geometry(inactive_sf))
  }

  # --- D. Combine & Save ---
  
  final_sf <- NULL
  final_df <- NULL

  if (!is.null(active_sf) && !is.null(inactive_sf)) {
    final_sf <- rbind(active_sf, inactive_sf)
    final_df <- rbind(st_drop_geometry(active_sf), st_drop_geometry(inactive_sf))
  } else if (!is.null(active_sf)) {
    final_sf <- active_sf
    final_df <- st_drop_geometry(active_sf)
  } else if (!is.null(inactive_sf)) {
    final_sf <- inactive_sf
    final_df <- st_drop_geometry(inactive_sf)
  }

  if (!is.null(final_sf)) {
    out_shp <- here(output_dir, paste0(state, "_Land_Mapped.shp"))
    out_csv <- here(output_dir, paste0(state, "_Land_Attributes.csv"))
    
    # --- Combine pond detections from all states ---
message("\nCombining pond detections...")
all_ponds <- list()
for (state in states) {
  pf <- here(input_dir, paste0(state, "_Land_Ponds.gpkg"))
  if (file.exists(pf)) {
    all_ponds[[state]] <- st_read(pf, quiet = TRUE)
    message(paste("  ✓ Loaded", nrow(all_ponds[[state]]), "ponds from", state))
  }
}

if (length(all_ponds) > 0) {
  national_ponds <- do.call(rbind, all_ponds)
  if (st_crs(national_ponds) != st_crs("EPSG:3577")) {
    national_ponds <- st_transform(national_ponds, "EPSG:3577")
  }
  out_ponds_national <- here(input_dir, "Australia_Land_Aquaculture_Ponds_National.gpkg")
  st_write(national_ponds, out_ponds_national, delete_dsn = TRUE, quiet = TRUE)
  message(paste("✓ Saved", nrow(national_ponds), "pond detections to",
                basename(out_ponds_national)))
  message(paste("  Total pond area:",
                round(sum(national_ponds$Infra_Area_KM, na.rm = TRUE), 2), "km²"))
}

    st_write(final_sf, out_shp, delete_dsn = TRUE, quiet = TRUE)
    write.csv(final_df, out_csv, row.names = FALSE)

    message(paste("Saved:", basename(out_shp), 
                  "| Active:", ifelse(is.null(active_sf), 0, nrow(active_sf)),
                  "| Inactive:", ifelse(is.null(inactive_sf), 0, nrow(inactive_sf))))
  } else {
    message("No data generated for this state.")
  }
  
  message("---")
}
 
message("All States Processed.")
```

#Making the Australian land aquaculture Natinal datasets
```{r}
library(sf)
library(here)
library(dplyr)

# --- COMBINE ALL STATE LAND-BASED AQUACULTURE DATA ---

message("=== Creating National Land-Based Aquaculture Dataset ===\n")

# Directory with state-level outputs
input_dir <- here("output_data", "Land_Aquaculture_Vectors")

# Output files
output_shp <- here(input_dir, "Australia_Land_Aquaculture_National.shp")
output_csv <- here(input_dir, "Australia_Land_Aquaculture_National.csv")

# List of states
states <- c("QLD", "NT", "NSW", "TAS", "SA", "WA", "VIC")

# Initialize list to store state data
all_states_sf <- list()
all_states_df <- list()

# Read each state file
for (state in states) {
  state_file <- here(input_dir, paste0(state, "_Land_Mapped.shp"))
  
  if (file.exists(state_file)) {
    message(paste("Loading", state, "..."))
    
    # Read shapefile
    state_sf <- st_read(state_file, quiet = TRUE)
    
    # Also read CSV to preserve full column names (shapefiles truncate names)
    state_csv <- here(input_dir, paste0(state, "_Land_Attributes.csv"))
    if (file.exists(state_csv)) {
      state_df <- read.csv(state_csv, stringsAsFactors = FALSE)
      all_states_df[[state]] <- state_df
    }
    
    all_states_sf[[state]] <- state_sf
    
    message(paste("  ✓ Loaded", nrow(state_sf), "sites from", state))
  } else {
    message(paste("  ✗ Warning: File not found for", state))
  }
}

# Combine all states
if (length(all_states_sf) > 0) {
  message("\nCombining all states...")
  
  # Combine shapefiles
  national_sf <- do.call(rbind, all_states_sf)
  
  # Combine CSVs (preserves full column names)
  national_df <- do.call(rbind, all_states_df)
  
  # Ensure CRS is Australian Albers
  if (st_crs(national_sf) != st_crs("EPSG:3577")) {
    message("Ensuring Australian Albers CRS...")
    national_sf <- st_transform(national_sf, "EPSG:3577")
  }
  
  # Save combined dataset
  message("\nSaving national dataset...")
  st_write(national_sf, output_shp, delete_dsn = TRUE, quiet = TRUE)
  write.csv(national_df, output_csv, row.names = FALSE)
  
  message(paste("\n=== SUMMARY ==="))
  message(paste("Total sites:", nrow(national_sf)))
  message(paste("  Active sites (with ponds):", sum(national_df$Infra_status == "active")))
  message(paste("  Inactive sites (no ponds):", sum(national_df$Infra_status == "inactive")))
  message(paste("\nBreakdown by state:"))
  
  state_summary <- national_df %>%
    group_by(State, Infra_status) %>%
    summarise(n_sites = n(), .groups = "drop") %>%
    tidyr::pivot_wider(names_from = Infra_status, values_from = n_sites, values_fill = 0)
  
  print(state_summary)
  
  message(paste("\nTotal pond area:", round(sum(national_df$Infra_Area_KM, na.rm = TRUE), 2), "km²"))
  message(paste("Total site area:", round(sum(national_df$Lease_Area_KM, na.rm = TRUE), 2), "km²"))
  
  message(paste("\n✓ Saved national shapefile:", basename(output_shp)))
  message(paste("✓ Saved national CSV:", basename(output_csv)))
  
} else {
  message("Error: No state files found!")
}

message("\n=== National Dataset Creation Complete ===")
```

```{r}
library(sf)
library(here)

# Check the new national file
national_sf <- st_read(here("output_data", "Land_Aquaculture_Vectors", 
                            "Australia_Land_Aquaculture_National.shp"))

cat("Species data check:\n")
cat("Rows with Taxon:", sum(!is.na(national_sf$Taxon)), "of", nrow(national_sf), "\n")
cat("Rows with Species:", sum(!is.na(national_sf$Spcs_cs)), "of", nrow(national_sf), "\n")

cat("\nUnique species:\n")
print(table(national_sf$Spcs_cs, useNA = "always"))

cat("\nUnique taxon/groups:\n")
print(table(national_sf$Taxon, useNA = "always"))
```


