---
title: "08_mapping active and icative aquaculture/presence of infrastructure"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: "28/01/2026 to 24/04/2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#In this code we aim to make a map with the active inactive land based and marine aquaculture in viridis 
#This is Figure 2 of the manuscript 

#upload libraries
```{r}
library(tmaptools)
library(gridExtra)
library(cowplot)
library(tidyverse)
library(here)
library(tmap)
library(sf)
library(ggplot2)
library(dplyr)

```

#upload all leases for Australia together 
#land
```{r}
AU_land_aqua <- st_read(here("output_data/Land_Aquaculture_Vectors/Australia_Land_Aquaculture_National.shp")) %>%
  st_transform("EPSG:3577")%>%
  st_make_valid()

# Convert to points
AU_land_aqua_points <- st_centroid(AU_land_aqua)

```

```{r}
unique(AU_land_aqua$Infr_st)
```
#Marine sites with species
```{r}
# Load marine infrastructure — runs independently of script 4
 AU_marine_leases <- st_read(
  here("output_data/marine_leases_status_results/all_marine_leases_with_status.shp"))%>%
   st_transform("EPSG:3577")%>%
  st_make_valid()

# Convert marine leases to points (centroids)
AU_marine_leases_points <- st_centroid(AU_marine_leases)


```
## ===== LOAD AUSTRALIAN BOUNDARIES =====
```{r}
message("Loading Australian boundary data...")

AUS_states <- st_read(here("raw_data/AustralianStates_SHP/STE_2021_AUST_GDA2020.shp")) %>%
  st_transform(3577) %>%
  st_make_valid()
```
#acronyms for states for mapping
```{r}
state_labels <- AUS_states %>%
  st_make_valid() %>%
  filter(!STE_NAME21 %in% c("Other Territories", "Outside Australia",
                             "Australian Capital Territory")) %>%
  st_centroid() %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2],
    acronym = case_when(
      STE_NAME21 == "New South Wales"    ~ "NSW",
      STE_NAME21 == "Victoria"           ~ "VIC",
      STE_NAME21 == "Queensland"         ~ "QLD",
      STE_NAME21 == "South Australia"    ~ "SA",
      STE_NAME21 == "Western Australia"  ~ "WA",
      STE_NAME21 == "Tasmania"           ~ "TAS",
      STE_NAME21 == "Northern Territory" ~ "NT",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(acronym)) %>%
  st_drop_geometry()
```


#Create the combined status column for land based and marine leases 
```{r}
# ===== CREATE COMBINED STATUS — 4 CATEGORIES =====
# Land-based: distinguish active vs inactive
# And the case_when mutations
AU_land_aqua_points <- AU_land_aqua_points %>%
  mutate(combined_status = case_when(
    Infr_st == "active" ~ "Land-based: active",
    TRUE                ~ "Land-based: inactive"
  ))

AU_marine_leases_points <- AU_marine_leases_points %>%
  mutate(combined_status = case_when(
    Infr_st == "present" ~ "Marine: active",
    TRUE                 ~ "Marine: inactive"
  ))

# ===== FACTOR ORDER (controls legend order) =====
status_levels <- c(
  "Marine: active",
  "Marine: inactive",
  "Land-based: active",
  "Land-based: inactive"
)

AU_land_aqua_points <- AU_land_aqua_points %>%
  mutate(combined_status = factor(combined_status, levels = status_levels))

AU_marine_leases_points <- AU_marine_leases_points %>%
  mutate(combined_status = factor(combined_status, levels = status_levels))

# ===== 4-COLOR PALETTE (viridis-family, colorblind-safe) =====
# Blues for marine (active = deep, inactive = muted)
# Greens for land  (active = deep, inactive = muted)
status_colors <- c(
  "Marine: active"      = "#2166AC",
  "Marine: inactive"    = "#92C5DE",
  "Land-based: active"  = "#1B7837",
  "Land-based: inactive" = "#A6DBA0"
)

```

#new maps Marine and Land based 
```{r}
# ===== MAP =====
AU_aqua_map_4color <- ggplot() +
  geom_sf(data = AUS_states, fill = "grey95", color = "grey80") +
  
  # Inactive layers first (rendered underneath)
  geom_sf(
    data  = AU_land_aqua_points %>% filter(combined_status == "Land-based: inactive"),
    aes(color = combined_status),
    size  = 1.0, alpha = 0.9, shape = 16
  ) +
  geom_sf(
    data  = AU_marine_leases_points %>% filter(combined_status == "Marine: inactive"),
    aes(color = combined_status),
    size  = 1.0, alpha = 0.9, shape = 16
  ) +

  # Active layers on top
  geom_sf(
    data  = AU_land_aqua_points %>% filter(combined_status == "Land-based: active"),
    aes(color = combined_status),
    size  = 1.0, alpha = 0.9, shape = 16
  ) +
  geom_sf(
    data  = AU_marine_leases_points %>% filter(combined_status == "Marine: active"),
    aes(color = combined_status),
    size  = 1.0, alpha = 0.9, shape = 16
  ) +
  
  # State label text
  geom_text(
    data   = state_labels,
    aes(x = lon, y = lat, label = acronym),
    size   = 3, colour = "grey50", fontface = "bold"
  ) +
  
  scale_color_manual(
    values = status_colors,
    name   = NULL,
    breaks = status_levels
  ) +
  
  guides(
    color = guide_legend(
      override.aes = list(size = 3, alpha = 1),
      ncol = 1
    )
  ) +
  
  theme_void() +
  theme(
    legend.position  = "bottom",
    legend.direction = "vertical",
    legend.text      = element_text(size = 10),
    legend.key       = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA)
  )

# Save
ggsave(
  here("figures", "AU_land_marine_aqua_4color.png"),
  plot   = AU_aqua_map_4color,
  width  = 10, height = 8, dpi = 300,
  bg     = "transparent"
)

plot(AU_aqua_map_4color)
```

#other possibility for the legends 
#theme(legend.position  = "right",   legend.direction = "vertical",   # stacks items on top of each other   legend.title     = element_text(face = "bold", size = 11),   legend.text      = element_text(size = 10),   panel.background = element_rect(fill = "transparent", colour = NA),   plot.background  = element_rect(fill = "transparent", colour = NA) )

#other maps only of land based
#land_based active and inactive sites Theme_minimal
```{r}
library(sf)
library(ggplot2)


# Convert polygons to centroids (point locations)
AU_land_aqua_points <- st_centroid(AU_land_aqua)

# Map with light grey background and small bubbles
ggplot() +
  geom_sf(data = AUS_states, fill = "grey95", color = "grey80") +
  geom_sf(data = AU_land_aqua_points, aes(color = Infr_st), 
          size = 1.5, alpha = 0.7) +  # adjust size for bubble size
  scale_color_manual(values = c("active" = "#0072B2", 
                                "inactive" = "#D55E00"),
                     name = "Infrastructure Status") +
  theme_minimal() +
  theme(legend.position = "bottom")
```



#land_based active and inactive sites theme_void green and orange
```{r}
library(sf)
library(ggplot2)

# Convert to points
AU_land_aqua_points <- st_centroid(AU_land_aqua)

# Map without gridlines and with green/red alternatives
ggplot() +
  geom_sf(data = AUS_states, fill = "grey95", color = "grey80") +
  geom_sf(data = AU_land_aqua_points, aes(color = Infr_st), 
          size = 1.5, alpha = 0.7) +
  scale_color_manual(values = c("active" = "#009E73",    # bluish green
                                "inactive" = "#D55E00"),  # red-orange
                     name = "Infrastructure Status") +
  theme_void() +  # removes all gridlines and axes
  theme(legend.position = "bottom")
```

```{r}
library(sf)
library(ggplot2)
library(here)


# Create the map
land_aqua_map <- ggplot() +
  geom_sf(data = AUS_states, fill = "grey95", color = "grey80") +
  geom_sf(data = AU_land_aqua_points, aes(color = Infr_st), 
          size = 1.5, alpha = 0.7) +
  scale_color_manual(
    values = c("active" = "#FDE725",    # viridis yellow (end of scale)
               "inactive" = "#440154"),  # viridis dark purple/blue (start of scale)
    name = "Infrastructure Status"
  ) +
  theme_void() +
  theme(legend.position = "bottom")

# Save the map
ggsave(here("figures", "AU_land_aqua_status.png"), 
       plot = land_aqua_map, 
       width = 10, height = 8, dpi = 300)
```
