---
title: "09_production distribution model - refactored"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: "2026-02-19 to April 2026"
---
```{r}
options(encoding = "UTF-8")
Sys.setlocale("LC_ALL", "en_US.UTF-8")  # or "English_Australia.UTF-8" on Windows
```

# Refactored production distribution model
# Single parameter table drives all species calculations.
#
# TO UPDATE FOR A NEW YEAR:
#   1. Change area_km2 in species_params (new GEE detection run)
#   2. Change actual_prod_t in species_params (new ABARES statistics)
#   3. Re-knit — everything else updates automatically.
#
# TO ADD A NEW SPECIES:
#   1. Add a row to species_params
#   2. If system type is new, add a branch in the dispatch loop (Step 3)
#   3. Add variable extraction aliases in Step 4
#   4. Add manuscript text chunk and plot data entry

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r packages}
library(tidyverse)
library(here)
library(tmap)
library(sf)
library(ggplot2)
library(dplyr)
library(scales)
library(cowplot)
library(gridExtra)
library(viridisLite)
library(tmaptools)
```

# ===== INFRASTRUCTURE SCALING FACTOR: MEASUREMENT DATA =====
# Empirical pen measurements from Google Earth Pro vs satellite-detected polygons.
# These are field measurements — update fpr_data when new leases are sampled.
# mean_pen_fraction feeds into run_corrected_volumetric() for salmon.

```{r isf-data}
fpr_data <- data.frame(
  Lease       = c(220,        201,        141),
  Detected_m2 = c(182418.40,  362625.60,  542620.70),
  Pen_area_m2 = c(17926.81,   45641.38,   59771.19)
)
fpr_data$FPR         <- fpr_data$Detected_m2 / fpr_data$Pen_area_m2
fpr_data$Pen_fraction <- fpr_data$Pen_area_m2 / fpr_data$Detected_m2

mean_fpr          <- mean(fpr_data$FPR)
mean_pen_fraction <- mean(fpr_data$Pen_fraction)

cat(sprintf("ISF data loaded: mean pen fraction = %.1f%% (mean scaling factor = %.2f×)\n",
            mean_pen_fraction * 100, mean_fpr))
```

# ===== STEP 1: SPECIES PARAMETER TABLE =====
# Single source of truth for all species parameters.
# ANNUAL UPDATE: change area_km2 and actual_prod_t columns only.

```{r species-params}
species_params <- tibble(

  # --- Identity ---
  species      = c("atl_salmon",
                   "sbt",
                   "prawn",
                   "barramundi",
                   "oyster"),

  label        = c("Atlantic Salmon",
                   "S. Bluefin Tuna",
                   "Black Tiger Prawn",
                   "Barramundi",
                   "Oysters"),

  label_sci    = c("Atlantic Salmon (Salmo salar)",
                   "Southern Bluefin Tuna (Thunnus maccoyii)",
                   "Black Tiger Prawn (Penaeus monodon)",
                   "Barramundi (Lates calcarifer)",
                   "Oysters (mixed spp.)"),

  system       = c("marine_pens",       # empirical + volumetric + corrected volumetric
                   "marine_pontoons",   # empirical + volumetric
                   "pond",              # area-yield
                   "pond",              # area-yield
                   "marine_longline"),  # area-yield (no volumetric)

  state        = c("Tasmania", "South Australia", "National", "National",
                   "NSW, QLD, SA, TAS, WA"),

  # ---- ANNUAL UPDATE: detection area (km²) and ABARES production (tonnes) ----
  # Oyster state breakdown: NSW 5.898, QLD 0.409, SA 6.789, TAS 3.128, WA 0.476
  area_km2     = c(6.18,   0.209,  10.325, 2.466,  16.700),  # land-based area for barramundi
  actual_prod_t = c(79271, 8322,   9450,   4742,   11735),    # ABARES 2021-22

  # --- Marine systems: stocking density (kg/m³) — NA for pond/longline systems ---
  density_low  = c(10,   2,  NA, NA, NA),
  density_mid  = c(12.2, 3,  NA, NA, NA),
  density_high = c(15,   4,  NA, NA, NA),

  # --- Marine systems: net depth (m) — NA for pond/longline systems ---
  # Salmon: pre-derived from pen geometry (pen_volume / pen_area = 11600 / 1145.9 = 10.12 m)
  # SBT: net wall depth of static ranching pontoons (Ellis & Kiessling, 2014)
  net_depth_m  = c(10.12, 10, NA, NA, NA),

  # --- Mortality range (proportion) ---
   # Oyster: 20% per growing season (Nell & Perkins, 2006; Green et al., 2008;
  #         Walton et al., 2013; Petrolia et al., 2024)
  mortality_low  = c(0.14, 0.02, NA,   NA,   NA),
  mortality_mid  = c(0.15, 0.03, 0.20, 0.10, 0.20),
  mortality_high = c(0.16, 0.04, NA,   NA,   NA),

  # --- Salmon pen geometry (for manuscript text; net_depth_m above is pre-derived) ---
  # Source: Tassal Group, 2021
  pen_circ_m  = c(120,   NA, NA, NA, NA),   # circumference (m)
  pen_vol_m3  = c(11600, NA, NA, NA, NA),   # volume (m³)

  # --- Infrastructure scaling factor ---
  # TRUE = apply mean_pen_fraction from fpr_data (salmon only)
  # FALSE = no correction (isolated pontoons or land-based or longline)
  apply_isf   = c(TRUE, FALSE, FALSE, FALSE, FALSE),

  # --- Empirical scaling baseline ---
  # "mid"        → use density_mid as baseline (salmon: reported prod ≈ density_mid)
  # "back_calc"  → back-calculate effective density from reported production
  density_baseline = c("mid", "back_calc", NA, NA, NA),

# Pond/longline systems: yield range (t/km²) — NA for marine pen/pontoon systems
  # Prawn:      25-40 PL/m² × 30-35g × 80% survival (Lovatelli et al., 2008)
  # Barramundi: 15-30 t/ha = 1,500-3,000 t/km² (de Schryver et al., 2008)
  # Oyster:     500-1,000 t/km² = 5-10 t/ha (annualised whole-farm rate)
  #             Derived from single-cycle capacity ÷ grow-out duration.
  #             FAO extensive lease yield: ~25 t/ha/yr (Helm, 2009).
  #             SRO (S. glomerata): 25 ÷ 3yr = ~8.3 t/ha (Nell, 1993;
  #               Sapphire Coast Oysters, 2024).
  #             Pacific oyster (C. gigas): 25 ÷ 1.5-2.5yr = ~10-17 t/ha
  #               (Helm, 2009; grow-out 18-30 months).
  #             Range spans both species; anchored by empirical rate 7.03 t/ha.
  #             Sources: Helm (2009) FAO C. gigas fact sheet; Nell (1993);
  #               Sapphire Coast Oysters (2024)
  yield_low_tkm2  = c(NA, NA,  600, 1500, 500),
  yield_high_tkm2 = c(NA, NA, 1120, 3000, 1000),

  # --- Prawn stocking parameters (for manuscript text derivation) ---
  prawn_pl_low_m2  = c(NA, NA, 25, NA, NA),   # post-larvae/m²
  prawn_pl_high_m2 = c(NA, NA, 40, NA, NA),
  prawn_hw_low_g   = c(NA, NA, 30, NA, NA),   # harvest weight (g)
  prawn_hw_high_g  = c(NA, NA, 35, NA, NA),

  # --- Plot label offset for actual value (tweak if labels overlap) ---
  actual_hjust = c(-0.3, -0.6, -0.3, -0.3, -0.3),

  # --- Detection method ---
  detection_method = c(
    "Alpha Earth Foundation embeddings",
    "Alpha Earth Foundation embeddings",
    "NDWI annual composite (Sentinel-2)",
    "NDWI annual composite (Sentinel-2)",
    "Sentinel 1 SAR + Alpha Earth Foundation embeddings"
  )
)

cat("Species parameter table created:\n")
print(select(species_params, species, system, area_km2, actual_prod_t))
```

# ===== STEP 2: SYSTEM-BASED MODEL FUNCTIONS =====

```{r model-functions}

# --- Empirical (top-down): realised yield from ABARES-reported production --------
# density_baseline: density level that reported production corresponds to
#   salmon  → density_mid (reported production ≈ seasonal average density)
#   SBT     → back-calculated effective density
run_empirical_marine <- function(actual_prod, area_km2,
                                  density_low, density_baseline, density_high) {
  yield_per_km2 <- actual_prod / area_km2
  prod_low      <- actual_prod * (density_low  / density_baseline)
  prod_high     <- actual_prod * (density_high / density_baseline)
  list(
    yield_per_km2 = yield_per_km2,
    prod_low      = prod_low,
    prod_high     = prod_high,
    yield_low     = prod_low  / area_km2,
    yield_high    = prod_high / area_km2
  )
}

# --- Volumetric (bottom-up, Halpern-compatible 3D): theoretical capacity ---------
# Yield (t/km²) = density (kg/m³) × net_depth (m) × survival × 1,000
# Survival is paired inversely with density (conservative at low, optimistic at high)
run_volumetric_marine <- function(area_km2,
                                   density_low, density_mid, density_high,
                                   mortality_low, mortality_mid, mortality_high,
                                   net_depth_m) {
  survival_low  <- 1 - mortality_high   # paired with low density (conservative)
  survival_mid  <- 1 - mortality_mid
  survival_high <- 1 - mortality_low    # paired with high density (optimistic)
  vol_yield_low  <- density_low  * net_depth_m * survival_low  * 1000
  vol_yield_mid  <- density_mid  * net_depth_m * survival_mid  * 1000
  vol_yield_high <- density_high * net_depth_m * survival_high * 1000
  list(
    yield_low  = vol_yield_low,
    yield_mid  = vol_yield_mid,
    yield_high = vol_yield_high,
    prod_low   = vol_yield_low  * area_km2,
    prod_mid   = vol_yield_mid  * area_km2,
    prod_high  = vol_yield_high * area_km2,
    survival_low  = survival_low,
    survival_mid  = survival_mid,
    survival_high = survival_high
  )
}

# --- Corrected volumetric: apply infrastructure scaling factor (pen fraction) ----
# Applies only where satellite-detected area includes substantial non-pen components
# pen_fraction: proportion of detected area that is productive pen water surface
run_corrected_volumetric <- function(vol, area_km2, pen_fraction) {
  pen_area_km2 <- area_km2 * pen_fraction
  list(
    prod_low     = vol$yield_low  * pen_area_km2,
    prod_mid     = vol$yield_mid  * pen_area_km2,
    prod_high    = vol$yield_high * pen_area_km2,
    pen_area_km2 = pen_area_km2
  )
}

# --- Area-yield (pond systems): P = A × Y ----------------------------------------
run_area_yield <- function(area_km2, yield_low_tkm2, yield_high_tkm2, actual_prod) {
  list(
    prod_low   = yield_low_tkm2  * area_km2,
    prod_high  = yield_high_tkm2 * area_km2,
    yield_low  = yield_low_tkm2,
    yield_high = yield_high_tkm2,
    yield_rate = actual_prod / area_km2   # empirical yield rate for validation
  )
}

cat("✓ Model functions defined:\n")
cat("  run_empirical_marine()\n")
cat("  run_volumetric_marine()\n")
cat("  run_corrected_volumetric()\n")
cat("  run_area_yield()\n")
```

# ===== STEP 3: COMPUTE ALL SPECIES =====
# Dispatch by system type. To add a species, add a row to species_params
# and handle its system type here.

```{r compute-all-species}
results <- list()

for (i in seq_len(nrow(species_params))) {
  sp <- as.list(species_params[i, ])

  if (sp$system %in% c("marine_pens", "marine_pontoons")) {

    # Determine empirical scaling baseline
    if (!is.na(sp$density_baseline) && sp$density_baseline == "back_calc") {
      # Back-calculate effective density from reported production and pen geometry
      density_baseline <- sp$actual_prod_t / (sp$area_km2 * sp$net_depth_m * 1000)
    } else {
      # Reported production corresponds to density_mid (e.g. salmon seasonal average)
      density_baseline <- sp$density_mid
    }

    emp <- run_empirical_marine(
      sp$actual_prod_t, sp$area_km2,
      sp$density_low, density_baseline, sp$density_high
    )

    vol <- run_volumetric_marine(
      sp$area_km2,
      sp$density_low, sp$density_mid, sp$density_high,
      sp$mortality_low, sp$mortality_mid, sp$mortality_high,
      sp$net_depth_m
    )

    r <- list(
      emp              = emp,
      vol              = vol,
      effective_density = density_baseline,
      actual_prod      = sp$actual_prod_t,
      area_km2         = sp$area_km2
    )

    # Corrected volumetric: only where ISF divergence is observed (salmon)
    if (isTRUE(sp$apply_isf)) {
      cor <- run_corrected_volumetric(vol, sp$area_km2, mean_pen_fraction)
      r$cor_vol            <- cor
      r$productive_fraction <- emp$yield_per_km2 / vol$yield_mid
    }

  } else if (sp$system == "marine_longline") {
    # Marine longline (oysters): area-yield, no volumetric — same maths as pond
    # but kept as a distinct system type for semantic clarity and future species
    pond <- run_area_yield(
      sp$area_km2, sp$yield_low_tkm2, sp$yield_high_tkm2, sp$actual_prod_t
    )
    r <- list(
      pond        = pond,
      actual_prod = sp$actual_prod_t,
      area_km2    = sp$area_km2
    )

  } else {
    # Pond system: area-yield model
    pond <- run_area_yield(
      sp$area_km2, sp$yield_low_tkm2, sp$yield_high_tkm2, sp$actual_prod_t
    )
    r <- list(
      pond        = pond,
      actual_prod = sp$actual_prod_t,
      area_km2    = sp$area_km2
    )
  }

  results[[sp$species]] <- r
  cat(sprintf("✓ %s computed (%s)\n", sp$label, sp$system))
}
```

# ===== STEP 4: EXTRACT NAMED VARIABLES =====
# Creates backward-compatible aliases so all manuscript text, display tables,
# and plots work without modification.
# When updating for a new year, nothing below this chunk needs to change.

```{r extract-variables}

# ---- Atlantic Salmon ----
sal_p <- filter(species_params, species == "atl_salmon")
sal_r <- results[["atl_salmon"]]

area_km2             <- sal_r$area_km2
actual_production    <- sal_r$actual_prod
density_low          <- sal_p$density_low
density_mid          <- sal_p$density_mid
density_high         <- sal_p$density_high
mortality_low        <- sal_p$mortality_low
mortality_mid        <- sal_p$mortality_mid
mortality_high       <- sal_p$mortality_high
pen_circumference    <- sal_p$pen_circ_m
pen_volume           <- sal_p$pen_vol_m3
pen_radius           <- pen_circumference / (2 * pi)
pen_area             <- pi * pen_radius^2              # m²
pen_depth            <- pen_volume / pen_area          # m (derived from pen geometry)
survival_low         <- sal_r$vol$survival_low
survival_mid         <- sal_r$vol$survival_mid
survival_high        <- sal_r$vol$survival_high
actual_yield_per_km2 <- sal_r$emp$yield_per_km2
emp_prod_low         <- sal_r$emp$prod_low
emp_prod_high        <- sal_r$emp$prod_high
emp_yield_low        <- sal_r$emp$yield_low
emp_yield_high       <- sal_r$emp$yield_high
vol_yield_low        <- sal_r$vol$yield_low
vol_yield_mid        <- sal_r$vol$yield_mid
vol_yield_high       <- sal_r$vol$yield_high
vol_prod_low         <- sal_r$vol$prod_low
vol_prod_mid         <- sal_r$vol$prod_mid
vol_prod_high        <- sal_r$vol$prod_high
cor_vol_prod_low     <- sal_r$cor_vol$prod_low
cor_vol_prod_mid     <- sal_r$cor_vol$prod_mid
cor_vol_prod_high    <- sal_r$cor_vol$prod_high
pen_area_km2         <- sal_r$cor_vol$pen_area_km2
productive_fraction  <- sal_r$productive_fraction
prod_low             <- emp_prod_low    # primary model = empirical
prod_high            <- emp_prod_high

# ---- Southern Bluefin Tuna ----
sbt_p <- filter(species_params, species == "sbt")
sbt_r <- results[["sbt"]]

sbt_area_km2              <- sbt_r$area_km2
sbt_actual_production     <- sbt_r$actual_prod
sbt_net_depth             <- sbt_p$net_depth_m
sbt_density_low           <- sbt_p$density_low
sbt_density_mid           <- sbt_p$density_mid
sbt_density_high          <- sbt_p$density_high
sbt_mortality_low         <- sbt_p$mortality_low
sbt_mortality_mid         <- sbt_p$mortality_mid
sbt_mortality_high        <- sbt_p$mortality_high
sbt_survival_low          <- sbt_r$vol$survival_low
sbt_survival_mid          <- sbt_r$vol$survival_mid
sbt_survival_high         <- sbt_r$vol$survival_high
sbt_effective_density     <- sbt_r$effective_density
sbt_actual_yield_per_km2  <- sbt_r$emp$yield_per_km2
sbt_emp_prod_low          <- sbt_r$emp$prod_low
sbt_emp_prod_high         <- sbt_r$emp$prod_high
sbt_emp_yield_low         <- sbt_r$emp$yield_low
sbt_emp_yield_high        <- sbt_r$emp$yield_high
sbt_vol_yield_low         <- sbt_r$vol$yield_low
sbt_vol_yield_mid         <- sbt_r$vol$yield_mid
sbt_vol_yield_high        <- sbt_r$vol$yield_high
sbt_vol_prod_low          <- sbt_r$vol$prod_low
sbt_vol_prod_mid          <- sbt_r$vol$prod_mid
sbt_vol_prod_high         <- sbt_r$vol$prod_high
sbt_prod_low              <- sbt_emp_prod_low    # primary model = empirical
sbt_prod_high             <- sbt_emp_prod_high
sbt_yield_low             <- sbt_emp_yield_low
sbt_yield_high            <- sbt_emp_yield_high

# ---- Black Tiger Prawn ----
prawn_p <- filter(species_params, species == "prawn")
prawn_r <- results[["prawn"]]

prawn_area_km2          <- prawn_r$area_km2
prawn_actual_production <- prawn_r$actual_prod
prawn_prod_low          <- prawn_r$pond$prod_low
prawn_prod_high         <- prawn_r$pond$prod_high
prawn_yield_rate        <- prawn_r$pond$yield_rate
prawn_yield_low_tkm2    <- prawn_r$pond$yield_low
prawn_yield_high_tkm2   <- prawn_r$pond$yield_high
prawn_pl_low            <- prawn_p$prawn_pl_low_m2
prawn_pl_high           <- prawn_p$prawn_pl_high_m2
prawn_hw_low            <- prawn_p$prawn_hw_low_g
prawn_hw_high           <- prawn_p$prawn_hw_high_g
prawn_survival          <- 1 - prawn_p$mortality_mid

# ---- Barramundi ----
barra_p <- filter(species_params, species == "barramundi")
barra_r <- results[["barramundi"]]

barra_total_area_km2    <- 2.521   # total detected (land + marine)
barra_land_area_km2     <- barra_r$area_km2     # ponds only (production distributed here)
barra_marine_area_km2   <- barra_total_area_km2 - barra_land_area_km2
barra_actual_production <- barra_r$actual_prod
barra_prod_low          <- barra_r$pond$prod_low
barra_prod_high         <- barra_r$pond$prod_high
barra_yield_rate        <- barra_r$pond$yield_rate
barra_yield_low_tkm2    <- barra_r$pond$yield_low
barra_yield_high_tkm2   <- barra_r$pond$yield_high
barra_survival          <- 1 - barra_p$mortality_mid

# ---- Oysters (mixed spp.) ----
oyster_p <- filter(species_params, species == "oyster")
oyster_r <- results[["oyster"]]

oyster_area_km2          <- oyster_r$area_km2
oyster_actual_production <- oyster_r$actual_prod
oyster_prod_low          <- oyster_r$pond$prod_low
oyster_prod_high         <- oyster_r$pond$prod_high
oyster_yield_rate        <- oyster_r$pond$yield_rate
oyster_yield_low_tkm2    <- oyster_r$pond$yield_low
oyster_yield_high_tkm2   <- oyster_r$pond$yield_high
oyster_survival          <- 1 - oyster_p$mortality_mid

cat("✓ Variables extracted for all species\n\n")
cat(sprintf("  Atlantic Salmon:    %s t | %.2f km² | yield %.0f t/km²\n",
            format(actual_production, big.mark = ","), area_km2, actual_yield_per_km2))
cat(sprintf("  S. Bluefin Tuna:    %s t | %.3f km² | yield %.0f t/km² | eff. density %.2f kg/m³\n",
            format(sbt_actual_production, big.mark = ","), sbt_area_km2,
            sbt_actual_yield_per_km2, sbt_effective_density))
cat(sprintf("  Black Tiger Prawn:  %s t | %.3f km² | yield %.0f t/km²\n",
            format(prawn_actual_production, big.mark = ","), prawn_area_km2, prawn_yield_rate))
cat(sprintf("  Barramundi:         %s t | %.3f km² | yield %.0f t/km²\n",
            format(barra_actual_production, big.mark = ","), barra_land_area_km2, barra_yield_rate))
cat(sprintf("  Oysters (mixed):    %s t | %.3f km² | yield %.0f t/km²\n",
            format(oyster_actual_production, big.mark = ","), oyster_area_km2, oyster_yield_rate))
```

# ===== DISPLAY RESULTS: ATLANTIC SALMON =====

```{r display-salmon}
cat("\n==========================================\n")
cat("ATLANTIC SALMON PRODUCTION MODELING\n")
cat("==========================================\n")

cat("\nPen specifications (Tassal Group, 2021):\n")
cat(sprintf("  Circumference: %g m | Radius: %.2f m\n", pen_circumference, pen_radius))
cat(sprintf("  Surface area: %.0f m² | Volume: %s m³ | Depth: %.2f m\n",
            pen_area, format(pen_volume, big.mark = ","), pen_depth))

cat("\n--- APPROACH 1: EMPIRICAL (top-down) ---\n")
cat(sprintf("Empirical yield rate: %s t/km²\n",
            format(round(actual_yield_per_km2, 0), big.mark = ",")))

sal_emp_table <- data.frame(
  Scenario      = c("Low Density (Summer)", "Actual (Seasonal Average)", "High Density (Winter)"),
  Density_kg_m3 = c(density_low, density_mid, density_high),
  Yield_t_km2   = c(round(emp_yield_low, 0), round(actual_yield_per_km2, 0), round(emp_yield_high, 0)),
  Production_t  = c(round(emp_prod_low, 0), actual_production, round(emp_prod_high, 0))
)
print(sal_emp_table)

cat("\n--- APPROACH 2: VOLUMETRIC (bottom-up, 3D) ---\n")
cat(sprintf("Formula: P = A × D × d × S × 1,000\n"))
cat(sprintf("Pen depth: %.2f m | Area: %.2f km²\n\n", pen_depth, area_km2))

sal_vol_table <- data.frame(
  Scenario      = c("Low (10 kg/m³, 16% mort)", "Mid (12.2 kg/m³, 15% mort)", "High (15 kg/m³, 14% mort)"),
  Density_kg_m3 = c(density_low, density_mid, density_high),
  Survival      = c(survival_low, survival_mid, survival_high),
  Yield_t_km2   = c(round(vol_yield_low, 0), round(vol_yield_mid, 0), round(vol_yield_high, 0)),
  Production_t  = c(round(vol_prod_low, 0), round(vol_prod_mid, 0), round(vol_prod_high, 0))
)
print(sal_vol_table)

cat("\n--- INFRASTRUCTURE SCALING FACTOR ---\n")
cat("Satellite-detected vectors vs Google Earth pen measurements:\n\n")
isf_display <- fpr_data
isf_display$Detected_m2  <- format(round(isf_display$Detected_m2, 0), big.mark = ",")
isf_display$Pen_area_m2  <- format(round(isf_display$Pen_area_m2, 0), big.mark = ",")
isf_display$FPR          <- round(fpr_data$FPR, 2)
isf_display$Pen_fraction <- paste0(round(fpr_data$Pen_fraction * 100, 1), "%")
names(isf_display)[names(isf_display) == "FPR"] <- "Scaling_Factor"
print(isf_display)
cat(sprintf("\nMean scaling factor: %.2f× (detected area is ~%g× larger than pen surface)\n",
            mean_fpr, round(mean_fpr, 0)))
cat(sprintf("Mean pen fraction: %.1f%% of detected area is pen water surface\n",
            mean_pen_fraction * 100))

cat("\n--- APPROACH 3: CORRECTED VOLUMETRIC (scaling-factor-adjusted 3D) ---\n")
cat(sprintf("Effective pen area: %.2f km² × %.3f = %.3f km²\n",
            area_km2, mean_pen_fraction, pen_area_km2))
cat(sprintf("Formula: P = (A × f) × D × d × S × 1,000\n"))
cat(sprintf("  where f = measured pen fraction (%.3f)\n\n", mean_pen_fraction))

sal_cor_table <- data.frame(
  Scenario      = c("Low (10 kg/m³, 16% mort)", "Mid (12.2 kg/m³, 15% mort)", "High (15 kg/m³, 14% mort)"),
  Density_kg_m3 = c(density_low, density_mid, density_high),
  Survival      = c(survival_low, survival_mid, survival_high),
  Production_t  = c(round(cor_vol_prod_low, 0), round(cor_vol_prod_mid, 0), round(cor_vol_prod_high, 0))
)
print(sal_cor_table)

cat("\n--- PRODUCTIVE FRACTION: 3-STEP CROSS-VALIDATION ---\n")

cat("\nSTEP 1 — Raw volumetric (all detected area treated as pen surface):\n")
cat(sprintf("  If 100%% of %.2f km² were pen water surface:\n", area_km2))
cat(sprintf("  Volumetric range = %s – %s t\n",
            format(round(vol_prod_low, 0), big.mark = ","),
            format(round(vol_prod_high, 0), big.mark = ",")))
cat(sprintf("  Mid estimate     = %s t  [ABARES reported: %s t]\n",
            format(round(vol_prod_mid, 0), big.mark = ","),
            format(actual_production,     big.mark = ",")))
cat(sprintf("  → Raw volumetric OVERESTIMATES by %.1f× at mid scenario\n",
            vol_prod_mid / actual_production))

cat("\nSTEP 2 — Back-calculation: what pen fraction does the actual production imply?\n")
cat(sprintf("  Empirical yield rate = %s t / %.2f km² = %s t/km²\n",
            format(actual_production, big.mark = ","),
            area_km2,
            format(round(actual_yield_per_km2, 0), big.mark = ",")))
cat(sprintf("  Volumetric yield (mid scenario, full area) = %s t/km²\n",
            format(round(vol_yield_mid, 0), big.mark = ",")))
cat(sprintf("  Production-implied pen fraction = %s / %s = %.1f%%\n",
            format(round(actual_yield_per_km2, 0), big.mark = ","),
            format(round(vol_yield_mid, 0), big.mark = ","),
            productive_fraction * 100))
cat(sprintf("  Interpretation: only ~%.0f%% of detected area is pen water surface;\n", productive_fraction * 100))
cat( "  the remainder is inter-pen spacing, walkways, feed barges, and\n")
cat( "  the spatial extent of pen movement over time within the lease.\n")

cat("\nSTEP 3 — Direct measurement (Google Earth Pro pen digitising):\n")
cat(sprintf("  42 pens digitised across 3 leases\n"))
cat(sprintf("  Measured pen fraction = %.1f%%\n", mean_pen_fraction * 100))
cat(sprintf("  (range: %.1f%% – %.1f%% across leases)\n",
            min(fpr_data$Pen_fraction) * 100, max(fpr_data$Pen_fraction) * 100))

cat(sprintf("\nCONVERGENCE: back-calculated %.1f%% vs directly measured %.1f%%  (Δ = %.1f pp)\n",
            productive_fraction * 100, mean_pen_fraction * 100,
            abs(productive_fraction - mean_pen_fraction) * 100))
cat("Two independent methods agree → validates both the detection and the model.\n")

cat("\nCorrected volumetric range (pen fraction applied):\n")
cat(sprintf("  %s – %s t  → WITHIN range of ABARES reported %s t ✓\n",
            format(round(cor_vol_prod_low, 0), big.mark = ","),
            format(round(cor_vol_prod_high, 0), big.mark = ","),
            format(actual_production,           big.mark = ",")))

cat("\n--- VALIDATION SUMMARY ---\n")
cat(sprintf("ABARES reported: %s t\n", format(actual_production, big.mark = ",")))
cat(sprintf("Empirical range:        %s – %s t → %s\n",
            format(round(emp_prod_low, 0), big.mark = ","),
            format(round(emp_prod_high, 0), big.mark = ","),
            ifelse(emp_prod_low <= actual_production & actual_production <= emp_prod_high,
                   "WITHIN RANGE ✓", "OUTSIDE RANGE")))
cat(sprintf("Corrected Vol. range:   %s – %s t → %s\n",
            format(round(cor_vol_prod_low, 0), big.mark = ","),
            format(round(cor_vol_prod_high, 0), big.mark = ","),
            ifelse(cor_vol_prod_low <= actual_production & actual_production <= cor_vol_prod_high,
                   "WITHIN RANGE ✓", "OUTSIDE RANGE")))
cat(sprintf("Raw Volumetric range:   %s – %s t → ABOVE (expected: includes non-pen area)\n",
            format(round(vol_prod_low, 0), big.mark = ","),
            format(round(vol_prod_high, 0), big.mark = ",")))
```

# ===== DISPLAY RESULTS: SOUTHERN BLUEFIN TUNA =====

```{r display-sbt}
cat("\n========================================\n")
cat("SOUTHERN BLUEFIN TUNA PRODUCTION MODELING\n")
cat("========================================\n")

cat("\n--- APPROACH 1: EMPIRICAL (top-down) ---\n")
cat(sprintf("Empirical yield rate: %s t/km²\n",
            format(round(sbt_actual_yield_per_km2, 0), big.mark = ",")))
cat(sprintf("Back-calculated effective density: %.2f kg/m³\n", sbt_effective_density))
cat(sprintf("  (within regulatory range of %g–%g kg/m³)\n\n", sbt_density_low, sbt_density_high))

sbt_emp_table <- data.frame(
  Scenario      = c("Low Density", "Actual (Effective)", "High Density"),
  Density_kg_m3 = c(sbt_density_low, round(sbt_effective_density, 2), sbt_density_high),
  Yield_t_km2   = c(round(sbt_emp_yield_low, 0), round(sbt_actual_yield_per_km2, 0), round(sbt_emp_yield_high, 0)),
  Production_t  = c(round(sbt_emp_prod_low, 0), sbt_actual_production, round(sbt_emp_prod_high, 0))
)
print(sbt_emp_table)

cat("\n--- APPROACH 2: VOLUMETRIC (bottom-up, 3D) ---\n")
cat(sprintf("Formula: P = A × D × d × S × 1,000\n"))
cat(sprintf("Net depth: %g m | Area: %.3f km²\n\n", sbt_net_depth, sbt_area_km2))

sbt_vol_table <- data.frame(
  Scenario      = c("Low (2 kg/m³, 4% mort)", "Mid (3 kg/m³, 3% mort)", "High (4 kg/m³, 2% mort)"),
  Density_kg_m3 = c(sbt_density_low, sbt_density_mid, sbt_density_high),
  Survival      = c(sbt_survival_low, sbt_survival_mid, sbt_survival_high),
  Yield_t_km2   = c(round(sbt_vol_yield_low, 0), round(sbt_vol_yield_mid, 0), round(sbt_vol_yield_high, 0)),
  Production_t  = c(round(sbt_vol_prod_low, 0), round(sbt_vol_prod_mid, 0), round(sbt_vol_prod_high, 0))
)
print(sbt_vol_table)

cat("\n--- CONVERGENCE CHECK ---\n")
cat(sprintf("Empirical yield rate:               %s t/km²\n",
            format(round(sbt_actual_yield_per_km2, 0), big.mark = ",")))
cat(sprintf("Volumetric yield at effective density\n"))
cat(sprintf("  (%.2f kg/m³ × %gm × %.2f × 1,000):  %s t/km²\n",
            sbt_effective_density, sbt_net_depth, sbt_survival_mid,
            format(round(sbt_effective_density * sbt_net_depth * sbt_survival_mid * 1000, 0), big.mark = ",")))
cat(sprintf("Agreement: %.1f%%\n",
            (sbt_effective_density * sbt_net_depth * sbt_survival_mid * 1000) / sbt_actual_yield_per_km2 * 100))

cat("\n--- VALIDATION ---\n")
cat(sprintf("ABARES reported: %s t\n", format(sbt_actual_production, big.mark = ",")))
cat(sprintf("Empirical range:   %s – %s t → %s\n",
            format(round(sbt_emp_prod_low, 0), big.mark = ","),
            format(round(sbt_emp_prod_high, 0), big.mark = ","),
            ifelse(sbt_emp_prod_low <= sbt_actual_production & sbt_actual_production <= sbt_emp_prod_high,
                   "WITHIN RANGE ✓", "OUTSIDE RANGE")))
cat(sprintf("Volumetric range:  %s – %s t → %s\n",
            format(round(sbt_vol_prod_low, 0), big.mark = ","),
            format(round(sbt_vol_prod_high, 0), big.mark = ","),
            ifelse(sbt_vol_prod_low <= sbt_actual_production & sbt_actual_production <= sbt_vol_prod_high,
                   "WITHIN RANGE ✓", "OUTSIDE RANGE")))
```

# ===== DISPLAY RESULTS: PRAWNS AND BARRAMUNDI =====

```{r display-prawn-barra}
cat("\n========================================\n")
cat("PRAWN & BARRAMUNDI PRODUCTION MODELING\n")
cat("========================================\n")

cat("\n--- BLACK TIGER PRAWN ---\n")
cat(sprintf("Model: Area-yield (P = A × Y)\n"))
cat(sprintf("Yield range: %.0f–%.0f t/km² (%.1f–%.1f t/ha)\n",
            prawn_yield_low_tkm2, prawn_yield_high_tkm2,
            prawn_yield_low_tkm2 / 100, prawn_yield_high_tkm2 / 100))
cat(sprintf("Production range: %s–%s t\n",
            format(round(prawn_prod_low, 0), big.mark = ","),
            format(round(prawn_prod_high, 0), big.mark = ",")))
cat(sprintf("ABARES reported: %s t → %s\n",
            format(prawn_actual_production, big.mark = ","),
            ifelse(prawn_prod_low <= prawn_actual_production & prawn_actual_production <= prawn_prod_high,
                   "WITHIN RANGE ✓", "OUTSIDE RANGE")))
cat(sprintf("Empirical yield rate: %.0f t/km² (%.2f t/ha)\n",
            prawn_yield_rate, prawn_yield_rate / 100))

cat("\n--- BARRAMUNDI ---\n")
cat(sprintf("Model: Area-yield (P = A × Y), land-based ponds only\n"))
cat(sprintf("Yield range: %.0f–%.0f t/km² (%.0f–%.0f t/ha)\n",
            barra_yield_low_tkm2, barra_yield_high_tkm2,
            barra_yield_low_tkm2 / 100, barra_yield_high_tkm2 / 100))
cat(sprintf("Production range: %s–%s t\n",
            format(round(barra_prod_low, 0), big.mark = ","),
            format(round(barra_prod_high, 0), big.mark = ",")))
cat(sprintf("ABARES reported: %s t → %s\n",
            format(barra_actual_production, big.mark = ","),
            ifelse(barra_prod_low <= barra_actual_production & barra_actual_production <= barra_prod_high,
                   "WITHIN RANGE ✓", "OUTSIDE RANGE")))
cat(sprintf("Empirical yield rate: %.0f t/km² (%.2f t/ha)\n",
            barra_yield_rate, barra_yield_rate / 100))
```

# ===== DISPLAY RESULTS: OYSTERS =====

```{r display-oysters}
cat("\n========================================\n")
cat("OYSTERS (MIXED SPP.) PRODUCTION MODELING\n")
cat("========================================\n")

cat(sprintf("\nModel: Area-yield (P = A × Y), marine longline systems\n"))
cat(sprintf("Total detected infrastructure: %.3f km²\n", oyster_area_km2))
cat(sprintf("  NSW: 5.898 km² | QLD: 0.409 km² | SA: 6.789 km² | TAS: 3.128 km² | WA: 0.476 km²\n"))
cat(sprintf("\nYield range: %.0f-%.0f t/km² (%.0f-%.0f t/ha)\n",
            oyster_yield_low_tkm2, oyster_yield_high_tkm2,
            oyster_yield_low_tkm2 / 100, oyster_yield_high_tkm2 / 100))
cat(sprintf("  Sources: FAO extensive lease yield ~25 t/ha (Helm, 2009) divided by\n"))
cat(sprintf("  grow-out duration to reflect multi-cohort steady-state farm structure.\n"))
cat(sprintf("  SRO: 25/3yr = ~8.3 t/ha (Nell, 1993; Sapphire Coast Oysters, 2024).\n"))
cat(sprintf("  Pacific oyster: 25/1.5-2.5yr = ~10-17 t/ha (Helm, 2009).\n"))
cat(sprintf("Production range: %s - %s t\n",
            format(round(oyster_prod_low, 0), big.mark = ","),
            format(round(oyster_prod_high, 0), big.mark = ",")))
cat(sprintf("ABARES reported: %s t → %s\n",
            format(oyster_actual_production, big.mark = ","),
            ifelse(oyster_prod_low <= oyster_actual_production &
                     oyster_actual_production <= oyster_prod_high,
                   "WITHIN RANGE ✓", "OUTSIDE RANGE")))
cat(sprintf("Empirical yield rate: %.0f t/km² (%.2f t/ha)\n",
            oyster_yield_rate, oyster_yield_rate / 100))
cat(sprintf("Survival: %.0f%% (mortality %.0f%% per growing season)\n",
            oyster_survival * 100, (1 - oyster_survival) * 100))
cat(sprintf("\nNote: ABARES reports all oyster species in a single category.\n"))
cat(sprintf("Detected infrastructure encompasses primarily Pacific Oysters (SA, TAS)\n"))
cat(sprintf("and Sydney Rock Oysters (NSW), with minor production in QLD and WA.\n"))
```

# ===== INFRASTRUCTURE SCALING FACTOR TABLE =====

```{r infrastructure-scaling-factor-table}
Table_isf <- data.frame(
  Lease_ID = c(220, 201, 141, NA),
  Number_of_Pens = c(12, 18, 12, 42),
  Measured_Pen_Area_m2 = c(17926.81, 45641.38, 59771.19, 123339.38),
  Detected_Infrastructure_m2 = c(182418.40, 362625.60, 542620.70, 1087664.70),
  Infrastructure_Scaling_Factor = c(
    182418.40 / 17926.81,
    362625.60 / 45641.38,
    542620.70 / 59771.19,
    mean_fpr
  ),
  Pen_Fraction_pct = c(
    17926.81 / 182418.40 * 100,
    45641.38 / 362625.60 * 100,
    59771.19 / 542620.70 * 100,
    mean_pen_fraction * 100
  )
)

Table_isf_display <- Table_isf
Table_isf_display$Lease_ID <- ifelse(is.na(Table_isf_display$Lease_ID), "Mean",
                                      as.character(Table_isf_display$Lease_ID))
Table_isf_display$Measured_Pen_Area_m2 <- format(round(Table_isf_display$Measured_Pen_Area_m2, 2),
                                                   big.mark = ",")
Table_isf_display$Detected_Infrastructure_m2 <- format(round(Table_isf_display$Detected_Infrastructure_m2, 2),
                                                         big.mark = ",")
Table_isf_display$Infrastructure_Scaling_Factor <- round(Table_isf_display$Infrastructure_Scaling_Factor, 2)
Table_isf_display$Pen_Fraction_pct <- round(Table_isf_display$Pen_Fraction_pct, 1)

cat("\n=== TABLE X: INFRASTRUCTURE SCALING FACTOR ANALYSIS ===\n")
cat("Atlantic salmon lease sites, Tasmania\n")
cat("Pen areas measured from Google Earth Pro; detected areas from Alpha Earth Foundation embeddings\n\n")
print(Table_isf_display, row.names = FALSE)

isf_output_path <- here("output_data/SSRG_GEE_results/Table_Infrastructure_Scaling_Factor.csv")
write.csv(Table_isf, isf_output_path, row.names = FALSE)
cat(sprintf("\nTable saved to: %s\n", isf_output_path))
```

# ===== MANUSCRIPT TEXT: GENERAL FRAMEWORK =====

```{r general-framework-manuscript-text, results='asis'}
cat("
## Species-specific production distribution

## Marine finfish systems

For marine finfish cultured in floating net pens or pontoons, production was 
estimated using two complementary approaches consistent with the global 
aquaculture spatial framework of Halpern et al. (2022): (i) an empirical 
(top-down) approach and (ii) a volumetric (bottom-up) approach.

The empirical approach derived realised yield (t km⁻²) by dividing 
ABARES-reported production for the reference financial year (2021–2022) by the 
total satellite-detected aquaculture infrastructure area in the producing 
state(s). This empirical yield implicitly incorporates operational 
characteristics such as fallowing rotations, inter-pen spacing, and 
non-productive infrastructure within the detected aquaculture infrastructure.

The volumetric approach estimated theoretical production capacity from 
three-dimensional pen or pontoon volume using stocking density (kg m⁻³), 
net depth (m), survival rate, and a unit conversion factor (×1,000).

Where empirical and volumetric estimates diverged, an infrastructure scaling 
factor was quantified to reconcile differences between satellite-detected 
infrastructure area and effective productive water surface. This correction 
factor was applied to adjust the volumetric estimate where appropriate, and 
was independently cross-validated against the production-derived productive 
fraction.

Species-specific parameterisation is described in the following sections.
")
```

# ===== MANUSCRIPT TEXT: ATLANTIC SALMON =====

```{r salmon-manuscript-text, results='asis'}
cat("
## Atlantic salmon (Salmo salar)

Atlantic salmon production occurs almost exclusively in Tasmania (>98% of 
national salmonid production).

Empirical yield was calculated by dividing ABARES-reported salmonid production 
(2021–2022) by the total detected infrastructure area in Tasmania. Production 
ranges were generated using reported seasonal stocking densities (", density_low, 
" kg m⁻³ in summer to ", density_high, " kg m⁻³ in winter; seasonal average ", 
density_mid, " kg m⁻³; Tassal Group, 2015; 2021). We applied mortality estimations for Atlantic Salmon based on a study in Norway, ranging from ", 
mortality_low * 100, "to", mortality_high * 100, "%(Bang Jensen et al., 2020), as we were unable to find public and straightforward numerical annual mortality or survival rates in Australian literature.


For the volumetric approach, pen geometry (", pen_circumference, " m circumference; 
", format(pen_volume, big.mark = ','), " m³ volume; ", round(pen_depth, 2), 
" m net depth) was derived from industry reporting (Tassal Group, 2021). 
Because satellite-detected infrastructure polygons include non-productive 
components (e.g., inter-pen spacing, walkways, feed barges, fallow positions), 
an infrastructure scaling factor ('pen fraction') was quantified from a 
stratified sample of lease sites (n = 42 pens across leases 141, 201, and 220). 
Individual pen boundaries were manually digitised from high-resolution imagery 
in Google Earth Pro and compared against satellite-detected infrastructure 
polygons. This correction was applied to generate a reconciled volumetric 
estimate: P = (A × f) × D × d × S × 1,000.


## RESULTS — Atlantic salmon production distribution model

Satellite detection identified ", sprintf('%.2f', area_km2), " km² of active 
aquaculture infrastructure for Atlantic salmon (Salmo salar) within Tasmania's 
25.88 km² of finfish leases (23.87% of leased area). The empirical yield rate 
was ", format(round(actual_yield_per_km2, 0), big.mark = ','), " t/km² 
of detected aquaculture infrastructure, or ", 
format(round(actual_production / 25.88, 0), big.mark = ','), " t/km² of total 
finfish lease area. Density-scaled production ranged from ", 
format(round(emp_prod_low, 0), big.mark = ','), " to ", 
format(round(emp_prod_high, 0), big.mark = ','), " t, encompassing the 
ABARES-reported salmonid production of ", format(actual_production, big.mark = ','), " t 
(AUD 1.15B, 2021–2022), of which >98% comprises Atlantic salmon produced 
exclusively in Tasmanian marine lease areas (the salmonid category includes 
minor freshwater trout production from mainland states).

The uncorrected volumetric approach yielded substantially higher estimates 
(", format(round(vol_yield_low, 0), big.mark = ','), "–", 
format(round(vol_yield_high, 0), big.mark = ','), " t/km²), reflecting the 
assumption that the entire detected aquaculture infrastructure represents 
productive pen water surface. Infrastructure scaling factor analysis across 
the three sampled leases (Table X) revealed a mean scaling factor of ", 
sprintf('%.2f', mean_fpr), "× (range: ", sprintf('%.2f', min(fpr_data$FPR)), 
"–", sprintf('%.2f', max(fpr_data$FPR)), "×), corresponding to a mean pen 
fraction of ", sprintf('%.1f', mean_pen_fraction * 100), "% (Table X). A 
total of 42 pens were digitised across three leases, encompassing ", 
format(round(sum(fpr_data$Pen_area_m2), 0), big.mark = ','), " m² of pen 
water surface within ", format(round(sum(fpr_data$Detected_m2), 0), big.mark = ','), 
" m² of detected aquaculture infrastructure.

When the measured pen fraction was applied to correct the volumetric estimates, 
the resulting production range (", format(round(cor_vol_prod_low, 0), big.mark = ','), 
"–", format(round(cor_vol_prod_high, 0), big.mark = ','), " t) converged with 
the empirical range and encompassed reported production. Notably, the directly 
measured pen fraction (", sprintf('%.1f', mean_pen_fraction * 100), "%) closely 
agreed with the production-derived productive fraction (", 
sprintf('%.1f', productive_fraction * 100), "%), calculated independently as 
the ratio of empirical to volumetric yield rates. This convergence between 
measurement-based and production-based estimates provides cross-validation of 
both the satellite detection methodology and the production distribution model, 
confirming that the empirical yield rate approach correctly accounts for the 
non-productive components captured within satellite-detected aquaculture 
infrastructure (consistent with the disturbance framework of Halpern et al., 2022).
")
```
#results Salmon shorter with corrections
```{r}
cat("
If all ", sprintf('%.2f', area_km2), " km² of detected aquaculture infrastructure 
had been treated as productive pen water surface — as would be assumed without 
an infrastructure scaling factor — the volumetric model would estimate ", 
format(round(vol_prod_low, 0), big.mark = ','), "–", 
format(round(vol_prod_high, 0), big.mark = ','), " t, a ", 
sprintf('%.1f', vol_prod_mid / actual_production), 
"× overestimate of ABARES-reported production at the mid scenario. This 
discrepancy arises because satellite-detected infrastructure polygons capture 
not only pen water surface but also inter-pen spacing, walkways, feed barges, 
and the spatial envelope of pen movement within the lease over time.

To quantify what fraction of detected area is truly productive, we 
back-calculated the implied pen fraction from reported production: the empirical 
yield rate of ", format(round(actual_yield_per_km2, 0), big.mark = ','), 
" t km⁻² divided by the volumetric yield of ", 
format(round(vol_yield_mid, 0), big.mark = ','), 
" t km⁻² implies that ", sprintf('%.1f', productive_fraction * 100), 
"% of detected area represents pen water surface. This 
production-derived estimate was independently verified by manually digitising 
individual pen boundaries from high-resolution imagery in Google Earth Pro 
across three lease sites (n = 42 pens; Table X). Direct measurement yielded 
a mean pen fraction of ", sprintf('%.1f', mean_pen_fraction * 100), 
"% (range: ", sprintf('%.1f', min(fpr_data$Pen_fraction) * 100), "–", 
sprintf('%.1f', max(fpr_data$Pen_fraction) * 100), 
"% across leases), closely agreeing with the production-implied estimate 
(difference: ", sprintf('%.1f', abs(productive_fraction - mean_pen_fraction) * 100), 
" percentage points). This convergence between two independent approaches — 
one derived from production statistics, one from direct spatial measurement — 
confirms that approximately ", sprintf('%.0f', mean_pen_fraction * 100), 
"% of satellite-detected aquaculture infrastructure corresponds to pen water 
surface, with the remaining ~", sprintf('%.0f', (1 - mean_pen_fraction) * 100), 
"% comprising non-productive components.

Applying the measured pen fraction to the volumetric model yielded a corrected 
production range of ", format(round(cor_vol_prod_low, 0), big.mark = ','), 
"–", format(round(cor_vol_prod_high, 0), big.mark = ','), 
" t, which encompassed the ABARES-reported production of ", 
format(actual_production, big.mark = ','), 
" t and converged with the empirical range of ", 
format(round(emp_prod_low, 0), big.mark = ','), "–", 
format(round(emp_prod_high, 0), big.mark = ','), 
" t (consistent with the disturbance framework of Halpern et al., 2022).
")
```

# ===== MANUSCRIPT TEXT: SOUTHERN BLUEFIN TUNA =====

```{r sbt-manuscript-text, results='asis'}
cat("
## Southern Bluefin Tuna (Thunnus maccoyii)

Southern Bluefin Tuna production consists of marine-based capture ranching 
confined to the Lincoln Offshore Aquaculture Zone in Spencer Gulf, South 
Australia.

Empirical yield was calculated by dividing ABARES-reported production 
(2021–2022) by the detected aquaculture infrastructure area in South Australia. 
Production ranges were generated using the regulatory stocking density range of 
", sbt_density_low, "–", sbt_density_high, " kg m⁻³ (Ma, 2021), with 
capture-to-harvest mortality of ", sbt_mortality_low * 100, "–", 
sbt_mortality_high * 100, "% (Ma, 2021).

For the volumetric approach, pontoon geometry (45 m diameter; ", sbt_net_depth, 
" m net depth; Ellis and Kiessling, 2014) was used to estimate theoretical 
production capacity. No infrastructure scaling factor was applied, as the 
isolated pontoon configuration means detected infrastructure is expected to 
closely approximate productive water surface. The effective stocking density 
implied by ABARES-reported production was back-calculated to assess internal 
model consistency.


## RESULTS — Southern Bluefin Tuna production distribution model

Southern Bluefin Tuna (Thunnus maccoyii) aquaculture infrastructure was 
detected exclusively within South Australia's Lincoln Offshore Aquaculture 
Zone in Spencer Gulf, comprising ", sprintf('%.3f', sbt_area_km2), " km² of 
marine infrastructure within 36.627 km² of lease area. All production is 
marine-based capture ranching, where wild-caught juveniles are transferred to 
sea pens for grow-out. The empirical yield rate was ", 
format(round(sbt_actual_yield_per_km2, 0), big.mark = ','), " t/km² of 
detected aquaculture infrastructure, substantially higher than Atlantic 
salmon (", format(round(actual_yield_per_km2, 0), big.mark = ','), 
" t/km²), reflecting the high production density of SBT ranching: ", 
format(sbt_actual_production, big.mark = ','), " t produced from only ", 
sprintf('%.3f', sbt_area_km2), " km² of detected infrastructure, compared to ", 
format(actual_production, big.mark = ','), " t from ", sprintf('%.2f', area_km2), 
" km² for Atlantic salmon. Density-scaled empirical production ranged from ", 
format(round(sbt_emp_prod_low, 0), big.mark = ','), " to ", 
format(round(sbt_emp_prod_high, 0), big.mark = ','), " t, encompassing the 
ABARES-reported production of ", format(sbt_actual_production, big.mark = ','), 
" t (AUD 110.4M, 2021–22).

The volumetric approach produced convergent estimates: at the back-calculated 
effective stocking density of ", sprintf('%.2f', sbt_effective_density), 
" kg/m³ across ", sbt_net_depth, " m net depth with ", sprintf('%.0f', sbt_survival_mid * 100), 
"% survival, the volumetric yield was approximately ", 
format(round(sbt_effective_density * sbt_net_depth * sbt_survival_mid * 1000, 0), big.mark = ','), 
" t/km², validating the empirical rate (", 
sprintf('%.1f', (sbt_effective_density * sbt_net_depth * sbt_survival_mid * 1000) / sbt_actual_yield_per_km2 * 100), 
"% agreement). The full volumetric production range (", 
format(round(sbt_vol_prod_low, 0), big.mark = ','), "–", 
format(round(sbt_vol_prod_high, 0), big.mark = ','), 
" t) also encompassed reported production. The back-calculated effective density of ", 
sprintf('%.2f', sbt_effective_density), " kg/m³ falls within the upper end of the 
regulatory range of ", sbt_density_low, "–", sbt_density_high, 
" kg/m³ (Ma, 2021), consistent with the intensive nature of southern bluefin tuna 
ranching operations within the Lincoln Offshore Aquaculture Zone. Unlike Atlantic 
salmon, where the pen fraction was ", sprintf('%.1f', mean_pen_fraction * 100), 
"%, meaning that only ~", round(mean_pen_fraction * 100), 
"% of detected aquaculture infrastructure represented productive pen water surface, 
the close agreement between empirical and volumetric yields for Southern Bluefin 
Tuna suggests that the detected aquaculture infrastructure closely approximates 
productive pontoon water surface, consistent with the isolated pontoon configuration 
of tuna ranching operations.
")
```

# ===== MANUSCRIPT TEXT: PRAWNS AND BARRAMUNDI =====

```{r prawn-barra-manuscript-text, results='asis'}
cat("
## Land-based pond systems

For pond-based systems, production was estimated using an area-based yield 
model. Species-specific annual yield rates (t km⁻²) were applied to the 
productive water footprint detected through Earth observation. Where yield 
ranges were reported, low and high values were used to generate a predicted 
production range. Production was calculated as:

P = A × Y

where A is the detected productive pond area (km²) and Y is the annual yield 
rate (t km⁻²). Mortality adjustments were incorporated where available.

## Barramundi (Lates calcarifer)

Barramundi production was distributed across detected land-based pond 
infrastructure using a yield range of ", 
sprintf('%.0f', barra_yield_low_tkm2), "–", sprintf('%.0f', barra_yield_high_tkm2), 
" t km⁻² (", sprintf('%.0f', barra_yield_low_tkm2 / 100), "–", 
sprintf('%.0f', barra_yield_high_tkm2 / 100), 
" t ha⁻¹), consistent with reported pond-based production rates (Business Queensland, 2025). As ABARES does not report marine barramundi production 
separately, production was distributed across terrestrial pond infrastructure 
only. The area-based approach reflects the predominantly pond-based production 
structure of barramundi aquaculture in Australia and does not require 
volumetric parameterisation.

## Black Tiger Prawn (Penaeus monodon)

Prawn production was estimated using annualised yield rates applied to detected 
pond area. Yield was derived from stocking density (", prawn_pl_low, "–", 
prawn_pl_high, " post-larvae m⁻²; Lovatelli et al., 2008), harvest weight (", 
prawn_hw_low, "–", prawn_hw_high, " g), and survival (", prawn_survival * 100, 
"%), yielding a production range of ", 
sprintf('%.0f', prawn_yield_low_tkm2), "–", sprintf('%.0f', prawn_yield_high_tkm2), 
" t km⁻². Yield assumptions reflect typical seasonal stocking cycles within 
the financial year. No volumetric modelling or infrastructure scaling factor 
was required, as production in pond-based prawn aquaculture is directly 
proportional to cultivated surface area.


## RESULTS — Prawn production distribution model

The area occupied by prawn aquaculture infrastructure nationally (", 
sprintf('%.3f', prawn_area_km2), " km²), all in land-based sites, could 
generate production between ", format(round(prawn_prod_low, 0), big.mark = ','), " and ", 
format(round(prawn_prod_high, 0), big.mark = ','), " tonnes. ABARES-reported production 
was ", format(prawn_actual_production, big.mark = ','), " t (AUD 180.8M, 
2021–2022), falling within the predicted range and corresponding to an empirical 
yield rate of ", format(round(prawn_yield_rate, 0), big.mark = ','), " t/km² 
(", sprintf('%.2f', prawn_yield_rate / 100), " t/ha). It remains below the typical range for 
semi-intensive operations of 1,000 to 1,750 t/km², suggesting that not all 
detected pond area is simultaneously in production.


## RESULTS — Barramundi production distribution model

Total detected barramundi aquaculture infrastructure was ", 
sprintf('%.3f', barra_total_area_km2), " km², comprising ", 
sprintf('%.3f', barra_land_area_km2), " km² of terrestrial ponds and ", 
sprintf('%.3f', barra_marine_area_km2), " km² of marine cage infrastructure. 
As ABARES did not report barramundi production from marine environments 
separately, production was distributed across land-based farms only. The 
expected production range of ", format(round(barra_prod_low, 0), big.mark = ','), " to ", 
format(round(barra_prod_high, 0), big.mark = ','), " tonnes encompasses the 
ABARES-reported production of ", format(barra_actual_production, big.mark = ','), 
" t (AUD 60.6M, 2021–2022), corresponding to an empirical yield rate of ", 
format(round(barra_yield_rate, 0), big.mark = ','), " t/km² (", 
sprintf('%.2f', barra_yield_rate / 100), " t/ha). This exceeds typical 
Queensland pond-based production rates of 1,000 to 1,500 t/km² but remains below 
maximum reported yields of 3,000 t/km², suggesting moderately intensified 
operations.
")
```

# ===== MANUSCRIPT TEXT: OYSTERS =====

```{r oyster-manuscript-text, results='asis'}
cat("
## Oysters (mixed spp.)

Oyster production was modelled as a single aggregated category consistent with 
ABARES reporting, which does not disaggregate production by species. Detected 
longline infrastructure encompasses farms producing primarily Pacific Oysters 
(*Crassostrea gigas*) in Tasmania and South Australia, and Sydney Rock Oysters 
(*Saccostrea glomerata*) in New South Wales, with minor production in Queensland 
and Western Australia.

An annual yield range of ", 
sprintf('%.0f', oyster_yield_low_tkm2), "–", sprintf('%.0f', oyster_yield_high_tkm2), 
" t km⁻² (", sprintf('%.0f', oyster_yield_low_tkm2 / 100), "–", 
sprintf('%.0f', oyster_yield_high_tkm2 / 100), 
" t ha⁻¹) was applied to the total detected longline infrastructure area. 
Extensive Pacific oyster lease areas yield approximately 25 t ha⁻¹ yr⁻¹ 
(Helm, 2009), with oysters harvested at a market size of 70–100+ g live weight. 
These single-cycle productivity figures are not directly applicable to 
satellite-detected infrastructure because remote sensing captures total farm 
area across all grow-out stages simultaneously. Sydney Rock Oysters — the 
dominant species in NSW, the largest producing state by detected area — require 
a minimum of three years to reach market size (Nell, 1993; Sapphire Coast 
Oysters, 2024). In a steady-state farm operating continuous recruitment, only 
approximately one cohort in three is harvestable in any given year, yielding an 
expected annualised whole-farm rate of approximately 25 ÷ 3 ≈ 8.3 t ha⁻¹. 
Pacific Oysters, which dominate detected infrastructure in South Australia and 
Tasmania and reach market size in 18–30 months (Helm, 2009), produce higher 
annualised rates of approximately 10–17 t ha⁻¹. The applied range of ", 
sprintf('%.0f', oyster_yield_low_tkm2 / 100), "–", 
sprintf('%.0f', oyster_yield_high_tkm2 / 100), 
" t ha⁻¹ therefore spans the expected annualised productivity of both species 
systems and is anchored by the empirically derived national yield rate of ", 
sprintf('%.2f', oyster_yield_rate / 100), 
" t ha⁻¹, which integrates the mixed-species, mixed-cohort composition of the 
national fleet. A background mortality rate of ", 
sprintf('%.0f', (1 - oyster_survival) * 100), 
"% per growing season was assumed, consistent with reported non-disease 
cumulative mortality for *S. glomerata* in commercial conditions (Nell and 
Perkins, 2006; Green et al., 2008) and within the range reported for 
off-bottom longline culture internationally (Walton et al., 2013; 
Petrolia et al., 2024). No volumetric modelling or infrastructure scaling 
factor was required, as longline production is directly proportional to 
cultivated surface area.

## RESULTS — Oyster production distribution model

Across the ", sprintf('%.3f', oyster_area_km2), " km² of oyster longline 
infrastructure detected nationally, estimated production ranged from ", 
format(round(oyster_prod_low, 0), big.mark = ','), " to ", 
format(round(oyster_prod_high, 0), big.mark = ','), 
" t. ABARES-reported oyster production was ", 
format(oyster_actual_production, big.mark = ','), 
" t (AUD 149.7 M, 2021–2022), falling within the predicted range and 
corresponding to an empirical yield rate of ", 
format(round(oyster_yield_rate, 0), big.mark = ','), " t km⁻² (", 
sprintf('%.2f', oyster_yield_rate / 100), " t ha⁻¹).
")
```

# ===== TABLE S6: PRODUCTION MODELING SUMMARY =====
# Now built programmatically from species_params + results.
# Adding a new species requires only a new row in species_params above.

```{r create-table-s6}
Table_S6 <- data.frame(
  Species = species_params$label_sci,

  Farming_Method = c(
    "Sea cages/pens",
    "Sea pens (capture-based ranching)",
    "Earthen ponds",
    "Ponds/sea cages",
    "Marine longlines (intertidal/subtidal)"
  ),

  Production_Model = c(
    "Empirical + Volumetric (3D) + Corrected Volumetric",
    "Empirical + Volumetric (3D)",
    "Area-yield (P = A × Y)",
    "Area-yield (P = A × Y)",
    "Area-yield (P = A × Y)"
  ),

  Mathematical_Expression = c(
  "Emp: P = A * Y(D) | Vol: P = A * D * d * S * 1,000 | Cor.Vol: P = (A * f) * D * d * S * 1,000",
  "Emp: P = A * Y(D) | Vol: P = A * D * d * S * 1,000",
  "P = A * Y",
  "P = A * Y",
  "P = A * Y"
),

  Productive_Area_km2 = sprintf("%.3f", species_params$area_km2),

  Stocking_Density = c(
  sprintf("%.0f-%.0f kg/m3 (avg %.1f kg/m3)", density_low, density_high, density_mid),
  sprintf("%.0f-%.0f kg/m3 (effective %.2f kg/m3)", sbt_density_low, sbt_density_high, sbt_effective_density),
  sprintf("%g-%g PL/m2 (midpoint %.1f PL/m2)", prawn_pl_low, prawn_pl_high, mean(c(prawn_pl_low, prawn_pl_high))),
  "N/A",
  "250,000-600,000 oysters/ha (longline baskets)"
),

  Survival_Rate = c(
  sprintf("%.0f-%.0f%%", survival_low * 100, survival_high * 100),
  sprintf("%.0f-%.0f%%", sbt_survival_low * 100, sbt_survival_high * 100),
  sprintf("%.0f%%", prawn_survival * 100),
  sprintf("~%.0f%%", barra_survival * 100),
  sprintf("%.0f%%", oyster_survival * 100)
),

 Mortality_Rate = c(
  sprintf("%.0f-%.0f%%", mortality_low * 100, mortality_high * 100),
  sprintf("%.0f-%.0f%%", sbt_mortality_low * 100, sbt_mortality_high * 100),
  sprintf("%.0f%%", (1 - prawn_survival) * 100),
  "Incorporated in Y",
  sprintf("%.0f%% per growing season", (1 - oyster_survival) * 100)
),

 Calculated_Yield_t_km2 = c(
  sprintf("Emp: %s-%s | Cor.Vol: %s-%s | Raw Vol: %s-%s t/km2",
          format(round(emp_yield_low, 0), big.mark = ","),
          format(round(emp_yield_high, 0), big.mark = ","),
          format(round(vol_yield_low * mean_pen_fraction, 0), big.mark = ","),
          format(round(vol_yield_high * mean_pen_fraction, 0), big.mark = ","),
          format(round(vol_yield_low, 0), big.mark = ","),
          format(round(vol_yield_high, 0), big.mark = ",")),
  sprintf("Emp: %s-%s | Vol: %s-%s t/km2",
          format(round(sbt_emp_yield_low, 0), big.mark = ","),
          format(round(sbt_emp_yield_high, 0), big.mark = ","),
          format(round(sbt_vol_yield_low, 0), big.mark = ","),
          format(round(sbt_vol_yield_high, 0), big.mark = ",")),
  sprintf("%.0f-%.0f t/km2 (%.1f-%.1f t/ha)",
          prawn_yield_low_tkm2, prawn_yield_high_tkm2,
          prawn_yield_low_tkm2 / 100, prawn_yield_high_tkm2 / 100),
  sprintf("%.0f-%.0f t/km2 (%.0f-%.0f t/ha)",
          barra_yield_low_tkm2, barra_yield_high_tkm2,
          barra_yield_low_tkm2 / 100, barra_yield_high_tkm2 / 100),
  sprintf("%.0f-%.0f t/km2 (%.0f-%.0f t/ha)",
          oyster_yield_low_tkm2, oyster_yield_high_tkm2,
          oyster_yield_low_tkm2 / 100, oyster_yield_high_tkm2 / 100)
),
  Production_Range_t = c(
  sprintf("Emp: %s-%s | Cor.Vol: %s-%s | Raw Vol: %s-%s",
          format(round(emp_prod_low, 0), big.mark = ","),
          format(round(emp_prod_high, 0), big.mark = ","),
          format(round(cor_vol_prod_low, 0), big.mark = ","),
          format(round(cor_vol_prod_high, 0), big.mark = ","),
          format(round(vol_prod_low, 0), big.mark = ","),
          format(round(vol_prod_high, 0), big.mark = ",")),
  sprintf("Emp: %s-%s | Vol: %s-%s",
          format(round(sbt_emp_prod_low, 0), big.mark = ","),
          format(round(sbt_emp_prod_high, 0), big.mark = ","),
          format(round(sbt_vol_prod_low, 0), big.mark = ","),
          format(round(sbt_vol_prod_high, 0), big.mark = ",")),
  sprintf("%s-%s", format(round(prawn_prod_low, 0), big.mark = ","),
                   format(round(prawn_prod_high, 0), big.mark = ",")),
  sprintf("%s-%s", format(round(barra_prod_low, 0), big.mark = ","),
                   format(round(barra_prod_high, 0), big.mark = ",")),
  sprintf("%s-%s", format(round(oyster_prod_low, 0), big.mark = ","),
                   format(round(oyster_prod_high, 0), big.mark = ","))
),

  Actual_Production_ABARES_t = c(
    format(actual_production, big.mark = ","),
    format(sbt_actual_production, big.mark = ","),
    format(prawn_actual_production, big.mark = ","),
    format(barra_actual_production, big.mark = ","),
    format(oyster_actual_production, big.mark = ",")
  ),

  Validation = c(
    "Within predicted range (all 3 approaches)",
    "Within predicted range",
    "Within predicted range",
    "Within predicted range",
    ifelse(oyster_prod_low <= oyster_actual_production &
             oyster_actual_production <= oyster_prod_high,
           "Within predicted range", "Outside predicted range")
  ),

  Infrastructure_Scaling_Factor = c(
    sprintf("%.2f* (n=3 leases; pen fraction = %.1f%%)", mean_fpr, mean_pen_fraction * 100),
    "Not applied (isolated pontoons; detected area ≈ productive surface)",
    "N/A (land-based)",
    "N/A (land-based)",
    "N/A (longline area = productive surface)"
  ),

  Detection_Method = species_params$detection_method,

  stringsAsFactors = FALSE
)

cat("\n=== TABLE S6: PRODUCTION MODELING SUMMARY ===\n")
print(Table_S6)

output_path <- here("output_data/SSRG_GEE_results/Table_S6_Production_Modeling.csv")
write.csv(Table_S6, output_path, row.names = FALSE, fileEncoding = "UTF-8")
cat(sprintf("\nTable S6 saved to: %s\n", output_path))


# Write a plain-ASCII version to CSV
Table_S6_csv <- as.data.frame(lapply(Table_S6, function(x) {
  gsub("\u2013", "-", gsub("\u00b3", "^3", x))
}))
```

# ===== PRODUCTION COMPARISON PLOT =====

```{r production-comparison-plot, fig.width=12, fig.height=6}
library(ggplot2)

production_data <- data.frame(
  Species  = rep(c("Atlantic Salmon", "S. Bluefin Tuna", "Black Tiger Prawn",
                   "Barramundi", "Oysters"), each = 3),
  Scenario = rep(c("Minimum", "Actual", "Maximum"), 5),
  Production = c(
    round(prod_low, 0),          actual_production,       round(prod_high, 0),
    round(sbt_prod_low, 0),      sbt_actual_production,   round(sbt_prod_high, 0),
    round(prawn_prod_low, 0),    prawn_actual_production,  round(prawn_prod_high, 0),
    round(barra_prod_low, 0),    barra_actual_production,  round(barra_prod_high, 0),
    round(oyster_prod_low, 0),   oyster_actual_production, round(oyster_prod_high, 0)
  )
)

production_data$Species  <- factor(production_data$Species,
                                    levels = c("Atlantic Salmon", "S. Bluefin Tuna",
                                               "Black Tiger Prawn", "Barramundi", "Oysters"))
production_data$Scenario <- factor(production_data$Scenario,
                                    levels = c("Minimum", "Actual", "Maximum"))

production_plot <- ggplot(production_data, aes(x = Species, y = Production, fill = Scenario)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  scale_fill_manual(values = c("Minimum" = "#0072B2",   # Wong CB-safe blue
                               "Actual"  = "#E69F00",   # Wong CB-safe orange
                               "Maximum" = "#D55E00")) + # Wong CB-safe vermillion
  scale_y_continuous(labels = scales::comma) +
  labs(
    title    = "Production Range Estimates vs. Actual Production",
    subtitle = "Based on species-specific stocking density and yield parameters",
    x = NULL, y = "Production (tonnes)", fill = "Scenario"
  ) +
  theme_minimal() +
  theme(
    plot.title    = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray40"),
    axis.text.x   = element_text(size = 10),
    axis.text.y   = element_text(size = 9),
    axis.title.y  = element_text(size = 11, face = "bold"),
    legend.position = "top",
    legend.title    = element_text(size = 10, face = "bold"),
    panel.grid.minor = element_blank()
  )

print(production_plot)
ggsave(here("figures/Production_Range_Comparison.png"),
       plot = production_plot, width = 12, height = 6, dpi = 300)
cat("\nProduction comparison plot saved\n")
```

# ===== INDIVIDUAL SPECIES VALIDATION PLOT =====

```{r individual-species-validation-clean, fig.width=20, fig.height=4.5}
library(ggplot2)
library(scales)
library(cowplot)

# Build species data frames from extracted variables
salmon_data <- data.frame(Species = "Atlantic Salmon",
                          Minimum = round(prod_low, 0),
                          Maximum = round(prod_high, 0),
                          Actual  = actual_production)

sbt_data    <- data.frame(Species = "S. Bluefin Tuna",
                          Minimum = round(sbt_prod_low, 0),
                          Maximum = round(sbt_prod_high, 0),
                          Actual  = sbt_actual_production)

prawn_data  <- data.frame(Species = "Black Tiger Prawn",
                          Minimum = round(prawn_prod_low, 0),
                          Maximum = round(prawn_prod_high, 0),
                          Actual  = prawn_actual_production)

barra_data  <- data.frame(Species = "Barramundi",
                          Minimum = round(barra_prod_low, 0),
                          Maximum = round(barra_prod_high, 0),
                          Actual  = barra_actual_production)

oyster_data <- data.frame(Species = "Oysters",
                          Minimum = round(oyster_prod_low, 0),
                          Maximum = round(oyster_prod_high, 0),
                          Actual  = oyster_actual_production)

# Plot function
# actual_hjust: negative = right of point, positive = left; tweak per species if labels overlap
create_clean_species_plot <- function(data, color_actual = "#E69F00", actual_hjust = -0.3) {
  ggplot(data, aes(x = Species, y = Actual)) +
    geom_errorbar(aes(ymin = Minimum, ymax = Maximum),
                  width = 0.25, size = 2, color = "gray50", alpha = 0.8) +
    geom_point(aes(y = Minimum), color = "#0072B2", size = 5, shape = 95, stroke = 4) +
    geom_point(aes(y = Maximum), color = "#D55E00", size = 5, shape = 95, stroke = 4) +
    geom_point(color = color_actual, size = 10, shape = 19) +
    geom_point(color = "white", size = 4, shape = 19) +
    geom_text(aes(y = Minimum, label = comma(Minimum)),
              vjust = 1.5, hjust = 0.5, size = 4.5, fontface = "bold", color = "#0072B2") +
    geom_text(aes(y = Maximum, label = comma(Maximum)),
              vjust = -0.5, hjust = 0.5, size = 4.5, fontface = "bold", color = "#D55E00") +
    geom_text(aes(y = Actual,  label = comma(Actual)),
              hjust = actual_hjust, size = 5, fontface = "bold", color = color_actual) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0.15, 0.15))) +
    labs(title = data$Species, x = NULL, y = "Production (tonnes)") +
    theme_minimal() +
    theme(
      plot.title        = element_text(size = 13, face = "bold", hjust = 0.5),
      axis.text.x       = element_blank(),
      axis.text.y       = element_text(size = 10),
      axis.title.y      = element_text(size = 11, face = "bold"),
      panel.grid.major.x = element_blank(),
      panel.grid.minor  = element_blank(),
      panel.background  = element_rect(fill = "white", color = NA),
      plot.background   = element_rect(fill = "white", color = NA),
      plot.margin       = margin(15, 15, 10, 15)
    )
}

# Per-species hjust from parameter table (tweak there if needed)
hjust_vals <- setNames(species_params$actual_hjust, species_params$species)

p_salmon <- create_clean_species_plot(salmon_data, actual_hjust = hjust_vals["atl_salmon"])
p_sbt    <- create_clean_species_plot(sbt_data,    actual_hjust = hjust_vals["sbt"])
p_prawn  <- create_clean_species_plot(prawn_data,  actual_hjust = hjust_vals["prawn"])
p_barra  <- create_clean_species_plot(barra_data,  actual_hjust = hjust_vals["barramundi"])
p_oyster <- create_clean_species_plot(oyster_data, actual_hjust = hjust_vals["oyster"])

combined_plot <- plot_grid(p_salmon, p_sbt, p_prawn, p_barra, p_oyster, ncol = 5)

title_row    <- ggdraw() +
  draw_label("Production Validation: Actual vs. Modeled Range by Species",
             fontface = "bold", size = 14, x = 0, hjust = 0) +
  theme(plot.margin = margin(0, 0, 0, 7))

subtitle_row <- ggdraw() +
  draw_label("Vertical bars show modeled min-max range | Blue = Minimum | Red = Maximum | Gold = Actual ABARES production",
             size = 10, color = "gray40", x = 0, hjust = 0) +
  theme(plot.margin = margin(0, 0, 8, 7))

final_plot <- plot_grid(title_row, subtitle_row, combined_plot,
                        ncol = 1, rel_heights = c(0.08, 0.06, 1))

print(final_plot)
ggsave(here("figures/Production_Validation_Individual_Species.png"),
       plot = final_plot, width = 20, height = 4.5, dpi = 300, bg = "white")
cat("\n✓ Individual species validation plot saved\n")
```
#make a horizontal plot starting at zero for all species

```{r validation-horizontal, fig.width=13, fig.height=5}
library(ggplot2)
library(scales)
library(shadowtext)  

val_data <- data.frame(
  species = factor(
    c("Atlantic Salmon", "S. Bluefin Tuna", "Black Tiger Prawn",
      "Barramundi", "Oysters"),
    levels = rev(c("Atlantic Salmon", "S. Bluefin Tuna", "Black Tiger Prawn",
                   "Barramundi", "Oysters"))
  ),
  minimum = c(round(prod_low, 0), round(sbt_prod_low, 0),
              round(prawn_prod_low, 0), round(barra_prod_low, 0),
              round(oyster_prod_low, 0)),
  actual  = c(actual_production, sbt_actual_production,
              prawn_actual_production, barra_actual_production,
              oyster_actual_production),
  maximum = c(round(prod_high, 0), round(sbt_prod_high, 0),
              round(prawn_prod_high, 0), round(barra_prod_high, 0),
              round(oyster_prod_high, 0))
)

col_min    <- "#0072B2"
col_actual <- "#E69F00"
col_max    <- "#D55E00"
col_range  <- "grey75"
tick_h     <- 0.25

p_horiz <- ggplot(val_data, aes(y = species)) +
  
  # Zero-anchored range bar
  geom_segment(aes(x = 0, xend = maximum, yend = species),
               colour = col_range, linewidth = 2.5, lineend = "butt") +
  
  # Minimum tick
  geom_segment(aes(x = minimum, xend = minimum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_min, linewidth = 2, lineend = "round") +
  
  # Maximum tick
  geom_segment(aes(x = maximum, xend = maximum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_max, linewidth = 2, lineend = "round") +
  
 # Actual — shape no outline
  geom_point(aes(x = actual),
             size = 8, shape = 5, fill = col_actual,
             colour = col_actual, stroke = 0) +
  
  # Minimum label — further below tick to clear the marker
  geom_text(aes(x = minimum, label = comma(minimum)),
            vjust = 3.8, hjust = 0.5, size = 2.8,
            fontface = "bold", colour = col_min) +
  
  # Maximum label — above bar
  geom_text(aes(x = maximum, label = comma(maximum)),
            vjust = -2.1, hjust = 0.5, size = 2.8,
            fontface = "bold", colour = col_max) +
  
 # Actual label — above circle with black outline
  geom_shadowtext(aes(x = actual, label = comma(actual)),
                  vjust = 0.5, hjust = 0.5, size = 2.9,
                  fontface = "bold", colour = col_actual,
                  bg.colour = "grey20", bg.r = 0.18) +
  
  scale_x_continuous(labels = comma,
                     expand = expansion(mult = c(0.02, 0.08))) +
  scale_y_discrete(expand = expansion(add = c(0.6, 0.6))) +
  
  labs(
    title    = "Production Validation: Actual vs. Modelled Range by Species",
    subtitle = paste0(
      "\u2502 Min (blue)   \u25cf Actual ABARES (gold)   \u2502 Max (red)",
      "   \u2015\u2015 Modelled range (zero-anchored)"),
    x = "Production (tonnes)",
    y = NULL
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13, hjust = 0),
    plot.subtitle      = element_text(size = 9.5, colour = "grey40", hjust = 0,
                                      margin = margin(b = 12)),
    axis.text.y        = element_text(size = 11, face = "bold"),
    axis.text.x        = element_text(size = 9.5),
    axis.title.x       = element_text(size = 11, face = "bold",
                                      margin = margin(t = 8)),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey92"),
    plot.background    = element_rect(fill = "white", colour = NA),
    panel.background   = element_rect(fill = "white", colour = NA),
    plot.margin        = margin(15, 20, 15, 15)
  )

print(p_horiz)

ggsave(here("figures/Production_Validation_Horizontal.png"),
       plot = p_horiz, width = 13, height = 5, dpi = 300, bg = "white")
cat("\n✓ Horizontal validation plot saved\n")
```

#figure with lines for all values with 2D calculations

```{r validation-horizontal, fig.width=13, fig.height=5}
library(ggplot2)
library(scales)
library(shadowtext)  

val_data <- data.frame(
  species = c("Atl.Salmon", "S.B.Tuna", "Prawn",
              "Barramundi", "Oysters"),
  minimum = c(round(prod_low, 0), round(sbt_prod_low, 0),
              round(prawn_prod_low, 0), round(barra_prod_low, 0),
              round(oyster_prod_low, 0)),
  actual  = c(actual_production, sbt_actual_production,
              prawn_actual_production, barra_actual_production,
              oyster_actual_production),
  maximum = c(round(prod_high, 0), round(sbt_prod_high, 0),
              round(prawn_prod_high, 0), round(barra_prod_high, 0),
              round(oyster_prod_high, 0))
)

# Order by actual production: smallest at top, largest at bottom
val_data$species <- factor(val_data$species,
                           levels = val_data$species[order(val_data$actual, decreasing = TRUE)])

col_min    <- "#0072B2"
col_actual <- "#009E73"
col_max    <- "#D55E00"
col_range  <- "grey75"
tick_h     <- 0.25

p_horiz <- ggplot(val_data, aes(y = species)) +

  # Zero-anchored range bar
  geom_segment(aes(x = 0, xend = maximum, yend = species),
               colour = col_range, linewidth = 2.5, lineend = "butt") +

  # Minimum tick
  geom_segment(aes(x = minimum, xend = minimum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_min, linewidth = 2, lineend = "round") +

  # Maximum tick
  geom_segment(aes(x = maximum, xend = maximum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_max, linewidth = 2, lineend = "round") +

  # Actual tick
  geom_segment(aes(x = actual, xend = actual,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_actual, linewidth = 2, lineend = "round") +

  # Minimum label — BELOW the tick
  geom_text(aes(x = minimum, label = comma(minimum)),
          vjust = -0.2, hjust = 1, nudge_x = -400, size = 4,
          fontface = "bold", colour = col_min)+

  # Maximum label — RIGHT of tick, vertically centred
  geom_text(aes(x = maximum, label = comma(maximum)),
            vjust = 0.5, hjust = -0.15, size = 4,
            fontface = "bold", colour = col_max) +

  # Actual label — ABOVE the tick, shadow for legibility
  geom_shadowtext(aes(x = actual, label = comma(actual)),
                  vjust = -1.8, hjust = 0.5, size = 4,
                  fontface = "bold", colour = col_actual,
                  bg.colour = "white", bg.r = 0.15) +

  scale_x_continuous(labels = comma,
                     expand = expansion(mult = c(0.02, 0.12))) +  # extra right margin for max labels
  scale_y_discrete(expand = expansion(add = c(0.6, 0.6))) +

  labs(
    title    = "Production Validation: Actual vs. Modelled Range by Species",
    subtitle = paste0(
      "\u2502 Min (blue)   \u25cf Actual ABARES (gold)   \u2502 Max (red)",
      "   \u2015\u2015 Modelled range (zero-anchored)"),
    x = "Production (tonnes)",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13, hjust = 0),
    plot.subtitle      = element_text(size = 9.5, colour = "grey40", hjust = 0,
                                      margin = margin(b = 12)),
    axis.text.y        = element_text(size = 12, face = "bold"),
    axis.text.x        = element_text(size = 11),
    axis.title.x       = element_text(size = 12, face = "bold",
                                      margin = margin(t = 8)),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey92"),
    plot.background    = element_rect(fill = "white", colour = NA),
    panel.background   = element_rect(fill = "white", colour = NA),
    plot.margin        = margin(15, 25, 15, 15)
  )

print(p_horiz)

ggsave(here("figures/Production_Validation_Horizontalv2.png"),
       plot = p_horiz, width = 13, height = 5, dpi = 300, bg = "white")
cat("\n✓ Horizontal validation v2 plot saved\n")
```

#figure with lines for all values with 2D calculations for other species and 3D calculations for Salmon and Tuna
# ===== PRODUCTION VALIDATION PLOT: EMPIRICAL (pond/longline) + VOLUMETRIC 3D (finfish) =====

```{r validation-horizontal-2d-3d, fig.width=13, fig.height=5}
library(ggplot2)
library(scales)
library(shadowtext)

val_data <- data.frame(
  species = c("Atl.Salmon", "S.B.Tuna", "Prawn",
              "Barramundi", "Oysters"),
  minimum = c(
    round(cor_vol_prod_low, 0),     # salmon:  corrected volumetric (3D)
    round(sbt_vol_prod_low, 0),     # SBT:     volumetric (3D)
    round(prawn_prod_low, 0),       # prawn:   area-yield (2D)
    round(barra_prod_low, 0),       # barra:   area-yield (2D)
    round(oyster_prod_low, 0)       # oyster:  area-yield (2D)
  ),
  actual = c(
    actual_production, sbt_actual_production,
    prawn_actual_production, barra_actual_production,
    oyster_actual_production
  ),
  maximum = c(
    round(cor_vol_prod_high, 0),    # salmon:  corrected volumetric (3D)
    round(sbt_vol_prod_high, 0),    # SBT:     volumetric (3D)
    round(prawn_prod_high, 0),      # prawn:   area-yield (2D)
    round(barra_prod_high, 0),      # barra:   area-yield (2D)
    round(oyster_prod_high, 0)      # oyster:  area-yield (2D)
  ),
  approach = c("3D", "3D",
               "2D", "2D", "2D")
)

# Order by actual production: smallest at bottom, largest at top
val_data$species <- factor(val_data$species,
                           levels = val_data$species[order(val_data$actual)])

# Approach label for right-hand annotation (matched to factor order)
approach_labels <- val_data[order(val_data$actual), c("species", "approach")]

col_min    <- "#0072B2"
col_actual <- "#009E73"
col_max    <- "#D55E00"
col_range  <- "grey75"
tick_h     <- 0.25

p_horiz_3d <- ggplot(val_data, aes(y = species)) +

  # Zero-anchored range bar
  geom_segment(aes(x = 0, xend = maximum, yend = species),
               colour = col_range, linewidth = 2.5, lineend = "butt") +

  # Minimum tick
  geom_segment(aes(x = minimum, xend = minimum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_min, linewidth = 2, lineend = "round") +

  # Maximum tick
  geom_segment(aes(x = maximum, xend = maximum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_max, linewidth = 2, lineend = "round") +

  # Actual tick
  geom_segment(aes(x = actual, xend = actual,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_actual, linewidth = 2, lineend = "round") +

  # Minimum label — below tick
  geom_text(aes(x = minimum, label = comma(minimum)),
          vjust = -0.2, hjust = 1, nudge_x = -400, size = 4,
          fontface = "bold", colour = col_min)+

  # Maximum label — right of tick
  geom_text(aes(x = maximum, label = comma(maximum)),
            vjust = 0.5, hjust = -0.15, size = 4,
            fontface = "bold", colour = col_max) +

  # Actual label — above tick with shadow
  geom_shadowtext(aes(x = actual, label = comma(actual)),
                  vjust = -1.8, hjust = 0.5, size = 4,
                  fontface = "bold", colour = col_actual,
                  bg.colour = "white", bg.r = 0.15) +

  # Approach annotation on the left margin
 geom_text(data = approach_labels,
          aes(x = 0, y = species, label = approach),
          hjust = 1.08, vjust = 0.5, size = 4,
          colour = ifelse(approach_labels$approach == "3D",
                          "#555555", "#999999"),
          fontface = "bold.italic") +

  scale_x_continuous(labels = comma,
                   expand = expansion(mult = c(0.04, 0.08)))+  # left margin for labels
  scale_y_discrete(expand = expansion(add = c(0.6, 0.6))) +

  labs(
    title    = "Production Validation: Actual vs. Modelled Range by Species",
    subtitle = NULL,
    x = "Production (tonnes)",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13, hjust = 0),
    plot.subtitle      = element_text(size = 9, colour = "grey40", hjust = 0,
                                      margin = margin(b = 12)),
    axis.text.y        = element_text(size = 12, face = "bold"),
    axis.text.x        = element_text(size = 11),
    axis.title.x       = element_text(size = 12, face = "bold",
                                      margin = margin(t = 8)),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey92"),
    plot.background    = element_rect(fill = "white", colour = NA),
    panel.background   = element_rect(fill = "white", colour = NA),
    plot.margin        = margin(15, 25, 15, 15)
  )

print(p_horiz_3d)

ggsave(here("figures/Production_Validation_Horizontal2D_3D.png"),
       plot = p_horiz_3d, width = 13, height = 5, dpi = 300, bg = "white")
cat("\n✓ 2D/3D horizontal validation plot saved\n")
```

#notes:

#We used the corrected volumetric for salmon (not raw vol)because the raw volumetric massively overestimates because it treats the entire detected polygon as pen water surface. The corrected version (`cor_vol_prod_*`) already applies our measured pen fraction (11.1%).  

# the *raw* volumetric could be used to illustrate the overestimation, the easiest way would be to add a second, lighter grey bar behind the corrected one but I havent added that. 

#2D and 3D added
```{r}
library(ggplot2)
library(scales)
library(shadowtext)
library(dplyr)

# ── Main data (2D ranges for all species) ──────────────────────────────────────
val_data <- data.frame(
  species = c("Atl.Salmon", "S.B.Tuna", "Prawn",
              "Barramundi", "Oysters"),
  minimum = c(round(prod_low, 0), round(sbt_prod_low, 0),
              round(prawn_prod_low, 0), round(barra_prod_low, 0),
              round(oyster_prod_low, 0)),
  actual  = c(actual_production, sbt_actual_production,
              prawn_actual_production, barra_actual_production,
              oyster_actual_production),
  maximum = c(round(prod_high, 0), round(sbt_prod_high, 0),
              round(prawn_prod_high, 0), round(barra_prod_high, 0),
              round(oyster_prod_high, 0))
)

# Highest production at bottom (first factor level = bottom of y-axis)
val_data$species <- factor(val_data$species,
                           levels = val_data$species[order(val_data$actual,
                                                           decreasing = TRUE)])

# ── 3D corrected volumetric ranges (finfish only) ─────────────────────────────
val_data_3d <- data.frame(
  species     = factor(c("Atl.Salmon", "S.B.Tuna"),
                       levels = levels(val_data$species)),
  minimum_3d  = c(round(cor_vol_prod_low, 0),  round(sbt_vol_prod_low, 0)),
  maximum_3d  = c(round(cor_vol_prod_high, 0), round(sbt_vol_prod_high, 0))
)

col_min    <- "#0072B2"
col_actual <- "#009E73"
col_max    <- "#D55E00"
col_range  <- "grey75"
col_3d     <- "grey50"    # darker grey for 3D range bar
tick_h     <- 0.25

p_horiz <- ggplot(val_data, aes(y = species)) +

  # 3D corrected volumetric range bar (finfish only, behind 2D bar)
  geom_segment(data = val_data_3d,
               aes(x = 0, xend = maximum_3d, yend = species),
               colour = col_3d, linewidth = 1.2,
               lineend = "butt", linetype = "dashed") +

  # 3D minimum tick
  geom_segment(data = val_data_3d,
               aes(x = minimum_3d, xend = minimum_3d,
                   y = as.numeric(species) - tick_h * 0.7,
                   yend = as.numeric(species) + tick_h * 0.7),
               colour = col_min, linewidth = 1.2,
               lineend = "round", linetype = "dashed") +

  # 3D maximum tick
  geom_segment(data = val_data_3d,
               aes(x = maximum_3d, xend = maximum_3d,
                   y = as.numeric(species) - tick_h * 0.7,
                   yend = as.numeric(species) + tick_h * 0.7),
               colour = col_max, linewidth = 1.2,
               lineend = "round", linetype = "dashed") +

  # Zero-anchored 2D range bar (all species, on top)
  geom_segment(aes(x = 0, xend = maximum, yend = species),
               colour = col_range, linewidth = 2.5, lineend = "butt") +

  # 2D minimum tick
  geom_segment(aes(x = minimum, xend = minimum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_min, linewidth = 2, lineend = "round") +

  # 2D maximum tick
  geom_segment(aes(x = maximum, xend = maximum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_max, linewidth = 2, lineend = "round") +

  # Actual tick
  geom_segment(aes(x = actual, xend = actual,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_actual, linewidth = 2, lineend = "round") +

  # Minimum label — below tick
  geom_text(aes(x = minimum, label = comma(minimum)),
            vjust = -0.2, hjust = 1, nudge_x = -400, size = 4,
            fontface = "bold", colour = col_min) +

  # Maximum label — right of tick
  geom_text(aes(x = maximum, label = comma(maximum)),
            vjust = 0.5, hjust = -0.15, size = 4,
            fontface = "bold", colour = col_max) +

  # Actual label — above tick with shadow
  geom_shadowtext(aes(x = actual, label = comma(actual)),
                  vjust = -1.8, hjust = 0.5, size = 4,
                  fontface = "bold", colour = col_actual,
                  bg.colour = "white", bg.r = 0.15) +

  scale_x_continuous(labels = comma,
                     expand = expansion(mult = c(0.02, 0.12))) +
  scale_y_discrete(expand = expansion(add = c(0.6, 0.6))) +

  labs(
    title    = "Production Validation: Actual vs. Modelled Range by Species",
    subtitle = paste0(
      "\u2502 Min/Max (solid) = empirical 2D range   ",
      "\u2502 Min/Max (dashed) = corrected volumetric 3D range (finfish only)   ",
      "\u25cf Actual ABARES (green)"),
    x = "Production (tonnes)",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13, hjust = 0),
    plot.subtitle      = element_text(size = 9, colour = "grey40", hjust = 0,
                                      margin = margin(b = 12)),
    axis.text.y        = element_text(size = 12, face = "bold"),
    axis.text.x        = element_text(size = 11),
    axis.title.x       = element_text(size = 12, face = "bold",
                                      margin = margin(t = 8)),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey92"),
    plot.background    = element_rect(fill = "white", colour = NA),
    panel.background   = element_rect(fill = "white", colour = NA),
    plot.margin        = margin(15, 25, 15, 15)
  )

print(p_horiz)

ggsave(here("figures/Production_Validation_Horizontalv3.png"),
       plot = p_horiz, width = 13, height = 5, dpi = 300, bg = "white")
cat("\n✓ Horizontal validation v3 plot saved\n")
```
#only 3D salmon and SBT
```{r}
library(ggplot2)
library(scales)
library(shadowtext)

# ── 3D corrected volumetric data (finfish only) ───────────────────────────────
val_data_3d <- data.frame(
  species    = c("Atl.Salmon", "S.B.Tuna"),
  minimum    = c(round(cor_vol_prod_low, 0),  round(sbt_vol_prod_low, 0)),
  actual     = c(actual_production,            sbt_actual_production),
  maximum    = c(round(cor_vol_prod_high, 0), round(sbt_vol_prod_high, 0))
)

# Highest production at bottom
val_data_3d$species <- factor(val_data_3d$species,
                              levels = val_data_3d$species[order(val_data_3d$actual,
                                                                 decreasing = TRUE)])

col_min    <- "#0072B2"
col_actual <- "#009E73"
col_max    <- "#D55E00"
col_range  <- "grey75"
tick_h     <- 0.25

p_3d <- ggplot(val_data_3d, aes(y = species)) +

  # Zero-anchored range bar
  geom_segment(aes(x = 0, xend = maximum, yend = species),
               colour = col_range, linewidth = 2.5, lineend = "butt") +

  # Minimum tick
  geom_segment(aes(x = minimum, xend = minimum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_min, linewidth = 2, lineend = "round") +

  # Maximum tick
  geom_segment(aes(x = maximum, xend = maximum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_max, linewidth = 2, lineend = "round") +

  # Actual tick
  geom_segment(aes(x = actual, xend = actual,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_actual, linewidth = 2, lineend = "round") +

  # Minimum label
  geom_text(aes(x = minimum, label = comma(minimum)),
            vjust = -0.2, hjust = 1, nudge_x = -200, size = 4,
            fontface = "bold", colour = col_min) +

  # Maximum label
  geom_text(aes(x = maximum, label = comma(maximum)),
            vjust = 0.5, hjust = -0.15, size = 4,
            fontface = "bold", colour = col_max) +

  # Actual label
  geom_shadowtext(aes(x = actual, label = comma(actual)),
                  vjust = -1.8, hjust = 0.7, size = 4,
                  fontface = "bold", colour = col_actual,
                  bg.colour = "white", bg.r = 0.15) +

  scale_x_continuous(labels = comma,
                     expand = expansion(mult = c(0.08, 0.12))) +
  scale_y_discrete(expand = expansion(add = c(0.6, 0.6))) +

  labs(
    title    = NULL,  # title goes on the combined figure
    subtitle = paste0(
      "\u2502 Min (blue)   \u25cf Actual ABARES (green)   \u2502 Max (red)",
      "   \u2015\u2015 Corrected volumetric 3D range"),
    x = "Production (tonnes)",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.subtitle      = element_text(size = 9, colour = "grey40", hjust = 0,
                                      margin = margin(b = 12)),
    axis.text.y        = element_text(size = 12, face = "bold"),
    axis.text.x        = element_text(size = 11),
    axis.title.x       = element_text(size = 12, face = "bold",
                                      margin = margin(t = 8)),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey92"),
    plot.background    = element_rect(fill = "white", colour = NA),
    panel.background   = element_rect(fill = "white", colour = NA),
    plot.margin        = margin(15, 25, 15, 15)
  )

print(p_3d)

ggsave(here("figures/Production_Validation_3D_Finfish.png"),
       plot = p_3d, width = 13, height = 3, dpi = 300, bg = "white")
cat("\n✓ 3D finfish validation plot saved\n")
```
#7 rows
```{r}
library(ggplot2)
library(scales)
library(shadowtext)
library(dplyr)

# ── Data: 7 rows — 5 species 2D + 2 finfish 3D ───────────────────────────────
# Add spacer rows between species groups
spacers <- data.frame(
  species   = c("_spacer1", "_spacer2"),
  minimum   = c(NA, NA),
  actual    = c(NA, NA),
  maximum   = c(NA, NA),
  approach  = c("2D", "2D"),
  show_actual_label = c(FALSE, FALSE)
)

val_data_all <- data.frame(
  species  = c("Atl.Salmon (2D)",  "Atl.Salmon (3D)",
               "_spacer1",
               "Oysters (2D)",
               "Prawn (2D)",
               "_spacer2",
               "S.B.Tuna (2D)",    "S.B.Tuna (3D)",
               "Barramundi (2D)"),
  minimum  = c(round(prod_low, 0),          round(cor_vol_prod_low, 0),
               NA,
               round(oyster_prod_low, 0),
               round(prawn_prod_low, 0),
               NA,
               round(sbt_prod_low, 0),       round(sbt_vol_prod_low, 0),
               round(barra_prod_low, 0)),
  actual   = c(actual_production,            actual_production,
               NA,
               oyster_actual_production,
               prawn_actual_production,
               NA,
               sbt_actual_production,         sbt_actual_production,
               barra_actual_production),
  maximum  = c(round(prod_high, 0),          round(cor_vol_prod_high, 0),
               NA,
               round(oyster_prod_high, 0),
               round(prawn_prod_high, 0),
               NA,
               round(sbt_prod_high, 0),       round(sbt_vol_prod_high, 0),
               round(barra_prod_high, 0)),
  approach = c("2D", "3D",
               "2D",
               "2D",
               "2D",
               "2D",
               "2D", "3D",
               "2D"),
  show_actual_label = c(TRUE,  FALSE,
                        FALSE,
                        TRUE,
                        TRUE,
                        FALSE,
                        TRUE,  FALSE,
                        TRUE)
)

val_data_all$species <- factor(val_data_all$species,
                               levels = c("Atl.Salmon (2D)",
                                          "Atl.Salmon (3D)",
                                          "_spacer1",
                                          "Oysters (2D)",
                                          "Prawn (2D)",
                                          "_spacer2",
                                          "S.B.Tuna (2D)",
                                          "S.B.Tuna (3D)",
                                          "Barramundi (2D)"))

col_min    <- "#0072B2"
col_actual <- "#009E73"
col_max    <- "#D55E00"
col_2d     <- "grey75"
col_3d     <- "grey45"
tick_h     <- 0.25

p_all <- ggplot(val_data_all, aes(y = species)) +

  # Range bar — thickness and colour by approach
  geom_segment(aes(x = 0, xend = maximum, yend = species,
                   colour = approach, linewidth = approach),
               lineend = "butt") +

  scale_colour_manual(values = c("2D" = col_2d, "3D" = col_3d),
                      guide = "none") +
  scale_linewidth_manual(values = c("2D" = 2.5, "3D" = 1.5),
                         guide = "none") +

  # Minimum tick
  geom_segment(aes(x = minimum, xend = minimum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_min, linewidth = 2, lineend = "round") +

  # Maximum tick
  geom_segment(aes(x = maximum, xend = maximum,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_max, linewidth = 2, lineend = "round") +

  # Actual tick
  geom_segment(aes(x = actual, xend = actual,
                   y = as.numeric(species) - tick_h,
                   yend = as.numeric(species) + tick_h),
               colour = col_actual, linewidth = 2, lineend = "round") +

  # Minimum label — LEFT of tick
  geom_text(aes(x = minimum, label = comma(minimum)),
            vjust = 0.1, hjust = 1.15, size = 3.8,
            fontface = "bold", colour = col_min) +

  # Maximum label — RIGHT of tick
  geom_text(aes(x = maximum, label = comma(maximum)),
            vjust = 0.5, hjust = -0.15, size = 3.8,
            fontface = "bold", colour = col_max) +

  # Actual label — ABOVE tick
  geom_shadowtext(data = filter(val_data_all, show_actual_label),
                  aes(x = actual, label = comma(actual)),
                  vjust = -1.8, hjust = 0.5, size = 3.8,
                  fontface = "bold", colour = col_actual,
                  bg.colour = "white", bg.r = 0.15) +
  
 scale_x_continuous(labels = comma,
                     expand = expansion(mult = c(0.08, 0.12))) +
 scale_y_discrete(
    expand = expansion(add = c(0.6, 0.6)),
    labels = function(x) ifelse(startsWith(x, "_spacer"), "", x)
  ) +

  labs(
    title    = "Production Validation: Actual vs. Modelled Range by Species",
    subtitle = paste0(
      "\u2502 Min (blue)   \u25cf Actual ABARES (green)   \u2502 Max (red)   ",
      "\u2015\u2015 thick = 2D empirical   thin = 3D corrected volumetric (finfish only)"),
    x = "Production (tonnes)",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13, hjust = 0),
    plot.subtitle      = element_text(size = 9, colour = "grey40", hjust = 0,
                                      margin = margin(b = 12)),
    axis.text.y        = element_text(size = 11, face = "bold"),
    axis.text.x        = element_text(size = 11),
    axis.title.x       = element_text(size = 12, face = "bold",
                                      margin = margin(t = 8)),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey92"),
    plot.background    = element_rect(fill = "white", colour = NA),
    panel.background   = element_rect(fill = "white", colour = NA),
    plot.margin        = margin(15, 25, 15, 15)
  )

print(p_all)

ggsave(here("figures/Production_Validation_All_2D3D.png"),
       plot = p_all, width = 13, height = 6, dpi = 300, bg = "white")
cat("\n✓ Combined 2D/3D validation plot saved\n")
```

