#Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture


# ===== PRODUCTION DISTRIBUTION MAP =====
# Faceted map of production per site for the five key species.
# Source routing:
#   Marine (Salmon, SBT, Oysters) → infrastructure_with_species
#   Land   (Prawn, Barramundi)    → Australia_Land_Aquaculture_National.shp
#
# Production per site = (site_area / total_species_area) × national_actual_prod

```{r production-distribution-map, fig.width=16, fig.height=20}
library(sf)
library(ggplot2)
library(dplyr)
library(stringr)
library(scales)
library(here)

# ── 0. Australia basemap (consistent with other SSRG maps) ───────────────────
AUS_states <- st_read(here("raw_data/AustralianStates_SHP/STE_2021_AUST_GDA2020.shp"),
                      quiet = TRUE) %>%
  st_transform(4326) %>%   # reproject to match points layer
  st_make_valid()


# ── 1. Helper: distribute national production to sites by area ────────────────
distribute_production <- function(infra_sf,
                                   species_col,
                                   species_match,
                                   species_label,
                                   actual_prod_t,
                                   area_col = "infra_area_km2") {

  sites <- infra_sf %>%
    filter(str_detect(get(species_col), regex(species_match, ignore_case = TRUE)))

  if (nrow(sites) == 0) {
    warning(sprintf("No features matched '%s' for %s — check species_match string",
                    species_match, species_label))
    return(NULL)
  }

  # Empirical yield rate: national production / total detected infrastructure area
  # Identical to script 11: actual_yield_per_km2 = actual_prod / area_km2
  total_area        <- sum(st_drop_geometry(sites)[[area_col]], na.rm = TRUE)
  empirical_yield   <- actual_prod_t / total_area   # t/km²

  cat(sprintf("  %-20s empirical yield = %.1f t/km²  (%.2f t/ha)  across %.3f km² (%d sites)\n",
              species_label, empirical_yield, empirical_yield / 100,
              total_area, nrow(sites)))

  sites <- sites %>%
    mutate(
      species_label = species_label,
      site_area_km2 = .data[[area_col]],
      site_prod_t   = site_area_km2 * empirical_yield
    ) %>%
    select(species_label, site_area_km2, site_prod_t) %>%
    st_make_valid() %>%   # ← repair per-species subset before centroid
    st_buffer(0)          # ← zero-buffer as backstop

  suppressWarnings(st_centroid(sites))
}

# ── 2. Load and prepare the land-based shapefile ──────────────────────────────
land_sf <- st_read(
  here("output_data/Land_Aquaculture_Vectors/Australia_Land_Aquaculture_National.shp"),
  quiet = TRUE
) %>%
  st_transform(4326)%>%
st_make_valid()

# Load marine infrastructure — runs independently of script 4
 infrastructure_with_species <- st_read(
  here("output_data/infrastructure_by_species/infrastructure_with_species.gpkg"),
  quiet = TRUE
)
cat("✓ infrastructure_with_species loaded:", nrow(infrastructure_with_species), "features\n")

marine_sf <- infrastructure_with_species %>%
  st_transform(4326) %>%
  st_make_valid() %>%
  st_buffer(0)

# Column names confirmed from your shapefile:
#   Taxon    = species group/scientific name
#   Spcs_cs  = common name
#   In_A_KM  = infrastructure area in km²
species_col_land <- "Taxon"    # switch to "Spcs_cs" if Taxon values don't match below
area_col_land    <- "In_A_KM"

# Diagnostic — shows you exactly what strings to use in species_match
cat("Unique Taxon values in land shapefile:\n")
print(sort(unique(land_sf$Taxon)))
cat("\nUnique Spcs_cs values in land shapefile:\n")
print(sort(unique(land_sf$Spcs_cs)))

# ── 3. Prepare the marine infrastructure layer ────────────────────────────────
marine_sf <- infrastructure_with_species %>%
  st_transform(4326)

# Confirm species column in marine layer
species_col_marine <- names(marine_sf)[
  str_detect(names(marine_sf), regex("species|taxon|common|name", ignore_case = TRUE))
][1]

cat("\nMarine layer — using species column:", species_col_marine, "\n")
cat("Unique values:\n")
print(sort(unique(marine_sf[[species_col_marine]])))

# Ensure area column exists in marine layer
if (!"infra_area_km2" %in% names(marine_sf)) {
  marine_sf <- marine_sf %>%
    mutate(infra_area_km2 = as.numeric(st_area(geometry)) / 1e6)
  cat("Note: computed infra_area_km2 from marine geometry\n")
}

# ── 4. Distribute production per site for each species ────────────────────────
# After running section 2-3, check the printed unique values above and update
# species_match strings here if needed.
# ── 4a. Production totals — pulled from species_params (must be loaded from script 11)
# ── 4a. ABARES 2021-22 production totals (hardcoded)
actual_production        <- 79271
sbt_actual_production    <- 8322
prawn_actual_production  <- 9450
barra_actual_production  <- 4742
oyster_actual_production <- 11735

# ── 4b. Use Spcs_cs for land (Taxon is too broad — only Fish/Crustacean etc.)
species_col_land <- "Spcs_cs"

# ── 4. Distribute production per site ─────────────────────────────────────────

# MARINE
sites_salmon <- distribute_production(
  infra_sf      = marine_sf,
  species_col   = species_col_marine,
  species_match = "Atlantic Salmon|atl_salmon",
  species_label = "Atlantic Salmon",
  actual_prod_t = actual_production
)

sites_sbt <- distribute_production(
  infra_sf      = marine_sf,
  species_col   = species_col_marine,
  species_match = "Southern Bluefin|bluefin|sbt",
  species_label = "S. Bluefin Tuna",
  actual_prod_t = sbt_actual_production
)

sites_oyster <- distribute_production(
  infra_sf      = marine_sf %>%
    filter(!str_detect(get(species_col_marine),
                       regex("pearl", ignore_case = TRUE))),
  species_col   = species_col_marine,
  species_match = "oyster",          # catches "Edible Oyster (aggregated)", excludes Pearl
  species_label = "Oysters",
  actual_prod_t = oyster_actual_production
)

# LAND — Spcs_cs values confirmed:
#   Prawn:      "Prawn", "Multiple (Barramundi, Prawn)", "Multiple (Red Claw Crayfish, Prawn)"
#   Barramundi: "Barramundi", "Multiple (Barramundi, Prawn)", "Multiple (Barramundi, Red Claw Crayfish)"
sites_prawn <- distribute_production(
  infra_sf      = land_sf,
  species_col   = species_col_land,
  species_match = "prawn",           # catches Prawn + all Multi entries containing Prawn
  species_label = "Black Tiger Prawn",
  actual_prod_t = prawn_actual_production,
  area_col      = area_col_land
)

sites_barra <- distribute_production(
  infra_sf      = land_sf,
  species_col   = species_col_land,
  species_match = "barramundi",      # catches Barramundi + all Multi entries containing Barramundi
  species_label = "Barramundi",
  actual_prod_t = barra_actual_production,
  area_col      = area_col_land
)

# ── 5. Combine all species ────────────────────────────────────────────────────

# Normalise geometry column name before rbind (marine and land layers may differ)
normalise_geom <- function(x) {
  if (is.null(x)) return(NULL)
  st_geometry(x) <- "geometry"   # force consistent name
  x
}

all_sites <- do.call(rbind, Filter(Negate(is.null), lapply(
  list(sites_salmon, sites_sbt, sites_oyster, sites_prawn, sites_barra),
  normalise_geom
)))

all_sites$species_label <- factor(
  all_sites$species_label,
  levels = c("Atlantic Salmon", "S. Bluefin Tuna",
             "Black Tiger Prawn", "Barramundi", "Oysters")
)

st_write(all_sites,
         here("output_data/Production_by_Detection.gpkg"),
         delete_dsn = TRUE)

# Extract coordinates for ggplot
all_sites_coords <- all_sites %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  st_drop_geometry()

# Sanity check — totals should equal ABARES figures
cat("\nProduction summary by species (should match ABARES totals):\n")
all_sites_coords %>%
  group_by(species_label) %>%
  summarise(n_sites = n(), total_prod_t = round(sum(site_prod_t)), .groups = "drop") %>%
  print()
```

#acronyms for states for mapping
```{r}
state_labels <- AUS_states %>%
  st_make_valid() %>%
  filter(!STE_NAME21 %in% c("Other Territories", "Outside Australia",
                             "Australian Capital Territory")) %>%
  st_centroid() %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2],
    acronym = case_when(
      STE_NAME21 == "New South Wales"    ~ "NSW",
      STE_NAME21 == "Victoria"           ~ "VIC",
      STE_NAME21 == "Queensland"         ~ "QLD",
      STE_NAME21 == "South Australia"    ~ "SA",
      STE_NAME21 == "Western Australia"  ~ "WA",
      STE_NAME21 == "Tasmania"           ~ "TAS",
      STE_NAME21 == "Northern Territory" ~ "NT",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(acronym)) %>%
  st_drop_geometry()
```

#mapping aquaculture production
# production totals = ABARES 2021.  Circle area  site production (tonnes).\n",      "Method: site_prod = site_area  empirical yield rate (t/km)

```{r production-distribution-map, fig.width=16, fig.height=20}
# ── 6. Species colour palette ─────────────────────────────────────────────────
species_colours <- c(
  "Atlantic Salmon"   = "#D55E00",  # vermillion 
  "S. Bluefin Tuna"   = "#0072B2",    # blue 
  "Black Tiger Prawn" = "#CC79A7",  #golden "#E69F00"
  "Barramundi"        = "#009E73",   # bluish green
  "Oysters"           = "#5C0099"    #  purple
)

# ── 7. Faceted production bubble map ─────────────────────────────────────────

prod_map <- ggplot() +
  geom_sf(data = AUS_states, fill = "grey90", colour = "grey60", linewidth = 0.3) +
  geom_text(
  data = state_labels,
  aes(x = lon, y = lat, label = acronym),
  size     = 3,
  colour   = "grey50",
  fontface = "bold"
) +
  geom_point(
    data  = all_sites_coords,
    aes(x = lon, y = lat, size = site_prod_t, colour = species_label),
    alpha = 0.72, shape = 16
  ) +
  scale_size_area(
    name     = "Estimated production\nat site (tonnes)",
    max_size = 10,
    breaks   = c(10, 100, 500, 2000, 5000, 20000),
    labels   = scales::comma
  ) +
  scale_colour_manual(values = species_colours, guide = "none") +
  facet_wrap(~ species_label, ncol = 2) +
  coord_sf(xlim = c(113, 155), ylim = c(-44, -10), expand = FALSE) +
  labs(x = NULL, y = NULL) +
   
  theme_void(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, hjust = 0),
    plot.subtitle    = element_text(size = 9.5, colour = "grey35", hjust = 0,
                                    margin = margin(b = 6)),
    plot.caption     = element_text(size = 8, colour = "grey50", hjust = 0),
    strip.text       = element_text(face = "bold", size = 11, colour = "white"),
    strip.background = element_rect(fill = "grey25", colour = NA),
    legend.position  = "bottom",
    legend.title     = element_text(size = 9, face = "bold"),
    legend.text      = element_text(size = 8),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    plot.margin      = margin(10, 10, 10, 10)
  ) +
  guides(
    size = guide_legend(
      title.position = "top",
      override.aes   = list(colour = "grey30", alpha = 0.8)
    )
  )

print(prod_map)

ggsave(
  here("figures/Production_Distribution_Map_Faceted.png"),
  plot = prod_map, width = 16, height = 20, dpi = 300,
  bg = "transparent"
)


cat("\n✓ Faceted production map saved to figures/Production_Distribution_Map_Faceted.png\n")
```


# ===== OPTIONAL: COMBINED SINGLE-PANEL MAP =====

```{r production-map-combined, fig.width=14, fig.height=9}

combined_map <- ggplot() +
  geom_sf(data = AUS_states, fill = "grey90", colour = "grey60", linewidth = 0.3) +
  geom_text(
  data = state_labels,
  aes(x = lon, y = lat, label = acronym),
  size     = 3,
  colour   = "grey50",
  fontface = "bold"
) +
  geom_point(
    data  = all_sites_coords,
    aes(x = lon, y = lat, size = site_prod_t,
        colour = species_label, fill = species_label),
    alpha = 0.72, shape = 21, stroke = 0.4
  ) +
  scale_size_area(
    name     = "Estimated\nproduction (t)",
    max_size = 10,
    breaks   = c(10, 100, 500, 2000, 10000, 30000),
    labels   = scales::comma
  ) +
  scale_colour_manual(name = "Species", values = species_colours) +
  scale_fill_manual(  name = "Species", values = scales::alpha(species_colours, 0.7)) +
  coord_sf(xlim = c(113, 155), ylim = c(-44, -10), expand = FALSE) +
  labs(x = NULL, y = NULL) +
  
  theme_void(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle    = element_text(size = 10, colour = "grey40", hjust = 0),
    plot.caption     = element_text(size = 8, colour = "grey50", hjust = 0),
    legend.position  = "right",
    legend.title     = element_text(size = 9, face = "bold"),
    legend.text      = element_text(size = 8),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    plot.margin      = margin(10, 10, 10, 10)
  ) +
  guides(
    colour = guide_legend(override.aes = list(size = 5, alpha = 0.9), order = 1),
    fill   = guide_legend(override.aes = list(size = 5, alpha = 0.9), order = 1),
    size   = guide_legend(
      title.position = "top",
      override.aes   = list(colour = "grey30", fill = "grey60"),
      order = 2
    )
  )

print(combined_map)

ggsave(
  here("figures/Production_Distribution_Map_Combinedv2.png"),
  plot = combined_map, width = 14, height = 9, dpi = 300, bg = "white"
)
cat("✓ Combined production map saved\n")
```
#colorblind safe map of all species for supplementary materials with no production distribution quantities, just locations hotspots

```{r}
# ===== SUPPLEMENTARY FIGURE: All-Species Aquaculture Site Distribution =====

# ── 0. Packages ───────────────────────────────────────────────────────────────
library(sf)
library(ggplot2)
library(dplyr)
library(stringr)
library(here)

# ── 1. Standardise marine species labels ──────────────────────────────────────
marine_supp <- infrastructure_with_species %>%
  st_transform(4326) %>%
  st_make_valid() %>%
  st_buffer(0) %>%
  rename(raw_species = all_of(species_col_marine)) %>%
  mutate(
    species_label = case_when(
      str_detect(raw_species, regex("atlantic salmon|atl.salmon",      ignore_case = TRUE)) ~ "Atlantic Salmon",
      str_detect(raw_species, regex("southern bluefin|bluefin|\\bsbt\\b", ignore_case = TRUE)) ~ "S. Bluefin Tuna",
      str_detect(raw_species, regex("pearl oyster",                    ignore_case = TRUE)) ~ "Pearl Oyster",
      str_detect(raw_species, regex("oyster",                          ignore_case = TRUE)) ~ "Edible Oyster",
      str_detect(raw_species, regex("mussel",                          ignore_case = TRUE)) ~ "Mussel",
      str_detect(raw_species, regex("abalone",                         ignore_case = TRUE)) ~ "Abalone",
      str_detect(raw_species, regex("prawn|shrimp",                    ignore_case = TRUE)) ~ "Prawn",
      str_detect(raw_species, regex("barramundi",                      ignore_case = TRUE)) ~ "Barramundi",
      TRUE ~ str_to_title(str_trim(raw_species))
    ),
    source = "Marine"
  ) %>%
  select(species_label, source)

# ── 2. Standardise land species labels ────────────────────────────────────────
expand_land_species <- function(sf_obj, species_col = "Spcs_cs") {
  sf_obj %>%
    mutate(row_id = row_number(),
           is_multi = str_detect(.data[[species_col]], regex("multiple", ignore_case = TRUE)),
           tokens   = if_else(is_multi,
                              str_extract_all(.data[[species_col]], "(?<=\\().*(?=\\))"),
                              list(NULL))) %>%
    {
      single <- filter(., !is_multi) %>%
        mutate(Spcs_clean = .data[[species_col]]) %>%
        select(row_id, Spcs_clean)
      multi <- filter(., is_multi) %>%
        tidyr::unnest(tokens) %>%
        tidyr::separate_rows(tokens, sep = ",\\s*") %>%
        mutate(Spcs_clean = str_trim(tokens)) %>%
        select(row_id, Spcs_clean)
      bind_rows(single, multi)
    } %>%
    left_join(st_drop_geometry(sf_obj) %>% mutate(row_id = row_number()), by = "row_id") %>%
    st_as_sf(crs = st_crs(sf_obj))
}

land_supp <- land_sf %>%
  st_transform(4326) %>%
  expand_land_species(species_col = "Spcs_cs") %>%
  mutate(
    species_label = case_when(
      str_detect(Spcs_clean, regex("atlantic salmon",                        ignore_case = TRUE)) ~ "Atlantic Salmon",
     str_detect(Spcs_clean, regex("atlantic salmon|\\bsalmon\\b|atl[._]salmon|salmonid", 
                              ignore_case = TRUE)) ~ "Atlantic Salmon",
      str_detect(Spcs_clean, regex("barramundi",                             ignore_case = TRUE)) ~ "Barramundi",
      str_detect(Spcs_clean, regex("prawn",                                  ignore_case = TRUE)) ~ "Prawn",
      str_detect(Spcs_clean, regex("abalone",                                ignore_case = TRUE)) ~ "Abalone",
      str_detect(Spcs_clean, regex("mussel",                                 ignore_case = TRUE)) ~ "Mussel",
      str_detect(Spcs_clean, regex("cherax|red claw|crayfish|yabby|marron",  ignore_case = TRUE)) ~ "Cherax / Red Claw",
      str_detect(Spcs_clean, regex("trout",                                  ignore_case = TRUE)) ~ "Trout",
      str_detect(Spcs_clean, regex("tuna",                                   ignore_case = TRUE)) ~ "S. Bluefin Tuna",
      str_detect(Spcs_clean, regex("murray cod|silver perch|jade perch",     ignore_case = TRUE)) ~ "Native Freshwater Fish",
      str_detect(Spcs_clean, regex("native finfish|mangrove jack|sea mullet|cobia|grouper", ignore_case = TRUE)) ~ "Native Finfish",
      TRUE ~ str_to_title(str_trim(Spcs_clean))
    ),
    source = "Land"
  ) %>%
  select(species_label, source)

# ── 3. Merge, centroid, filter ────────────────────────────────────────────────
normalise_geom <- function(x) { st_geometry(x) <- "geometry"; x }

all_supp_centroids <- bind_rows(
  normalise_geom(marine_supp),
  normalise_geom(land_supp)
) %>%
  st_make_valid() %>%
  st_buffer(0) %>%
  st_cast("MULTIPOLYGON") %>%
  st_cast("POLYGON")

# Centroid outside the pipe to avoid the recursive evaluation error
all_supp_centroids <- suppressWarnings(st_centroid(all_supp_centroids)) %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(lon), !is.na(lat))

# Exclude non-species categories
exclude_labels <- c("Hatchery", "Aquarium", "Tourism", "Subtidal",
                    "Production Lease", "Miscellaneous", "Finfish",
                    "Lobster", "Pearl Oyster", "Snapper", "Aquarium")

all_supp_centroids <- all_supp_centroids %>%
  filter(!species_label %in% exclude_labels,
         !is.na(species_label))

cat("Species present:\n")
all_supp_centroids %>% count(species_label, sort = TRUE) %>% print(n = 30)

# ── 4. Hardcoded palette ──────────────────────────────────────────────────────
# Anchor species = exact production map colours
# Secondary species = tonal relatives within same hue family

species_pal <- c(
  # Vermillion — Salmon anchor
  "Atlantic Salmon"        = "#D55E00",

  # Blue — Tuna anchor
  "S. Bluefin Tuna"        = "#0072B2",

  # Purple — Oyster anchor, Mussel lighter
  "Edible Oyster"          = "#5C0099",
  "Mussel"                 = "#9B4DD4",

  # Golden — Prawn anchor, Cherax darker amber
  "Prawn"                  = "#E69F00",
  "Cherax / Red Claw"      = "#CC6677",

  # Teal — Barramundi anchor, other land finfish lighter
  "Barramundi"             = "#009E73",
  "Native Finfish"         = "#4DC4A3",
  "Native Freshwater Fish" = "#A8E6D5",
  "Trout"                  = "#005240",

  # Magenta — Abalone
  "Abalone"                = "#AE017E"
)

# ── 5. Build all_supp_facet ───────────────────────────────────────────────────
all_supp_facet <- all_supp_centroids %>%
  filter(species_label %in% names(species_pal)) %>%
  mutate(species_label = factor(species_label, levels = names(species_pal)))

n_species_facet <- nlevels(droplevels(all_supp_facet$species_label))
n_cols <- 3
n_rows <- ceiling(n_species_facet / n_cols)
fig_h  <- max(14, n_rows * 5)

cat("Species in faceted map:", n_species_facet, "\n")
print(levels(droplevels(all_supp_facet$species_label)))

# ── 6. Faceted presence map ───────────────────────────────────────────────────
supp_map <- ggplot() +
  geom_sf(data = AUS_states, fill = "grey93", colour = "grey60", linewidth = 0.25) +
  geom_text(data = state_labels, aes(x = lon, y = lat, label = acronym),
            size = 1.5, colour = "grey55", fontface = "bold") +
  geom_point(data = all_supp_facet, aes(x = lon, y = lat, colour = species_label),
             size = 1.0, alpha = 0.75, shape = 16) +
  scale_colour_manual(values = species_pal, guide = "none") +
  facet_wrap(~ species_label, ncol = n_cols) +
  coord_sf(xlim = c(113, 155), ylim = c(-44, -10), expand = FALSE) +
  labs(x = NULL, y = NULL) +
  theme_void(base_size = 11) +
  theme(
    strip.text       = element_text(face = "bold", size = 9.5, colour = "white"),
    strip.background = element_rect(fill = "grey25", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "white", colour = NA),
    plot.margin      = margin(10, 12, 10, 12)
  )

print(supp_map)

ggsave(
  here("figures/Supplementary_AllSpecies_Distribution_Map.png"),
  plot = supp_map, width = 14, height = fig_h, dpi = 300, bg = "white"
)
cat(sprintf("\n✓ Faceted supplementary map saved — %d species, %d × %d grid\n",
            n_species_facet, n_rows, n_cols))

# ── 7. Combined single-panel map ──────────────────────────────────────────────
# Subset palette to only species actually present in all_supp_facet
facet_pal <- species_pal[levels(droplevels(all_supp_facet$species_label))]

cat("Colours assigned:\n")
print(facet_pal)

supp_combined <- ggplot() +
  geom_sf(data = AUS_states, fill = "grey93", colour = "grey60", linewidth = 0.15) +
  geom_text(data = state_labels, aes(x = lon, y = lat, label = acronym),
            size = 3, colour = "grey55", fontface = "bold") +
  geom_point(data = all_supp_facet, aes(x = lon, y = lat, colour = species_label),
             size = 1.0, alpha = 0.80, shape = 16) +
  scale_colour_manual(name = "Species", values = facet_pal) +
  coord_sf(xlim = c(113, 155), ylim = c(-44, -10), expand = FALSE) +
  labs(x = NULL, y = NULL) +
  theme_void(base_size = 11) +
  theme(
    legend.position = "right",
    legend.title    = element_text(size = 9, face = "bold"),
    legend.text     = element_text(size = 7.5),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin     = margin(10, 10, 10, 10)
  ) +
  guides(colour = guide_legend(ncol = 1,
                               override.aes = list(size = 4, alpha = 0.9, shape = 16)))

print(supp_combined)

ggsave(
  here("figures/Supplementary_AllSpecies_Distribution_Map_Combinedv2.png"),
  plot = supp_combined, width = 14, height = 9, dpi = 300, bg = "transparent"
)
cat("✓ Combined supplementary map saved\n")
```

