---
title: "13_v2_Build_clean_combined_dataset_and_summaries"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

date: "2026-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# About

I Made this script in order to replace the original Scripts 11 and 13. The aim i sto build a single, internally consistent
set of outputs for downstream reporting and the Shiny app.

**Reads:**

- Land (from Script 7):
  `output_data/Land_Aquaculture_Vectors/Australia_Land_Aquaculture_National.shp`
  (+ CSV for untruncated column names)
- Marine leases (from Script 4):
  `output_data/all_marine_leases_with_status.shp` — the **single canonical**
  marine lease source. Pearl + NT already excluded upstream.
- Marine detections (from Script 4):
  `output_data/SSRG_GEE_results/SSRG_shellfish_longlines_marine.shp`
  `output_data/SSRG_GEE_results/SSRG_finfish_cages_marine.shp`

**Writes:**

- `output_data/summaries/AU_aquaculture_clean.gpkg`
   - layer `leases`     — one row per lease (land + marine)
   - layer `detections` — one row per marine detection, with `lease_id` FK
- `output_data/summaries/combined_aquaculture_species_summaries.xlsx`
   - All summary sheets regenerated from the same cleaned data.

# 0. Packages

```{r packages}
library(sf)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(purrr)
library(here)
library(writexl)
```

# 1. Standardisation helpers

```{r helpers}
# --- State -------------------------------------------------------------------
state_to_abbr <- c(
  "new south wales" = "NSW", "queensland" = "QLD",
  "south australia" = "SA",  "tasmania" = "TAS",
  "victoria" = "VIC",        "western australia" = "WA",
  "northern territory" = "NT", "australian capital territory" = "ACT",
  "nsw"="NSW","qld"="QLD","sa"="SA","tas"="TAS",
  "vic"="VIC","wa"="WA","nt"="NT","act"="ACT"
)
abbr_to_full <- c(
  NSW = "New South Wales", QLD = "Queensland",
  SA  = "South Australia", TAS = "Tasmania",
  VIC = "Victoria",        WA  = "Western Australia",
  NT  = "Northern Territory", ACT = "Australian Capital Territory"
)
clean_state <- function(x) {
  k <- tolower(trimws(gsub("_", " ", as.character(x))))
  out <- unname(state_to_abbr[k])
  out[is.na(out) | k %in% c("", "na", "unknown")] <- "Unknown"
  out
}

# --- Realm / status ---------------------------------------------------------
clean_status <- function(x) {
  k <- tolower(trimws(as.character(x)))
  case_when(
    k %in% c("active", "present") ~ "Active",
    k %in% c("inactive", "absent") ~ "Inactive",
    TRUE ~ NA_character_
  )
}

# --- Infrastructure type ----------------------------------------------------
clean_infra_type <- function(x) {
  k <- tolower(trimws(as.character(x)))
  case_when(
    k %in% c("pond", "ponds")                          ~ "Pond",
    k %in% c("cage", "cages", "pen", "pens",
             "finfish (cages)")                         ~ "Cage",
    k %in% c("longline", "longlines", "long-line",
             "shellfish (longlines)")                   ~ "Longline",
    k %in% c("tank","tanks","ras","tanks/ras",
             "tank-ras","tank/ras")                     ~ "Tank-RAS",
    is.na(k) | k == "" | k == "na"                     ~ NA_character_,
    TRUE                                                ~ str_to_title(k)
  )
}

# --- Taxonomic group: Fish / Mollusc / Crustacean / Algae / Unknown ---------
# Bivalves are collapsed into Mollusc (bivalves ARE molluscs). The finer
# Bivalve / Gastropod split lives in the subgroup column.
clean_group <- function(x) {
  k <- tolower(trimws(as.character(x)))
  case_when(
    k %in% c("bivalve","bivalves","mollusc","molluscs") ~ "Mollusc",
    k %in% c("crustacean","crustaceans")                ~ "Crustacean",
    k %in% c("fish","finfish","fishes")                 ~ "Fish",
    k %in% c("echinoderm","echinoderms")                ~ "Echinoderm",
    k %in% c("algae","seaweed","seaweeds")              ~ "Algae",
    is.na(k) | k == "" | k == "na"                      ~ "Unknown",
    TRUE                                                ~ str_to_title(k)
  )
}

# --- Subgroup: only for Molluscs --------------------------------------------
# Bivalve: oysters, mussels, scallops, clams, pipis, cockles
# Gastropod: abalone
compute_subgroup <- function(group_clean, species) {
  sp <- tolower(as.character(species))
  case_when(
    group_clean == "Mollusc" & grepl("abalone", sp) ~ "Gastropod",
    group_clean == "Mollusc" &
      grepl("oyster|mussel|scallop|clam|pipi|cockle", sp) ~ "Bivalve",
    TRUE ~ NA_character_
  )
}

clean_species <- function(x) {
  k <- trimws(as.character(x))
  ifelse(is.na(k) | k == "" | tolower(k) == "unknown", "Unknown", k)
}

# --- Helper: compute area in km^2 using EPSG:3577 (Australian Albers) -------
area_km2 <- function(sf_obj, crs = 3577) {
  suppressWarnings(
    as.numeric(sf::st_area(sf::st_transform(sf_obj, crs)))
  ) / 1e6
}
```

# 2. Load land data

```{r load-land}
land_shp_path <- here("output_data", "Land_Aquaculture_Vectors",
                      "Australia_Land_Aquaculture_National.shp")
land_csv_path <- here("output_data", "Land_Aquaculture_Vectors",
                      "Australia_Land_Aquaculture_National.csv")

land_sf <- st_read(land_shp_path, quiet = TRUE) %>%
  st_transform(4326) %>%
  st_make_valid()

# CSV has untruncated column names (e.g. Species_custodian instead of Spcs_cs).
# Prefer CSV-sourced species where available.
if (file.exists(land_csv_path)) {
  land_csv <- read_csv(land_csv_path, show_col_types = FALSE)

  # Join on SiteID (which is unique per site)
  # Pick a minimal set of full-name columns we want to preserve.
  land_attrs_full <- land_csv %>%
    select(SiteID,
           any_of(c("Species_custodian", "Species_verified",
                    "Status_custodian", "Taxon")))

  land_sf <- land_sf %>%
    left_join(land_attrs_full, by = "SiteID", suffix = c("_shp", ""))
}

message("Land features: ", nrow(land_sf))
message("Land columns:  ", paste(names(land_sf), collapse = ", "))
```

# 3. Build standardised `land_leases`

```{r land-leases}
# Column helpers that cope with truncated shapefile names
pick_col <- function(df, candidates) {
  hit <- candidates[candidates %in% names(df)]
  if (length(hit) > 0) hit[1] else NA_character_
}

col_land_species <- pick_col(land_sf, c("Species_custodian", "Spcs_cs"))
col_land_taxon   <- pick_col(land_sf, c("Taxon"))
col_land_status  <- pick_col(land_sf, c("Infr_status", "Infr_st"))
col_land_infra   <- pick_col(land_sf, c("Infr_type", "Infr_ty"))
col_land_state   <- pick_col(land_sf, c("State"))
col_land_ls_km   <- pick_col(land_sf, c("Lease_Area_KM", "Ls_A_KM"))
col_land_in_km   <- pick_col(land_sf, c("Infra_Area_KM", "In_A_KM"))
col_land_siteid  <- pick_col(land_sf, c("SiteID"))

# Build a normalised tibble, recompute lease area from geometry,
# preserve infra area (which for land comes from pond geometry upstream).
land_leases <- land_sf %>%
  transmute(
    source_site_id = as.character(.data[[col_land_siteid]]),
    realm          = "Land-based",
    state          = clean_state(.data[[col_land_state]]),
    status         = clean_status(.data[[col_land_status]]),
    species_raw    = .data[[col_land_species]],
    group_raw      = .data[[col_land_taxon]],
    infra_type_raw = .data[[col_land_infra]],
    infra_area_km2 = suppressWarnings(as.numeric(.data[[col_land_in_km]])),
    geometry       = geometry
  ) %>%
  mutate(
    species        = clean_species(species_raw),
    group          = clean_group(group_raw),
    subgroup       = compute_subgroup(group, species),
    infra_type     = clean_infra_type(infra_type_raw),
    # For inactive land leases set infra_type = "None"; otherwise default
    # any remaining NAs to "Pond" (the dominant land type) or "Unknown"
    infra_type = case_when(
      status == "Inactive"            ~ "None",
      is.na(infra_type)               ~ "Pond",
      TRUE                            ~ infra_type
    ),
    state_full     = unname(abbr_to_full[state])
  )

# Recompute lease area from geometry (EPSG:3577)
land_leases$lease_area_km2 <- area_km2(land_leases)

# Assign lease_id
land_leases <- land_leases %>%
  mutate(lease_id = paste0("LL_", sprintf("%04d", row_number()))) %>%
  select(lease_id, source_site_id, realm, state, state_full,
         group, subgroup, species, status, infra_type,
         lease_area_km2, infra_area_km2, geometry)

message("Land leases assembled: ", nrow(land_leases))
message("  active:   ", sum(land_leases$status == "Active",   na.rm = TRUE))
message("  inactive: ", sum(land_leases$status == "Inactive", na.rm = TRUE))
message(sprintf("  total lease area: %.2f km^2 | total infra area: %.2f km^2",
                sum(land_leases$lease_area_km2, na.rm = TRUE),
                sum(land_leases$infra_area_km2, na.rm = TRUE)))
```

# 4. Load + standardise marine leases (single canonical source)

```{r load-marine-leases}
marine_shp_path <- here("output_data", "all_marine_leases_with_status.shp")

marine_raw <- st_read(marine_shp_path, quiet = TRUE) %>%
  st_transform(4326) %>%
  st_make_valid()

message("Marine leases loaded: ", nrow(marine_raw))
message("Columns: ", paste(names(marine_raw), collapse = ", "))

# Safety filters (redundant with Script 4 but defensive)
marine_raw <- marine_raw %>%
  filter(!grepl("pearl", Spcs_st, ignore.case = TRUE)) %>%
  filter(!grepl("^Northern", stat_ls, ignore.case = TRUE)) %>%
  filter(stat_ls != "NT")

message("After pearl + NT safety filter: ", nrow(marine_raw))
```

```{r marine-leases}
col_m_species <- pick_col(marine_raw, c("Species_standardized", "Spcs_st"))
col_m_group   <- pick_col(marine_raw, c("Group"))
col_m_state   <- pick_col(marine_raw, c("state_lease", "stat_ls"))
col_m_status  <- pick_col(marine_raw, c("Infra_status", "Infr_st"))
col_m_leaseid <- pick_col(marine_raw, c("lease_id", "lease_d"))
col_m_region  <- pick_col(marine_raw, c("region"))

marine_leases_pre <- marine_raw %>%
  transmute(
    source_lease_id = as.character(.data[[col_m_leaseid]]),
    realm    = "Marine",
    state    = clean_state(.data[[col_m_state]]),
    status_upstream  = clean_status(.data[[col_m_status]]),
    species_raw = .data[[col_m_species]],
    group_raw   = .data[[col_m_group]],
    region   = if (!is.na(col_m_region)) as.character(.data[[col_m_region]])
               else NA_character_,
    geometry = geometry
  ) %>%
  mutate(
    species    = clean_species(species_raw),
    group      = clean_group(group_raw),
    subgroup   = compute_subgroup(group, species),
    state_full = unname(abbr_to_full[state])
  )

# Recompute lease area from geometry
marine_leases_pre$lease_area_km2 <- area_km2(marine_leases_pre)

marine_leases_pre <- marine_leases_pre %>%
  mutate(lease_id = paste0("ML_", sprintf("%05d", row_number())))

message(sprintf("Marine leases standardised: %d | total lease area: %.2f km^2",
                nrow(marine_leases_pre),
                sum(marine_leases_pre$lease_area_km2, na.rm = TRUE)))
```

# 5. Load marine detections

```{r load-detections}
det_ll_path <- here("output_data", "SSRG_GEE_results",
                    "SSRG_shellfish_longlines_marine.shp")
det_cg_path <- here("output_data", "SSRG_GEE_results",
                    "SSRG_finfish_cages_marine.shp")

det_ll <- st_read(det_ll_path, quiet = TRUE) %>%
  st_transform(4326) %>% st_make_valid() %>%
  mutate(infra_type_raw = "Longline")

det_cg <- st_read(det_cg_path, quiet = TRUE) %>%
  st_transform(4326) %>% st_make_valid() %>%
  mutate(infra_type_raw = "Cage")

detections_raw <- bind_rows(
  det_ll %>% select(infra_type_raw, geometry),
  det_cg %>% select(infra_type_raw, geometry)
) %>%
  st_make_valid()

detections_raw$infra_area_km2 <- area_km2(detections_raw)
detections_raw$detection_id   <- paste0("MD_",
                                 sprintf("%05d", seq_len(nrow(detections_raw))))

message(sprintf("Detections loaded: %d | total infra area: %.2f km^2",
                nrow(detections_raw),
                sum(detections_raw$infra_area_km2, na.rm = TRUE)))
```
#5B . load land-based pond detections
```{r load-land-detections}
land_ponds_path <- here("output_data", "Land_Aquaculture_Vectors",
                        "Australia_Land_Aquaculture_Ponds_National.gpkg")

land_ponds_raw <- st_read(land_ponds_path, quiet = TRUE)
st_geometry(land_ponds_raw) <- "geometry"   # standardise col name from gpkg
land_ponds_raw <- land_ponds_raw %>%
  st_transform(4326) %>%
  st_make_valid()

message(sprintf("Land ponds loaded: %d | total infra area: %.2f km^2",
                nrow(land_ponds_raw),
                sum(land_ponds_raw$Infra_Area_KM, na.rm = TRUE)))
```

# 6. Spatial join: each detection → its containing lease

```{r spatial-join}

# Each detection gets the lease_id of the lease it falls inside.
# Ties (detection overlapping two leases) are broken by largest overlap area.
det_lease_idx <- st_intersects(detections_raw, marine_leases_pre)

n_matched <- sum(lengths(det_lease_idx) > 0)
message(sprintf("  Detections intersecting a lease: %d of %d (%.1f%%)",
                n_matched, nrow(detections_raw),
                100 * n_matched / nrow(detections_raw)))

# Resolve ties: for detections overlapping multiple leases, pick the lease
# whose intersection area is largest.
resolve_lease_id <- function(det_i, candidate_ix) {
  if (length(candidate_ix) == 0) return(NA_character_)
  if (length(candidate_ix) == 1) return(marine_leases_pre$lease_id[candidate_ix])
  # compute overlap area against each candidate
  det_geom <- st_geometry(detections_raw)[det_i]
  overlaps <- vapply(candidate_ix, function(j) {
    inter <- suppressWarnings(st_intersection(det_geom,
                                   st_geometry(marine_leases_pre)[j]))
    if (length(inter) == 0) 0 else
      as.numeric(sum(st_area(st_transform(inter, 3577))))
  }, numeric(1))
  marine_leases_pre$lease_id[candidate_ix[which.max(overlaps)]]
}

detections_raw$lease_id <- mapply(resolve_lease_id,
                                  seq_len(nrow(detections_raw)),
                                  det_lease_idx,
                                  USE.NAMES = FALSE)

message(sprintf("  Detections with resolved lease_id: %d of %d",
                sum(!is.na(detections_raw$lease_id)),
                nrow(detections_raw)))
```

# 7. Aggregate detections back onto marine leases

```{r aggregate-marine}
# Per-lease aggregates from detections
lease_det_agg <- detections_raw %>%
  st_drop_geometry() %>%
  filter(!is.na(lease_id)) %>%
  group_by(lease_id) %>%
  summarise(
    infra_area_km2 = sum(infra_area_km2, na.rm = TRUE),
    # modal infra_type
    infra_type = {
      tps <- infra_type_raw[!is.na(infra_type_raw)]
      if (length(tps) == 0) NA_character_
      else names(sort(table(tps), decreasing = TRUE))[1]
    },
    n_detections = n(),
    .groups = "drop"
  )

marine_leases <- marine_leases_pre %>%
  left_join(lease_det_agg, by = "lease_id") %>%
  mutate(
    infra_area_km2 = coalesce(infra_area_km2, 0),
    # Status is recomputed from detection overlap for consistency with detections
    status         = if_else(infra_area_km2 > 0, "Active", "Inactive"),
    infra_type     = if_else(is.na(infra_type), "None", infra_type)
  )

# Consistency check against upstream status
if ("status_upstream" %in% names(marine_leases)) {
  mismatch <- sum(marine_leases$status != marine_leases$status_upstream,
                  na.rm = TRUE)
  if (mismatch > 0) {
    message(sprintf(
      "  NOTE: %d marine leases have status recomputed differently from upstream.",
      mismatch))
  }
}

marine_leases <- marine_leases %>%
  select(lease_id, source_lease_id, realm, state, state_full,
         group, subgroup, species, status, infra_type,
         lease_area_km2, infra_area_km2, geometry)

message(sprintf("\nMarine leases finalised: %d", nrow(marine_leases)))
message(sprintf("  Active:   %d (lease area %.2f km^2, infra %.2f km^2)",
                sum(marine_leases$status == "Active"),
                sum(marine_leases$lease_area_km2[marine_leases$status == "Active"]),
                sum(marine_leases$infra_area_km2[marine_leases$status == "Active"])))
message(sprintf("  Inactive: %d (lease area %.2f km^2)",
                sum(marine_leases$status == "Inactive"),
                sum(marine_leases$lease_area_km2[marine_leases$status == "Inactive"])))
```

# 8. Build final `leases` layer

```{r combine-leases}

# Unify columns: both tables must have the same schema, but they use
# different source-id column names. Drop those and keep the canonical lease_id.
common_cols <- c("lease_id", "realm", "state", "state_full",
                 "group", "subgroup", "species", "status", "infra_type",
                 "lease_area_km2", "infra_area_km2", "geometry")

leases <- rbind(
  land_leases[,   common_cols],
  marine_leases[, common_cols]
)

message(sprintf("\nCombined leases layer: %d (Land %d + Marine %d)",
                nrow(leases),
                sum(leases$realm == "Land-based"),
                sum(leases$realm == "Marine")))
```

# 9. Build final detections layer

```{r detections-layer}
# --- Marine detections ------------------------------------------------------
lease_attrs_marine <- marine_leases %>%
  st_drop_geometry() %>%
  select(lease_id, state, state_full, group, subgroup, species)

marine_detections <- detections_raw %>%
  left_join(lease_attrs_marine, by = "lease_id") %>%
  mutate(
    realm      = "Marine",
    infra_type = clean_infra_type(infra_type_raw),
    state      = coalesce(state, "Unknown"),
    state_full = coalesce(state_full, NA_character_),
    group      = coalesce(group, "Unknown"),
    species    = coalesce(species, "Unknown")
  ) %>%
  select(detection_id, lease_id, realm, state, state_full,
         group, subgroup, species, infra_type,
         infra_area_km2, geometry)

# --- Land-based detections --------------------------------------------------
# Link each pond to its parent land lease via SiteID ↔ source_site_id
lease_attrs_land <- land_leases %>%
  st_drop_geometry() %>%
  select(lease_id, source_site_id,
         state, state_full, group, subgroup, species)

land_detections <- land_ponds_raw %>%
  mutate(
    detection_id   = as.character(DetectID),
    source_site_id = as.character(SiteID),
    infra_type_raw = Infra_type,
    infra_area_km2 = as.numeric(Infra_Area_KM)
  ) %>%
  left_join(lease_attrs_land, by = "source_site_id") %>%
  mutate(
    realm      = "Land-based",
    infra_type = clean_infra_type(infra_type_raw),
    state      = coalesce(state, "Unknown"),
    state_full = coalesce(state_full, NA_character_),
    group      = coalesce(group, "Unknown"),
    species    = coalesce(species, "Unknown")
  ) %>%
  select(detection_id, lease_id, realm, state, state_full,
         group, subgroup, species, infra_type,
         infra_area_km2, geometry)

# --- Combine ----------------------------------------------------------------
detections <- rbind(marine_detections, land_detections)

message(sprintf("Detections layer: %d total | Marine %d + Land-based %d",
                nrow(detections),
                sum(detections$realm == "Marine"),
                sum(detections$realm == "Land-based")))
message(sprintf("  with lease_id: %d (%.1f%%)",
                sum(!is.na(detections$lease_id)),
                100 * mean(!is.na(detections$lease_id))))
```

# 10. Save clean GeoPackage

```{r save-gpkg}

output_dir <- here("output_data", "summaries")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

gpkg_path <- file.path(output_dir, "AU_aquaculture_clean.gpkg")
if (file.exists(gpkg_path)) file.remove(gpkg_path)

st_write(leases,     gpkg_path, layer = "leases",     quiet = TRUE)
st_write(detections, gpkg_path, layer = "detections", append = TRUE, quiet = TRUE)

message("Saved: ", gpkg_path)
message(sprintf("  leases layer:     %d features", nrow(leases)))
message(sprintf("  detections layer: %d features", nrow(detections)))
```

# 11. Build summary sheets from the cleaned data

All sheets below are built from the same `leases` + `detections` in-memory
objects, so they cannot drift from the gpkg.

```{r summaries}
lg <- leases     %>% st_drop_geometry()
dg <- detections %>% st_drop_geometry()

# --- Sheet: Area by Status ---------------------------------------------------
land_rows   <- lg %>% filter(realm == "Land-based")
marine_rows <- lg %>% filter(realm == "Marine")

area_by_status <- bind_rows(
  tibble(Sector = "Land-based", Status = "Active",
         N = sum(land_rows$status == "Active"),
         Lease_area_km2 = round(sum(land_rows$lease_area_km2[
                           land_rows$status == "Active"]), 3),
         Detected_infra_km2 = round(sum(land_rows$infra_area_km2[
                           land_rows$status == "Active"]), 3)),
  tibble(Sector = "Land-based", Status = "Inactive",
         N = sum(land_rows$status == "Inactive"),
         Lease_area_km2 = round(sum(land_rows$lease_area_km2[
                           land_rows$status == "Inactive"]), 3),
         Detected_infra_km2 = 0),
  tibble(Sector = "Marine", Status = "Infrastructure detected",
         N = sum(marine_rows$status == "Active"),
         Lease_area_km2 = round(sum(marine_rows$lease_area_km2[
                           marine_rows$status == "Active"]), 3),
         Detected_infra_km2 = round(sum(marine_rows$infra_area_km2[
                           marine_rows$status == "Active"]), 3)),
  tibble(Sector = "Marine", Status = "No infrastructure detected",
         N = sum(marine_rows$status == "Inactive"),
         Lease_area_km2 = round(sum(marine_rows$lease_area_km2[
                           marine_rows$status == "Inactive"]), 3),
         Detected_infra_km2 = NA_real_)
)

# --- Sheet: Land by State (active/inactive) ---------------------------------
land_by_state_full <- land_rows %>%
  group_by(State = state) %>%
  summarise(
    n_active_sites     = sum(status == "Active"),
    active_lease_km2   = round(sum(lease_area_km2[status == "Active"]), 3),
    active_infra_km2   = round(sum(infra_area_km2[status == "Active"]), 3),
    n_inactive_sites   = sum(status == "Inactive"),
    inactive_lease_km2 = round(sum(lease_area_km2[status == "Inactive"]), 3),
    total_sites        = n(),
    .groups = "drop"
  ) %>%
  mutate(
    inactive_pct_sites = round(n_inactive_sites / total_sites * 100, 1),
    inactive_pct_area  = round(inactive_lease_km2 /
                           (active_lease_km2 + inactive_lease_km2) * 100, 1)
  ) %>%
  arrange(desc(inactive_lease_km2))

# --- Sheet: Marine by State (active/inactive) -------------------------------
marine_by_state_full <- marine_rows %>%
  group_by(State = state) %>%
  summarise(
    n_active_leases   = sum(status == "Active"),
    active_area_km2   = round(sum(lease_area_km2[status == "Active"]), 3),
    n_inactive_leases = sum(status == "Inactive"),
    inactive_area_km2 = round(sum(lease_area_km2[status == "Inactive"]), 3),
    total_leases      = n(),
    total_area_km2    = round(sum(lease_area_km2), 3),
    .groups = "drop"
  ) %>%
  mutate(
    inactive_pct_leases = round(n_inactive_leases / total_leases * 100, 1),
    inactive_pct_area   = round(inactive_area_km2 / total_area_km2 * 100, 1)
  ) %>%
  arrange(desc(inactive_area_km2))

# --- Sheet: Land National (species-level) -----------------------------------
land_national <- land_rows %>%
  filter(status == "Active", infra_area_km2 > 0) %>%
  group_by(species_display = species) %>%
  summarise(
    n_sites       = n(),
    land_area_km2 = round(sum(infra_area_km2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(desc(land_area_km2))

# --- Sheet: Land by State (species × state) ---------------------------------
land_by_state <- land_rows %>%
  filter(status == "Active", infra_area_km2 > 0) %>%
  group_by(species_display = species, State = state) %>%
  summarise(
    n_sites       = n(),
    land_area_km2 = round(sum(infra_area_km2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(species_display, desc(land_area_km2))

# --- Sheet: Marine Infra National (species-level, detection-based) ----------
marine_infra_national <- dg %>%
  filter(realm == "Marine") %>%     
  group_by(Species_standardized = species) %>%
  summarise(
    n_detections    = n(),
    total_infra_km2 = round(sum(infra_area_km2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(desc(total_infra_km2))

# --- Sheet: Marine Infra by State (species × state) -------------------------
marine_infra_by_state <- dg %>%
  filter(realm == "Marine") %>%     
  group_by(state_lease = state, Species_standardized = species) %>%
  summarise(
    n_detections    = n(),
    total_infra_km2 = round(sum(infra_area_km2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(state_lease, desc(total_infra_km2))

# --- Sheet: Marine Lease National (species-level) ---------------------------
marine_lease_national <- marine_rows %>%
  group_by(species) %>%
  summarise(
    n_leases       = n(),
    lease_area_km2 = round(sum(lease_area_km2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(desc(lease_area_km2))

# --- Sheet: Marine Lease by State -------------------------------------------
marine_lease_by_state <- marine_rows %>%
  group_by(species, state) %>%
  summarise(
    n_leases       = n(),
    lease_area_km2 = round(sum(lease_area_km2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(species, desc(lease_area_km2))
```

# 12. Build the Manuscript Table

```{r manuscript-table}

# --- ABARES 2021-22 production (unchanged from Script 11) -------------------
abares_production <- tribble(
  ~species_manuscript,        ~production_aud, ~production_tonnes,
  "Southern Bluefin Tuna",    110400,          8322,
  "Other finfish (marine)",   NA,              NA,
  "Oysters",                  149671,          11735,
  "Mussels",                  4984,            2173,
  "Abalone",                  26285,           689,
  "Yabbies",                  752,             21,
  "Marron",                   4511,            113,
  "Cherax spp.",              NA,              NA,
  "Red claw",                 850,             31,
  "Prawns",                   180821,          9450,
  "Barramundi",               60572,           4741,
  "Salmon",                   1149408,         79271,
  "Native finfish (land)",    NA,              NA,
  "Murray Cod",               NA,              NA,
  "Lobster",                  NA,              NA,
  "Unknown (land-based)",     NA,              NA,
  "Hatcheries (land-based)",  NA,              NA,
  "Multiple species (land)",  NA,              NA,
  "Aquarium",                 NA,              NA,
  "Rainbow Trout",            NA,              NA
)

# --- Species mappings (unchanged) -------------------------------------------
land_species_mapping <- tribble(
  ~species_display,    ~species_manuscript,
  "Prawn",             "Prawns",
  "Cherax spp.",       "Cherax spp.",
  "Unknown species",   "Unknown (land-based)",
  "Unknown",           "Unknown (land-based)",
  "Barramundi",        "Barramundi",
  "Red Claw Crayfish", "Red claw",
  "Native finfish",    "Native finfish (land)",
  "Hatchery",          "Hatcheries (land-based)",
  "Multiple species",  "Multiple species (land)",
  "Abalone",           "Abalone",
  "Murray Cod",        "Murray Cod",
  "aquarium",          "Aquarium",
  "Salmon",            "Salmon",
  "Yabby",             "Yabbies",
  "Rainbow Trout",     "Rainbow Trout",
  "lobster",           "Lobster",
  "Marine Finfish (Cobia, Groupers)", "Other finfish (marine)"
)

marine_species_mapping <- tribble(
  ~marine_species,                                ~species_manuscript,
  "Edible Oyster (aggregated)",                   "Oysters",
  "Pacific Oyster",                               "Oysters",
  "Pacific Oyster or Blue Mussel",                "Oysters",
  "Sydney Rock Oyster (predominant)",             "Oysters",
  "Edible Oyster (Albany/Oyster Harbour)",        "Oysters",
  "Mussel (aggregated)",                          "Mussels",
  "Blue Mussel",                                  "Mussels",
  "Atlantic Salmon",                              "Salmon",
  "Southern Bluefin Tuna (aggregated)",           "Southern Bluefin Tuna",
  "Southern Bluefin Tuna or Yellowtail Kingfish", "Southern Bluefin Tuna",
  "Finfish",                                      "Other finfish (marine)",
  "Barramundi (Cone Bay - Marine)",               "Barramundi",
  "Abalone",                                      "Abalone"
)

# --- Build per-species states strings ---------------------------------------
land_states_str <- land_by_state %>%
  group_by(species_display) %>%
  summarise(
    land_states_detail = paste0(State, " (", land_area_km2, " km2)",
                                collapse = ", "),
    land_states_list   = paste0(sort(unique(State)), collapse = ", "),
    .groups = "drop"
  )

marine_infra_states_str <- marine_infra_by_state %>%
  group_by(Species_standardized) %>%
  summarise(
    marine_infra_states_detail = paste0(state_lease, " (",
                                   total_infra_km2, " km2)", collapse = ", "),
    marine_infra_states_list   = paste0(sort(unique(state_lease)),
                                         collapse = ", "),
    .groups = "drop"
  )

# --- Build the manuscript table ---------------------------------------------
land_for_table <- land_national %>%
  left_join(land_species_mapping, by = "species_display") %>%
  left_join(land_states_str,      by = "species_display") %>%
  filter(!is.na(species_manuscript)) %>%
  group_by(species_manuscript) %>%
  summarise(
    land_n_sites       = sum(n_sites,       na.rm = TRUE),
    land_infra_km2     = round(sum(land_area_km2, na.rm = TRUE), 3),
    land_states_detail = paste(na.omit(unique(land_states_detail)),
                                collapse = " | "),
    land_states_list   = paste(na.omit(unique(land_states_list)),
                                collapse = ", "),
    .groups = "drop"
  )

marine_infra_for_table <- marine_infra_national %>%
  left_join(marine_species_mapping,
            by = c("Species_standardized" = "marine_species")) %>%
  filter(!is.na(species_manuscript)) %>%
  group_by(species_manuscript) %>%
  summarise(
    marine_infra_n   = sum(n_detections,    na.rm = TRUE),
    marine_infra_km2 = round(sum(total_infra_km2, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  left_join(
    marine_infra_states_str %>%
      left_join(marine_species_mapping,
                by = c("Species_standardized" = "marine_species")) %>%
      filter(!is.na(species_manuscript)) %>%
      group_by(species_manuscript) %>%
      summarise(
        marine_infra_states_detail = paste(
          na.omit(unique(marine_infra_states_detail)), collapse = " | "),
        marine_infra_states_list = paste(
          na.omit(unique(marine_infra_states_list)),   collapse = ", "),
        .groups = "drop"
      ),
    by = "species_manuscript"
  )

marine_lease_for_table <- marine_lease_national %>%
  left_join(marine_species_mapping,
            by = c("species" = "marine_species")) %>%
  filter(!is.na(species_manuscript)) %>%
  group_by(species_manuscript) %>%
  summarise(
    marine_lease_n   = sum(n_leases,       na.rm = TRUE),
    marine_lease_km2 = round(sum(lease_area_km2, na.rm = TRUE), 3),
    .groups = "drop"
  )

manuscript_table <- abares_production %>%
  left_join(land_for_table,         by = "species_manuscript") %>%
  left_join(marine_infra_for_table, by = "species_manuscript") %>%
  left_join(marine_lease_for_table, by = "species_manuscript") %>%
  mutate(
    infra_type = case_when(
      !is.na(land_infra_km2) & !is.na(marine_infra_km2) ~ "Land + Marine",
      !is.na(land_infra_km2)                             ~ "Land-based",
      !is.na(marine_infra_km2)                           ~ "Marine",
      TRUE                                               ~ NA_character_
    ),
    total_infra_km2 = round(
      coalesce(land_infra_km2, 0) + coalesce(marine_infra_km2, 0), 3
    ),
    total_infra_km2 = if_else(total_infra_km2 == 0,
                              NA_real_, total_infra_km2),
    states_detail = case_when(
      !is.na(land_states_detail) & !is.na(marine_infra_states_detail) ~
        paste0("Land: ", land_states_detail,
               " | Marine: ", marine_infra_states_detail),
      !is.na(land_states_detail)         ~ land_states_detail,
      !is.na(marine_infra_states_detail) ~ marine_infra_states_detail,
      TRUE ~ NA_character_
    ),
    all_states_list = map2_chr(
      coalesce(land_states_list,        ""),
      coalesce(marine_infra_states_list, ""),
      ~ {
        all <- unique(trimws(unlist(str_split(paste(.x, .y, sep = ", "), ","))))
        all <- all[all != ""]
        if (length(all) == 0) NA_character_ else paste(sort(all), collapse = ", ")
      }
    ),
    Notes = case_when(
      species_manuscript == "Cherax spp." ~
        "Combined Yabbies + Marron (not distinguishable by satellite)",
      species_manuscript %in% c("Yabbies", "Marron") ~
        "Subset of Cherax spp. (ABARES only)",
      TRUE ~ NA_character_
    )
  ) %>%
  select(
    Species                    = species_manuscript,
    `Infra type`               = infra_type,
    `Total infra area (km2)`   = total_infra_km2,
    `Land infra (km2)`         = land_infra_km2,
    `Marine infra (km2)`       = marine_infra_km2,
    `Lease area (km2)`         = marine_lease_km2,
    `Production 2021-22 (AUD)` = production_aud,
    `Production 2021-22 (t)`   = production_tonnes,
    States                     = all_states_list,
    `States detail`            = states_detail,
    Notes
  )

print(manuscript_table, n = 30, width = 150)
```

# 13. Save xlsx summaries

```{r save-xlsx}
xlsx_path <- file.path(output_dir, "combined_aquaculture_species_summaries.xlsx")

write_xlsx(
  list(
    "Manuscript Table"                   = as.data.frame(manuscript_table),
    "Area by Status"                     = as.data.frame(area_by_status),
    "Land by State (active-inactive)"    = as.data.frame(land_by_state_full),
    "Marine by State (active-inactive)"  = as.data.frame(marine_by_state_full),
    "Land National"                      = as.data.frame(land_national),
    "Land by State"                      = as.data.frame(land_by_state),
    "Marine Infra National"              = as.data.frame(marine_infra_national),
    "Marine Infra by State"              = as.data.frame(marine_infra_by_state),
    "Marine Lease National"              = as.data.frame(marine_lease_national),
    "Marine Lease by State"              = as.data.frame(marine_lease_by_state)
  ),
  path = xlsx_path
)

message("\nSaved: ", xlsx_path)
```


# 14. Validation — print totals from the cleaned objects

```{r validate}
# --- Helpers -----------------------------------------------------------------
# Count leases matching realm + status
n_leases <- function(realm_val = NULL, status_val = NULL) {
  x <- leases
  if (!is.null(realm_val))  x <- x[x$realm  == realm_val, ]
  if (!is.null(status_val)) x <- x[x$status == status_val, ]
  nrow(x)
}

# Sum a numeric column on leases matching realm + status
sum_col <- function(col, realm_val = NULL, status_val = NULL) {
  x <- leases
  if (!is.null(realm_val))  x <- x[x$realm  == realm_val, ]
  if (!is.null(status_val)) x <- x[x$status == status_val, ]
  sum(x[[col]], na.rm = TRUE)
}

cat("\n========= HEADLINE NUMBERS (from cleaned gpkg) =========\n")

# --- Lease counts ------------------------------------------------------------
cat("\n--- LEASE COUNTS ---\n")
cat(sprintf("Total leases:            %d\n",  n_leases()))
cat(sprintf("  Land:                  %d\n",  n_leases("Land-based")))
cat(sprintf("  Marine:                %d\n",  n_leases("Marine")))

cat(sprintf("Active leases:           %d\n",  n_leases(status_val = "Active")))
cat(sprintf("  Land active:           %d\n",  n_leases("Land-based", "Active")))
cat(sprintf("  Marine active:         %d\n",  n_leases("Marine",     "Active")))

cat(sprintf("Inactive leases:         %d\n",  n_leases(status_val = "Inactive")))
cat(sprintf("  Land inactive:         %d\n",  n_leases("Land-based", "Inactive")))
cat(sprintf("  Marine inactive:       %d\n",  n_leases("Marine",     "Inactive")))

# --- Lease area --------------------------------------------------------------
cat("\n--- LEASE AREA (km^2) ---\n")
cat(sprintf("Total lease area:        %.2f\n", sum_col("lease_area_km2")))
cat(sprintf("  Land:                  %.2f\n", sum_col("lease_area_km2", "Land-based")))
cat(sprintf("  Marine:                %.2f\n", sum_col("lease_area_km2", "Marine")))

cat(sprintf("Active lease area:       %.2f\n",
            sum_col("lease_area_km2", status_val = "Active")))
cat(sprintf("  Land active:           %.2f\n",
            sum_col("lease_area_km2", "Land-based", "Active")))
cat(sprintf("  Marine active:         %.2f\n",
            sum_col("lease_area_km2", "Marine",     "Active")))

cat(sprintf("Inactive lease area:     %.2f\n",
            sum_col("lease_area_km2", status_val = "Inactive")))
cat(sprintf("  Land inactive:         %.2f\n",
            sum_col("lease_area_km2", "Land-based", "Inactive")))
cat(sprintf("  Marine inactive:       %.2f\n",
            sum_col("lease_area_km2", "Marine",     "Inactive")))

# --- Infra area (from leases table) ------------------------------------------
cat("\n--- INFRA AREA (km^2, from leases table) ---\n")
cat(sprintf("Total infra area:        %.2f\n", sum_col("infra_area_km2")))
cat(sprintf("  Land:                  %.2f\n", sum_col("infra_area_km2", "Land-based")))
cat(sprintf("  Marine:                %.2f\n", sum_col("infra_area_km2", "Marine")))

# --- Detections --------------------------------------------------------------
cat("\n--- DETECTIONS ---\n")
cat(sprintf("Total detections:        %d\n", nrow(detections)))
cat(sprintf("  Land:                  %d\n",
            sum(detections$realm == "Land-based")))
cat(sprintf("  Marine:                %d\n",
            sum(detections$realm == "Marine")))
cat(sprintf("  with lease_id:         %d\n",
            sum(!is.na(detections$lease_id))))
cat(sprintf("  orphaned:              %d\n",
            sum(is.na(detections$lease_id))))
cat(sprintf("Total detection area:    %.2f km^2\n",
            sum(detections$infra_area_km2, na.rm = TRUE)))
cat(sprintf("  Land detection area:   %.2f km^2\n",
            sum(detections$infra_area_km2[detections$realm == "Land-based"],
                na.rm = TRUE)))
cat(sprintf("  Marine detection area: %.2f km^2\n",
            sum(detections$infra_area_km2[detections$realm == "Marine"],
                na.rm = TRUE)))

# --- Cross-check: detection area should ≈ lease-side infra area by realm -----
cat("\n--- CROSS-CHECK (detection-side vs lease-side infra area) ---\n")
for (r in c("Land-based", "Marine")) {
  d_side <- sum(detections$infra_area_km2[detections$realm == r], na.rm = TRUE)
  l_side <- sum_col("infra_area_km2", r)
  cat(sprintf("  %s: detections=%.2f km^2, leases=%.2f km^2, diff=%.3f\n",
              r, d_side, l_side, d_side - l_side))
}
```
```{r}
# --- Productivity: detected infra as % of leased area -----------------------
cat("\n--- PRODUCTIVITY (detected infra / lease area) ---\n")
for (r in c("Land-based", "Marine")) {
  infra_area        <- sum(detections$infra_area_km2[detections$realm == r],
                           na.rm = TRUE)
  all_lease_area    <- sum_col("lease_area_km2", r)
  active_lease_area <- sum_col("lease_area_km2", r, "Active")
  cat(sprintf("  %s:\n", r))
  cat(sprintf("    infra / all leased:    %7.2f / %7.2f km^2 = %5.2f%%\n",
              infra_area, all_lease_area,    100 * infra_area / all_lease_area))
  cat(sprintf("    infra / active leased: %7.2f / %7.2f km^2 = %5.2f%%\n",
              infra_area, active_lease_area, 100 * infra_area / active_lease_area))
}

# --- Detection density (mean detections per active lease) -------------------
cat("\n--- DETECTION DENSITY (detections per active lease) ---\n")
for (r in c("Land-based", "Marine")) {
  n_det    <- sum(detections$realm == r, na.rm = TRUE)
  n_active <- n_leases(r, "Active")
  cat(sprintf("  %s: %d detections / %d active leases = %.2f per lease\n",
              r, n_det, n_active, n_det / n_active))
}

# --- Activity rate (% of leases that are active) ----------------------------
cat("\n--- ACTIVITY RATE (% of leases with detected infra) ---\n")
for (r in c("Land-based", "Marine")) {
  n_total     <- n_leases(r)
  n_active    <- n_leases(r, "Active")
  area_total  <- sum_col("lease_area_km2", r)
  area_active <- sum_col("lease_area_km2", r, "Active")
  cat(sprintf("  %s:\n", r))
  cat(sprintf("    by count: %d / %d = %.1f%% active\n",
              n_active, n_total, 100 * n_active / n_total))
  cat(sprintf("    by area:  %.2f / %.2f km^2 = %.1f%% active\n",
              area_active, area_total, 100 * area_active / area_total))
}
```

# 15. Data dictionary

| Layer        | Column          | Type   | Description |
|--------------|-----------------|--------|-------------|
| leases       | lease_id        | char   | Unique ID (`LL_*` land, `ML_*` marine) |
| leases       | realm           | char   | Marine / Land-based |
| leases       | state           | char   | State abbreviation |
| leases       | state_full      | char   | State full name |
| leases       | group           | char   | Fish / Mollusc / Crustacean / Algae / Echinoderm / Unknown |
| leases       | subgroup        | char   | Bivalve / Gastropod / NA (only for Molluscs) |
| leases       | species         | char   | Common species name or "Unknown" |
| leases       | status          | char   | Active (infra detected) / Inactive |
| leases       | infra_type      | char   | Pond / Cage / Longline / Tank-RAS / None / Unknown |
| leases       | lease_area_km2  | num    | Lease polygon area, EPSG:3577 |
| leases       | infra_area_km2  | num    | Detected infrastructure area inside the lease |
| detections   | detection_id    | char   | Unique ID (`MD_*`) |
| detections   | lease_id        | char   | FK to `leases.lease_id` (NA if orphaned) |
| detections   | infra_type      | char   | Cage / Longline |
| detections   | infra_area_km2  | num    | Detection polygon area, EPSG:3577 |
