---
title: "Manuscript tables — aquaculture production mapping"
author: "Marina"
date: "`r Sys.Date()`"
Author: Marina Christofidis (marina.christofidis@griffithuni.edu.au and marina.christofidis@gmail.com)
#https://www.sustainableseasgroup.com/
#These Scripts are relative to Version 1 of Australia's Aquaculture Mapping 
#GitHub: Marina6578/Australias_Aquaculture

output: html_document
---

## Overview

This script builds all main-text and supplementary tables for the aquaculture
production mapping manuscript from a single set of source files:

- `combined_aquaculture_species_summaries.xlsx` — sector/species summaries
- `concentration_gini_by_species_clean.csv` — species concentration (Gini)

All outputs are written to `output_data/summaries/`. Rerun the chunks below
whenever the underlying data is updated; the tables regenerate automatically.

### Main-text table decision

Two options are provided for the species-level main-text table(s). Run
both, eyeball the outputs in Word, and pick one:

- **Option A** — separate Table 3 (concentration) and Table 4 (production)
- **Option B** — single combined Table 3 (concentration + production, 9 rows)

Both are built below; keep whichever you prefer in the final manuscript and
delete the other chunk(s).

## Setup

```{r setup, message=FALSE, warning=FALSE}
library(here)
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(tibble)

# Source file paths — adjust if your project layout differs
xlsx_path <- here("output_data", "summaries",
                  "combined_aquaculture_species_summaries.xlsx")
gini_path <- here("output_data", "summaries",
                  "concentration_gini_by_species_clean.csv")

# Output folder
out_dir <- here("output_data", "summaries")
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
```

## Table 3 — Species concentration (Option A main text)

Paired with Figure 3. Gini values rounded to 2 dp; species capitalisation
normalised for publication.

```{r table3-concentration, message=FALSE}
table3_concentration <- read_csv(gini_path, show_col_types = FALSE) %>%
  mutate(
    gini_count = round(gini_count, 2),
    gini_area  = round(gini_area, 2),
    species = recode(species,
      "Edible Oyster (aggregated)"          = "Edible oyster (aggregated)",
      "Atlantic Salmon"                     = "Atlantic salmon",
      "Mussel (aggregated)"                 = "Mussel (aggregated)",
      "Red Claw Crayfish"                   = "Red claw crayfish",
      "Marine Finfish (Cobia, Groupers)"    = "Marine finfish (cobia, groupers)",
      "Multiple (Barramundi, Prawn)"        = "Polyculture (barramundi + prawn)",
      "Southern Bluefin Tuna (aggregated)"  = "Southern bluefin tuna (aggregated)",
      "Murray Cod"                          = "Murray cod",
      "aquarium"                            = "Aquarium",
      "Barramundi (Cone Bay - Marine)"      = "Barramundi (Cone Bay)",
      "Salmon"                              = "Salmon (land-based)"
    )
  ) %>%
  arrange(desc(total_infra_km2))

write_csv(table3_concentration, here(out_dir, "table3.csv"))
```

## Table 4 — Species × production summary (Option A main text)

Species-level integration of detected infrastructure area, lease area, and
ABARES-reported 2021–2022 production (tonnes and AUD).

```{r table4-species-production, message=FALSE}
species_production <- read_excel(xlsx_path, sheet = "Manuscript Table") %>%
  select(
    species              = Species,
    infra_type           = `Infra type`,
    total_infra_km2      = `Total infra area (km2)`,
    land_infra_km2       = `Land infra (km2)`,
    marine_infra_km2     = `Marine infra (km2)`,
    lease_area_km2       = `Lease area (km2)`,
    production_aud_2122  = `Production 2021-22 (AUD)`,
    production_t_2122    = `Production 2021-22 (t)`,
    states               = States,
    states_detail        = `States detail`,
    notes                = Notes
  ) %>%
  arrange(desc(total_infra_km2))

# Sum ABARES-reported yabbies + marron production into Cherax spp.
cherax_totals <- species_production %>%
  filter(species %in% c("Yabbies", "Marron")) %>%
  summarise(
    aud_sum = sum(production_aud_2122, na.rm = TRUE),
    t_sum   = sum(production_t_2122,   na.rm = TRUE)
  )

species_production <- species_production %>%
  mutate(
    production_aud_2122 = if_else(species == "Cherax spp.",
                                  cherax_totals$aud_sum,
                                  production_aud_2122),
    production_t_2122   = if_else(species == "Cherax spp.",
                                  cherax_totals$t_sum,
                                  production_t_2122),
    notes = if_else(
      species == "Cherax spp.",
      paste("Combined yabbies + marron (not distinguishable by satellite);",
            "ABARES production summed across both species"),
      notes
    )
  )

write_csv(species_production, here(out_dir, "table4.csv"))
```

## Combined Table 3 (Option B main text — alternative to A)

Single main-text table integrating concentration and production for the
9 commercial species with matching data in both sources. Barramundi is
land-based only (marine Cone Bay site had no 2021–22 ABARES production).

Run this chunk as an *alternative* to the separate Table 3 + Table 4 above.

```{r table3-combined, message=FALSE}
# --- Harmonise species names across the two sources ---

canonical_names <- tribble(
  ~canonical_species,       ~table3_species,                      ~production_species,
  "Edible oyster",          "Edible oyster (aggregated)",         "Oysters",
  "Atlantic salmon",        "Atlantic salmon",                    "Salmon",
  "Southern bluefin tuna",  "Southern bluefin tuna (aggregated)", "Southern Bluefin Tuna",
  "Mussel",                 "Mussel (aggregated)",                "Mussels",
  "Prawn",                  "Prawn",                              "Prawns",
  "Barramundi",             "Barramundi",                         "Barramundi",
  "Red claw crayfish",      "Red claw crayfish",                  "Red claw",
  "Cherax spp.",            "Cherax spp.",                        "Cherax spp.",
  "Abalone",                "Abalone",                            "Abalone"
)

# Pull concentration columns
t3_sub <- table3_concentration %>%
  inner_join(canonical_names, by = c("species" = "table3_species")) %>%
  select(canonical_species, realm,
         n_states_present, gini_count, gini_area,
         dominant_state, dominant_share_pct,
         total_infra_km2, n_detections_total)

# Pull production columns

t4_sub <- species_production %>%
  inner_join(canonical_names, by = c("species" = "production_species")) %>%
  select(canonical_species,
         lease_area_km2,
         production_t_2122,
         production_aud_2122,
         states_detail)            # <- added

table3_combined <- t3_sub %>%
  left_join(t4_sub, by = "canonical_species") %>%
  rename(species = canonical_species) %>%
  mutate(
    notes = if_else(
      species == "Barramundi",
      "Land-based only; marine site (Cone Bay, WA) excluded — no 2021–22 ABARES production",
      NA_character_
    )
  ) %>%
  select(realm, species,
         n_states_present, dominant_state, dominant_share_pct,
         gini_count, gini_area,
         total_infra_km2, lease_area_km2,
         states_detail,              # <- added, placed near the state columns
         production_t_2122, production_aud_2122,
         n_detections_total, notes) %>%
  arrange(desc(total_infra_km2))

# Combine and format
# ... canonical_names, t3_sub, t4_sub unchanged ...

table3_combined <- t3_sub %>%
  left_join(t4_sub, by = "canonical_species") %>%
  rename(species = canonical_species) %>%
  mutate(
    notes = if_else(
      species == "Barramundi",
      "Land-based only; marine site (Cone Bay, WA) excluded — no 2021–22 ABARES production",
      NA_character_
    )
  ) %>%
  # round all numeric columns to 2 dp
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
  # final column order + display names
  select(
    `realm`                          = realm,
    `species`                        = species,
    `number of states`               = n_states_present,
    `gini count`                     = gini_count,
    `gini area`                      = gini_area,
    `dominant state`                 = dominant_state,
    `dominant share pct`             = dominant_share_pct,
    `infra area (km2)`               = total_infra_km2,
    `lease area km2`                 = lease_area_km2,
    `production weight (tonnes) 21-22` = production_t_2122,
    `production AUD 21-22`           = production_aud_2122,
    `active leases`                  = n_detections_total,
    `states detail`                  = states_detail,
    `notes`                          = notes
  ) %>%
  arrange(desc(`infra area (km2)`))

write_csv(table3_combined,
          here(out_dir, "table3_combined_species_summary.csv"))
```


```{r table3-combined, message=FALSE}
## Supplementary — Sector × status lease summary

#realm level summary of leases and detected infrastructure by activity status.
#"Active" is defined as leases containing detected infrastructure (per Methods).
```


```{r table-s-sector-summary, message=FALSE}
sector_summary <- read_excel(xlsx_path, sheet = "Area by Status") %>%
  rename(realm = Sector) %>%
  mutate(
    Status = case_when(
      Status == "Infrastructure detected"    ~ "Active",
      Status == "No infrastructure detected" ~ "Inactive",
      TRUE                                   ~ Status
    ),
    Detected_infra_km2 = replace_na(Detected_infra_km2, 0)
  ) %>%
  select(
    realm,
    status             = Status,
    n_leases           = N,
    lease_area_km2     = Lease_area_km2,
    detected_infra_km2 = Detected_infra_km2
  )

totals <- sector_summary %>%
  group_by(realm) %>%
  summarise(
    status             = "All",
    n_leases           = sum(n_leases),
    lease_area_km2     = sum(lease_area_km2),
    detected_infra_km2 = sum(detected_infra_km2),
    .groups = "drop"
  )

sector_summary_out <- bind_rows(sector_summary, totals) %>%
  arrange(realm, factor(status, levels = c("Active", "Inactive", "All")))

write_csv(sector_summary_out,
          here(out_dir, "table_s_sector_lease_summary.csv"))
```

## Supplementary — Dormant lease capacity by state

Dormant capacity = leased area with no detected infrastructure. Computed
from the state-level active/inactive sheets for land-based and marine leases.

```{r table-s-dormant-capacity, message=FALSE}
land_state <- read_excel(xlsx_path, sheet = "Land by State (active-inactiv") %>%
  select(state = State, land_inactive_km2 = inactive_lease_km2)

marine_state <- read_excel(xlsx_path, sheet = "Marine by State (active-inact") %>%
  select(state = State, marine_inactive_km2 = inactive_area_km2)

dormant_capacity <- full_join(land_state, marine_state, by = "state") %>%
  mutate(across(c(land_inactive_km2, marine_inactive_km2),
                ~replace_na(.x, 0))) %>%
  mutate(total_dormant_km2 = land_inactive_km2 + marine_inactive_km2) %>%
  arrange(desc(total_dormant_km2))

dormant_capacity %>%
  mutate(pct_national = round(100 * total_dormant_km2 / sum(total_dormant_km2), 1))

dormant_capacity_out <- dormant_capacity %>%
  bind_rows(
    summarise(
      dormant_capacity,
      state = "Total",
      across(where(is.numeric), sum)
    )
  )

write_csv(dormant_capacity_out,
          here(out_dir, "dormant_lease_capacity_by_state.csv"))
```

## Supplementary — Land-based infrastructure by state × species

```{r table-s-land-by-state, message=FALSE}
land_by_state <- read_excel(xlsx_path, sheet = "Land by State") %>%
  select(
    species  = species_display,
    state    = State,
    n_sites,
    land_area_km2
  ) %>%
  arrange(state, desc(land_area_km2))

write_csv(land_by_state,
          here(out_dir, "table_s_land_by_state_species.csv"))
```

## Supplementary — Marine infrastructure by state × species

```{r table-s-marine-by-state, message=FALSE}
marine_by_state <- read_excel(xlsx_path, sheet = "Marine Infra by State") %>%
  select(
    state           = state_lease,
    species         = Species_standardized,
    n_detections,
    total_infra_km2
  ) %>%
  arrange(state, desc(total_infra_km2))

write_csv(marine_by_state,
          here(out_dir, "table_s_marine_by_state_species.csv"))
```

## Output summary

After running all chunks, `output_data/summaries/` contains:

| File                                      | Purpose                                  |
|-------------------------------------------|------------------------------------------|
| `table3.csv`                              | Option A — main Table 3 (concentration)  |
| `table4.csv`                              | Option A — main Table 4 (production)     |
| `table3_combined_species_summary.csv`     | Option B — single combined main table    |
| `table_s_sector_lease_summary.csv`        | Supplementary (sector × status)          |
| `dormant_lease_capacity_by_state.csv`     | Supplementary                            |
| `table_s_land_by_state_species.csv`       | Supplementary                            |
| `table_s_marine_by_state_species.csv`     | Supplementary                            |

Once the main-text option is chosen, delete the unused chunk(s) and the
corresponding CSV file(s) to keep the output folder tidy.

## Notes

- Species names with biological italics (e.g. *Cherax* spp.) are stored as
  plain text in CSVs; apply italic formatting when building the final Word
  tables.
- Marine figures exclude pearl oyster leases (see Methods — outside the
  study's food-aquaculture scope).
- If source spatial data (`AU_aquaculture_clean.gpkg`) is updated, rebuild
  the Excel summary first, then rerun this document.
